### 2014-08-15 drafts/ApproximatePCFG/ApproximatePCFG.tex
Initial draft commits.

\documentclass{article}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{amsmath}

\usepackage{algorithm}
\usepackage{algpseudocode}

\author{Eyal Dechter}

\begin{document}
The Inside-Outside algorithm is a popular algorithm for computing the
MAP marginals of PCFG rules given a sentence or collection of
sentences. The algorithm's computational complexity is $O(n^3 K)$
where $n$ is the length of the sentence and $K$ is the the number of
rules in the PCFG. The cubic algorithm makes inference tractable for
relatively long sentences -- containing perhaps a few dozen terminals
-- but makes it impossible for this algorithm to be used in an online
fashion, or to be used on grammars that span multiple levels of
linguistic abstration (for example, imagine trying to learn
morphological and syntactic grammar at the same time with the
Inside-Outside algorithm).

Our goal here is to develop an algorithm that approximates the MAP
marginals of a PCFG given data but that can run much faster; further,
we would like to be able to put bounds on the approximation. The core
intuition is that if we know that the probability of some symbol
dominating a span cannot be more than some very small amount, we
shouldn't have to calculate that amount accurately; instead, we should
be able to substitute this small upperbound in our calculations,
absolving us of exploring parts of the space of parses that are very
unlikely to be relevant.

In what follows we will assume that we have a PCFG $G$ in Chomsky
normal form with start symbol $S$, non-terminals $\mathcal{A}$,
terminals $\mathcal{T}$, and rules $\mathcal{R}$, where we associate
with each rule $r \in \mathcal{R}$ a potential $\psi(r)$. We also
assume that we are given a sentence $x = x_1, \dots, x_N$.

For every triple $(A, i, j)$ where $A$ is a nonterminal and $1 \leq i
\leq j \leq N$, the Inside-Outside algorithm computes an alpha and
beta value. The alpha and beta values are defined according to the
following recurrence:

\begin{align}
\alpha(A, i, i) &= \psi(A \rightarrow x_i)\\
\alpha(A, i, j) &= \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)\\
\beta(A, 1, N) &= 1 \\
\beta(A, i, j) &= \sum_{k=1}^{i-1} \sum_{B \rightarrow C A \in \mathcal{R}}
  \beta(B, k, j) \psi(B \rightarrow C A) \alpha(C, k, i-1)\\ 
  &+  \sum_{k=j+1}^{N} \sum_{B \rightarrow A C \in \mathcal{R} : A \neq C}
  \beta(B, i, k) \psi(B \rightarrow A C) \alpha(C, j+1, k) 
\end{align}

Suppose we have an approximate $\alpha'$ of $\alpha$ with the
property that for some small $\epsilon > 0$ and for all $A, i, j$:

\begin{align}
\alpha'(A, i, i) &= \alpha(A, i, i)\\ 
\alpha(A, i, j) &\leq \alpha'(A, i, j) \leq (1+\epsilon)  \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
  \psi(A \rightarrow B C ) \alpha'(B, i, k) \alpha'(C, k+1, j).\\
\end{align}

By how much can $\alpha'$ over estimate $\alpha$? Let $E(A, i,
j)=\frac{\alpha'(A, i, j)}{\alpha(A, i, j)}$ be the measure of this
over-estimate. We have that

\begin{align}
E_{\alpha}(A, i, i) &= 1\\
E_{\alpha}(A, i,j) &= \frac{\alpha'(A, i, j)}{\alpha(A, i, j)}\\
          &\leq (1+\epsilon) \frac{ \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
  \psi(A \rightarrow B C ) \alpha'(B, i, k) \alpha'(C, k+1, j)}
           {\sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)}\\
          &=  (1+\epsilon) \frac{ \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
  \psi(A \rightarrow B C ) E_{\alpha}(B, i, k) \alpha(B, i, k) E_{\alpha}(C, k+1, j) \alpha(C, k+1, j)}
           {\sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)}\\
         &= (1+\epsilon) \max_{k\in \{i, \dots, j-1\}, A \rightarrow B C \in \mathcal{R} }
            E_{\alpha}(B, i, k) E_{\alpha}(C, k+1, j).
\end{align}

Let $\hat{E}_{\alpha}(d) = \max_{A\in\mathcal{A}, i \in \{1, \dots, N-d \}}
E_{\alpha}(A, i, i+d)$ be the maximum upper bound above over all spans of
length $d$. Then we have that

\begin{align}
\hat{E}_{\alpha}(1) &= 1\\
\hat{E}_{\alpha}(d) &= (1+\epsilon) \max_{s \in \{1, \dots, d-1\}} \hat{E}_{\alpha}(s) \hat{E}_{\alpha}(d-s).\\
\end{align}

Solving this recurrence, gives $\hat{E}_{\alpha}(d) = (1 + \epsilon)^{d-1}$. To
re-iterate, what this means is that if we have such an $\alpha'$ then
for all $A, i, j$, $\alpha'(A, i, j) \leq (1 + \epsilon)^{j - i}
\alpha(A, i, j)$.

We can now develop an analog $\beta'$ of $\beta$ which has the same
dependency on $\alpha'$ that $\beta$ has on $\alpha$. Let
$\hat{E}_{\beta}(d)$ be an upper bound on $\frac{\beta'(A, i,
  i+d)}{\beta(A, i, i+d)}$ for all $A$ and $i$. Then a similar
analysis as above shows that one such $\hat{E}_\beta$ is

\begin{align}
\hat{E}_{\beta}(N) &= 1\\
\hat{E}_{\beta}(N-k) &= (1+\epsilon)^{N-k-1} \max_{s \in \{N-k+1, \dots, N\}} \hat{E}_{\beta}(s).
\end{align}

Solving this recurrence yields $\hat{E}_{\beta}(d) = (1+\epsilon)^{\frac{(N-d)(N+d-1)}{2}}$.

In the Inside-Outside algorithm, the merit $\mu(A, i, j) = \alpha(A,
i, j) \beta(A, i, j)$ is the unnormalized probability that nonterminal
$A$ yields span $(i, j)$. Define $\mu'(A, i, j) = \alpha'(A, i, j)
\beta(A, i, j)$. By the results above, we have that the ratio of the
actual merit to our approximation is upper bounded by
$(1+\epsilon)^{\frac{(N+d-2)(N-d+1)}{2}}$, where, as before, $d$ is
the length of the span $(i, j)$.

\section{Approximate Inside}
Here we describe the algorithm ApproxInside which computes the
$\alpha'$ values.

\begin{algorithm}
\caption{The Approximate Inside Algorithm \label{alg:approxInside}}
\begin{algorithmic}[1]
\Procedure {ApproxInside}{$s$, $G=(S, bRules, uRules)$, $U$, $\epsilon$}
  \State \Comment $s$ sentences of length $n$
  \State \Comment $G$ grammar 
  \State \Comment $S$ start symbol
  \State \Comment $U$ upper bound, function from $(A, i, j) \mapsto \mathbf{R}$
  \State $\alpha \leftarrow \Call{InitAlpha}{G, 1, n}$ \Comment Initialize table of $\alpha$ 
  \State \Return \Call{$GO_{OR}$}{$S$, $1$, $n$}
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{The Approximate Inside Algorithm (cont)}
\begin{algorithmic}[2]
\Procedure {$GO_{OR}$}{$A$, $i$, $j$}
  \State \Comment $A$ nonterminal symbol
  \State $children \leftarrow \Call{GetChildren}{A, i, k}$
  \State $lb \leftarrow 0 $ \Comment Initialize lower bound on OR node to 0
  \For{$node \in children$}{}{}
    \If{$node$ is unary of form $A \rightarrow t, i, i$ and $t = s_i$}
      \State $v \leftarrow \psi(A \rightarrow t)$
    \ElsIf{$node$ is binary}
      \State $v \leftarrow $\;\Call{$GO_{AND}$}{$node$, $lb$}
    \EndIf
    \State $lb \leftarrow lb + v$ \Comment Update lower bound
  \EndFor
  \State $\alpha[A, i, j] \leftarrow lb$ \Comment Insert $\alpha$ value into table.
\EndProcedure
\Statex
\Procedure {$GO_{AND}$}{$A \rightarrow B\, C$, $i$, $k$,$j$, $lb$}
  \State $w \leftarrow \psi(A \rightarrow B\, C)$ 
  \State $v_{\ell} \leftarrow \alpha[B, i, k]$ 
  \State $v_{r} \leftarrow \alpha[C, k+1, j]$
  \State $u_{\ell} \leftarrow U[B, i, k]$ 
  \State $u_{r} \leftarrow U[C, k+1, j]$
  \If{neither $v_{\ell}$ nor $v_r$ is $NIL$} \Comment If both values are already computed, just use them
    \State \Return $w * v_{\ell} * v_r$
  \ElsIf{ $v_r \neq NIL$}  \Comment If the right value is already computed, compute left value
    \If {$u_\ell \leq \frac{\epsilon * lb}{w * v_r}$}
      \State $v_\ell \leftarrow u_\ell$
    \Else
      \State $v_\ell \leftarrow $\;\Call{$GO_{OR}$}{$B$, $i$, $k$}
    \EndIf
  \ElsIf{ $v_\ell \neq NIL$}  \Comment If the left value is already computed, compute right value
    \If {$u_r \leq \frac{\epsilon * lb}{w * v_\ell}$}
      \State $v_r \leftarrow u_r$
    \Else
      \State $v_r \leftarrow $\;\Call{$GO_{OR}$}{$C$, $k+1$, $j$}
    \EndIf
  \Else \Comment If neither values yet computed, compute both
    \If {$u_\ell u_r \leq \frac{\epsilon * lb}{w}$}
      \State $v_r \leftarrow u_r$
      \State $v_\ell \leftarrow u_\ell$
    \Else
      \State $v_\ell \leftarrow $\;\Call{$GO_{OR}$}{$B$, $i$, $k$}
      \State $v_r \leftarrow $\;\Call{$GO_{OR}$}{$C$, $k+1$, $j$}
    \EndIf
  \EndIf
  \State \Return $w * v_{\ell} * v_r$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{The Approximate Outside Algorithm.\label{alg:approxOutside}}
\begin{algorithmic}[2]
\Procedure {$ApproxOutside$}{$s$, $G$, $S$}
  \State $n \leftarrow length(s)$
  \State $\beta[S, 1, n] \leftarrow 1$ \Comment initialize $\beta$ recursion, unless $n==1$ in which case we are done
  \State $open \leftarrow [(S, 1, n)]$ \Comment initialize list that will hold nodes to be processed
  \While {$open \neq NIL$}
  \State $children \leftarrow []$ \Comment initialize list that holds next level of nodes to be processed
   \For {$(A, i, j) \in open$ \Comment $i \neq j$}
    \For {$A \leftarrow B\, C \in rules(G)$} \Comment iterate over binary rules headed by $A$
      \For {$k \leftarrow i, \dots, j - 1$}
        \If {$\alpha[B, i, k] \neq 0$}
          \State if $\beta[C, k+1, j] = NIL$, set to $0$
          \State $\beta[C, k+1, j] \leftarrow \beta[C, k+1, j] + \psi(A \rightarrow B C) \alpha[B, i, k] \beta[A, i, j]$
          \State add $(C, k+1, j)$ to $children$
        \EndIf

        \If {$\alpha[C, k+1, j] \neq 0$}
          \State if $\beta[B, i, k] = NIL$, set to $0$
          \State $\beta[B, i, k] \leftarrow \beta[B, i, k] + \psi(A \rightarrow B C) \alpha[C, k+1, j] \beta[A, i, j]$
          \State add $(B, i, k)$ to $children$
        \EndIf 
      \EndFor 
   \EndFor
   \EndFor
   \State $open \leftarrow children$
  \EndWhile    
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{The Approximate Outside Algorithm.\label{alg:approxOutside}}
\begin{algorithmic}[2]
\Procedure {$ApproxOutside$}{$s$, $G$, $S$}
  \State $OPEN \leftarrow []$ \Comment initialize OPEN FIFO queue
  \State $n \leftarrow length(s)$
  \State $\beta(S, 1, n) \leftarrow 1$
  \State \Call{AddChildrenToOpen}{$S, 1, n, \beta(S, 1, n)$}
  \While {$OPEN$ is not empty}
    \State $node \leftarrow pop(OPEN)$ where $rule_{node} = (A \rightarrow B C, i, k, j) = fst(node)$ and $pa_\beta= snd(node)$
    \State $a_\ell \leftarrow \alpha[B, i, k]$
    \State $a_r \leftarrow \alpha[C, k+1, j]$
    \If {$a_r \neq NIL$}
      \If {$\beta[B, i, k] = NIL$}
         $\beta[B, i, k] \leftarrow 0$
      \EndIf
      \State $\beta[B, i, k] \leftarrow \beta[B, i, k] + pa_\beta a_\ell \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
      \State \Call{AddChildrenToOpen}{$B, i, k, \beta(B, i, k)$}
    \EndIf
    \If {$a_l \neq NIL$}
      \If {$\beta[C, k+1, j] = NIL$}
         $\beta[C, k+1, j] \leftarrow 0$
      \EndIf
      \State $\beta[C, k+1, j] \leftarrow \beta[C, k+1, j] + pa_\beta a_r \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
      \State \Call{AddChildrenToOpen}{$C, k+1, j, \beta(C, k+1, j)$}
    \EndIf
  \EndWhile  
\EndProcedure

\Statex
\Procedure {AddChildrenToOPEN}{A, i, j, betaVal}
  \State $\{c_i\}\leftarrow \Call{GetChildren}{A, i, j}$ \Comment Get AND-node children
  \State insert pairs $\{(c_i, betaVal\}_i)$ into OPEN
\EndProcedure
\end{algorithmic}
\end{algorithm}

\noindent \textbf{Claim: Algorithm~\ref{alg:approxInside} correctly computes the $\beta'$ values.}
\noindent TODO

\noindent \textbf{Claim: Algorithm~\ref{alg:approxOutside} correctly computes the $\beta'$ values.}
\noindent TODO

\section{A simple upper bound}
We will present a simple upper bound on a slightly restricted set of
PCFGs which is based on the following intuition: if I am a nonterminal
$A$ and I know that I need to yield a string $x_1 \dots x_n$, then
this is going to be at least as hard as starting at $A$ and generating
any one of the elements in $x_1 \dost x_n$. That is, the highest
probability single word in my yield will have a higher probability of
being produced, then any longer span.

Suppose our PCFG has potentials less than or equal to one. This is the
case in the commons situation in which the PCFG is normalized, or if,
as in variational approximations to the Hierarchical Diirichlet PCFG,
our PCFG is unnormalized with values summing to less than one. Now we
define a function $g(A, w, m)$ for nonterminal $A$, terminal $w$ and
integer $w>0$, which we claim is greater the weight our PCFG would
assign to any parse tree rooted at $A$ and yielding any span of length
$m$ containing $w$. Specifically, consider any such parse tree, and
set all weights not on the path between $A$ and $w$ to 1. The total
weight assigned to the parse tree is the product of the weights on its
edges, so the resulting parse tree will have a strictly greater score.

\begin{algorithm}
\caption{An O(n^2}
\begin{algorithmic}[2]
\Procedure {$GO_{OR}$}{$A$, $i$, $j$}

\EndProcedure
\end{algorithmic}
\end{algorithm}

\end{document}

### 2014-08-15 drafts/writeup.tex
Initial draft commits.

\documentclass{article}

\begin{document}

\section*{Introduction}
What enables humans to learn -- from a finite number of examples --
systems of concepts that contain a potentially unbounded number of
individuals elements?

How do people learn to combine new conceptual knowledge with core knowledge?

Is the acquisiton of conceptual knowledge driven by generalization
from examples or by the application of high-level conceptual schemas?

Does the structure of language or the learning of language guide
the learning of concepts?

How does one learn what counts as evidence of conceptual meaning?

What is the nature of conceptual knowledge? Is it declarative, a
collection of facts from which questions are answered, best actions
are inferred, etc.? Or is it procedural, a collection of routines or
productions? Or both? If both, how do these two forms of knowledge
interact?

\section*{AAAI Abstract}
Title: Learning the many meanings of number words. 
or   : Grammar induction for multi-faceted concept learning.

Title: Scaffolding meaning on unsupervised grammar induction.

What comes first, learning the structure of a perceptual input or
learning its meaning? Evidence from psychology and linguistics
suggests that children jointly learn the compositional structure of
words and phrases and how to map meanings onto those structural
components. But little is known about whether jointly learn structure
and meaning can improve AI and machine learning performance, because
most previous research has focused on supervised learning (mapping
perceptual inputs directly to task-relevant outputs) or unsupervised
learning (asserting that the learned structure *is* the output). Here,
we combine probabilistic grammar induction with supervised learning

When does someone *understand* a concept? Intuitively, understanding a
concept requires more than being able to answer a specific query
(e.g. is an element a member of that concept). Rather, we say that a
person understands a concept when they are able to leverage their
knowledge of that concept to perform a wide set of (perhaps
unspecified) tasks. Take, for example, a child's knowledge of number,
and the various kinds of queries that a mastery of the concept
ultimately supports support: counting (of various kinds, up, down, by
fives); comparing numbers (which is greater?); approximate numerical
understanding (is ten approximately half of a hundred? how about forty
five?); converting numbers to digits (e.g. ``four hundred and five''
to 405)l; converting digits to numbers (e.g. 1243 to ``twelve hundred
and forty three''); and an unbounded set of conceptual queries
(e.g. Is their a largest/smallest number? If I take N marbles out of a
bowl and put N back in, are there more, less, or the same number of
marbles in the bowl? If I start at five and start counting up by one
will I every encounter the same number twice? What's one more than a
plutpyltillion and one?).

What would it take for computers to able to learn a similar kind of
conceptual understanding? One possibility is to attempt to learn ev to
answer every query independently. At the other extreme, typified by
inductive logic programming, the goal is to learn a coplletel theory
of the domain of interest in the form a logical theory, such that
queries are expressed as querying of a logical thoery. However, But it
is more But people's actual competency is somewhere in between these
two approaches.  Clearly, people can answer arbitrary new
questions. But they cannot perform any manipultation. Fro example,
people cannot cannot add arbitrarily large numbers in their
heads. People don't know whether there ais a largest prime.
Therefore, we propose an intermediate kgoalswhich is that people learn
what will term ``conceptual routines''. This might include counting,
comparing numbers, approximate magnitutde routines, etc. And that
these are each learned on a the basis of common representation. We
learn this base repsenetation in an unsupervised manner, and learn
these conceptual routines on top of the learned representation.

In this paper, we focus on a subset of the kinds of queries How can
computers learn a similar kind of conceptual understanding as this?
Although Psychological and linguistic investigations indicate that
children's concept of number develops over many years.

-----------------------------------------------------------------

Title: Domain specific grammar induction for semi-supervised
programming by demonstration.

Abstract: For programming by demonstration (PDB) systems to be useful
to non-expert users, these systems must be able to understand and
manipulate data that is implicitly structured in a domain-specific
way. For example, the data might be email addresses or street
addresses, a company-specific record format for book-keeping, an ad
hoc organization relying on visual separation for demaraction, or a
user's shorthand notation. While recent success in PBD has been driven
by the careful construction of domain-specific languages, those
languages generally have fixed or user-specified primitives and
features. Here, we show that a hierarchical Bayesian approach to
grammar induction can be used as a general and robust tool for
learning the latent structure of user data, and show that such
learning can help automate tasks that would otherwise require
task-specific programming to solve.

------------------------------------------------------------

One of the distinguishing aspects of human societies is their use of
symbol systems to express useful systems of concepts. One particularly
successful such symbol system is the Arabic numerals, which are used
to represent things as varied as sizes, lengths, times, dates, and
cardinality. What is it that makes this a successful symbol system for
these purposes, and what is it that enables people to learn the
various meanings of these symbols from a small fraction of the
infinite possible elements? Our proposal is that what makes this a
successful symbol system is that it bridges the gap between important
and innate domains of semantics, i.e., the compositional semantics of
language and the geometric semantics of our core cognition of
magnitude, space and time. We provide a model that is based on this
proposal. The use of the Arabic numerals provides the model access to
both kinds of semantic representations, and allow the model to learn
the mappings that would otherwise require many more examples or would
eb impossible.

\end{document}

### 2014-08-15 src/GIJoe/ApproximateInference/InsideOutside.hs

@@ -15,7 +15,10 @@ import GIJoe.Parse
 
 import Numeric.Log
 
+import Control.Monad.Loop
+
 import Control.Monad
+import Control.Monad.Trans
 import Control.Monad.ST
 import Control.Exception (assert)
 
@@ -108,7 +111,7 @@ approxInside gr start xs ubFunc epsilon = do
             rules = binaryRulesHeadedBy gr sym
 
 -- each element of Open is an AND node and one of its parent OR nodes
- type Open = LazyDequeue ((Rule, Int, Int, Int), (Symbol, Int, Int))
+ -- type Open = LazyDequeue ((Rule, Int, Int, Int), (Symbol, Int, Int))
 -- set of elements already expanded and whose children have already been placed on the Open list
 -- type Expanded = Set ((Symbol, Int, Int))
             
@@ -121,44 +124,28 @@ approxOutside gr start xs alphaTable = {-# SCC approxOutsideMainLoop #-} do
   let n = length xs 
   !betaTable <- newSized $ htSize n (numSymbols gr) -- generate empty hashtable
   insert betaTable (start, 1, n) 1.0 -- initialize start node cell
-  -- initialize open 
-  let open  = [(start, 1, n)] 
-  loop betaTable Set.empty open  
-  return betaTable
-  where
-    loop _ children [] = return ()
-    loop betaTable children ((_A, i, j):open') = do
-      debugM $ "(_A, i, j): " ++ show (_A, i, j)
-      debugM $ "open': " ++ show open'                 
-      debugM $ "children: " ++ show children
-      
-      let go children [] = return children
-          go children ((r, k):splits) = do
-            let (BinaryRule _ _B _C w) = r
-            Just pa_beta <- lookup betaTable (_A, i, j)
+  -- initialize open
+  exec_ $ do
+    ell <- forEach [n-x | x <- [1..n-1]] -- span length
+    i   <- forEach [1..n-ell]
+    let j = i + ell
+    _A  <- forEach $ allNonTerminals gr
+    lift (getAlphaValue (_A, i, j)) >>= \case
+      Nothing -> return ()
+      Just _  -> do
+        k <- forEach [i..j-1]
+        (BinaryRule _ _B _C w) <- forEach $ binaryRulesHeadedBy gr _A
+        Just pa_beta <- lift $ lookup betaTable (_A, i, j)
         
-            children' <- getAlphaValue (_B, i, k) >>= \case
[…]

### 2014-08-19 drafts/ApproximatePCFG/ApproximatePCFG.tex
Implemented Boundary Probabilties but is way too slow. Need to optimize with HashTable.

@@
 \usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
-\usepackage{amsmath}
+\usepackage{amssymb,amsmath}
 
@@
 \usepackage{algpseudocode}
+
+\usepackage{natbib}
+
+\makeatletter
+\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}
+\makeatother
+
+\newcommand{\derives}{\xRightarrow{*}}
 
@@
   \State $\beta[S, 1, n] \leftarrow 1$ \Comment initialize $\beta$ recursion, unless $n==1$ in which case we are done
+  \For {$\ell \gets 1, \dots, n-1$}
+    \For {$i \gets \ell$}
+      \State $j \leftarrow i + \ell$
+    \EndFor
+  \EndFor
+
   \State $open \leftarrow [(S, 1, n)]$ \Comment initialize list that will hold nodes to be processed
@@
 
-\begin{algorithm}
-\caption{The Approximate Outside Algorithm.\label{alg:approxOutside}}
-\begin{algorithmic}[2]
-\Procedure {$ApproxOutside$}{$s$, $G$, $S$}
-  \State $OPEN \leftarrow []$ \Comment initialize OPEN FIFO queue
-  \State $n \leftarrow length(s)$
-  \State $\beta(S, 1, n) \leftarrow 1$
-  \State \Call{AddChildrenToOpen}{$S, 1, n, \beta(S, 1, n)$}
-  \While {$OPEN$ is not empty}
-    \State $node \leftarrow pop(OPEN)$ where $rule_{node} = (A \rightarrow B C, i, k, j) = fst(node)$ and $pa_\beta= snd(node)$
-    \State $a_\ell \leftarrow \alpha[B, i, k]$
-    \State $a_r \leftarrow \alpha[C, k+1, j]$
-    \If {$a_r \neq NIL$}
-      \If {$\beta[B, i, k] = NIL$}
-         $\beta[B, i, k] \leftarrow 0$
-      \EndIf
-      \State $\beta[B, i, k] \leftarrow \beta[B, i, k] + pa_\beta a_\ell \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
-      \State \Call{AddChildrenToOpen}{$B, i, k, \beta(B, i, k)$}
-    \EndIf
-    \If {$a_l \neq NIL$}
-      \If {$\beta[C, k+1, j] = NIL$}
-         $\beta[C, k+1, j] \leftarrow 0$
-      \EndIf
-      \State $\beta[C, k+1, j] \leftarrow \beta[C, k+1, j] + pa_\beta a_r \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
-      \State \Call{AddChildrenToOpen}{$C, k+1, j, \beta(C, k+1, j)$}
-    \EndIf
-  \EndWhile  
-\EndProcedure
-
-\Statex
-\Procedure {AddChildrenToOPEN}{A, i, j, betaVal}
-  \State $\{c_i\}\leftarrow \Call{GetChildren}{A, i, j}$ \Comment Get AND-node children
-  \State insert pairs $\{(c_i, betaVal\}_i)$ into OPEN
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
 \noindent \textbf{Claim: Algorithm~\ref{alg:approxInside} correctly computes the $\beta'$ values.}
@@
 
-\section{A simple upper bound}
-We will present a simple upper bound on a slightly restricted set of
-PCFGs which is based on the following intuition: if I am a nonterminal
-$A$ and I know that I need to yield a string $x_1 \dots x_n$, then
-this is going to be at least as hard as starting at $A$ and generating
-any one of the elements in $x_1 \dost x_n$. That is, the highest
-probability single word in my yield will have a higher probability of
-being produced, then any longer span.
+The algorithm presented provides two degrees of freedom to the user: the first is a choice of upperbound function $U$, and the second is a way to order the children of a node.
+The tighter the upperbound, the better an approximation we get.
+The better the ordering of the children, the less often we need to explore derivations of low relative weight.
 
-Suppose our PCFG has potentials less than or equal to one. This is the
-case in the commons situation in which the PCFG is normalized, or if,
-as in variational approximations to the Hierarchical Diirichlet PCFG,
-our PCFG is unnormalized with values summing to less than one. Now we
-define a function $g(A, w, m)$ for nonterminal $A$, terminal $w$ and
-integer $w>0$, which we claim is greater the weight our PCFG would
-assign to any parse tree rooted at $A$ and yielding any span of length
-$m$ containing $w$. Specifically, consider any such parse tree, and
-set all weights not on the path between $A$ and $w$ to 1. The total
-weight assigned to the parse tree is the product of the weights on its
-edges, so the resulting parse tree will have a strictly greater score.
+\section{Ordering heuristics}
+How we order the children of a node is critical to the speed-up this
+aproximate algorithm gets. After all, if there is a child whose
+derivation dominates all others and we accidentally put it last, then
+we will not be able to short-circuit at all; if, instead, we put it
+first, then we will likely only have to explore that child.
 
-\begin{algorithm}
-\caption{An O(n^2}
-\begin{algorithmic}[2]
-\Procedure {$GO_{OR}$}{$A$, $i$, $j$}
+One can image various choices of ordering functions, with perhaps the
+simplest being to pick the rules in order of decreasing weight. This,
+however, only gives us an ordering on the rules, not on where we
+should split the span of the child. One reasonable suggestion,
+especially for right branching structures like English parse trees, is
+to start with the smallest left child spans first and increase from
+there.
 
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
+These heuristics, however, are very weak, and do not take into
+consideration the content span at all. In the following we propose a
+simple heuristic with child ordering that takes into account the
+bottom-up probability of the content but whose complexity is constant
+in the size of the span. The basic intutiion behind this heuristic is
+that if one wants to evaluate the probability that, e.g., a noun
+phrase will yield that fragmant ``saw the boy with,'' one can gain a
+lot of information by calculating the probability that the phrase will
+start and end with ``saw'' and ``with,'' respectively.
+
+Some notation: let's denote by $\gamma(A, x)$ the probability that a
+nonterminal $A$ derives any string with prefix $x$. Likewise, let's
+denote by $\theta(A, x)$ the proabability that a nonterminal $A$
+derives any string with suffix $x$. We will write this using the
+following notation:
+
+\begin{align}
+  \gamma(A, x) &\triangleq p(A \xRightarrow{*} x\dots) \label{eq:gammaDef}\\
+  \theta(A, x) &\triangleq p(A \xRightarrow{*} \dots x)  \label{eq:thetaDef}
+\end{align}
+
+$\gamma(A, x)$ is just the prefix probability defined for a single
+lexical item. This can be defined as:
+\begin{align}
+  p(A \derives x\dots) &= \sum_{A \rightarrow BC} p(B \derives x \dots) p(A \rightarrow B C) 
+    + p(A \rightarrow x)\\
+  p(A \derives \dots x) &= \sum_{A \rightarrow CB} p(B \derives \dots x) p(A \rightarrow C B) 
+    + p(A \rightarrow x)
+\end{align}
+
+If we define the matrices $\Gamma$ and $\Theta$ where $\Gamma_{A, B} = \gamma(A, B)$
+and $\Theta_{A, B} = \theta(A, B)$. then we can rewrite these recursions
+
+\begin{align}
+  \Gamma &= \Phi^\ell \Gamma + \Psi \\
+  \Theta &= \Phi^r \Gamma  + \Psi \\
+  \text{where }\Phi^\ell_{A, B} &\triangleq \sum_{A \rightarrow B C} p(A \rightarrow B C)\\
+               \Phi^r_{A, B} &\triangleq \sum_{A \rightarrow C B} p(A \rightarrow C B)\\
+               \Psi_{A, x} &\triangleq p(A \rightarrow x).
+\end{align}
+
+This means that we can calculate $\Gamma$ and $\Phi$ by solving the linear systems
+
+\begin{align}
+  (1-\Phi_\ell) \Gamma &= \Psi\\
+  (1-\Phi_r) \Theta &= \Psi.\\
+\end{align}
+
+The complexity of these calculations is $K^3$, where $K$ is the number
+of symbols in the grammar.
+
+From $\Gamma$ and $\Theta$ we can compute the boundary probability: 
+\begin{align}
+  p(A \derives x \dots y) &= p(A \rightarrow x) \mathbf{1}(x = y) 
+     + \sum_{A \rightarrow B C} p(A \rightarrow B C) p(B \derives x \dots) p(C \derives \dots y)\\
+  p(A \derives x \dots y) &= p(A \rightarrow x) \mathbf{1}(x = y) 
+     + \sum_{A \rightarrow B C} p(A \rightarrow B C) \Gamma_{B, x} \Theta_{y, C}
+\end{align}
+
+Here, as is often the case in computations that make use of heuristic
+scores, it may be advantageous to precompute all the heuristic values
+so that lookup is $O(1)$.
+
+\section{Appendix}
+\subsection{Existence of $\Gamma$ and $\Theta$}
+\textbf{Claim: } The matrices $\Gamma$ and $\Theta$ defined by
+Equations~\ref{eq:gammaDef}-\ref{eq:thetaDef} exist.  \textbf{Proof: }
+The $\Gamma$ and $\Theta$ matrices will not exist only if
+$1-\Phi^\ell$ and $1-\Phi^r$ are singular. To show that they are not
+singular we will show that there is no non-zero linear combination of
+their columns that sums to zero. The proofs are nearly identical for
+$1-\Phi^\ell$ and $1-Phi^r$.
+
+Suppose, by contradiction, that their is some linear combination $a_1,
+..., a_K$ such that $(1-\Phi^\ell) a = 0$. This implies that each $a_i
+= \sum_{j} \Phi^\ell_{i, j} a_j$. Recall that the element of
+$\Phi^\ell_{A, B} = \sum_{A \rightarrow B C} \p(A \rightarrow B C)$,
+where $A$ and $B$ are nonterminals. For the PCFG defined by these
+probabilities to be valid, $\Phi^\ell_{A,B}$ must be less than
+one. Otherwise, there would be a nonzero probability of an infinite
+string. Taking the absolute values of both sides of the equation
+above, we see that $|{a_i}| \leq \sum_{j} \Phi^\ell_{i, j} |
+a_j|$. But if this is true, then  TODO
+
+\bibliographystyle{bibliographystyle}
+\bibliography{main}
 

### 2014-08-19 src/GIJoe/ApproximateInference/InsideOutside.hs

@@ -1,4 +1,4 @@
-{-# Language BangPatterns, ParallelListComp, FlexibleInstances, DeriveGeneric, LambdaCase #-}
+{-# Language BangPatterns,  FlexibleInstances, LambdaCase #-}
 
 module GIJoe.ApproximateInference.InsideOutside where
 
@@ -9,6 +9,7 @@ import GIJoe.Types
 import GIJoe.Grammar
 import GIJoe.ApproximateInference.HashTable
 import GIJoe.Parse
+import GIJoe.ApproximateInference.BoundaryProbabilities
 
 -- import GIJoe.ApproximateInference.LazyDequeue
 -- import qualified GIJoe.ApproximateInference.LazyDequeue as LazyDequeue
@@ -25,17 +26,49 @@ import Control.Exception (assert)
 import Data.Sequence (Seq, (><), (|>), (<|), ViewL(..))
 import qualified Data.Sequence as Seq
 
+import Data.HashMap.Strict (HashMap)
+import qualified Data.HashMap.Strict as HashMap
+
+
 import Data.Set (Set)
 import qualified Data.Set as Set
 
-import Debug.Trace
-_DEBUG = True
-debugM x = if _DEBUG then trace x $ return () else return ()
+import Data.Function (on)
+import Data.List (sortBy)
 
--- type Set = Set
+
+import System.IO.Unsafe
+import Data.IORef 
+
+------------------------------------------------------------
+-- DEBUG
+import Debug.Trace
+_DEBUG = False
+debugM x = when _DEBUG $ do !_ <- trace x $ return ()
+                            return ()
+
+-- *TOP LEVEL GlOBAL* Counter for debugging
+_globalCounter :: IORef Int
+{-# NOINLINE _globalCounter #-}
+_globalCounter = unsafePerformIO (newIORef 17) 
+
+_zeroCounter :: IORef Int -> ()
+{-# NOINLINE _zeroCounter #-}
+_zeroCounter ref = let x = unsafePerformIO $ writeIORef ref 0 in x
+
+_incrCounter :: IORef Int -> ()
+{-# NOINLINE _incrCounter #-}
+_incrCounter ref = let x = unsafePerformIO $ atomicModifyIORef' ref $ \a -> (a+1, ()) in x
+
+_readCounter :: IORef Int -> Int
+{-# NOINLINE _readCounter #-}
+_readCounter ref = let x = unsafePerformIO $ readIORef ref in x
[…]

### 2014-08-19 src/GIJoe/Grammar.hs

@@ -13,9 +13,11 @@ import Numeric.Log
 
 import Control.Monad
 import Data.List.Split (splitOn)
-import Data.List (foldl', sortBy, sort, maximumBy, foldl1')
+import Data.List (foldl', sortBy, sort, maximumBy, foldl1', nub)
 import Control.Monad.Logic
 import Data.Function (on)
+
+import Control.DeepSeq
 
 import Control.Monad.State.Strict
 import Control.Monad.Random
@@ -44,6 +46,7 @@ import Data.Tree.Pretty
 
 import Debug.Trace
 
+import Data.Sequence (Seq, (><))
 import qualified Data.Sequence as Seq
 
 import qualified Data.MemoCombinators as Memo
@@ -57,7 +60,12 @@ instance Show Symbol where
   show (N i Nothing) = "N" ++ show i
   show (T s) = s
 
-instance Hashable Symbol where 
+instance NFData Symbol where
+  rnf (N i t) = i `seq` t `seq` N i t `seq` ()
+  rnf (T s) = s `seq` T s `seq` ()
+
+
+instance Hashable Symbol where
 
 memoSymbol :: Memo.Memo Symbol
 memoSymbol f = table (m (f . uncurry N)) (m' (f . T)) f  
@@ -235,8 +243,14 @@ grammarFromRules rs = foldl' insertRule emptyGrammar (zip rs [0..])
 allNonTerminals :: Grammar -> [Symbol]
 allNonTerminals gr = HashMap.keys . headSymbolTable $  gr
 
+nonterminals :: Grammar -> [Symbol]
+nonterminals = allNonTerminals
+
 numSymbols :: Grammar -> Int
 numSymbols gr = HashMap.size . headSymbolTable $  gr
+
+allSymbols :: Grammar -> [Symbol]
+allSymbols gr = allNonTerminals gr ++ lexicon gr
 
 modifyRuleWeightWithId :: Grammar -> Int -> (Double' -> Double') -> Grammar
 modifyRuleWeightWithId gr i f = gr{ruleIndex=IntMap.adjust g i (ruleIndex gr)}
@@ -268,6 +282,12 @@ normalizeGrammar gr = foldl' go gr (allNonTerminals gr)
 
 type Lexicon = [Symbol]
 
+lexicon :: Grammar -> Lexicon
+lexicon gr = nub $ map child unaryRules
+  where unaryRules = getRulesById gr (unaryRuleIds gr)
+
+terminals = lexicon
+
[…]

### 2014-09-17 drafts/ApproximatePCFG/ApproximatePCFG.tex
Added log untracked files for Approximate inference and stochastic rewrite system (SRS).

@@
 \usepackage{algpseudocode}
+
+\usepackage{graphicx}
 
@@
 
+\subsection{A serious problem with boundary probabilities}
+Unfortunately, there is are serious problems with using boundary
+probabilities. First, the dependency between the first and last word
+of a constituent decreases rapidly with the length of the
+constitutent. Second, since calculating the boundary probability
+involves taking the expectation over yields of all lengths, the value
+assigned does not take into accoutn the length of the constitutent,
+and produces values that upper bound the insid probability of a
+consituent even for two word constitutent. But even an unlikely pair
+of adjacent words is going to have a higher probability than a typical
+long sentence. Therefore, we need to come up with a scoring function
+that takes into account the length of the constitutent.
+
+\section{Using constituent length to upper bound inside probabilities}
+The probability that symbol $A$ yields any constitutent of length $n$
+is greater than or equal to the probability that symbol $A$ yields any
+specific string of words of that length.
+
+Let $\lambda(A, n) \triangleq p(A \text{ yields a string of lengths }
+n)$. This probability can be calculated recursively: 
+\begin{align}
+\lambda(A, n=1) =& \sum_{A \leftarrow t} \psi(A \leftarrow t)\\ 
+\lambda(A, n>1) =& \sum_{A \leftarrow B \,C} \sum_{k=1}^{n-1} \psi(A \leftarrow B\, C) 
+  \lambda(B, k) \lambda(C, k=1)\\ 
+\end{align}
+
+This recursive definition leads directly to an $O(Kn^2)$ algorithm for
+$\lambda$.
+
+\begin{figure}
+\begin{minipage}[t]{\textwidth}
+  \includegraphics[width=\linewidth]{figures/toy_grammar_approx_vs_exact.pdf}
+\end{minipage}
+\begin{minipage}[t]{4cm}
+  \begin{align*}
+    S \rightarrow& \,L\, S :: 0.1 \, | \,a :: 0.9\\
+    L \rightarrow& \,L\, S :: 0.0001\, | \, a :: 0.999\\
+  \end{align*}
+  \centering (a)\par
+\end{minipage}
+
+\begin{minipage}[t]{\textwidth}
+  \caption{A comparison of the approximate and exact ``Inside''
+    algorithms on the toy grammar (c). (a) The number of edges seen by
+    the algorithm. The exact Inside algorithm has a predictable $n^3$
+    behavior because it looks once at every edge $(A, i, k, j)$. The
+    approximate algorithm ignores edges that are gauranteed to be
+    insignificant, so that the number of inspected edges is
+    dramatically reduced. (b) The amount by which the approximate
+    Inside algorithm over estimates the loglikelihood of the
+    sentence. }
+\end{minipage}
+\end{figure}
+
 \section{Appendix}
@@
 = \sum_{j} \Phi^\ell_{i, j} a_j$. Recall that the element of
-$\Phi^\ell_{A, B} = \sum_{A \rightarrow B C} \p(A \rightarrow B C)$,
+$\Phi^\ell_{A, B} = \sum_{A \rightarrow B C} p(A \rightarrow B C)$,
 where $A$ and $B$ are nonterminals. For the PCFG defined by these

### 2014-09-17 src/GIJoe/ApproximateInference/BoundaryProbabilities.hs

@@ -1,0 +1,151 @@
+{-# Language BangPatterns, LambdaCase, RankNTypes #-}
+
+module GIJoe.ApproximateInference.BoundaryProbabilities
+       (mkOmegaTable,
+        mkGammaThetaTables,
+        mkLambdaTable, 
+        OmegaTable,
+        GammaTable,
+        ThetaTable,
+        LambdaTable
+       ) where
+
+import Prelude hiding (sum)
+import GIJoe.Grammar
+import GIJoe.Types
+
+import GIJoe.ApproximateInference.HashTable (HashTable)
+import qualified GIJoe.ApproximateInference.HashTable as HashTable
+
+import Data.Maybe
+
+import Control.Monad
+import Control.Monad.Trans
+import Control.Monad.ST
+
+import Numeric.LinearAlgebra.HMatrix
+import Data.Packed.Matrix
+import Numeric.Log
+
+import Data.HashMap.Strict (HashMap)
+import qualified Data.HashMap.Strict as HashMap
+
+import Control.Monad.Loop
+
+import Data.List hiding (sum)
+
+import Debug.Trace
+
+type OmegaTable s = HashTable s (Symbol, Symbol, Symbol) Double'
+type GammaTable s = HashTable s (Symbol, Symbol) Double'
+type ThetaTable s = HashTable s (Symbol, Symbol) Double'
+
+
+mkOmegaTable :: Grammar -> ST s (HashTable s (Symbol, Symbol, Symbol) Double')
+mkOmegaTable gr = trace ("Making OmegaTable") $ do
+  (gamma_ht, theta_ht) <- mkGammaThetaTables_ gr si
+  omega_ht <- HashTable.newSized $ (numSymbols gr)^3
+  exec_ $ do 
+    _A <- forEach $ nonterminals gr 
+    x  <- forEach $ terminals gr
+    y  <- forEach $ terminals gr
+    let !v0 = if x == y
+                 then sum $ map weight $ getUnaryRulesBySymbols gr _A x
+                 else 0
+    lift $ HashTable.insertWith omega_ht (+) (_A, x, y) v0
+    let rs = binaryRulesHeadedBy gr _A
+    BinaryRule _ _B _C w <- forEach rs
+    lift (HashTable.lookup gamma_ht (_B, x)) >>= \case
+      Nothing -> return ()
+      Just gamma_Bx ->
+        lift (HashTable.lookup theta_ht (_C, y) ) >>= \case
+          Nothing -> return ()
+          Just theta_Cy ->
[…]

### 2014-09-17 src/GIJoe/ApproximateInference/DList.hs

@@ -1,0 +1,60 @@
+
+module GIJoe.ApproximateInference.DList where
+
+import Prelude hiding (head, reverse, concat, break, null)
+import qualified Prelude
+
+import Debug.Trace
+
+data BTree a = Node (BTree a) (BTree a) |
+               Leaf a deriving (Show, Ord, Eq)
+
+type DList a = BTree [a] 
+
+empty :: DList a
+empty = Leaf []
+
+null :: DList a -> Bool
+null (Node l r) = null l && null r
+null (Leaf []) = True
+null _ = False
+
+-- | Add element
+cons :: a -> DList a -> DList a
+cons x (Leaf xs) = Leaf (x:xs)
+cons x (Node l r) = Node (cons x l) r
+
+-- | Pop off first element
+pop :: DList a -> (a, DList a)
+pop dlist = case pop' dlist of
+  Nothing -> error "pop: empty list"
+  Just res -> res
+  where pop' (Leaf (x:xs)) = Just (x, Leaf xs)
+        pop' (Node l r) = case pop' l of
+          Just (x, l') -> Just (x, concat l' r)
+          Nothing -> pop' r
+        pop' (Leaf []) = Nothing
+
+        -- debug x = trace "pop'!" x
+
+
+-- | Remove a single element from head of list
+head :: DList a -> a
+head (Leaf xs) = Prelude.head xs
+head (Node l r) = head l
+
+-- | Return everything but first element
+tail :: DList a -> DList a
+tail xs = let (_, xs) = pop xs in xs
+
+-- | Concatenate two DLists
+concat :: DList a -> DList a -> DList a
+concat = Node
+
+-- | Reverse a DList
+reverse :: DList a -> DList a
+reverse (Leaf xs) = Leaf (Prelude.reverse xs)
+reverse (Node l r) = Node (reverse r) (reverse l) -- this enables reverse without evaluating intermediate lists
+
+fromList :: [a] -> DList a
+fromList = Leaf

### 2014-09-17 src/GIJoe/ApproximateInference/HashTable.hs

@@ -1,0 +1,45 @@
+{-# Language BangPatterns, LambdaCase #-}
+
+module GIJoe.ApproximateInference.HashTable
+       (HashTable,
+        new,
+        newSized,
+        insert,
+        insertWith,
+        delete,
+        lookup,
+        lookupDefault,        
+        foldM,
+        mapM_,
+        fromList,
+        toList,
+        showTable
+        )
+       where
+
+import Prelude hiding (lookup, mapM_)
+import Data.HashTable.Class hiding (HashTable)
+import qualified Data.HashTable.ST.Basic as H
+import Control.Monad.ST
+import Data.Hashable (Hashable(..))
+
+type HashTable s k v = H.HashTable s k v
+
+insertWith :: (Eq k, Hashable k) => HashTable s k v -> (v -> v -> v) -> k -> v -> ST s ()
+insertWith htRef f k v = do
+  res <- lookup htRef k
+  case res of
+    Nothing -> insert htRef k v
+    Just w -> let !v' = f w v in insert htRef k v'
+
+lookupDefault :: (Eq k, Hashable k) => HashTable s k v -> v -> k -> ST s v
+lookupDefault h d k = lookup h k >>= \case
+  Nothing -> return d
+  Just x -> return x
+    
+
+
+showTable :: (Eq k, Hashable k, Show k, Show v) => HashTable s k v -> ST s String
+showTable htRef = do rows <- toList htRef
+                     return $ unlines $ map show rows
+

### 2014-09-17 src/GIJoe/ApproximateInference/InsideOutside.hs

@@ -1,18 +1,20 @@
 {-# Language BangPatterns,  FlexibleInstances, LambdaCase #-}
+{-# LANGUAGE ScopedTypeVariables #-}
+{-# LANGUAGE RankNTypes #-}
 
 module GIJoe.ApproximateInference.InsideOutside where
 
 import Prelude hiding (lookup, sum)
 import qualified Prelude
 
-import GIJoe.Types
-import GIJoe.Grammar
-import GIJoe.ApproximateInference.HashTable
-import GIJoe.Parse
 import GIJoe.ApproximateInference.BoundaryProbabilities
+import GIJoe.ApproximateInference.HashTable
+import GIJoe.ApproximateInference.MaxPath
+import GIJoe.Grammar
+import GIJoe.Parse
 
--- import GIJoe.ApproximateInference.LazyDequeue
--- import qualified GIJoe.ApproximateInference.LazyDequeue as LazyDequeue
+import GIJoe.Types
+import GIJoe.Utils
 
 import Numeric.Log
 
@@ -43,9 +45,9 @@ import Data.IORef
 ------------------------------------------------------------
 -- DEBUG
 import Debug.Trace
-_DEBUG = False
-debugM x = when _DEBUG $ do !_ <- trace x $ return ()
-                            return ()
+_DEBUG = 0
+debugM level x = when (_DEBUG >= level) $ do !_ <- trace x $ return ()
+                                             return ()
 
 -- *TOP LEVEL GlOBAL* Counter for debugging
 _globalCounter :: IORef Int
@@ -66,9 +68,9 @@ _readCounter ref = let x = unsafePerformIO $ readIORef ref in x
 ------------------------------------------------------------
                             
 
-type UpperBound = (Symbol, Int, Int) -> Double'
+type UpperBound s = (Symbol, Int, Int) -> ST s Double'
 type Epsilon = Double'
-type EdgeGenerator = (Symbol, Int, Int) -> [(Rule, Int, Int, Int)]
+type EdgeGenerator s = (Symbol, Int, Int) -> ST s [(Rule, Int, Int, Int)]
 type AlphaTable s = HashTable s (Symbol, Int, Int) Double'
 type BetaTable s = HashTable s (Symbol, Int, Int) Double'
 
@@ -78,37 +80,38 @@ htSize :: Int -- ^ length of sentence
 htSize n k = n * (n+1) `div` 2 * k 
 
 inside :: Grammar
-          -> Symbol -- ^ start symbol
[…]

### 2014-09-17 src/GIJoe/ApproximateInference/LazyDequeue.hs

@@ -1,0 +1,70 @@
+
+module GIJoe.ApproximateInference.LazyDequeue
+       (LazyDequeue(..),
+        empty,
+        enqueue,
+        enqueueMany,
+        dequeue,
+        fromList,
+        toList,
+        null)
+       where
+
+import Prelude hiding (null)
+import GIJoe.ApproximateInference.DList (DList)
+import qualified GIJoe.ApproximateInference.DList as DList
+
+import Debug.Trace
+
+data LazyDequeue a = LazyDequeue (DList a) (DList a) deriving (Show, Eq, Ord)
+
+-- | An empty queue
+empty :: LazyDequeue a
+{-# INLINE empty #-}
+empty = LazyDequeue DList.empty DList.empty
+
+null :: LazyDequeue a -> Bool
+null (LazyDequeue l r) = DList.null l && DList.null r
+
+enqueue :: a -> LazyDequeue a -> LazyDequeue a
+{-# INLINE enqueue #-}
+enqueue a (LazyDequeue l r) = LazyDequeue (DList.cons a l) r
+
+enqueueMany :: [a] -> LazyDequeue a -> LazyDequeue a
+{-# INLINE enqueueMany #-}
+enqueueMany xs q = let (LazyDequeue l r) = shift q 
+  in LazyDequeue l (DList.concat (DList.fromList xs) r)
+
+-- | Dequeue a single element. If the right list is empty, reverses
+-- the first list and sticks the result in the second list.
+dequeue :: LazyDequeue a -> (a, LazyDequeue a)
+dequeue (LazyDequeue l r) | DList.null r = dequeue $ shift (LazyDequeue l r)
+                          | otherwise    = (a, LazyDequeue l r')
+  where (a, r') = DList.pop r
+
+toList :: Show a => LazyDequeue a -> [a]
+toList q | null q = []
+         | otherwise = let (x, q') = dequeue q in x: toList q'
+
+fromList :: [a] -> LazyDequeue a
+{-# INLINE fromList #-}
+fromList xs = enqueueMany xs empty
+
+------------------------
+-- Private functions
+
+shift :: LazyDequeue a -> LazyDequeue a
+shift (LazyDequeue l r) = LazyDequeue DList.empty (DList.concat r (DList.reverse l))
+
+
+
+                                 
+
+
+
+
+
+
+
+
+

### 2014-09-17 src/GIJoe/ApproximateInference/MaxPath.hs

@@ -1,0 +1,123 @@
+{-# Language BangPatterns, LambdaCase #-}
+
+module GIJoe.ApproximateInference.MaxPath where
+
+import GIJoe.Grammar
+import GIJoe.Types
+import GIJoe.Parse
+
+import GIJoe.ApproximateInference.HashTable (HashTable)
+import qualified GIJoe.ApproximateInference.HashTable as HashTable
+
+import Numeric.Log
+
+import Data.Maybe
+import Data.List (nub)
+
+import Data.HashMap.Strict (HashMap)
+import qualified Data.HashMap.Strict as HashMap
+
+import Data.IntSet (IntSet)
+import qualified Data.IntSet as IntSet
+
+import Control.Monad.Loop
+
+import Control.Monad
+import Control.Monad.Trans
+import Control.Monad.ST
+import Control.Applicative
+import Control.Monad.State.Strict
+
+import Data.PSQueue (PSQ, Binding(..))
+import qualified Data.PSQueue as PSQ
+
+import Debug.Trace
+
+type WeightTable s = HashTable s (Symbol, Symbol) Double'
+
+
+mkChildWeights :: Grammar -> ST s (WeightTable s)
+-- | Return a hashtable which maps a (_A, _B) to the probability that
+-- either direct child of _A is a _B. We collapse all terminals into
+-- one special terminal (T "!").
+mkChildWeights gr = do
+  !ht <- HashTable.newSized $ k * (k + h) 
+  exec_ $ do
+    rule <- forEach $ grammarRules gr
+    case rule of
+      BinaryRule h l r w ->
+        do lift $ HashTable.insertWith ht (+) (h, l) w
+           when (l /= r) $ 
+             lift $ HashTable.insertWith ht (+) (h, r) w
+      UnaryRule h _ w -> lift $ HashTable.insertWith ht (+) (h, t) w
+  return ht 
+  where
+    -- because these are IntSets, we are not overcounting the rule _A
+    -- -> _B_B
+    nts = nonterminals gr
+    terms = terminals gr
+    k = length nts
+    h = length terms
+    t = T "!"
+
+type NegLogProb = Double
+data OpenElement = Root Int | NT (Symbol, Int) deriving (Ord, Show, Eq)
+type Open =  PSQ OpenElement NegLogProb
+type DistanceTable = HashMap (Symbol, Int) NegLogProb
+
+singleSourceMaxWeight gr _A n =
[…]

### 2014-09-17 src/GIJoe/ComplexityTest.hs

@@ -1,0 +1,32 @@
+
+module Main where
+
+import Prelude hiding (lookup)
+
+import GIJoe.Grammar
+import GIJoe.Parse
+import GIJoe.InsideOutside
+
+import System.Random
+import Control.Monad.Random
+import Control.Monad
+import Control.Monad.Trans
+import Data.List hiding (lookup)
+import Data.Function
+
+main = do
+  let n = 50
+  let xs = stringToSentence $ ('a':) $ concat $ replicate (n-1) " a"
+  let lexicon = [T "a"]
+  let _K = 5
+  let _S = N 0 Nothing
+  let gr = initGrammar _S lexicon _K
+  let alphaChart = {-# SCC "alphaChart" #-} alphas gr 1000000 xs
+  putStrLn "Complexity Test"
+  putStrLn $ "K: " ++ show _K
+  putStrLn $ "n: " ++ show n
+  putStrLn $ "xs: " ++ show xs
+  forM [(i, j, t) | i <-[1..n], j<-[i..n], t <- allNonTerminals gr] $ \(i, j, t) ->
+    putStrLn $ "alpha " ++ show i ++ " " ++ show j ++ " n " ++ show t  ++ " : " ++ show (lookup i j t alphaChart)
+  -- let betaChart = {-# SCC "betaChart" #-} betas gr _S 1000000 xs alphaChart      
+  -- forM [1..n] $ \i -> putStrLn $ "betas 1 n _S: " ++ show (lookup i i _S betaChart)

### 2014-09-17 src/GIJoe/EM.hs

@@ -1,0 +1,10 @@
+
+
+module GIJoe.EM where
+
+
+import GIJoe.Types
+import GIJoe.Utils
+import GIJoe.Chart
+import GIJoe.Grammar
+

### 2014-09-17 src/GIJoe/Grammar.hs

@@ -121,6 +121,12 @@ data Grammar = Grammar {headSymbolTable :: HashMap Symbol IntSet,
 grammarRules :: Grammar -> [Rule]
 grammarRules gr = IntMap.elems . ruleIndex $ gr
 
+unaryRules :: Grammar -> [Rule]
+unaryRules gr = getRulesById gr $ unaryRuleIds gr
+
+binaryRules :: Grammar -> [Rule]
+binaryRules gr = getRulesById gr $ binaryRuleIds gr
+
 getRuleById :: Grammar -> Int -> Maybe Rule
 getRuleById gr i = IntMap.lookup i (ruleIndex gr)
 

### 2014-09-17 src/GIJoe/InsideOutside.hs

@@ -1,0 +1,520 @@
+{-# Language BangPatterns, ParallelListComp, FlexibleInstances, DeriveGeneric #-}
+
+module GIJoe.InsideOutside where
+
+import GIJoe.Types
+import GIJoe.Utils
+import GIJoe.Chart
+import GIJoe.Grammar
+
+
+import Prelude hiding (sum, lookup)
+import qualified Prelude
+
+import Numeric.Log 
+
+import Control.Monad
+import Data.List.Split (splitOn)
+import Data.List (foldl', sortBy, sort, maximumBy, foldl1')
+import qualified Data.Foldable as Foldable (toList)
+import Control.Monad.Logic
+import Data.Function (on)
+
+import Control.Monad.State.Strict
+import Control.Monad.Random
+import Control.Monad.Identity
+import Control.Parallel.Strategies 
+import qualified Control.DeepSeq as DeepSeq (NFData, rnf)
+import Numeric.SpecFunctions (digamma)
+
+import Data.HashMap.Strict (HashMap, (!))
+import qualified Data.HashMap.Strict as HashMap
+
+import Data.IntMap.Strict (IntMap)
+import qualified Data.IntMap.Strict as IntMap
+
+import Data.IntSet (IntSet) 
+import qualified Data.IntSet as IntSet
+
+import GHC.Generics (Generic)
+import Data.Hashable
+
+import Data.Tree
+import Data.Tree.Pretty 
+
+import Debug.Trace
+
+import qualified Data.Sequence as Seq
+
+type Charts a = [Chart a]
+newtype Chart a = Chart {unChart :: HashMap (Int, Int) (HashMap Symbol a)}
+                    deriving Show
+
+lookup :: Int -> Int -> Symbol -> Chart a -> Maybe a
+lookup i j sym (Chart m) = do m_sym <- HashMap.lookup (i, j) m
+                              HashMap.lookup sym m_sym
+
+{-# INLINE lookupDefault #-}
+lookupDefault :: Int -> Int -> Symbol -> a -> Chart a  -> a
+lookupDefault i j sym def chart = case lookup i j sym chart of
+  Nothing -> def
+  Just a -> a
+
+insert :: Show a => Int -> Int -> Symbol -> a -> Chart a -> Chart a
+insert i j sym a (Chart m_ij) = Chart m_ij'
+  where m_ij' =
+          case HashMap.lookup (i, j) m_ij of
[…]

### 2014-09-17 src/GIJoe/MCFG.hs

@@ -1,0 +1,41 @@
+{-# LANGUAGE TemplateHaskell #-}
+
+module GIJoe.MCFG where
+
+import Control.Lens
+import Data.List (intercalate)
+import Numeric.Log
+
+type Double' = Log Double
+
+data Var = Var {_varInt :: Int} deriving (Eq, Ord)
+makeLenses ''Var
+
+instance Show Var where
+  show (Var i) = "x" ++ show i
+  
+type Symbol = String
+data Atom = Atom {_atomPred :: Symbol,
+                  _atomVars :: [Var]} deriving (Eq, Ord)
+makeLenses ''Atom                                               
+instance Show Atom where
+  show (Atom s vs) = s ++ "(" ++ intercalate " " (map show vs) ++ ")"
+  
+data Rule = Rule {_lhs :: Atom,
+                  _rhs :: [Atom]} deriving (Eq, Ord)
+makeLenses ''Rule                                           
+instance Show Rule where
+  show (Rule l rs) = show l ++ " <- " ++ intercalate " " (map show rs)
+
+data WeightedRule = WRule {_wRule :: Rule,
+                           _wRuleWeight :: Double'} deriving Show
+makeLenses ''WeightedRule
+
+data MCFG = MCFG {_start :: Symbol,
+                  _rules :: [WeightedRule]} deriving Show
+makeLenses ''MCFG
+
+ 
+
+                                                     
+

### 2014-09-17 src/GIJoe/RunEmail.hs

@@ -1,0 +1,43 @@
+
+module Main where
+
+import GIJoe.Grammar
+import GIJoe.Parse
+import GIJoe.InsideOutside
+
+import System.Random
+import Control.Monad.Random
+import Control.Monad
+import Control.Monad.Trans
+import Data.List
+import Data.Function
+
+main = do
+  xs <- readFile
+       "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/emails.txt"
+  let corpus = take 1 $  map stringToSentence . lines $ xs
+  let _N0 = N 0 Nothing
+  let lexicon = nub $ concat corpus
+      _K = 5
+      gr = initGrammar _N0 lexicon _K
+      prune_limit = 1000000
+  gr' <- evalRandIO $ randomizeGrammar gr
+  let emlog = em gr' _N0 100000 corpus 1 1e-3
+  print $ map snd emlog
+  let gr_final = fst . last $ emlog
+  let gr_out_path = "/Users/edechter/Dropbox/Projects/GrammarInduction/data/grammars/emailGrammar1.grammar"
+
+  writeGrammar gr_out_path "grammar induction, K=10, 30 iterations" gr_final
+
+
+  -- ys <- readFile
+  --      "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/emails.txt"
+  -- let test_corpus = map stringToSentence . lines $ ys
+  -- sequence $ do
+  --   xs <- test_corpus
+  --   let (prs, _) = withCharts gr_final _N0 prune_limit xs $ getMapParse
+  --   return $ print prs
+  -- return ()
+
+          
+

### 2014-09-17 src/GIJoe/RunNumberOld.hs

@@ -1,0 +1,39 @@
+
+module Main where
+
+import GIJoe.Grammar
+import GIJoe.Induction hiding (main)
+
+import System.Random
+import Control.Monad.Random
+import Control.Monad
+import Control.Monad.Trans
+import Data.List
+import Data.Function
+
+
+
+main = do
+  xs <- readFile
+        "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/simpleNumbers.txt"
+  let corpus = map stringToSentence . lines $ xs
+  print corpus
+  let lexicon = nub $ concat corpus
+      _K = 6
+      gr = initGrammar _ROOT lexicon _K
+      prune_limit = 1000000000
+  print lexicon
+  gr' <- evalRandIO $ randomizeGrammar gr
+  print gr'
+  let emlog = em gr' _ROOT prune_limit corpus 10 0
+  print $ map snd emlog
+  let gr_final = fst . last $ emlog
+
+  let test_corpus = corpus
+  sequence $ do
+    xs <- test_corpus
+    let (prs, _) = withCharts gr_final _ROOT pruneLimit xs $ getMapParse
+    return $ print prs
+  return ()
+
+          

### 2014-09-17 src/GIJoe/SRS/Parse.hs

@@ -1,0 +1,129 @@
+{-# Language LambdaCase #-}
+
+module GIJoe.SRS.Parse where
+
+import GIJoe.SRS.Type
+
+import Control.Monad
+import Numeric.Log
+import qualified Data.HashMap.Strict as HashMap
+
+import Data.Char
+
+import Data.Text (pack)
+
+import Text.Parsec.Prim hiding (State)
+import Text.Parsec hiding (parse, State)
+import Text.ParserCombinators.Parsec.Number
+
+
+readSystem :: FilePath -> IO RewriteSystem
+readSystem fpath = do xs <- readFile fpath
+                      case parseSystem fpath xs of
+                        Right sys -> return sys
+                        Left err -> error $ show err
+
+-- writeSystem fpath info sys = writeFile fpath (unlines (comment:ls))
+--   where ls = map go (grammarRules gr)
+--         go (BinaryRule h l r w) =
+--           "$" ++ show h ++ " --> " ++ "$" ++ show l ++ " " ++ "$" ++ show r ++ " :: " ++ show w
+--         go (UnaryRule h c w) =
+--           "$" ++ show h ++ " --> " ++ show c ++ " :: " ++ show w
+--         comment = '#':info
+                         
+parseSystem :: String -- ^ source name
+            -> String -- ^ system content
+            -> Either ParseError RewriteSystem
+parseSystem source xs = parse systemParser source xs
+
+comment = do spaces
+             char '#'
+             _ <- manyTill anyToken eol
+             return ()
+             
+emptyLine = inlineSpaces >> eol >> return ()
+
+eol = char '\n'
+
+inlineSpaces = let oneSpace = satisfy $ \c -> c == ' ' || c == '\t'
+               in many oneSpace
+                  
+systemParser = do rs <- manyTill lineParser eof
+                  let rs' = getJusts rs
+                  return $! systemFromRules rs'
+  where getJusts [] = []
+        getJusts ((Just x):xs) = x:(getJusts xs)
+        getJusts (Nothing:xs) = getJusts xs
+
+lineParser = do {try (try comment <|> emptyLine); return Nothing }
+             <|>
+             do {r <- ruleParser; return $ Just r}
+
+ruleParser   = do inlineSpaces
+                  h <- cterm
+                  inlineSpaces
+                  ss <- optionMaybe rhs >>= \case
+                    Just ss -> return ss
+                    Nothing -> return []
+                  w <- optionMaybe weightAnotation >>= \case
+                    Just w -> return w
[…]

### 2014-09-17 src/GIJoe/SRS/Query.hs

@@ -1,0 +1,15 @@
+
+module GIJoe.SRS.Query where 
+
+import GIJoe.SRS.Type
+import Control.Monad.Logic
+
+import Data.Map (Map)
+import qualified Data.Map as Map
+  
+debug = False 
+
+write :: a -> String -> a 
+write x  s = if debug then trace s x else x
+
+match :: MonadLogic m => Term -> Term -> m a 

### 2014-09-17 src/GIJoe/SRS/Sample.hs

@@ -1,0 +1,52 @@
+
+module GIJoe.SRS.Sample where
+
+import GIJoe.SRS.Type
+
+import Control.Monad
+import Control.Monad.Random
+import Control.Applicative ((<$>))
+
+import Data.Map (Map, (!))
+import qualified Data.Map as Map
+
+import Debug.Trace
+
+---- SAMPLING -------
+
+sampleRule :: MonadRandom m => RewriteSystem -> Rule -> m ComplexTerm
+sampleRule rs (headTerm :<-: []) = return headTerm
+sampleRule rs (headTerm :<-: bodyTerms) = do
+  subst <- Map.unions `liftM` mapM (sampleTermSubst rs) bodyTerms
+  return $ headTerm `substCTermVars` subst
+  where vars = cTermVars headTerm
+
+sampleTerm rs term@(SimpleTerm pred vars) = do
+  subst <- sampleTermSubst rs term
+  return $ (ComplexTerm pred (map (NTString . (:[]) . ElVar) vars))  `substCTermVars` subst
+  
+sampleTermSubst :: MonadRandom m => RewriteSystem -> SimpleTerm -> m (Map Var NonTerminalString)
+sampleTermSubst rs (SimpleTerm pred vars) = do
+    i <- sampleCategorical $ normalizeWeights weights
+    let (WRule rule w) = matchingRules !! i
+    (ComplexTerm _ xs) <- sampleRule rs rule
+    return $ Map.fromList (zip vars xs)
+  where matchingRules = filter ((== pred) . cTermPredicate . ruleHead . weightedRuleRule)
+                          $ rsRules rs
+        weights = map weightedRuleWeight matchingRules
+
+normalizeWeights ws= map (/_Z) ws
+  where _Z = sum ws
+
+sampleCategorical :: (MonadRandom m, Num a, Ord a, Random a) => [a] -> m Int
+sampleCategorical ws = do
+  r <- getRandomR (0, 1)
+  return $ go r $ zip (cumsum ws) [0..]
+  where go r ((w, i):[]) = i
+        go r ((w, i):ws) | r < w = i
+                         | otherwise = go r ws 
+        
+cumsum xs = reverse $ go [] xs
+  where go ys [] = ys
+        go [] (x:xs) = go [x] xs
+        go zs@(y:_) (x:xs) = go (x+y:zs) xs

### 2014-09-17 src/GIJoe/SRS/Type.hs

@@ -1,0 +1,104 @@
+{-# Language ScopedTypeVariables, TypeSynonymInstances, FlexibleInstances #-}
+
+module GIJoe.SRS.Type where
+
+import GIJoe.ApproximateInference.HashTable
+
+import Data.Text (Text)
+import qualified Data.Text as Text
+
+import Data.HashMap.Strict (HashMap)
+import qualified Data.HashMap.Strict as HashMap
+
+import Data.IntSet (IntSet)
+import qualified Data.IntSet as IntSet
+
+import Data.IntMap (IntMap)
+import qualified Data.IntMap as IntMap
+
+import Data.Map (Map)
+import qualified Data.Map as Map
+
+import Data.List (intercalate)
+
+import System.Random
+
+import Numeric.Log
+
+type Double' = Log Double
+
+data Predicate = Predicate {predName :: Text, predArity :: Int} deriving (Show, Ord, Eq)
+
+type Var = Text 
+
+type Symbol = Text
+
+data Element = ElVar {elToVar :: Var} | ElSym {elToSym :: Symbol} deriving (Ord, Eq)
+instance Show Element where
+  show (ElVar v) = Text.unpack v
+  show (ElSym s) = Text.unpack s
+
+newtype NonTerminalString = NTString {unNT :: [Element]} deriving (Ord, Eq)
+instance Show NonTerminalString where
+  show (NTString xs) = intercalate " " (map show xs)
+
+type TerminalString = [Symbol]
+
+data ComplexTerm = ComplexTerm {cTermPredicate :: Predicate, ctermArgs :: [NonTerminalString]} deriving (Ord, Eq)
+
+instance Show ComplexTerm where
+  show (ComplexTerm p as) = Text.unpack (predName p) ++ "(" ++ intercalate ", " (map show as) ++ ")"
+  
+data SimpleTerm  = SimpleTerm {sTermPredicate :: Predicate, stermArgs :: [Var]} deriving (Ord, Eq)
+instance Show SimpleTerm where
+  show (SimpleTerm p as) = Text.unpack (predName p) ++ "(" ++ intercalate ", " (map Text.unpack as) ++ ")"
+
+data Term = CTerm ComplexTerm | STerm SimpleTerm deriving (Show, Ord, Eq)
+
+data Rule = ComplexTerm :<-: [SimpleTerm] deriving (Ord, Eq)
+instance Show Rule where
+  show (h :<-: []) = show h ++ "."
[…]

### 2014-10-05 drafts/cfp/draft.tex
Added the cfp draft to version control.

\documentclass{article}

\usepackage{amsmath}
\title{The Latent Constructive Rewrite Model \\ or \\ The Latent Conceptual Grammar Model}
\author{Eyal Dechter \& Josh Rule \& Josh Tenenbaum}

\begin{document}
\maketitle

\section{Abstract}
How do people learn systems of related concepts from data? Consider,
for example, a school-age child's knowledge of number. This knowledge
supports a large number of related operations: the child knows which
words denote numbers, how to use those words to label cardinalities of
sets, how to translate between digit representations and word
representations of number, how to compare cardinalities, etc. The
child knows how to perform these operations over an unbounded set of
elements, even though the data that underlies this knowledge is small
and noisy. Here we propose that such systems of related concepts can
be learned via hierarchical probabilistic inference over probabilistic
string rewrite systems (PSRS) -- a mildly context-sensitive class of
probabilistic string gramamrs that corresponds to a restricted
fragment of stochastic logic programs (SLPs). Specifically, we extend
the variational expectation-maximization algorithm for Hierarchical
Dirichlet PCFGs to accomodate PSRSs. This algorithm we present enables
a tractible compromise between highly restrictive context-free
knowledge representations and the intractability of performing similar
inference procedures over general logic programs. Empirically, we show
how such restricted string rewrite systems can capture non-trivial
conceptual knowledge and how our algorithm is able to learn such
knowledge from data.

\section{Abstract 2}
What does it mean to understand a concept, and how do people come to
such an understanding.

\section{Abstract 3}
Evidence: 
  - ``the number after twenty seven is twenty six''
  -

\section{Introduction}
The vast majority of the models of concept acquisition that have been
proposed in the artificial intelligence, machine learning and
cognitive science literatures cannot represent the conceptual systems
that even young children learn. Take, for example, the school-age
child's knowledge of number. This child knows that there is no largest
number; that ``one million forty three'' and ``8439''denote numbers
but that ``one million forty billion'' and ``a82bd'' do not; that no
two cardinalities refer to the same quantity but that some number
words refer to the same cardinalities (e.g. ``two thousand'' and
``twenty-one hundred''). Moreover, the child knows how to guage the
approximate number of elements in a collection and how to count this
number exactly. In short, the child's knowledge of number consists of
a complex system of concepts over an unbounded set of entities
(i.e. the numbers).

Children are not born with this kind of knowledge, and, since number
is a relatively recent cultural invention, nor is it likely that
children possess innate domain specific background knowledge that
makes it easier to acquire. Children only ever see a tiny fraction of
all the numbers they can reason about and this evidence is biased,
noisy, and can be unreliable.

A fundamental challenge for cognitive science -- and for any field
interested in understanding human-like concept acquisition -- is to
provide a theory of how such complex systems of concepts can be
learned from sparse and noisy data.

People are able to learn about complex systems of related
concepts from small amounts of noisy data. A compelling approach to
understanding such learning is to frame it as Bayesian inference over
probabilistic knowledge bases. Advances in context free grammar
induction in the past couple of decades can be seen as an instance of
this approach, in which

Our goal is to understand how children can obtain semantic knowledge
about a system of concepts from the structure of relavent perceptual
stimuli. How does a child learn that prefixing ``great-'' corresponds
to adding an additional generation in kinship relationships? How does
a child learn that ``just before'' is after ``a while before'' which
is before ``a while after''? How does a child learn that if
Figure~\ref{pairOfApples} is ``a pair of apples'' then
Figure~\ref{singleApple} is a single apple?

Our proposal is that some interesting fragments of human conceptual
knowledge can be represented as structured relations between
linguistic tokens and other perceptual inputs; that the rules that
define these relations implicitly define the concepts that are
learned; and that these rules can be learned using the tools of
hierarchical Bayesian inference.

For example, the concept TWENTY SEVEN is defined as the collection of
representations ``27'', ``twenty seven'', ``3 times 9,'' along with a
system of symbolic rules that translate among these representations
and facilitate answered various queries. According to our hypothesis,
the difference between TWENTY EIGHT and TWENTY SEVEN is not that they
are represented in the mind by distinct mental symbols but rather that
the rules that operate on their various representations treat them in
systematically different ways.

This view implies that an attempt to construct a computational account
of concept learning in humans needs to ground out that learning in the
relavent perceptual stimuli.

Contribution:
\begin{itemize}
\item We present Generate Rewrite Systems (GRS) -- a formalism that borrows
  heavily from Linear Context-Free Rewriting Systems (CITE)
\item We show how GRS, though restricted in representational capactiy,
  can capture substantial cognitively plausible first-order conceptual knowledge 
\item We propose a heirarchical Bayesian model for GRS that provides a coherent framework in which a GRS can be learned given evidence
\item We show how GRS can be transformed into PRISM programs and learned within that probabilstic logic programming system
\item We provide computational experiments that provide evidence that
  our model can provide an account of some kinds of complex conceptual
  knowledge can be inferred from noisy and sparse data
\end{itemize}

\subsection{Related Work}

\section{The Model}
In this section we propose a hierarchical probabilistic model for
learning distributions over string relations. Suppose we are given
some positive instances $y_1, \dots, y_M$ of relations $R_1, \dots,
R_N$. Our model assumes that each positive instance
$y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent stochastic grammar
(LSG), $\mathcal{K}$. $\mathcal{K}$ consists of some fixed number $K$
of clauses (i.e. rules) $C_1, \dots, C_K$ over the $N$ observable
relations plus a fixed number of $H$ hidden relations $T_1, \dots,
T_H$. The \emph{architecture} of the model is the exact specification
of the parameters $K$ and $H$ and the clauses $mathcal{C}$. To each
clause $C_i$ is associated a positive real weights $\theta_i$.

The clauses of $\mathcal{K}$ are Horn clauses in a first order logic
with string concatenation. We will use PROLOG notation, in which lower
case letters are predicates and constants and uppercase letters are
variables. If two symbols are adjacent it means that the corresponding
strings are concatenated. For example, the following would is a valid
clause: $a(XY, XV) \leftarrow b(X, UV), a(X, X)$. In this rule $a$ and
$b$ are relations with arity two.

Let $C_1, \dots, C_{N(R)} = C(R)$ be the set of clauses whose head is
relation $R$ and let $\Theta_R$ be the set of corresponding
weights. Let $p_R(C_n) \sim \text{Discrete}(\Theta_R)$; that is, $p_R$
is a distribution over the clauses headed by $R$, such that each
clauses is sampled in proportion to its weight parameter.

A positive instance of this relation is sampled from $\mathcal{K}$ by
arbitrarily (i.e. uniformly at random) choosing a relation $R$ and
instantiating that relation by sampling a clause from $p_R$. Recursing
on this procedure for every atom in the body of this clause binds the
variables in the clause's head.

Given this assumed generative process for the evidence, learning a
program involves inferring the parameter $\theta$ given a prior
probability distribution over their values. We place an independent
Dirichlet distribution over the parameters of every relation. This is
a natural choice that enables us to promote sparse programs, i.e.,
programs in which most of the mass is placed on a small number of
predicates and rules.

\section{Inference}
Inference in hierarchical Dirichlet processes in stochastic grammars
and stochastic logic programs have been well studied. Variational
Expectation-Maximization (VB-EM) is a popular approach, and this is
the algorithm we adopt here. The algorithm is described extensively
elsewhere (CITATION). As we will show in the following section, our
latent stochastic grammar model can be encoded as a stochastic logic
program, so that any VB-EM algorithm for that domain will suffice.

\section{Implementation}
We implemented learning in LCG by encoding the model as a PRISM program.

TODO: Josh should write this section on how to translate an LCG into a
PRISM program.

\section{Experiments}
\section{Discussion}

\section{Outline}
1) Introduction
  - psychological motivation 
    - juxtaposing the kinds of knowledge that freebase has and the kinds that children learn
    - compositional semantic structure --> derive semantics from syntax
  - what is the contribution?
     - the use of string rewrite formalisms to capture conceptual knowledge and conceptual learning
     - to bring tools from computational linguistics to learning semantics
  - what kind of concept systems are we trying to capture?
  - why restrict the formalism? 
     - tractability of inference 
  - why go beyond context free grammars? 
     - if they are intractable to learn anyway, why try to go beyond them? 
  - Related Work
2) Model
- stochastic linear context free rewrite systems
  - model: what is the data? and what is being inferred? what is the prior? 
  - this can be written as PRISM program, or an SLP. 
  - inference: variational approximation 
3) Experiments
  a) worked example
  b) Number system (main result)
  3b) Results
4) Conclusion
5) Citations

\end{document}

### 2014-10-09 drafts/cfp/draft.tex
Changes to aaai symposium draft.

@@
 \subsection{Related Work}
+This work borrows heavily
 
 \section{The Model}
-In this section we propose a hierarchical probabilistic model for
-learning distributions over string relations. Suppose we are given
-some positive instances $y_1, \dots, y_M$ of relations $R_1, \dots,
-R_N$. Our model assumes that each positive instance
-$y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent stochastic grammar
-(LSG), $\mathcal{K}$. $\mathcal{K}$ consists of some fixed number $K$
-of clauses (i.e. rules) $C_1, \dots, C_K$ over the $N$ observable
-relations plus a fixed number of $H$ hidden relations $T_1, \dots,
-T_H$. The \emph{architecture} of the model is the exact specification
-of the parameters $K$ and $H$ and the clauses $mathcal{C}$. To each
-clause $C_i$ is associated a positive real weights $\theta_i$.
+An LCG is a hierarchical Bayesian model of string tuples. It is an
+extension of the Hierarhical Dirichlet PCFG model (HD-PCFG) to
+Probabilistic Range Concatenation Grammars (PRCGs). Whereas the
+HD-PCFG is a distribution over strings, an LCG is a distribution over
+string tuples belonging to one of a set of observable
+relations.
+
+Range Concatenation Grammars (RCGs) are a class of string grammars
+that represent all and only those languages that can be parsed in time
+polynomial in the length of the target
+string~\cite{boullier2005range}. An RCG $G=(N, T, V, P, S)$ is a
+5-tuple where $N$ is a finite set of predicate symbols, $T$ is a set
+of terminal symbols, $V$ is a set of variable symbols, P is a finite
+set of $M \geq 0$ clauses of the form $\psi_0 \rightarrow \psi_1 \dots
+\psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
+the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
+N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
+\cup V)^*$ is an argument of $\psi_m$. We call the left hand side term
+of any clause the \emph{head} of that clause and its predicate symbol
+is the \emph{head predicate}.
+
+\textbf{Example:} The following is an RCG for the 3-copy language
+$\{www | w \in {a, b}^*\}$:
+\begin{alignat*}{3}
+&S(XYZ) &&\rightarrow A(X, Y, Z)\\
+&A(aX, aY, aZ) &&\rightarrow A(X, Y, Z)\\
+&A(bX, bY, bZ) &&\rightarrow A(X, Y, Z)\\
+&A(\epsilon, \epsilon, \epsilon) &&\rightarrow 
+\end{alignat*}
+
+A string $x$ is in the language defined by an RCG if one can
+\emph{derive} $S(x)$. A derivation is a sequence of rewrite steps in
+which substrings of the left hand side argument string are bound to
+the variables of the head of some clause, thus determining the
+arguments in the clause body. If a clause has no body terms, then its
+head is derived; otherwise, its head is derived if its body clauses
+are derived~\footnote{This description of the language of an RCG
+  technically only holds for \emph{non-combinatory} RCGs, in which the
+  arguments of body terms can only contain single variables. Since any
+  \emph{combinatory} RCG can be converted into a non-combinatory RCG
+  and we only consider non-combinatory RCGs here, this description
+  suffices.}
+
+We extend RCGs to Probabilistic Range Concatenation Grammars (PRCGs)
+by annotating each clause $C_k \in P$ with probabilities $p_k$ such
+that for all predicates $A \in N$ $\sum_{k:head(C_k)=A} p_k = 1$. A
+PRCGs defines a distribution over strings $x$ by sampling from
+derivations of $S(x)$ according to the product of probabilities of
+clauses used in that derivation. This is a well defined distribution
+as long as no probability mass is placed on derivations of infinite
+length; in this paper, we only consider PRCGs with derivations of
+finite length, so we need not worry about this requirement.
+
+Now we can describe the model we introduce here, the Latent Concept
+Grammar (LCG). We introduce the LCG as a machine learning model to
+learn a distribution over instances of relations over strings. Unlike
+a PRCG, which places a distribution over terms $(
+
+RCGs consist of a set
+of rewrite rules, called clauses, each of which takes a termwhich
+operate on a set of terms (predicates applied to arguments) to produce
+new terms. Each rewrite rule is of the form:
+
+Suppose we are given some positive instances $y_1, \dots,
+y_M$ of relations $R_1, \dots, R_N$. Our model assumes that each
+positive instance $y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent
+stochastic grammar (LSG), $\mathcal{K}$. $\mathcal{K}$ consists of
+some fixed number $K$ of clauses (i.e. rules) $C_1, \dots, C_K$ over
+the $N$ observable relations plus a fixed number of $H$ hidden
+relations $T_1, \dots, T_H$. The \emph{architecture} of the model is
+the exact specification of the parameters $K$ and $H$ and the clauses
+$mathcal{C}$. To each clause $C_i$ is associated a positive real
+weights $\theta_i$.
 

### 2014-10-10 drafts/cfp/draft.tex
Changes to draft and initial bib file for draft.

@@
 \section{Abstract}
-How do people learn systems of related concepts from data? Consider,
-for example, a school-age child's knowledge of number. This knowledge
-supports a large number of related operations: the child knows which
-words denote numbers, how to use those words to label cardinalities of
-sets, how to translate between digit representations and word
-representations of number, how to compare cardinalities, etc. The
-child knows how to perform these operations over an unbounded set of
-elements, even though the data that underlies this knowledge is small
-and noisy. Here we propose that such systems of related concepts can
-be learned via hierarchical probabilistic inference over probabilistic
-string rewrite systems (PSRS) -- a mildly context-sensitive class of
-probabilistic string gramamrs that corresponds to a restricted
-fragment of stochastic logic programs (SLPs). Specifically, we extend
-the variational expectation-maximization algorithm for Hierarchical
-Dirichlet PCFGs to accomodate PSRSs. This algorithm we present enables
-a tractible compromise between highly restrictive context-free
-knowledge representations and the intractability of performing similar
-inference procedures over general logic programs. Empirically, we show
-how such restricted string rewrite systems can capture non-trivial
-conceptual knowledge and how our algorithm is able to learn such
-knowledge from data.
+A popular approach in statistical machine learning is to automatically
+find the latent \emph{features} of the data that are most relevant or
+informative. But a latent feature based view of concept learning does
+not match up naturally with the structured symbolic knowledge that
+seems to underlie people's understanding of many domains. How can we
+combine the flexibility, noise tolerance, and computational
+practicality of latent feature based statistical machine learning with
+the symbolic knowledge representations that best capture human
+understanding? Here present a learning model we dub Latent Concept
+Grammars (LCGs) as a proposal. An LCG is a densely connected set of
+probabilistic string rewrite rules over an initially uninterpreted set
+of predicates. When fit to data using a hierarchical Bayesian prior,
+these predicates assume the roles of informative latent concepts. This
+approach allows us to learn knowledge that is stochastic and
+recursively structured while maintaining computational
+tractability. We show how learning in LCGs can be implemented using
+the probabilistic logic programming system PRISM, and we present
+preliminary experiments.
+
+How do people learn systems of related concepts from
+data?  Consider, for example, a school-age child's knowledge of
+number. This knowledge supports a large number of related operations:
+the child knows which words denote numbers, how to use those words to
+label cardinalities of sets, how to translate between digit
+representations and word representations of number, how to compare
+cardinalities, etc. The child knows how to perform these operations
+over an unbounded set of elements, even though the data that underlies
+this knowledge is small and noisy. Here we propose that such systems
+of related concepts can be learned via hierarchical probabilistic
+inference over probabilistic string rewrite systems (PSRS) -- a mildly
+context-sensitive class of probabilistic string gramamrs that
+corresponds to a restricted fragment of stochastic logic programs
+(SLPs). Specifically, we extend the variational
+expectation-maximization algorithm for Hierarchical Dirichlet PCFGs to
+accomodate PSRSs. This algorithm we present enables a tractible
+compromise between highly restrictive context-free knowledge
+representations and the intractability of performing similar inference
+procedures over general logic programs. Empirically, we show how such
+restricted string rewrite systems can capture non-trivial conceptual
+knowledge and how our algorithm is able to learn such knowledge from
+data.
 
@@
 \section{Abstract 3}
-Evidence: 
-  - ``the number after twenty seven is twenty six''
-  -
 

### 2014-10-10 src/GIJoe/SRS/BuildNumber.hs
Various changes to BuildNumber.hs and Prismify.hs.
Prismify: fixed acyclicity check and changed main goal to take a list of target data goals rather than a file containing data goals.
BuildNumber: added a NextSentence rule. Also added functions for generating train/test data splits and spitting those out into files for prism consumption.

@@ -2,7 +2,8 @@ module GIJoe.SRS.BuildNumber where
 
 import Data.List
 import System.Random
-import System.Random.Shuffle (shuffle')
+import System.Random.Shuffle (shuffle', shuffleM)
+import Control.Monad.Random.Class
 
 -- for all i,j,k \in [1..n], w \in lexicon
 -- A_i(X,Y) <- A_j(X),A_k(Y).
@@ -58,14 +59,15 @@ grammarOfNumber k outpath = do
 
 grammarOfNumber2 :: Int -> FilePath -> IO ()
 grammarOfNumber2 n outpath = do
-  let lexicon = ones ++ teens ++ tens
+  let lexicon = ones ++ teens ++ tens ++ ["after", "comes"]
   let reorderRules = ["A" ++ show i ++ "(X Y, U V) <-- A" ++ show j ++ "(" ++ (ls !! 0) ++ "," ++ (ls !! 1) ++ "), A" ++ show k ++ "(" ++ (ls !! 2) ++ "," ++ (ls !! 3) ++ ")." | i <- [1..n], j <- [1..n], k <- [1..n], ls <- (permutations ["X","Y","U","V"]), j < k]
   let reorderRulesEqual = ["A" ++ show i ++ "(X Y, U V) <-- A" ++ show j ++ "(" ++ (ls !! 0) ++ "," ++ (ls !! 1) ++ "), A" ++ show j ++ "(" ++ (ls !! 2) ++ "," ++ (ls !! 3) ++ ")." | i <- [1..n], j <- [1..n], ls <- ((map (\x -> "X":x) (permutations ["Y","U","V"])) ++ (map (\y -> "Y":y) (permutations ["X","U","V"])))]
 --  let terminalRules = ["word" ++ "("++ w ++ ")."| w <- lexicon]
   let relationRules = ["A" ++ show i ++ "(" ++ w1 ++ "," ++ w2 ++ ")." | i <- [1..n], w1 <- lexicon ++ ["null"] , w2 <- lexicon ++ ["null"]]
   let numberRule = ["Number(X) <-- A1(X,X)."]
   let nextRule = ["Next(X,Y) <-- A2(X,Y)."]
-  writeFile outpath $ unlines $ reorderRules ++ reorderRulesEqual ++ relationRules ++ numberRule ++ nextRule
+  let nextSentenceRule = ["NextSentence(X) <-- A2(X,Y)."]      
[…]

### 2014-10-10 src/GIJoe/SRS/Prismify.hs

@@ -45,7 +45,8 @@ body slcfrs =
     sameHead =  (\(WRule ((ComplexTerm p1 _) :<-: _) _)
                   (WRule ((ComplexTerm p2 _) :<-: _) _)
                  -> p1 == p2)
-    mainClause = [ "main(Fin,Foutp,Fouta) :- open(Fin,read,S), read(S,Gs), close(S), set_prism_flag(restart,10), set_prism_flag(learn_mode,vb), set_prism_flag(viterbi_mode,vb), set_prism_flag(default_sw_a,uniform), set_prism_flag(log_scale,on), learn(Gs), save_sw(Foutp), save_sw_a(Fouta).", "", "show_num(X) :- set_prism_flag(rerank,20), n_viterbig(20,srs('Number_1'-[X])).", "", "show_next(X,Y) :- set_prism_flag(rerank,20), n_viterbig(20,srs('Next_2'-[X,Y]))."]
+    mainClause = [ "main(Gs,Foutp,Fouta) :- set_prism_flag(restart,1), set_prism_flag(learn_mode,vb), set_prism_flag(viterbi_mode,vb), set_prism_flag(default_sw_a,uniform), set_prism_flag(log_scale,on), learn(Gs), save_sw(Foutp), save_sw_a(Fouta).",
+                  "", "show_num(X) :- set_prism_flag(rerank,20), n_viterbig(20,srs('Number_1'-[X])).", "", "show_next(X,Y) :- set_prism_flag(rerank,20), n_viterbig(20,srs('Next_2'-[X,Y]))."]
     srsClause =  [ "srs(P-IN) :- msw(P,V), reduce(P-IN,V)." ]
     acyclicClause = ["acyclic([A,B],[C,D]) :- length(A,AL), length(C,CL), AL < CL; length(B,BL), length(D,DL), BL < DL." ]
     switchClauses = map makeSwitch groupedRules

### 2014-10-14 drafts/cfp/draft.tex
initial attempt at next sentences

@@
 learn a distribution over instances of relations over strings. Unlike
-a PRCG, which places a distribution over terms $(
-
-RCGs consist of a set
+a PRCG, which places a distribution over terms RCGs consist of a set
 of rewrite rules, called clauses, each of which takes a termwhich

### 2014-10-14 src/GIJoe/SRS/BuildNumber.hs

@@ -18,8 +18,6 @@ grammarOfNumber k outpath = do
   let constantRule = ["Number(X) <-- A(X)."]
   writeFile outpath $ unlines $ binaryRules ++ unaryRules ++ constantRule
 
-
-
 -- even if we generalize, we don't need every rule
 -- for all A,B,C
 -- A(X,Y) <- B(Y,X).          # #4# flip
@@ -66,8 +64,8 @@ grammarOfNumber2 n outpath = do
   let relationRules = ["A" ++ show i ++ "(" ++ (lexicon !! w1) ++ "," ++ (lexicon !! w2) ++ ")." | i <- [1..n], w1 <- [0..((length lexicon) - 1)] , w2 <- [0..((length lexicon) - 1)], w1 <= w2]
   let numberRule = ["Number(X,Y) <-- A1(X,Y)."]
   let nextRule = ["Next(X,Y) <-- A2(X,Y)."]
-  let nextSentenceRule = ["NextSentence(X) <-- A2(X,Y)."]      
-  writeFile outpath $ unlines $ reorderRules ++ reorderRulesEqual ++ relationRules ++ numberRule ++ nextRule ++ nextSentenceRule
+  let nextSentenceRule = ["Next(X,Y) <-- A2(X,Y)."]      
+  writeFile outpath $ unlines $ reorderRules ++ reorderRulesEqual ++ relationRules ++ numberRule ++ nextSentenceRule
 
 writeWrapper :: Int -> Int -> FilePath -> IO ()
 writeWrapper nNu nNe outpath = do
@@ -92,8 +90,6 @@ giveNumbers n c = do
     return $ map (\x -> concat ["count(srs('Number_2'-[", showCleanList x, ",[]]),", show c, ")"]) nShuffledNumbers
 
 showCleanList xs = "[" ++ intercalate "," xs ++ "]"
-
-
 
 giveNexts :: Int -> Int -> IO [String]
 giveNexts n c = do
@@ -142,15 +138,11 @@ mkTrainTestDataFile path specs = do
   let out = unlines ss
   writeFile path out
 
-numberExamples = [[x] | x <- numberLists]
+numberExamples = [[x,[""]] | x <- numberLists]
 nextExamples = [[a, b] | (a, b) <- zip (init numberLists) (tail numberLists)]
 
 -- Example: to generate file of 90% of the numbers and 20% of the nexts as training:
 --  mkTrainTestDataFile <path>
[…]

### 2014-10-14 src/GIJoe/SRS/Prismify.hs

@@ -46,7 +46,8 @@ body slcfrs =
                   (WRule ((ComplexTerm p2 _) :<-: _) _)
                  -> p1 == p2)
     mainClause = [ "main(Gs,Foutp,Fouta) :- set_prism_flag(restart,1), set_prism_flag(learn_mode,vb), set_prism_flag(viterbi_mode,vb), set_prism_flag(default_sw_a,uniform), set_prism_flag(log_scale,on), learn(Gs), save_sw(Foutp), save_sw_a(Fouta).",
-                  "", "show_num(X) :- set_prism_flag(rerank,20), n_viterbig(20,srs('Number_1'-[X])).", "", "show_next(X,Y) :- set_prism_flag(rerank,20), n_viterbig(20,srs('Next_2'-[X,Y]))."]
+                  "", "getTrain(F,Gs) :- load_clauses(F,ALL,[]), findall(X,member(train(X),ALL),Gs).",
+                  "", "getTest(F,Gs) :- load_clauses(F,ALL,[]), findall(X,member(test(X),ALL),Gs).", ""]
     srsClause =  [ "srs(P-IN) :- msw(P,V), reduce(P-IN,V)." ]
     acyclicClause = [ "acyclic([A,B],[C,D]) :- length(A,AL), length(B,BL), length(C,CL), length(D,DL), X is AL + BL, Y is CL + DL, X < Y." ]
     switchClauses = map makeSwitch groupedRules

### 2014-10-14 src/GIJoe/SRS/BuildNumber.hs
Continued modifications to code base

@@ -4,6 +4,12 @@ import Data.List
 import System.Random
 import System.Random.Shuffle (shuffle', shuffleM)
 import Control.Monad.Random.Class
+import Control.Monad
+
+import GIJoe.SRS.Type
+import GIJoe.SRS.Parse
+import GIJoe.SRS.Prismify
+import Data.Text (pack)
 
 -- for all i,j,k \in [1..n], w \in lexicon
 -- A_i(X,Y) <- A_j(X),A_k(Y).
@@ -69,6 +75,48 @@ grammarOfNumber2 n outpath = do
   let nextSentenceRule = ["NextSentence(X) <-- A2(X,Y)."]      
   writeFile outpath $ unlines $ reorderRules ++ reorderRulesEqual ++ relationRules ++ numberRule ++ nextRule ++ nextSentenceRule
 
+buildGrammar :: Int -- ^ number of hidden predicates
+             -> [String] -- ^ lexicon
+             -> [String] -- ^ list of observed unary predicates
+             -> FilePath
+             -> IO ()
+buildGrammar n lexicon observed outpath = do
+  writeFile outpath $ show rs
+  writeFile (outpath ++ ".psm") $ "srs(P-IN) :- reduce(P-IN,V), msw(P,V).\n\n" ++ prismClauses rs
+  where rs = fromRight $ parseSystem "buildGrammar" $ systemString
+        fromRight (Right x) = x
+        fromRight (Left err) = error $ show err
+        systemString = unlines $ 
+                       (concat [rules i j k | i <- [1..n], j <-[1..n], k <- [j..n]])++
+                       [ruleLex i w v | i <- [1..n],
+                        (wi, w) <- zip [0..] lexicon,
+                        (vi, v) <- zip [0..] lexicon,
+                        vi >= wi] ++ 
+                       [ruleLex i w "null" | i <- [1..n], w <- lexicon] ++ 
+                       [ruleLex i "null" w | i <- [1..n], w <- lexicon] ++
+                       [pred ++ "(X Y) <-- A" ++ show i ++ "(X, Y)." | (pred, i) <- zip observed [1..]]
+        rules i j k | j /= k  = do
+          [v1, v2, v3, v4] <- permutations ["X", "Y", "U", "V"]
+          return $ "A" ++ show i ++ "(X Y, U V) <-- " ++
[…]

### 2014-10-14 src/GIJoe/SRS/Prismify.hs

@@ -1,4 +1,4 @@
-module GIJoe.SRS.Prismify (prismify) where
+module GIJoe.SRS.Prismify (prismify, prismClauses) where
 
 import Data.Function (on)
 import Data.List
@@ -7,7 +7,16 @@ import Data.Text (unpack)
 import GIJoe.SRS.Type
 import GIJoe.SRS.Parse
 
--- utils
+
+-- type PrismScript = String
+-- type PrismFlag = (String, String)
+-- setPrismFlag :: PrismScript
+--                -> PrismFlag
+--                -> PrismScript
+-- setPrismFlag script (a, b) = script ++ [":- set_prism_flag(" ++
+--                                         show a ++ ", " ++ show b ++ ").\n"]
+
+
 
 fst3 (a,_,_) = a
 snd3 (_,b,_) = b
@@ -33,10 +42,27 @@ labeledMap f xs =  map f $ zip xs $ take (length xs) ['A'..'Z']
 --  where
 --    commented = map (\x -> "%% " ++ show x)
 
+
+
+prismClauses :: RewriteSystem -> String
+prismClauses rs = unlines $ 
+  switchClauses ++ [""] ++
+  probClauses ++ [""] ++
+  reductionClauses
+  where switchClauses = map makeSwitch groupedRules
+        probClauses = map makeProbs groupedRules
+        reductionClauses = concatMap makeReductions groupedRules
+        groupedRules = groupBy sameHead $
+                       sortBy (compare `on`
+                               (cTermPredicate . ruleHead . weightedRuleRule))
+                         (rsRules rs)
+        sameHead =  (\(WRule ((ComplexTerm p1 _) :<-: _) _)
+                  (WRule ((ComplexTerm p2 _) :<-: _) _)
+                 -> p1 == p2)
+
 body :: RewriteSystem -> [String]
 body slcfrs =
     mainClause ++ [""] ++
-    acyclicClause ++ [""] ++
     srsClause ++ [""] ++
     switchClauses ++ [""] ++
     probClauses ++ [""] ++ reductionClauses
@@ -48,7 +74,6 @@ body slcfrs =
     mainClause = [ "main(Gs,Foutp,Fouta) :- set_prism_flag(restart,1), set_prism_flag(learn_mode,vb), set_prism_flag(viterbi_mode,vb), set_prism_flag(default_sw_a,uniform), set_prism_flag(log_scale,on), learn(Gs), save_sw(Foutp), save_sw_a(Fouta).",
[…]

### 2014-10-14 src/GIJoe/SRS/BuildNumber.hs
compiling, but not learning, for next sentences

@@ -5,20 +5,7 @@ import System.Random
 import System.Random.Shuffle (shuffle', shuffleM)
 import Control.Monad.Random.Class
 
--- for all i,j,k \in [1..n], w \in lexicon
--- A_i(X,Y) <- A_j(X),A_k(Y).
--- A_i([w]).
--- Number <-- A_1(X).
-
-grammarOfNumber :: Int -> FilePath -> IO ()
-grammarOfNumber k outpath = do
-  let lexicon = ["one","two","three","four","five","six","seven","eight","nine","ten","eleven","twelve","thirteen","fourteen","fifteen","sixteen","seventeen","eighteen","nineteen","twenty","thirty","forty","fifty","sixty","seventy","eighty","ninety"]
-  let binaryRules = ["A" ++ show i ++ "(X Y) <-- A" ++ show j ++ "(X), A" ++ show ki ++ "(Y)." | i <- [1..k], j <- [1..k], ki <- [1..k]]
-  let unaryRules = ["A" ++ show i ++ "("++ w ++ ")."| i <- [1..k], w <- lexicon]
-  let constantRule = ["Number(X) <-- A(X)."]
-  writeFile outpath $ unlines $ binaryRules ++ unaryRules ++ constantRule
-
--- even if we generalize, we don't need every rule
+-- while we don't need every rule
 -- for all A,B,C
 -- A(X,Y) <- B(Y,X).          # #4# flip
 -- A(12,34) <- B(1,2),C(3,4). # #1# both from a single predicate
@@ -45,36 +32,25 @@ grammarOfNumber k outpath = do
 -- A(12,34) <- B(4,2),C(3,1). # =2, reordered, flip B, flip C
 -- A(12,34) <- B(4,3),C(1,2). # =1, reordered, flip B
 -- A(12,34) <- B(4,3),C(2,1). # =1, reordered, flip B, flip C
-
+--
+-- we use every rule to avoid cycles in the explanation graph from flipping
 -- so, the full grammar is:
--- for all i,j,k \in [1..n], w \in lexicon
--- A_i(12,34) <-- A_j(1,2),A_k(3,4).
--- A_i(12,34) <-- A_j(1,3),A_k(2,4).
--- A_i(12,34) <-- A_j(1,4),A_k(2,3).
--- A_i(X,Y) <-- A_j(Y,X).
[…]

### 2014-10-14 src/GIJoe/SRS/Prismify.hs

@@ -36,7 +36,6 @@ labeledMap f xs =  map f $ zip xs $ take (length xs) ['A'..'Z']
 body :: RewriteSystem -> [String]
 body slcfrs =
     mainClause ++ [""] ++
-    acyclicClause ++ [""] ++
     srsClause ++ [""] ++
     switchClauses ++ [""] ++
     probClauses ++ [""] ++ reductionClauses
@@ -49,7 +48,6 @@ body slcfrs =
                   "", "getTrain(F,Gs) :- load_clauses(F,ALL,[]), findall(X,member(train(X),ALL),Gs).",
                   "", "getTest(F,Gs) :- load_clauses(F,ALL,[]), findall(X,member(test(X),ALL),Gs).", ""]
     srsClause =  [ "srs(P-IN) :- msw(P,V), reduce(P-IN,V)." ]
-    acyclicClause = [ "acyclic([A,B],[C,D]) :- length(A,AL), length(B,BL), length(C,CL), length(D,DL), X is AL + BL, Y is CL + DL, X < Y." ]
     switchClauses = map makeSwitch groupedRules
     probClauses = map makeProbs groupedRules
     reductionClauses = concatMap makeReductions groupedRules
@@ -82,8 +80,8 @@ makeReduction (i,(h :<-: bs)) =
     theAppends = if (not $ null appendVars)
                  then "(" ++ appendVars ++ " -> " ++ appends1 ++ "; " ++ appends2 ++ ")"
                  else appendSRSs
-    appends1 = intercalate ", " $ filter (not . null) [appendSRSs, appendAppends, appendAcyclics]
-    appends2 = intercalate ", " $ filter (not . null) [appendAppends, appendAcyclics, appendSRSs]
+    appends1 = intercalate ", " $ filter (not . null) [appendSRSs, appendAppends]
+    appends2 = intercalate ", " $ filter (not . null) [appendAppends, appendSRSs]
     anyAppends = if null (theAppends) then "" else " :- " ++ theAppends
     collectTupleItems (ComplexTerm _ xs) = intercalate ", " $ filter (not . null) $ labeledMap collectTupleItem xs
     collectTupleItem ((NTString (x:[])),l) = ""

### 2014-10-17 drafts/cfp/draft.tex
Continued modifications to code base
translation and related work in draft, prismify and grammar updates

@@
 \usepackage{amsmath}
-\title{The Latent Constructive Rewrite Model \\ or \\ The Latent Conceptual Grammar Model}
+\usepackage{listings}
+\usepackage{anyfontsize}
+\usepackage{subcaption}
+\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
 \author{Eyal Dechter \& Josh Rule \& Josh Tenenbaum}
@@
 such an understanding.
-
-\section{Abstract 3}
 
@@
 \subsection{Related Work}
-This work borrows heavily
+Our core inference algorithm borrows heavily from recent algorithms for nonparametric PCFG induction, as well as work extending that algorithm to SLPs (CITATION Johnson; Liang; Cussens, 2001).
+Like Johnson and Liang, we are interested in inferring the most likely grammar given some collection of sentences, and in doing so without a pre-specified limit on the complexity of the grammar.
+Like Cussens, we are interested in efficiently extending this algorithm to more powerful and expressive formalisms than CFGs.
+
+This work is also closely related to the growing field of semantic parsing.
+Both are aimed at jointly learning syntactic and semantic information from a body of sentences.
+Semantic parsers, however, learn separate systems of syntax and semantics, often pairing Combinatory Categorial Grammars of syntax with an explicit logical language of semantics, often the untyped lambda calculus (CITATION Steedman; Liang; Zettlemoyer; etc.).
+By contrast, our approach uses a single formalism, the latent predicate network, to describe both which sentences are syntactically valid and what sorts of semantic information those sentence imply.
+Semantic parsers also typically learn models of semantics by pairing sentences with an explicit logical form or real-world denotation.
+In our model it is the grammar induction itself which reveals candidate relations between words that are key to capturing their meaning.
+
+More generally, we see this model as helping to strike a balance between expressivity and tractability, bridge the gap between more traditional forms of grammar induction (CITATION) and program induction over more powerful languages (CITATION E.C.
+paper, other P.I.
+papers).
 
@@
 \section{Implementation}
-We implemented learning in LCG by encoding the model as a PRISM program.
 
-TODO: Josh should write this section on how to translate an LCG into a
-PRISM program.
+\lstset{
+    language=Prolog,
+    basicstyle=\fontsize{5}{6}\selectfont\ttfamily
+}
+
+\begin{figure}[t]
+	\centering
+	\begin{minipage}[b]{0.45\textwidth}
+A(a, a) :: 0.25. \\
+A(b, b) :: 0.3. \\
+A(X a, Y a) $\leftarrow$ A(X, Y) :: 0.25. \\
+A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.\\
+		\label{fig:grammar}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.5\textwidth}
+		\begin{lstlisting}
+values('A', [1, 2, 3, 4],[0.25, 0.30, 0.25, 0.20]).
+
+reduce('A'-[[a],[a]],1).
+reduce('A'-[[b],[b]],2).
+reduce('A'-[A2,B2],3) :- lpn('A'-[X,Y]),
+    append(X,[a],A2), append(Y,[a],B2).
+reduce('A'-[A2,B2],4) :- lpn('A'-[X,Y]),
+    append(X,[b],A2), append(Y,[b],B2).
+
+lpn(P-IN) :- msw(P,V), reduce(P-IN,V).
+		\end{lstlisting}
+		\label{fig:prism}
+	\end{minipage}
+	\caption{The copy language as an LPN (left), and a PRISM Program (right).}
+	\label{fig:copy}
+\end{figure}
+
+LPNs can be encoded as a restricted and tractable subclass of Stochastic Logic Programs (SLP).
+Specifically, we encode LPNs as programs in PRISM PROLOG.
+PRISM provides probabilistic execution and Bayesian inference over SLPs.
+
+To illustrate the encoding, consider the copy language (CL) in Figure \ref{fig:copy}.
+CL is a toy language producing two identical strings of $a$s and $b$s, making it mildly context-sensitive.
+Here, we see four rules concerning a single symbol, $A$, whose total probability sums to unity.
+
+To translate LPNs into PRISM SLPs, we first define the distribution over each symbol.
+We use the {\tt values} predicate, which takes an atom, in this case the name of the symbol, the values that atom can take on, and the probability of each value.
+For inference, we initialize the distribution uniformly.
+For our encoding, each possible value for an atom (or symbol) corresponds to exactly one clause in CL where that symbol appears on the left-hand side.
+There are four clauses for $A$ in CL, so {\tt values} can take on one of four values, each of whose probabilities corresponds to a rule in CL.
+
+For each rule in the LPN, we also define a corresponding clause of the {\tt reduce} predicate in our PRISM program.
+These {\tt reduce} rules are where the actual relations between the left-hand and right-hand side of each rule are encoded.
+Because PROLOG isn't perfectly declarative, the actual {\tt reduce} clauses are slightly more complex than presented here, reversing the order of the {\tt append} and recursive {\tt lpn} calls based on whether we are sampling or parsing.
+
+To parse or sample a sentence, we call {\tt lpn(P-IN)}, where {\tt P} is the top-level symbol, {\tt IN} is a list of terminal lists, and {\tt -} denotes a pair ({\it e.g.} {\tt lpn('A'-[[a,a,
+b,a],[a,a,b,a]])}).
+A sample or parse begins by probabilistically choosing some rule for the left-hand side via {\tt msw(P,V)} and then recursively reducing the input term using the chosen rule ({\tt reduce(P-IN,V)}).
 

### 2014-10-17 src/GIJoe/SRS/BuildNumber.hs

@@ -11,6 +11,19 @@ import GIJoe.SRS.Parse
 import GIJoe.SRS.Prismify
 import Data.Text (pack)
 
+writeGrammar :: FilePath -> [String] -> IO ()
+writeGrammar outpath g = writeFile outpath $ unlines g
+
+-- copy language
+copyGrammar :: [String]
+copyGrammar = 
+    [ "A(a,a)."
+    , "A(b,b)."
+    , "A(X a, Y a) <-- A(X,Y)."
+    , "A(X b, Y b) <-- A(X,Y)."]
+
+-- Number Grammar
+
 -- while we don't need every rule
 -- for all A,B,C
 -- A(X,Y) <- B(Y,X).          # #4# flip
@@ -47,58 +60,28 @@ import Data.Text (pack)
 -- Number(X) <-- A_1(X,Y).
 -- Next(X) <-- A_2(X,Y).
 
-grammarOfNumber2 :: Int -> FilePath -> IO ()
-grammarOfNumber2 n outpath = do
-  let lexicon = ones ++ teens ++ tens ++ ["after", "comes"]
-  let reorderRules = ["A" ++ show i ++ "(X Y, U V) <-- A" ++ show j ++ "(" ++ (ls !! 0) ++ "," ++ (ls !! 1) ++ "), A" ++ show k ++ "(" ++ (ls !! 2) ++ "," ++ (ls !! 3) ++ ")." | i <- [1..n], j <- [1..n], k <- [1..n], ls <- (permutations ["X","Y","U","V"]), j < k]
-  let reorderRulesEqual = ["A" ++ show i ++ "(X Y, U V) <-- A" ++ show j ++ "(" ++ (ls !! 0) ++ "," ++ (ls !! 1) ++ "), A" ++ show j ++ "(" ++ (ls !! 2) ++ "," ++ (ls !! 3) ++ ")." | i <- [1..n], j <- [1..n], ls <- ((map (\x -> "X":x) (permutations ["Y","U","V"])) ++ (map (\y -> "Y":y) (permutations ["X","U","V"])))]
-  let binaryRelationRules = ["A" ++ show i ++ "(" ++ (lexicon !! w1) ++ "," ++ (lexicon !! w2) ++ ")." | i <- [1..n], w1 <- [0..((length lexicon) - 1)] , w2 <- [0..((length lexicon) - 1)], w1 <= w2]
[…]

### 2014-10-17 src/GIJoe/SRS/Prismify.hs

@@ -50,30 +50,31 @@ prismClauses rs = unlines $
 
 body :: RewriteSystem -> [String]
 body slcfrs =
-    mainClause ++ [""] ++
-    srsClause ++ [""] ++
-    switchClauses ++ [""] ++
-    probClauses ++ [""] ++ reductionClauses
+    switchClauses     ++ [""] ++
+    probClauses       ++ [""] ++
+    reductionClauses  ++ [""] ++
+    prismToSRSClauses ++ [""] ++
+    [ ":- p_not_table reduce/2."
+    , ":- include(vis)."
+    , ":- include(main)."]
   where
     groupedRules = groupBy sameHead $ sortBy (compare `on` (cTermPredicate . ruleHead . weightedRuleRule)) (rsRules slcfrs)
     sameHead =  (\(WRule ((ComplexTerm p1 _) :<-: _) _)
                   (WRule ((ComplexTerm p2 _) :<-: _) _)
                  -> p1 == p2)
-    mainClause = [ "main(Gs,Foutp,Fouta) :- set_prism_flag(restart,1)," ++
-                   " set_prism_flag(learn_mode,vb), set_prism_flag(viterbi_mode,vb)," ++
-                   " set_prism_flag(default_sw_a,uniform), set_prism_flag(log_scale,on)," ++
-                   " learn(Gs), save_sw(Foutp), save_sw_a(Fouta).",
-                  "", "getTrain(F,Gs) :- load_clauses(F,ALL,[]), findall(X,member(train(X),ALL),Gs).",
-                  "", "getTest(F,Gs) :- load_clauses(F,ALL,[]), findall(X,member(test(X),ALL),Gs).", ""]
-    srsClause =  [ "srs(P-IN) :- msw(P,V), reduce(P-IN,V)." ]
---    srsClause =  [ "srs(P-IN) :- reduce(P-IN,V), msw(P,V)." ]
     switchClauses = map makeSwitch groupedRules
     probClauses = map makeProbs groupedRules
+    prismToSRSClauses = concatMap makePrismToSRS groupedRules
     reductionClauses = concatMap makeReductions groupedRules
 
 makeSwitch :: [WeightedRule] -> String
 makeSwitch rs =
     "values(" ++ weightedRuleName (head rs) ++ "," ++ show [1..(length rs)] ++ ")."
+
+makePrismToSRS :: [WeightedRule] -> [String]
[…]

### 2014-10-19 experiments/prism/run_parallel.sh
More files from Ec2

@@ -1,7 +1,7 @@
 #!/bin/bash
 
-NumExamples=(700 900 1100 1300 1500 1700 1900 2100);
-NumReps=10;
+NumExamples=(1700);
+NumReps=1;
 NumPred=(4);
 SaveDir="data/saturday_"
 args=()
@@ -19,7 +19,7 @@ do
     done
 done
 
-CMD="parallel -n 3 --noswap --resume-failed --joblog ./parallel_log --delay 2 upprism srs {} $SaveDir &> logs/overnight/log_{1}_{2}_{3}.log ::: ${args[*]}"
+CMD="parallel -n 3 --noswap --resume-failed --joblog ./parallel_log --delay 2 upprism srs {} $SaveDir &> logs/overnight/saturday_log_{1}_{2}_{3}.log ::: ${args[*]}"
 echo $CMD
 `$CMD`
 

### 2014-10-19 prolog/main.pl
add vis.pl and main.pl

@@ -1,0 +1,22 @@
+main(Gs,Foutp,Fouta) :-
+    set_prism_flag(restart,1),
+    set_prism_flag(learn_mode,vb),
+    set_prism_flag(viterbi_mode,vb),
+    set_prism_flag(default_sw_a,uniform),
+    set_prism_flag(log_scale,on),
+    learn(Gs),
+    save_sw(Foutp),
+    save_sw_a(Fouta).
+
+getTrain(F,Gs) :-
+    load_clauses(F,ALL,[]),
+    findall(X,member(train(X),ALL),Gs).
+
+getTest(F,Gs) :-
+    load_clauses(F,ALL,[]),
+    findall(X,member(test(X),ALL),Gs).
+
+srs(P-IN) :-
+    msw(P,V),
+    reduce(P-IN,V).
+

### 2014-10-19 prolog/vis.pl

@@ -1,0 +1,130 @@
+%% %% Print a Parse Tree %% %%
+
+srsTree(X,[Olhs,Orule,Orhs]) :-
+    select(msw(P,V),X,X2),
+    select(srs(P-IN),X2,X3),
+    delete(X3,reduce(_,_),X4),
+    maplist(srsTree,X4,Orhs),
+    prismToSRS(P,V,Orule),
+    predicateToString(P-IN,Olhs).
+
+%% hack! This currently only works up to 9-ary predicates
+predicateToString(P-IN,S) :-
+    sub_atom(P,0,_,2,P2),
+    atom_lists_to_atoms(IN,A2),
+    atom_concat(P2,'(',S1),
+    atom_concat(A2,')',S2),
+    atom_concat(S1,S2,S).
+
+atom_lists_to_atoms([],'').
+atom_lists_to_atoms([X],O) :-
+    recursive_atom_concat_space(X,O).
+
+atom_lists_to_atoms([X|XS],O) :-
+    length(XS,XL), XL > 0,
+    atom_lists_to_atoms(XS,OS),
+    recursive_atom_concat_space(X,OS2),
+    atom_concat(OS2,', ',O1),
+    atom_concat(O1,OS,O).
+
+recursive_atom_concat([],'').
+recursive_atom_concat([X],X).
+recursive_atom_concat([X|XS],O) :-
+    length(XS,XL), XL > 0,
+    recursive_atom_concat(XS,OS),
+    atom_concat(X,OS,O).
+
+recursive_atom_concat_space([],'').
+recursive_atom_concat_space([X],X).
+recursive_atom_concat_space([X|XS],O) :-
+    length(XS,XL),
+    XL > 0,
+    recursive_atom_concat_space(XS,OS),
+    atom_concat(X,' ',O1),
+    atom_concat(O1,OS,O).
+
+formatSRSTree(X) :- format("~n",[]), formatSRSTree(X,'').
+formatSRSTree([LHS,RULE,RHS],SEP) :-
+    format("~a~a~n~a~a~n",[SEP,LHS,SEP,RULE]),
+    atom_concat(SEP,'    ',NEWSEP),
+    findall(X,(member(X,RHS),formatSRSTree(X,NEWSEP)),_).
+
+%% %% Print The Grammar %% %%
+
+%% srs Grammar: give grammar as [predicate,value,count] triplets
[…]

### 2014-10-19 drafts/cfp/draft.tex
Recreated beforeAndAfter system; changed examples a little; moved vis.pl and main.pl to experiments/prolog/.

@@
-\documentclass{article}
+\documentclass[11pt, twocolumn]{article}
 
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{listings}
@@
 \section{Experiments}
+
+\begin{tabular}{l l}
+Question & $N_p=4$ \\ \hline
+ after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
+ after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
+ after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
+ after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
+ after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
+ after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
+ after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
+ after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
+ after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
+ after sixty nine comes \underline{\hspace{1cm}}? & sixty ten \checkmark \\
+ after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
+ after seventy nine comes \underline{\hspace{1cm}}? & seventy ten $\times$ \\
+ after ninety five comes \underline{\hspace{1cm}}? & ninety six \checkmark \\
+ before twenty three comes \underline{\hspace{1cm}}? & twenty two \checkmark \\
+ before thirty comes \underline{\hspace{1cm}}? & thirty eighty $\times$ \\
+ before thirty eight comes \underline{\hspace{1cm}}? & thirty seven \checkmark \\
+ before forty one comes \underline{\hspace{1cm}}? & forty \checkmark \\
+ before fifty three comes \underline{\hspace{1cm}}? & fifty two \checkmark \\
+ before sixty eight comes \underline{\hspace{1cm}}? & sixty seven \checkmark \\
+ before seventy two comes \underline{\hspace{1cm}}? & seventy one \checkmark \\
+ before seventy three comes \underline{\hspace{1cm}}? & seventy two \checkmark \\
+ before eighty five comes \underline{\hspace{1cm}}? & eighty four \checkmark \\
+ before ninety two comes \underline{\hspace{1cm}}? & ninety one \checkmark \\
+ before ninety three comes \underline{\hspace{1cm}}? & ninety two \checkmark \\
+ before ninety five comes \underline{\hspace{1cm}}? &ninety four \checkmark \\
+
+\end{tabular}
+
+\begin{alignat*}{2}
+&S(X Y) \leftarrow A_1(X, Y) : 1.0000 \\
+&A_1(X Y, U V) \leftarrow A_2(X, U), A_3(V, Y) : 0.5002 \\
+&A_2(before, comes) : 0.7316 \\
+&A_4(after, comes) : 0.9990 \\
+&A_1(X Y, U V) \leftarrow A_3(Y, V), A_4(X, U) : 0.3428 \\
+&A_3(one, two) : 0.3993 \\
+&A_3(two, three) : 0.2063 \\
+&A_3(three, four) : 0.1093 \\
+&A_1(X Y, U V) \leftarrow A_1(V, Y), A_4(X, U) : 0.0796 \\
+&A_1(X Y, U V) \leftarrow A_1(Y, V), A_2(X, U) : 0.0712 \\
+&A_3(four, five) : 0.0734 \\
+&A_3(five, six) : 0.0502 \\
+&A_3(six, seven) : 0.0355 \\
+&A_3(eight, nine) : 0.0290 \\
+&A_3(seven, eight) : 0.0271 \\
+&A_2(fifty, fifty) : 0.0375 \\
+&A_2(thirty, thirty) : 0.0361 \\
+&A_2(eighty, eighty) : 0.0339 \\
+&A_3(null, one) : 0.0231 \\
+&A_2(forty, forty) : 0.0332 \\
+&A_2(twenty, twenty) : 0.0310 \\
+&A_2(seventy, seventy) : 0.0296 \\
+&A_2(sixty, sixty) : 0.0274 \\
+&A_2(ninety, ninety) : 0.0260 \\
+&A_3(eighteen, nineteen) : 0.0064 \\
+&A_3(sixteen, seventeen) : 0.0049 \\
+&A_3(eleven, twelve) : 0.0044 \\
+&A_3(nine, ten) : 0.0044 \\
+&A_3(thirteen, fourteen) : 0.0044 \\
+&A_3(fourteen, fifteen) : 0.0039 \\
+&A_3(ten, eleven) : 0.0034 \\
+&A_2(null, fifty) : 0.0043 \\
+&A_3(eighty, null) : 0.0030 \\
+&A_3(seventeen, eighteen) : 0.0030 \\
+&A_1(X Y, U V) \leftarrow A_2(Y, X), A_3(V, U) : 0.0021 \\
+&A_3(nine, sixty) : 0.0025 \\
+&A_2(nine, seventy) : 0.0029 \\
+&A_3(nine, forty) : 0.0020 \\
+&A_1(X Y, U V) \leftarrow A_1(V, Y), A_2(X, U) : 0.0013 \\
+&A_2(null, thirty) : 0.0022 \\
+&A_2(nine, ninety) : 0.0022 \\
+&A_3(twelve, thirteen) : 0.0015 \\
+&A_3(fifteen, sixteen) : 0.0015 \\
+&A_1(X Y, U V) \leftarrow A_2(V, X), A_3(Y, U) : 0.0008 \\
+&A_3(null, comes) : 0.0010 \\
+&A_1(X Y, U V) \leftarrow A_1(X, U), A_2(V, Y) : 0.0008 \\
+&A_1(X Y, U V) \leftarrow A_1(X, U), A_2(Y, V) : 0.0008 \\
+&A_2(twenty, after) : 0.0014 \\
+&A_2(sixty, before) : 0.0007 \\
+&A_3(nineteen, comes) : 0.0005 \\
+&A_4(nine, thirty) : 0.0010 \\
+&A_1(X Y, U V) \leftarrow A_1(X, U), A_4(Y, V) : 0.0004 \\
+\end{alignat*} \\
 \section{Discussion}

### 2014-10-19 experiments/prism/allExamples.pl

@@ -297,33 +297,58 @@ exampleCount( 96 , 13 ).
 exampleCount( 97 , 15 ).
 exampleCount( 98 , 10 ).
 exampleCount( 99 , 8 ).
+
 %% test set
 
-%% after test examples
-test( 45 ).
-test( 65 ).
-test( 61 ).
-test( 64 ).
-test( 63 ).
-test( 49 ).
-test( 20 ).
-test( 95 ).
-test( 79 ).
-test( 69 ).
-test( 47 ).
-test( 59 ).
-test( 73 ).
+% case 1
+test(20). % 20-21
+test(50). % 50-51
 
-%% before test examples
-test(150).
-test(190).
-test(192).
-test(169).
-test(120).
-test(165).
-test(138).
-test(170).
-test(135).
-test(189).
-test(182).
-test(127).
+% case 2
+test(128). % 31-30
+test(158). % 61-60
+
+% case 3
+test(75). % 75-76 
+test(54). % 54-55
+
+% case 4
+test(159). % 62-62 
+test(135). % 38-37
+
+% case 5
+test(29). % 29-30 
+test(89). % 89-90
+
+% case 6
+test(40). % 40-39
+test(80). % 80-79
+
+%% %% after test examples
+%% test( 45 ).
+%% test( 65 ).
+%% test( 61 ).
+%% test( 64 ).
+%% test( 63 ).
+%% test( 49 ).
+%% test( 20 ).
+%% test( 95 ).
+%% test( 79 ).
+%% 1test( 69 ).
+%% test( 47 ).
+%% test( 59 ).
+%% test( 73 ).
+
+%% %% before test examples
+%% test(150).
+%% test(190).
+%% test(192).
+%% test(169).
+%% test(120).
+%% test(165).
+%% test(138).
+%% test(170).
+%% test(135).
+%% test(189).
+%% test(182).
+%% test(127).

### 2014-10-19 experiments/prism/run_parallel.sh

@@ -19,7 +19,7 @@ do
     done
 done
 
-CMD="parallel -n 3 --noswap --resume-failed --joblog ./parallel_log --delay 2 upprism srs {} $SaveDir &> logs/overnight/saturday_log_{1}_{2}_{3}.log ::: ${args[*]}"
+CMD="parallel -n 3 --noswap --resume-failed --joblog ./parallel_log --delay 10 upprism srs {} $SaveDir &> logs/overnight/saturday_log_{1}_{2}_{3}.log ::: ${args[*]}"
 echo $CMD
 `$CMD`
 

### 2014-10-19 prolog/vis.pl

@@ -51,9 +51,19 @@ formatSRSTree([LHS,RULE,RHS],SEP) :-
 
 %% %% Print The Grammar %% %%
 
+%% findall([AS,B,C,DS],(get_sw_a(A,_,B,C), D is sum(C), length(B,BL), listOfN(A,BL,AS), listOfN(D,BL,DS), XS2).
+
 %% srs Grammar: give grammar as [predicate,value,count] triplets
 srsQuadGrammar(OS) :-
-    findall([AS,B,C,DS],(get_sw_a(A,_,B,C,_), D is sum(C), length(B,BL), listOfN(A,BL,AS), listOfN(D,BL,DS)),XS2),
+    write('hello'),
+    findall([AS,B,C,DS],(get_sw_a(A,_,B,C), 
+                         D is sum(C), 
+                         length(B,BL), 
+                         listOfN(A,BL,AS), 
+                         listOfN(D,BL,DS)),
+            XS2),
+    write('hello'), 
+    write(XS2), nl, 
     maplist(quadZip,XS2,XS3),
     concat(XS3,OS).
 

### 2014-10-19 src/GIJoe/SRS/BuildNumber.hs

@@ -148,6 +148,8 @@ nextSentences = [[["after"] ++ a ++ ["comes"] ++  b] | [a,b] <- nextExamples]
 prevDecadeSentences = [[["before"] ++ b ++ ["comes"] ++ a] | [a,b] <- nextDecades]
 nextDecadeSentences = [[["after"] ++ a ++ ["comes"] ++  b] | [a,b] <- nextDecades]
 beforeDecadeSentences = [[["before"] ++ b ++ ["comes"] ++  a] | [a,b] <- nextDecades]
+invertedNextSentences = [[b ++ ["comes"] ++  ["after"] ++ a] | [a,b] <- nextExamples]
+invertedBeforeSentences = [[a ++ ["comes"] ++  ["before"] ++ b] | [a,b] <- nextExamples]
 
 -- predicate networks
 (first19NextSentences, restNextSentences) = splitAt 19 nextSentences

### 2014-10-19 src/GIJoe/SRS/Prismify.hs

@@ -36,7 +36,10 @@ prismClauses :: RewriteSystem -> String
 prismClauses rs = unlines $ 
   switchClauses ++ [""] ++
   probClauses ++ [""] ++
-  reductionClauses
+  reductionClauses ++ [""] ++ 
+  prismToSRSClauses ++ [""] ++
+    [":- include(vis)."
+    , ":- include(main)."]
   where switchClauses = map makeSwitch groupedRules
         probClauses = map makeProbs groupedRules
         reductionClauses = concatMap makeReductions groupedRules
@@ -47,6 +50,7 @@ prismClauses rs = unlines $
         sameHead =  (\(WRule ((ComplexTerm p1 _) :<-: _) _)
                   (WRule ((ComplexTerm p2 _) :<-: _) _)
                  -> p1 == p2)
+        prismToSRSClauses = concatMap makePrismToSRS groupedRules                    
 
 body :: RewriteSystem -> [String]
 body slcfrs =

### 2014-10-19 drafts/cfp/draft.tex
Draft changes.

@@
 \usepackage{subcaption}
+
 \title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
@@
 \section{Abstract}
-A popular approach in statistical machine learning is to automatically
-find the latent \emph{features} of the data that are most relevant or
-informative. But a latent feature based view of concept learning does
-not match up naturally with the structured symbolic knowledge that
-seems to underlie people's understanding of many domains. How can we
-combine the flexibility, noise tolerance, and computational
-practicality of latent feature based statistical machine learning with
-the symbolic knowledge representations that best capture human
-understanding? Here present a learning model we dub Latent Concept
-Grammars (LCGs) as a proposal. An LCG is a densely connected set of
-probabilistic string rewrite rules over an initially uninterpreted set
-of predicates. When fit to data using a hierarchical Bayesian prior,
-these predicates assume the roles of informative latent concepts. This
-approach allows us to learn knowledge that is stochastic and
-recursively structured while maintaining computational
-tractability. We show how learning in LCGs can be implemented using
-the probabilistic logic programming system PRISM, and we present
-preliminary experiments.
+To what extent can the meaning of language be learned from linguistic
+data alone?  The sentences that people say contain a great deal of
+information about the meaning of linguistic concepts. Most approaches
+to semantics assume an explicit knowledge representation formalism
+that ties utterances to propositions about things in the world. The
+study of linguistic syntax, however, has largely assumed that the
+meaning of syntactic structures is encoded directly into the grammar
+of a language. For machine learning, extending the grammatical
+embedding of meaning to at least some of the linguistic concepts is
+attractive because it reduces the problem of automated language
+learning to the problem of learning a distribution over naturally
+occuring utterances. To what extent is this a feasible approach? Here,
+we present a first attempt at the automated induction of probabilistic
+context sensitive grammars which we call Latent Predicate Networks
+(LPNs), motivated by the idea that the predicates and rules that are
+learned correspond to the concepts expressed by utterances. We show
+that for a small fragment of sentences expressing relationships
+between English number words, we can use hierarchical Bayesian
+inference to learn LPNs that can answer simple queries about
+previously unseen linguistic concepts within this domain. These
+generalizations, we argue, demonstrate LPNs' promise as a tool for
+learning and representing conceptual knowledge in language.
 
-How do people learn systems of related concepts from
-data?  Consider, for example, a school-age child's knowledge of
-number. This knowledge supports a large number of related operations:
-the child knows which words denote numbers, how to use those words to
-label cardinalities of sets, how to translate between digit
-representations and word representations of number, how to compare
-cardinalities, etc. The child knows how to perform these operations
-over an unbounded set of elements, even though the data that underlies
-this knowledge is small and noisy. Here we propose that such systems
-of related concepts can be learned via hierarchical probabilistic
-inference over probabilistic string rewrite systems (PSRS) -- a mildly
-context-sensitive class of probabilistic string gramamrs that
-corresponds to a restricted fragment of stochastic logic programs
-(SLPs). Specifically, we extend the variational
-expectation-maximization algorithm for Hierarchical Dirichlet PCFGs to
-accomodate PSRSs. This algorithm we present enables a tractible
-compromise between highly restrictive context-free knowledge
-representations and the intractability of performing similar inference
-procedures over general logic programs. Empirically, we show how such
-restricted string rewrite systems can capture non-trivial conceptual
-knowledge and how our algorithm is able to learn such knowledge from
-data.
+% A popular approach in statistical machine learning is to automatically
+% find the latent \emph{features} of the data that are most relevant or
+% informative. But a latent feature based view of concept learning does
+% not match up naturally with the structured symbolic knowledge that
+% seems to underlie people's understanding of many domains. How can we
+% combine the flexibility, noise tolerance, and computational
+% practicality of latent feature based statistical machine learning with
+% the symbolic knowledge representations that best capture human
+% understanding? Here present a learning model we dub Latent Concept
+% Grammars (LCGs) as a proposal. An LCG is a densely connected set of
+% probabilistic string rewrite rules over an initially uninterpreted set
+% of predicates. When fit to data using a hierarchical Bayesian prior,
+% these predicates assume the roles of informative latent concepts. This
+% approach allows us to learn knowledge that is stochastic and
+% recursively structured while maintaining computational
+% tractability. We show how learning in LCGs can be implemented using
+% the probabilistic logic programming system PRISM, and we present
+% preliminary experiments.
 
-\section{Abstract 2}
-What does it mean to understand a concept, and how do people come to
-such an understanding.
+% How do people learn systems of related concepts from
+% data?  Consider, for example, a school-age child's knowledge of
+% number. This knowledge supports a large number of related operations:
+% the child knows which words denote numbers, how to use those words to
+% label cardinalities of sets, how to translate between digit
+% representations and word representations of number, how to compare
+% cardinalities, etc. The child knows how to perform these operations
+% over an unbounded set of elements, even though the data that underlies
+% this knowledge is small and noisy. Here we propose that such systems
+% of related concepts can be learned via hierarchical probabilistic
+% inference over probabilistic string rewrite systems (PSRS) -- a mildly
+% context-sensitive class of probabilistic string gramamrs that
+% corresponds to a restricted fragment of stochastic logic programs
+% (SLPs). Specifically, we extend the variational
+% expectation-maximization algorithm for Hierarchical Dirichlet PCFGs to
+% accomodate PSRSs. This algorithm we present enables a tractible
+% compromise between highly restrictive context-free knowledge
+% representations and the intractability of performing similar inference
+% procedures over general logic programs. Empirically, we show how such
+% restricted string rewrite systems can capture non-trivial conceptual
+% knowledge and how our algorithm is able to learn such knowledge from
+% data.
+
+% \section{Abstract 2}
+% What does it mean to understand a concept, and how do people come to
+% such an understanding.
 

### 2014-10-19 drafts/cfp/draft.tex
update draft with new figures

--- 2014-10-19 drafts/cfp/draft.tex
@@
 \usepackage{subcaption}
+%\usepackage{subfigure}
+\usepackage{graphicx}
 \title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
@@
 \subsection{Related Work}
-Our core inference algorithm borrows heavily from recent algorithms for nonparametric PCFG induction, as well as work extending that algorithm to SLPs (CITATION Johnson; Liang; Cussens, 2001).
+Our core inference algorithm borrows heavily from recent algorithms for nonparametric PCFG induction, as well as work extending those algorithms to SLPs~\cite{DBLP:conf/emnlp/LiangPJK07,goldwater2006contextual,johnson2006adaptor,cussens2001parameter}.
 Like Johnson and Liang, we are interested in inferring the most likely grammar given some collection of sentences, and in doing so without a pre-specified limit on the complexity of the grammar.
@@
 Both are aimed at jointly learning syntactic and semantic information from a body of sentences.
-Semantic parsers, however, learn separate systems of syntax and semantics, often pairing Combinatory Categorial Grammars of syntax with an explicit logical language of semantics, often the untyped lambda calculus (CITATION Steedman; Liang; Zettlemoyer; etc.).
+Semantic parsers, however, learn separate systems of syntax and semantics, often pairing Combinatory Categorial Grammars of syntax with an explicit logical language of semantics, often the untyped lambda calculus~\cite{berant2013semantic,liang2013learning,kwiatkowski2010inducing}.
 By contrast, our approach uses a single formalism, the latent predicate network, to describe both which sentences are syntactically valid and what sorts of semantic information those sentence imply.
@@
 
-More generally, we see this model as helping to strike a balance between expressivity and tractability, bridge the gap between more traditional forms of grammar induction (CITATION) and program induction over more powerful languages (CITATION E.C.
-paper, other P.I.
+More generally, we see this model as helping to strike a balance between expressivity and tractability, bridge the gap between more traditional PCFG grammar induction and program induction over more powerful languages~\cite{dechter2013bootstrap,} (CITATION other P.I.
 papers).
@@
 \section{The Model}
+
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
+		\caption{A schematic LPN where $m = |L|,\, n = |N|-1$,\, arity$(A_i) = 2$, and arity$(S) = 1$.}
+		\label{fig:schematic}
+\end{figure}
+
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
+		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+\end{figure}
+
 An LCG is a hierarchical Bayesian model of string tuples. It is an
@@
 the exact specification of the parameters $K$ and $H$ and the clauses
-$mathcal{C}$. To each clause $C_i$ is associated a positive real
+$\mathcal{C}$. To each clause $C_i$ is associated a positive real
 weights $\theta_i$.
@@
     language=Prolog,
-    basicstyle=\fontsize{5}{6}\selectfont\ttfamily
+    basicstyle=\fontsize{9}{10.5}\selectfont\ttfamily
 }
@@
 	\centering
-	\begin{minipage}[b]{0.45\textwidth}
+	\begin{minipage}[b]{\linewidth}
 A(a, a) :: 0.25. \\
@@
 A(X a, Y a) $\leftarrow$ A(X, Y) :: 0.25. \\
-A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.\\
+A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.
+		\subcaption{}
 		\label{fig:grammar}
 	\end{minipage}
-	\hfill
-	\begin{minipage}[b]{0.5\textwidth}
+	%\rule[0pt]{0.8\linewidth}{.5pt}
+	\begin{minipage}[b]{\linewidth}
 		\begin{lstlisting}
-values('A', [1, 2, 3, 4],[0.25, 0.30, 0.25, 0.20]).
+values('A', [1, 2, 3, 4],
+    [0.25, 0.30, 0.25, 0.20]).
 
@@
 		\end{lstlisting}
+		\subcaption{}
 		\label{fig:prism}
 	\end{minipage}
-	\caption{The copy language as an LPN (left), and a PRISM Program (right).}
+	\caption{Possible encodings of the 2-copy language as (\subref{fig:grammar}) an LPN, (\subref{fig:prism}) a PRISM program.}
 	\label{fig:copy}
@@
 
-To illustrate the encoding, consider the copy language (CL) in Figure \ref{fig:copy}.
-CL is a toy language producing two identical strings of $a$s and $b$s, making it mildly context-sensitive.
+To illustrate the encoding, consider the 2-copy language (CL2) in Figure \ref{fig:copy}.
+CL2 is a toy language producing two identical strings of $a$s and $b$s, making it mildly context-sensitive.
 Here, we see four rules concerning a single symbol, $A$, whose total probability sums to unity.
@@
 For inference, we initialize the distribution uniformly.
-For our encoding, each possible value for an atom (or symbol) corresponds to exactly one clause in CL where that symbol appears on the left-hand side.
-There are four clauses for $A$ in CL, so {\tt values} can take on one of four values, each of whose probabilities corresponds to a rule in CL.
+For our encoding, each possible value for an atom (or symbol) corresponds to exactly one clause in CL2 where that symbol appears on the left-hand side.
+There are four clauses for $A$ in CL2, so {\tt values} can take on one of four values, each of whose probabilities corresponds to a rule in CL2.
 
@@
 A sample or parse begins by probabilistically choosing some rule for the left-hand side via {\tt msw(P,V)} and then recursively reducing the input term using the chosen rule ({\tt reduce(P-IN,V)}).
+\section{Experiments}
 
-\section{Experiments}
+\begin{figure}
+	\rule{\linewidth}{3in}
+	\caption{The Results}
+	\label{fig:results}
+\end{figure}
 
@@
 
-\section{Outline}
-1) Introduction
-  - psychological motivation 
-    - juxtaposing the kinds of knowledge that freebase has and the kinds that children learn
-    - compositional semantic structure --> derive semantics from syntax
-  - what is the contribution?
-     - the use of string rewrite formalisms to capture conceptual knowledge and conceptual learning
-     - to bring tools from computational linguistics to learning semantics
-  - what kind of concept systems are we trying to capture?
-  - why restrict the formalism? 
-     - tractability of inference 
-  - why go beyond context free grammars? 
-     - if they are intractable to learn anyway, why try to go beyond them? 
-  - Related Work
-2) Model
-- stochastic linear context free rewrite systems
-  - model: what is the data? and what is being inferred? what is the prior? 
-  - this can be written as PRISM program, or an SLP. 
-  - inference: variational approximation 
-3) Experiments
-  a) worked example
-  b) Number system (main result)
-  3b) Results
-4) Conclusion
-5) Citations
+%1) Introduction
+%  - psychological motivation 
+%    - juxtaposing the kinds of knowledge that freebase has and the kinds that children learn
+%    - compositional semantic structure --> derive semantics from syntax
+%  - what is the contribution?
+%     - the use of string rewrite formalisms to capture conceptual knowledge and conceptual learning
+%     - to bring tools from computational linguistics to learning semantics
+%  - what kind of concept systems are we trying to capture?
+%  - why restrict the formalism? 
+%     - tractability of inference 
+%  - why go beyond context free grammars? 
+%     - if they are intractable to learn anyway, why try to go beyond them? 
+%  - Related Work
+%2) Model
+%- stochastic linear context free rewrite systems
+%  - model: what is the data? and what is being inferred? what is the prior? 
+%  - this can be written as PRISM program, or an SLP. 
+%  - inference: variational approximation 
+%3) Experiments
+%  a) worked example
+%  b) Number system (main result)
+%  3b) Results
+%4) Conclusion
+%5) Citations
 

### 2014-10-19 drafts/cfp/draft.tex

--- 2014-10-19 drafts/cfp/draft.tex
@@
 \usepackage{subcaption}
-
+\usepackage{graphicx}
 \title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}

### 2014-10-20 drafts/cfp/draft.tex
Changes to draft.

@@
 \title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
-\author{Eyal Dechter \& Josh Rule \& Josh Tenenbaum}
+\author{Eyal Dechter,  Josh Rule, Josh Tenenbaum}
 
@@
 \section{Abstract}
-To what extent can the meaning of language be learned from linguistic
-data alone?  The sentences that people say contain a great deal of
-information about the meaning of linguistic concepts. Most approaches
-to semantics assume an explicit knowledge representation formalism
-that ties utterances to propositions about things in the world. The
-study of linguistic syntax, however, has largely assumed that the
-meaning of syntactic structures is encoded directly into the grammar
-of a language. For machine learning, extending the grammatical
-embedding of meaning to at least some of the linguistic concepts is
-attractive because it reduces the problem of automated language
-learning to the problem of learning a distribution over naturally
-occuring utterances. To what extent is this a feasible approach? Here,
-we present a first attempt at the automated induction of probabilistic
-context sensitive grammars which we call Latent Predicate Networks
-(LPNs), motivated by the idea that the predicates and rules that are
-learned correspond to the concepts expressed by utterances. We show
-that for a small fragment of sentences expressing relationships
-between English number words, we can use hierarchical Bayesian
-inference to learn LPNs that can answer simple queries about
+How can machines learn about a system of concepts from sentences about
+those concepts? One approach is to estimate the distribution from
+which those sentences are drawn, and, in doing so, infer the latent
+concepts and relationships that best explain the data. We present
+early work on a learning framework called Latent Predicate Networks
+(LPNs) which attempts to learn concepts by inferring the parameters of
+probabilistic context-sensitive grammars that best describe sentences.
+We show that for a small fragment of sentences expressing
+relationships between English number words, we can use hierarchical
+Bayesian inference to learn LPNs that can answer simple queries about
 previously unseen linguistic concepts within this domain. These
@@
 % preliminary experiments.
-
+}
 % How do people learn systems of related concepts from
@@
 \section{Introduction}
-The vast majority of the models of concept acquisition that have been
-proposed in the artificial intelligence, machine learning and
-cognitive science literatures cannot represent the conceptual systems
-that even young children learn. Take, for example, the school-age
-child's knowledge of number. This child knows that there is no largest
-number; that ``one million forty three'' and ``8439''denote numbers
-but that ``one million forty billion'' and ``a82bd'' do not; that no
-two cardinalities refer to the same quantity but that some number
-words refer to the same cardinalities (e.g. ``two thousand'' and
-``twenty-one hundred''). Moreover, the child knows how to guage the
-approximate number of elements in a collection and how to count this
-number exactly. In short, the child's knowledge of number consists of
-a complex system of concepts over an unbounded set of entities
-(i.e. the numbers).
 
-Children are not born with this kind of knowledge, and, since number
-is a relatively recent cultural invention, nor is it likely that
-children possess innate domain specific background knowledge that
-makes it easier to acquire. Children only ever see a tiny fraction of
-all the numbers they can reason about and this evidence is biased,
-noisy, and can be unreliable.
+As Claude Shannon demonstrated with his guessing game experiment in
+1950~\cite{Shannon1950}, an adult English speaker's understanding of
+her language and her world allows her to reveal that the entropy of a
+character in English text is only a little more than one
+bit. Plausibly, any machine learning system that were able to predict
+natural language text with such accuracy will have learned much about
+the concepts expressed in that text. This intuition motivates us to
+invert Shannon's experiment and attempt to learn about linguistic
+concepts by learning a probabilistic generative model for natural
+language.
 
-A fundamental challenge for cognitive science -- and for any field
-interested in understanding human-like concept acquisition -- is to
-provide a theory of how such complex systems of concepts can be
-learned from sparse and noisy data.
+Probabilistic Context Free Grammars (PCFGs) are popular and much
+studied generative models for natural language
+syntax~\cite{Someone}. And though there ongoing debate about their
+ability to express the syntax of natural language, there is no doubt
+that they cannot express dependencies between words that depend upon
+their meaning. Consider an example from the domain of sentences
+addressed in this paper: ``after twenty eight comes twenty nine.''
+The second ``twenty'' depends on the first, and the ``nine'' depends
+on the ``eight,'' and this dependency pattern is a clear example of a
+\emph{cross-serial dependency}, a dependency pattern that context-free
+grammars do not support.
 
-People are able to learn about complex systems of related
-concepts from small amounts of noisy data. A compelling approach to
-understanding such learning is to frame it as Bayesian inference over
-probabilistic knowledge bases. Advances in context free grammar
-induction in the past couple of decades can be seen as an instance of
-this approach, in which
+Our approach, then, will be to explore the use of probabilistic
+context \emph{sensitive} grammars for modeling natural language,
+extending the tools of hierarchical grammar induction~\ref{Cite} from
+context-free grammars to ones that can express not only syntactic
+relationships but semantic ones as well. In the place of latent
+syntactic categories, we will refer to latent predicates -- relations
+between sentence components -- that, when learned, should correspond
+to the concepts that underlie sentence semantics.
+
+To explore this approach, we introduce a probabilisitic model called a
+Latent Predicate Network (LPN), and describe an inference procedure to
+learn its parameters using the probabilistic logic programming
+language PRISM. Finally, we study its performance when trained on
+sentences describing the relationships between numbers.
+
+\section{Latent Predicate Networks}
+
+An LPN is a hierarchical Bayesian model of strings. It is an
+extension of the Hierarhical Dirichlet PCFG model (HD-PCFG) to
+Probabilistic Range Concatenation Grammars (PRCGs).
+
+Range Concatenation Grammars (RCGs) are a class of string grammars
+that represent all and only those languages that can be parsed in time
+polynomial in the length of the target
+string~\cite{boullier2005range}. An RCG $G=(N, T, V, P, S)$ is a
+5-tuple where $N$ is a finite set of predicate symbols, $T$ is a set
+of terminal symbols, $V$ is a set of variable symbols, P is a finite
+set of $M \geq 0$ clauses of the form $\psi_0 \rightarrow \psi_1 \dots
+\psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
+the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
+N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
+\cup V)^*$ is an argument of $\psi_m$. We call the left hand side term
+of any clause the \emph{head} of that clause and its predicate symbol
+is the \emph{head predicate}.
+
+\textbf{Example:} The following is an RCG for the 3-copy language
+$\{www | w \in {a, b}^*\}$:
+\begin{alignat*}{3}
+&S(XYZ) &&\rightarrow A(X, Y, Z)\\
+&A(aX, aY, aZ) &&\rightarrow A(X, Y, Z)\\
+&A(bX, bY, bZ) &&\rightarrow A(X, Y, Z)\\
+&A(\epsilon, \epsilon, \epsilon) &&\rightarrow 
+\end{alignat*}
+
+A string $x$ is in the language defined by an RCG if one can
+\emph{derive} $S(x)$. A derivation is a sequence of rewrite steps in
+which substrings of the left hand side argument string are bound to
+the variables of the head of some clause, thus determining the
+arguments in the clause body. If a clause has no body terms, then its
+head is derived; otherwise, its head is derived if its body clauses
+are derived~\footnote{This description of the language of an RCG
+  technically only holds for \emph{non-combinatory} RCGs, in which the
+  arguments of body terms can only contain single variables. Since any
+  \emph{combinatory} RCG can be converted into a non-combinatory RCG
+  and we only consider non-combinatory RCGs here, this description
+  suffices.}
+
+We extend RCGs to Probabilistic Range Concatenation Grammars (PRCGs)
+by annotating each clause $C_k \in P$ with probabilities $p_k$ such
+that for all predicates ${A \in N$ $\sum_{k:head(C_k)=A} p_k = 1}$. A
+PRCGs defines a distribution over strings $x$ by sampling from
+derivations of $S(x)$ according to the product of probabilities of
+clauses used in that derivation. This is a well defined distribution
+as long as no probability mass is placed on derivations of infinite
+length; in this paper, we only consider PRCGs with derivations of
+finite length, so we need not worry about this requirement.
+
+% The vast majority of the models of concept acquisition that have been
+% proposed in the artificial intelligence, machine learning and
+% cognitive science literatures cannot represent the conceptual systems
+% that even young children learn. Take, for example, the school-age
+% child's knowledge of number. This child knows that there is no largest
+% number; that ``one million forty three'' and ``8439''denote numbers
+% but that ``one million forty billion'' and ``a82bd'' do not; that no
+% two cardinalities refer to the same quantity but that some number
+% words refer to the same cardinalities (e.g. ``two thousand'' and
+% ``twenty-one hundred''). Moreover, the child knows how to guage the
+% approximate number of elements in a collection and how to count this
+% number exactly. In short, the child's knowledge of number consists of
+% a complex system of concepts over an unbounded set of entities
+% (i.e. the numbers).
+
+% Children are not born with this kind of knowledge, and, since number
+% is a relatively recent cultural invention, nor is it likely that
+% children possess innate domain specific background knowledge that
+% makes it easier to acquire. Children only ever see a tiny fraction of
+% all the numbers they can reason about and this evidence is biased,
+% noisy, and can be unreliable.
+
+% A fundamental challenge for cognitive science -- and for any field
+% interested in understanding human-like concept acquisition -- is to
+% provide a theory of how such complex systems of concepts can be
+% learned from sparse and noisy data.
+
+% People are able to learn about complex systems of related
+% concepts from small amounts of noisy data. A compelling approach to
+% understanding such learning is to frame it as Bayesian inference over
+% probabilistic knowledge bases. Advances in context free grammar
+% induction in the past couple of decades can be seen as an instance of
+% this approach, in which
 
@@
 \subsection{Related Work}
-Our core inference algorithm borrows heavily from recent algorithms for nonparametric PCFG induction, as well as work extending those algorithms to SLPs~\cite{DBLP:conf/emnlp/LiangPJK07,goldwater2006contextual,johnson2006adaptor,cussens2001parameter}.
-Like Johnson and Liang, we are interested in inferring the most likely grammar given some collection of sentences, and in doing so without a pre-specified limit on the complexity of the grammar.
-Like Cussens, we are interested in efficiently extending this algorithm to more powerful and expressive formalisms than CFGs.
+Johnson and Liang, we are interested in inferring the most likely
+grammar given some collection of sentences, and in doing so without a
+pre-specified limit on the complexity of the grammar. Like Cussens, we
+are interested in efficiently extending this algorithm to more
+powerful and expressive formalisms than CFGs.
 
-This work is also closely related to the growing field of semantic parsing.
-Both are aimed at jointly learning syntactic and semantic information from a body of sentences.
-Semantic parsers, however, learn separate systems of syntax and semantics, often pairing Combinatory Categorial Grammars of syntax with an explicit logical language of semantics, often the untyped lambda calculus~\cite{berant2013semantic,liang2013learning,kwiatkowski2010inducing}.
-By contrast, our approach uses a single formalism, the latent predicate network, to describe both which sentences are syntactically valid and what sorts of semantic information those sentence imply.
-Semantic parsers also typically learn models of semantics by pairing sentences with an explicit logical form or real-world denotation.
-In our model it is the grammar induction itself which reveals candidate relations between words that are key to capturing their meaning.
+This work is also closely related to the growing field of semantic
+parsing. Both are aimed at jointly learning syntactic and semantic
+information from a body of sentences. Semantic parsers, however, learn
+separate systems of syntax and semantics, often pairing Combinatory
+Categorial Grammars of syntax with an explicit logical language of
+semantics, often the untyped lambda
+calculus~\cite{berant2013semantic,liang2013learning,kwiatkowski2010inducing}. By
+contrast, our approach uses a single formalism, the latent predicate
+network, to describe both which sentences are syntactically valid and
+what sorts of semantic information those sentence imply. Semantic
+parsers also typically learn models of semantics by pairing sentences
+with an explicit logical form or real-world denotation. In our model
+it is the grammar induction itself which reveals candidate relations
+between words that are key to capturing their meaning.
 
@@
 \end{figure}
-
-An LCG is a hierarchical Bayesian model of string tuples. It is an
-extension of the Hierarhical Dirichlet PCFG model (HD-PCFG) to
-Probabilistic Range Concatenation Grammars (PRCGs). Whereas the
-HD-PCFG is a distribution over strings, an LCG is a distribution over
-string tuples belonging to one of a set of observable
-relations.
-
-Range Concatenation Grammars (RCGs) are a class of string grammars
-that represent all and only those languages that can be parsed in time
-polynomial in the length of the target
-string~\cite{boullier2005range}. An RCG $G=(N, T, V, P, S)$ is a
-5-tuple where $N$ is a finite set of predicate symbols, $T$ is a set
-of terminal symbols, $V$ is a set of variable symbols, P is a finite
-set of $M \geq 0$ clauses of the form $\psi_0 \rightarrow \psi_1 \dots
-\psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
-the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
-N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
-\cup V)^*$ is an argument of $\psi_m$. We call the left hand side term
-of any clause the \emph{head} of that clause and its predicate symbol
-is the \emph{head predicate}.
-
-\textbf{Example:} The following is an RCG for the 3-copy language
-$\{www | w \in {a, b}^*\}$:
-\begin{alignat*}{3}
-&S(XYZ) &&\rightarrow A(X, Y, Z)\\
-&A(aX, aY, aZ) &&\rightarrow A(X, Y, Z)\\
-&A(bX, bY, bZ) &&\rightarrow A(X, Y, Z)\\
-&A(\epsilon, \epsilon, \epsilon) &&\rightarrow 
-\end{alignat*}
-
-A string $x$ is in the language defined by an RCG if one can
-\emph{derive} $S(x)$. A derivation is a sequence of rewrite steps in
-which substrings of the left hand side argument string are bound to
-the variables of the head of some clause, thus determining the
-arguments in the clause body. If a clause has no body terms, then its
-head is derived; otherwise, its head is derived if its body clauses
-are derived~\footnote{This description of the language of an RCG
-  technically only holds for \emph{non-combinatory} RCGs, in which the
-  arguments of body terms can only contain single variables. Since any
-  \emph{combinatory} RCG can be converted into a non-combinatory RCG
-  and we only consider non-combinatory RCGs here, this description
-  suffices.}
-
-We extend RCGs to Probabilistic Range Concatenation Grammars (PRCGs)
-by annotating each clause $C_k \in P$ with probabilities $p_k$ such
-that for all predicates $A \in N$ $\sum_{k:head(C_k)=A} p_k = 1$. A
-PRCGs defines a distribution over strings $x$ by sampling from
-derivations of $S(x)$ according to the product of probabilities of
-clauses used in that derivation. This is a well defined distribution
-as long as no probability mass is placed on derivations of infinite
-length; in this paper, we only consider PRCGs with derivations of
-finite length, so we need not worry about this requirement.
 
@@
 Inference in hierarchical Dirichlet processes in stochastic grammars
-and stochastic logic programs have been well studied. Variational
-Expectation-Maximization (VB-EM) is a popular approach, and this is
-the algorithm we adopt here. The algorithm is described extensively
-elsewhere (CITATION). As we will show in the following section, our
-latent stochastic grammar model can be encoded as a stochastic logic
-program, so that any VB-EM algorithm for that domain will suffice.
+and stochastic logic programs has been an active area of research in
+recent
+decades~\cite{DBLP:conf/emnlp/LiangPJK07,goldwater2006contextual,johnson2006adaptor,cussens2001parameter}.
+Variational inference is a popular approach in this domain and it is
+the approach we adopt here. As we explain in the next section, we
+implemented inference by translating LPNs into PRISM programs, and we
+use the builtin Variational Bayes Expectation-Maximization
+algorithm~\cite{sato2008variational}.
 
@@
 
+\bibliographystyle{amsplain}
+\bibliography{draft}
+
+\nocite{*}
+
 \end{document}

### 2014-10-20 drafts/cfp/draft.tex
Changes to draft.

@@
 \usepackage{graphicx}
+\usepackage{titling}
+
+\setlength{\droptitle}{-10em}   % This is your set screw
+
 \title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
@@
 \begin{document}
+\vspace{-10cm}
 \maketitle
@@
 
+\subsection{Probabilistic Range Concatentation Grammars}
 Range Concatenation Grammars (RCGs) are a class of string grammars
@@
 by annotating each clause $C_k \in P$ with probabilities $p_k$ such
-that for all predicates ${A \in N$ $\sum_{k:head(C_k)=A} p_k = 1}$. A
+that for all predicates ${A \in N, \, \sum_{k:head(C_k)=A} p_k = 1}$. A
 PRCGs defines a distribution over strings $x$ by sampling from
@@
 finite length, so we need not worry about this requirement.
+
+\subsection{Architecture}
+
+An LPN is an PRCG with $K$ binary predicates $A_1, \dots, A_K$ with a
+densely connected structure. The architecture of the LPNs used in this
+work is described in Figure~\ref{fig:architecture}. Specifically, for
+each latent predicate $A_k$,
+
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
+		\caption{A schematic LPN where $m = |L|,\, n = |N|-1$,\, arity$(A_i) = 2$, and arity$(S) = 1$.}
+		\label{fig:architecture}
+\end{figure}
+
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
+		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+\end{figure}
+
+% Now we can describe the model we introduce here, the Latent Concept
+% Grammar (LCG). We introduce the LCG as a machine learning model to
+% learn a distribution over instances of relations over strings. Unlike
+% a PRCG, which places a distribution over terms RCGs consist of a set
+% of rewrite rules, called clauses, each of which takes a termwhich
+% operate on a set of terms (predicates applied to arguments) to produce
+% new terms. Each rewrite rule is of the form:
 
@@
 
-Our goal is to understand how children can obtain semantic knowledge
-about a system of concepts from the structure of relavent perceptual
-stimuli. How does a child learn that prefixing ``great-'' corresponds
-to adding an additional generation in kinship relationships? How does
-a child learn that ``just before'' is after ``a while before'' which
-is before ``a while after''? How does a child learn that if
-Figure~\ref{pairOfApples} is ``a pair of apples'' then
-Figure~\ref{singleApple} is a single apple?
+% Our goal is to understand how children can obtain semantic knowledge
+% about a system of concepts from the structure of relavent perceptual
+% stimuli. How does a child learn that prefixing ``great-'' corresponds
+% to adding an additional generation in kinship relationships? How does
+% a child learn that ``just before'' is after ``a while before'' which
+% is before ``a while after''? How does a child learn that if
+% Figure~\ref{pairOfApples} is ``a pair of apples'' then
+% Figure~\ref{singleApple} is a single apple?
 
-Our proposal is that some interesting fragments of human conceptual
-knowledge can be represented as structured relations between
-linguistic tokens and other perceptual inputs; that the rules that
-define these relations implicitly define the concepts that are
-learned; and that these rules can be learned using the tools of
-hierarchical Bayesian inference.
+% Our proposal is that some interesting fragments of human conceptual
+% knowledge can be represented as structured relations between
+% linguistic tokens and other perceptual inputs; that the rules that
+% define these relations implicitly define the concepts that are
+% learned; and that these rules can be learned using the tools of
+% hierarchical Bayesian inference.
 
-For example, the concept TWENTY SEVEN is defined as the collection of
-representations ``27'', ``twenty seven'', ``3 times 9,'' along with a
-system of symbolic rules that translate among these representations
-and facilitate answered various queries. According to our hypothesis,
-the difference between TWENTY EIGHT and TWENTY SEVEN is not that they
-are represented in the mind by distinct mental symbols but rather that
-the rules that operate on their various representations treat them in
-systematically different ways.
+% For example, the concept TWENTY SEVEN is defined as the collection of
+% representations ``27'', ``twenty seven'', ``3 times 9,'' along with a
+% system of symbolic rules that translate among these representations
+% and facilitate answered various queries. According to our hypothesis,
+% the difference between TWENTY EIGHT and TWENTY SEVEN is not that they
+% are represented in the mind by distinct mental symbols but rather that
+% the rules that operate on their various representations treat them in
+% systematically different ways.
 
-This view implies that an attempt to construct a computational account
-of concept learning in humans needs to ground out that learning in the
-relavent perceptual stimuli.
+% This view implies that an attempt to construct a computational account
+% of concept learning in humans needs to ground out that learning in the
+% relavent perceptual stimuli.
 
-Contribution:
-\begin{itemize}
-\item We present Generate Rewrite Systems (GRS) -- a formalism that borrows
-  heavily from Linear Context-Free Rewriting Systems (CITE)
-\item We show how GRS, though restricted in representational capactiy,
-  can capture substantial cognitively plausible first-order conceptual knowledge 
-\item We propose a heirarchical Bayesian model for GRS that provides a coherent framework in which a GRS can be learned given evidence
-\item We show how GRS can be transformed into PRISM programs and learned within that probabilstic logic programming system
-\item We provide computational experiments that provide evidence that
-  our model can provide an account of some kinds of complex conceptual
-  knowledge can be inferred from noisy and sparse data
-\end{itemize}
+% Contribution:
+% \begin{itemize}
+% \item We present Generate Rewrite Systems (GRS) -- a formalism that borrows
+%   heavily from Linear Context-Free Rewriting Systems (CITE)
+% \item We show how GRS, though restricted in representational capactiy,
+%   can capture substantial cognitively plausible first-order conceptual knowledge 
+% \item We propose a heirarchical Bayesian model for GRS that provides a coherent framework in which a GRS can be learned given evidence
+% \item We show how GRS can be transformed into PRISM programs and learned within that probabilstic logic programming system
+% \item We provide computational experiments that provide evidence that
+%   our model can provide an account of some kinds of complex conceptual
+%   knowledge can be inferred from noisy and sparse data
+% \end{itemize}
 
@@
 
-\section{The Model}
+% \section{The Model}
 
-\begin{figure}[t]
-		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
-		\caption{A schematic LPN where $m = |L|,\, n = |N|-1$,\, arity$(A_i) = 2$, and arity$(S) = 1$.}
-		\label{fig:schematic}
-\end{figure}
+% Suppose we are given some positive instances $y_1, \dots,
+% y_M$ of relations $R_1, \dots, R_N$. Our model assumes that each
+% positive instance $y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent
+% stochastic grammar (LSG), $\mathcal{K}$. $\mathcal{K}$ consists of
+% some fixed number $K$ of clauses (i.e. rules) $C_1, \dots, C_K$ over
+% the $N$ observable relations plus a fixed number of $H$ hidden
+% relations $T_1, \dots, T_H$. The \emph{architecture} of the model is
+% the exact specification of the parameters $K$ and $H$ and the clauses
+% $\mathcal{C}$. To each clause $C_i$ is associated a positive real
+% weights $\theta_i$.
 
-\begin{figure}[t]
-		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
-		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
-\end{figure}
+% The clauses of $\mathcal{K}$ are Horn clauses in a first order logic
+% with string concatenation. We will use PROLOG notation, in which lower
+% case letters are predicates and constants and uppercase letters are
+% variables. If two symbols are adjacent it means that the corresponding
+% strings are concatenated. For example, the following would is a valid
+% clause: $a(XY, XV) \leftarrow b(X, UV), a(X, X)$. In this rule $a$ and
+% $b$ are relations with arity two.
 
-Now we can describe the model we introduce here, the Latent Concept
-Grammar (LCG). We introduce the LCG as a machine learning model to
-learn a distribution over instances of relations over strings. Unlike
-a PRCG, which places a distribution over terms RCGs consist of a set
-of rewrite rules, called clauses, each of which takes a termwhich
-operate on a set of terms (predicates applied to arguments) to produce
-new terms. Each rewrite rule is of the form:
+% Let $C_1, \dots, C_{N(R)} = C(R)$ be the set of clauses whose head is
+% relation $R$ and let $\Theta_R$ be the set of corresponding
+% weights. Let $p_R(C_n) \sim \text{Discrete}(\Theta_R)$; that is, $p_R$
+% is a distribution over the clauses headed by $R$, such that each
+% clauses is sampled in proportion to its weight parameter.
 
-Suppose we are given some positive instances $y_1, \dots,
-y_M$ of relations $R_1, \dots, R_N$. Our model assumes that each
-positive instance $y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent
-stochastic grammar (LSG), $\mathcal{K}$. $\mathcal{K}$ consists of
-some fixed number $K$ of clauses (i.e. rules) $C_1, \dots, C_K$ over
-the $N$ observable relations plus a fixed number of $H$ hidden
-relations $T_1, \dots, T_H$. The \emph{architecture} of the model is
-the exact specification of the parameters $K$ and $H$ and the clauses
-$\mathcal{C}$. To each clause $C_i$ is associated a positive real
-weights $\theta_i$.
+% A positive instance of this relation is sampled from $\mathcal{K}$ by
+% arbitrarily (i.e. uniformly at random) choosing a relation $R$ and
+% instantiating that relation by sampling a clause from $p_R$. Recursing
+% on this procedure for every atom in the body of this clause binds the
+% variables in the clause's head.
 
-The clauses of $\mathcal{K}$ are Horn clauses in a first order logic
-with string concatenation. We will use PROLOG notation, in which lower
-case letters are predicates and constants and uppercase letters are
-variables. If two symbols are adjacent it means that the corresponding
-strings are concatenated. For example, the following would is a valid
-clause: $a(XY, XV) \leftarrow b(X, UV), a(X, X)$. In this rule $a$ and
-$b$ are relations with arity two.
-
-Let $C_1, \dots, C_{N(R)} = C(R)$ be the set of clauses whose head is
-relation $R$ and let $\Theta_R$ be the set of corresponding
-weights. Let $p_R(C_n) \sim \text{Discrete}(\Theta_R)$; that is, $p_R$
-is a distribution over the clauses headed by $R$, such that each
-clauses is sampled in proportion to its weight parameter.
-
-A positive instance of this relation is sampled from $\mathcal{K}$ by
-arbitrarily (i.e. uniformly at random) choosing a relation $R$ and
-instantiating that relation by sampling a clause from $p_R$. Recursing
-on this procedure for every atom in the body of this clause binds the
-variables in the clause's head.
-
-Given this assumed generative process for the evidence, learning a
-program involves inferring the parameter $\theta$ given a prior
-probability distribution over their values. We place an independent
-Dirichlet distribution over the parameters of every relation. This is
-a natural choice that enables us to promote sparse programs, i.e.,
-programs in which most of the mass is placed on a small number of
-predicates and rules.
+% Given this assumed generative process for the evidence, learning a
+% program involves inferring the parameter $\theta$ given a prior
+% probability distribution over their values. We place an independent
+% Dirichlet distribution over the parameters of every relation. This is
+% a natural choice that enables us to promote sparse programs, i.e.,
+% programs in which most of the mass is placed on a small number of
+% predicates and rules.
 

### 2014-10-20 drafts/cfp/draft.tex
another pass through the draft

@@
 
-\section{Abstract}
-How can machines learn about a system of concepts from sentences about
+{\small For humans, learning abstract concepts and learning language
+  go hand in hand: we acquire abstract knowledge primarily through
+  linguistic experience, and acquiring abstract concepts is a crucial
+  step in learning the meanings of linguistic expressions. How,
+  though, can machines learn systems of concepts from sentences about
   those concepts? One approach is to estimate the distribution from
@@
   early work on a learning framework called Latent Predicate Networks
-(LPNs) which attempts to learn concepts by inferring the parameters of
-probabilistic context-sensitive grammars that best describe sentences.
-We show that for a small fragment of sentences expressing
-relationships between English number words, we can use hierarchical
-Bayesian inference to learn LPNs that can answer simple queries about
+  (LPNs) which learns concepts by inferring the parameters of
+  probabilistic context-sensitive grammars over sentences.  We show
+  that for a small fragment of sentences expressing relationships
+  between English number words, we can use hierarchical Bayesian
+  inference to learn LPNs that can answer simple queries about
   previously unseen linguistic concepts within this domain. These
-generalizations, we argue, demonstrate LPNs' promise as a tool for
-learning and representing conceptual knowledge in language.
-
-% A popular approach in statistical machine learning is to automatically
-% find the latent \emph{features} of the data that are most relevant or
-% informative. But a latent feature based view of concept learning does
-% not match up naturally with the structured symbolic knowledge that
-% seems to underlie people's understanding of many domains. How can we
-% combine the flexibility, noise tolerance, and computational
-% practicality of latent feature based statistical machine learning with
-% the symbolic knowledge representations that best capture human
-% understanding? Here present a learning model we dub Latent Concept
-% Grammars (LCGs) as a proposal. An LCG is a densely connected set of
-% probabilistic string rewrite rules over an initially uninterpreted set
-% of predicates. When fit to data using a hierarchical Bayesian prior,
-% these predicates assume the roles of informative latent concepts. This
-% approach allows us to learn knowledge that is stochastic and
-% recursively structured while maintaining computational
-% tractability. We show how learning in LCGs can be implemented using
-% the probabilistic logic programming system PRISM, and we present
-% preliminary experiments.
-}
-% How do people learn systems of related concepts from
-% data?  Consider, for example, a school-age child's knowledge of
-% number. This knowledge supports a large number of related operations:
-% the child knows which words denote numbers, how to use those words to
-% label cardinalities of sets, how to translate between digit
-% representations and word representations of number, how to compare
-% cardinalities, etc. The child knows how to perform these operations
-% over an unbounded set of elements, even though the data that underlies
-% this knowledge is small and noisy. Here we propose that such systems
-% of related concepts can be learned via hierarchical probabilistic
-% inference over probabilistic string rewrite systems (PSRS) -- a mildly
-% context-sensitive class of probabilistic string gramamrs that
-% corresponds to a restricted fragment of stochastic logic programs
-% (SLPs). Specifically, we extend the variational
-% expectation-maximization algorithm for Hierarchical Dirichlet PCFGs to
-% accomodate PSRSs. This algorithm we present enables a tractible
-% compromise between highly restrictive context-free knowledge
-% representations and the intractability of performing similar inference
-% procedures over general logic programs. Empirically, we show how such
-% restricted string rewrite systems can capture non-trivial conceptual
-% knowledge and how our algorithm is able to learn such knowledge from
-% data.
-
-% \section{Abstract 2}
-% What does it mean to understand a concept, and how do people come to
-% such an understanding.
+  generalizations demonstrate LPNs' promise as a tool for learning and
+  representing conceptual knowledge in language.}
 
@@
 1950~\cite{Shannon1950}, an adult English speaker's understanding of
-her language and her world allows her to reveal that the entropy of a
-character in English text is only a little more than one
-bit. Plausibly, any machine learning system that were able to predict
-natural language text with such accuracy will have learned much about
-the concepts expressed in that text. This intuition motivates us to
-invert Shannon's experiment and attempt to learn about linguistic
-concepts by learning a probabilistic generative model for natural
-language.
+her language and her world is such that the entropy of a character in
+English text is slightly more than one bit. Plausibly, any machine
+learning system able to predict natural language text with such
+accuracy will have learned much about the concepts expressed in that
+text. This intuition motivates us to invert Shannon's experiment and
+learn linguistic concepts by learning a probabilistic generative model
+for natural language.
 
@@
 studied generative models for natural language
-syntax~\cite{Someone}. And though there ongoing debate about their
-ability to express the syntax of natural language, there is no doubt
-that they cannot express dependencies between words that depend upon
+syntax~\cite{Someone}. While debate continues about their ability to
+express the syntax of natural language, it is agreed that they cannot
+express dependencies between words that depend upon each other for
 their meaning. Consider an example from the domain of sentences
 addressed in this paper: ``after twenty eight comes twenty nine.''
-The second ``twenty'' depends on the first, and the ``nine'' depends
-on the ``eight,'' and this dependency pattern is a clear example of a
-\emph{cross-serial dependency}, a dependency pattern that context-free
-grammars do not support.
+The second ``twenty'' depends on the first ``twenty'' and the
+``eight'', and the ``nine'' depends on the ``eight''. This pattern is
+a clear example of the sort of \emph{cross-serial dependencies} which
+context-free grammars cannot support.
 
@@
 context \emph{sensitive} grammars for modeling natural language,
-extending the tools of hierarchical grammar induction~\ref{Cite} from
+extending the tools of hierarchical grammar induction~\cite{SOMEONE} from
 context-free grammars to ones that can express not only syntactic
@@
 
-An LPN is a hierarchical Bayesian model of strings. It is an
-extension of the Hierarhical Dirichlet PCFG model (HD-PCFG) to
-Probabilistic Range Concatenation Grammars (PRCGs).
+An LPN is a hierarchical Bayesian model of strings
+extending the Hierarchical Dirichlet PCFG model (HD-PCFG) to
+Stochastic Range Concatenation Grammars (SRCGs).
 
@@
 
-We extend RCGs to Probabilistic Range Concatenation Grammars (PRCGs)
-by annotating each clause $C_k \in P$ with probabilities $p_k$ such
-that for all predicates ${A \in N, \, \sum_{k:head(C_k)=A} p_k = 1}$. A
-PRCGs defines a distribution over strings $x$ by sampling from
-derivations of $S(x)$ according to the product of probabilities of
-clauses used in that derivation. This is a well defined distribution
-as long as no probability mass is placed on derivations of infinite
-length; in this paper, we only consider PRCGs with derivations of
-finite length, so we need not worry about this requirement.
+We extend RCGs to SRCGs by annotating each clause $C_k \in P$ with
+probabilities $p_k$ such that for all predicates ${A \in N, \,
+  \sum_{k:head(C_k)=A} p_k = 1}$. An SRCG defines a distribution over
+strings $x$ by sampling from derivations of $S(x)$ according to the
+product of probabilities of clauses used in that derivation. This is a
+well defined distribution as long as no probability mass is placed on
+derivations of infinite length; in this paper, we only consider SRCGs
+with derivations of finite length, so we need not worry about this
+requirement.
 
@@
 
-An LPN is an PRCG with $K$ binary predicates $A_1, \dots, A_K$ with a
-densely connected structure. The architecture of the LPNs used in this
-work is described in Figure~\ref{fig:architecture}. Specifically, for
-each latent predicate $A_k$,
+An LPN is an SRCG with 1 unary predicate $S$, and $K$ densely
+connected binary predicates $A_1, \dots, A_K$ (Figure
+\ref{fig:architecture}).  Specifically, for each latent predicate
+$A_i$, we define rules such that $A_i(w_a,w_b)$ is true for each
+possible pair of terminals, $w_a,w_b \in T$. We also define rules such
+that $A_i(XY,UV)$ is true for each possible pair of predicates,
+$A_j,A_k$, and every possible ordering of the variables $X,Y,U$ and
+$V$ across $A_j$ and $A_k$. Finally, we define $S(XY)$ to be the
+concatenation of two strings generated by $A_1$.
 
@@
 		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
-		\caption{A schematic LPN where $m = |L|,\, n = |N|-1$,\, arity$(A_i) = 2$, and arity$(S) = 1$.}
+		\caption{A schematic LPN.}
 		\label{fig:architecture}
@@
 
-\subsection{Related Work}
-Johnson and Liang, we are interested in inferring the most likely
-grammar given some collection of sentences, and in doing so without a
-pre-specified limit on the complexity of the grammar. Like Cussens, we
-are interested in efficiently extending this algorithm to more
-powerful and expressive formalisms than CFGs.
-
-This work is also closely related to the growing field of semantic
-parsing. Both are aimed at jointly learning syntactic and semantic
-information from a body of sentences. Semantic parsers, however, learn
-separate systems of syntax and semantics, often pairing Combinatory
-Categorial Grammars of syntax with an explicit logical language of
-semantics, often the untyped lambda
-calculus~\cite{berant2013semantic,liang2013learning,kwiatkowski2010inducing}. By
-contrast, our approach uses a single formalism, the latent predicate
-network, to describe both which sentences are syntactically valid and
-what sorts of semantic information those sentence imply. Semantic
-parsers also typically learn models of semantics by pairing sentences
-with an explicit logical form or real-world denotation. In our model
-it is the grammar induction itself which reveals candidate relations
-between words that are key to capturing their meaning.
-
-More generally, we see this model as helping to strike a balance between expressivity and tractability, bridge the gap between more traditional PCFG grammar induction and program induction over more powerful languages~\cite{dechter2013bootstrap,} (CITATION other P.I.
-papers).
-
 % \section{The Model}
@@
 Inference in hierarchical Dirichlet processes in stochastic grammars
-and stochastic logic programs has been an active area of research in
+and stochastic logic programs (SLPs) has been an active area of research in
 recent
@@
 	\begin{minipage}[b]{\linewidth}
-A(a, a) :: 0.25. \\
-A(b, b) :: 0.3. \\
-A(X a, Y a) $\leftarrow$ A(X, Y) :: 0.25. \\
-A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.
+1) A(a, a) :: 0.25. \\
+2) A(b, b) :: 0.3. \\
+3) A(X a, Y a) $\leftarrow$ A(X, Y) :: 0.25. \\
+4) A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.
 		\subcaption{}
@@
 
-LPNs can be encoded as a restricted and tractable subclass of Stochastic Logic Programs (SLP).
-Specifically, we encode LPNs as programs in PRISM PROLOG.
-PRISM provides probabilistic execution and Bayesian inference over SLPs.
+LPNs can be encoded as a restricted and tractable subclass of
+SLPs. Specifically, we encode LPNs with PRISM PROLOG to obtain
+built-in predicates for probabilistic execution and Bayesian inference
+over SLPs (Figure \ref{fig:copy}).
 
-To illustrate the encoding, consider the 2-copy language (CL2) in Figure \ref{fig:copy}.
-CL2 is a toy language producing two identical strings of $a$s and $b$s, making it mildly context-sensitive.
-Here, we see four rules concerning a single symbol, $A$, whose total probability sums to unity.
+First, we define the distribution over each non-terminal symbol, $n
+\in N$, using the {\tt values} predicate. For each $n$ we define the
+values the symbol can take on ({\it i.e.} one for each rule), as well
+as the probability of each value. For each rule in the LPN, we also
+define a corresponding clause of the {\tt reduce} predicate to capture
+the actual relations between the left-hand and right-hand side of each
+rule. Sampling and parsing begins with a call to {\tt lpn(P-IN)},
+which chooses some rule for the left-hand side via {\tt msw(P,V)} and
+recursively reduces the input term using the chosen rule ({\tt
+reduce(P-IN,V)}).
 
-To translate LPNs into PRISM SLPs, we first define the distribution over each symbol.
-We use the {\tt values} predicate, which takes an atom, in this case the name of the symbol, the values that atom can take on, and the probability of each value.
-For inference, we initialize the distribution uniformly.
-For our encoding, each possible value for an atom (or symbol) corresponds to exactly one clause in CL2 where that symbol appears on the left-hand side.
-There are four clauses for $A$ in CL2, so {\tt values} can take on one of four values, each of whose probabilities corresponds to a rule in CL2.
-
-For each rule in the LPN, we also define a corresponding clause of the {\tt reduce} predicate in our PRISM program.
-These {\tt reduce} rules are where the actual relations between the left-hand and right-hand side of each rule are encoded.
-Because PROLOG isn't perfectly declarative, the actual {\tt reduce} clauses are slightly more complex than presented here, reversing the order of the {\tt append} and recursive {\tt lpn} calls based on whether we are sampling or parsing.
-
-To parse or sample a sentence, we call {\tt lpn(P-IN)}, where {\tt P} is the top-level symbol, {\tt IN} is a list of terminal lists, and {\tt -} denotes a pair ({\it e.g.} {\tt lpn('A'-[[a,a,
-b,a],[a,a,b,a]])}).
-A sample or parse begins by probabilistically choosing some rule for the left-hand side via {\tt msw(P,V)} and then recursively reducing the input term using the chosen rule ({\tt reduce(P-IN,V)}).
 \section{Experiments}
@@
 
-%1) Introduction
-%  - psychological motivation 
-%    - juxtaposing the kinds of knowledge that freebase has and the kinds that children learn
-%    - compositional semantic structure --> derive semantics from syntax
-%  - what is the contribution?
-%     - the use of string rewrite formalisms to capture conceptual knowledge and conceptual learning
-%     - to bring tools from computational linguistics to learning semantics
-%  - what kind of concept systems are we trying to capture?
-%  - why restrict the formalism? 
-%     - tractability of inference 
-%  - why go beyond context free grammars? 
-%     - if they are intractable to learn anyway, why try to go beyond them? 
-%  - Related Work
-%2) Model
-%- stochastic linear context free rewrite systems
-%  - model: what is the data? and what is being inferred? what is the prior? 
-%  - this can be written as PRISM program, or an SLP. 
-%  - inference: variational approximation 
-%3) Experiments
-%  a) worked example
-%  b) Number system (main result)
-%  3b) Results
-%4) Conclusion
-%5) Citations
+This work is also closely related to the growing field of semantic
+parsing. Both are aimed at jointly learning syntactic and semantic
+information from a body of sentences. Semantic parsers, however, learn
+separate systems of syntax and semantics, often pairing Combinatory
+Categorial Grammars of syntax with an explicit logical language of
+semantics, often the untyped lambda
+calculus~\cite{berant2013semantic,liang2013learning,kwiatkowski2010inducing}. By
+contrast, our approach uses a single formalism, the latent predicate
+network, to describe both which sentences are syntactically valid and
+what sorts of semantic information those sentence imply. Semantic
+parsers also typically learn models of semantics by pairing sentences
+with an explicit logical form or real-world denotation. In our model
+it is the grammar induction itself which reveals candidate relations
+between words that are key to capturing their meaning.
+
+More generally, we see this model as helping to strike a balance
+between expressivity and tractability, bridge the gap between more
+traditional PCFG grammar induction and program induction over more
+powerful languages~\cite{dechter2013bootstrap,} (CITATION other
+P.I. papers).
 

### 2014-10-20 drafts/cfp/draft.tex

@@
 
+\usepackage{fullpage}
 \usepackage{amsmath}
@@
 
-\setlength{\droptitle}{-10em}   % This is your set screw
+\setlength{\droptitle}{-4em}   % This is your set screw
 
@@
 
-\begin{figure}[t]
-		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
-		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
-\end{figure}
-
 % Now we can describe the model we introduce here, the Latent Concept
@@
 
-\section{Inference}
+\subsection{Learning Model}
+
+Given a collection of predicates, $\{A_k\}_{k=1}^{K}$, and a
+distribution over clauses, $\{\vec{w}_{A_k}\}$, the learning task is
+to model a set of utterances, $\{x_j\}_{j=1}^{J}$, as being generated
+according to the following distribution:
+
+\begin{align*}
+  \vec{w}_{A_k} &\sim Dir(\vec{\alpha}_{A_k})\\
+  x_j &\underset{iid}{\sim} p_{\text{\textsc{prcg}}}(S(x_j)\,|\,\{w_{A_k}\})
+\end{align*}
+
+\[ p(\{\vec{w}_{A_k}\}\,|\, \vec{x}, \{\vec{\alpha}_{A_k}\}) \propto\]\[ \prod_j
+p_{\text{\textsc{prcg}}}(x_j|\{\vec{w}_{A_k}\}) \prod_{A_k} p_{\text{\textsc{dir}}}(\vec{w}_{A_k}\,|\,\vec{\alpha}_{A_k}) \]
+
+In words, the weights of clauses sharing head predicate $A_k$ are drawn from a Dirichlet distribution defined by $\vec{\alpha}_{A_k}$. Sentences $x_j$ are then drawn from the resultant PRCG.
+
 Inference in hierarchical Dirichlet processes in stochastic grammars
-and stochastic logic programs (SLPs) has been an active area of research in
-recent
-decades~\cite{DBLP:conf/emnlp/LiangPJK07,goldwater2006contextual,johnson2006adaptor,cussens2001parameter}.
-Variational inference is a popular approach in this domain and it is
-the approach we adopt here. As we explain in the next section, we
-implemented inference by translating LPNs into PRISM programs, and we
-use the builtin Variational Bayes Expectation-Maximization
+and stochastic logic programs (SLPs) has been an active area of
+research in recent decades~\cite{DBLP:conf/emnlp/LiangPJK07,
+  goldwater2006contextual, johnson2006adaptor, cussens2001parameter}.
+Variational inference is a popular approach in this domain and the one
+we adopt here. As explained in the next section, we implemented
+inference by translating LPNs into PRISM programs and using the
+built-in Variational Bayes Expectation-Maximization
 algorithm~\cite{sato2008variational}.
@@
 \section{Experiments}
+
+See Figures \ref{fig:parseexample} and \ref{fig:results}.
+
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
+		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+                \label{fig:parseexample}
+\end{figure}
 

### 2014-10-20 drafts/cfp/draft.tex
another pass on the draft

@@
   linguistic experience, and acquiring abstract concepts is a crucial
-  step in learning the meanings of linguistic expressions. How,
-  though, can machines learn systems of concepts from sentences about
-  those concepts? One approach is to estimate the distribution from
-  which those sentences are drawn, and, in doing so, infer the latent
-  concepts and relationships that best explain the data. We present
-  early work on a learning framework called Latent Predicate Networks
-  (LPNs) which learns concepts by inferring the parameters of
-  probabilistic context-sensitive grammars over sentences.  We show
-  that for a small fragment of sentences expressing relationships
-  between English number words, we can use hierarchical Bayesian
-  inference to learn LPNs that can answer simple queries about
-  previously unseen linguistic concepts within this domain. These
-  generalizations demonstrate LPNs' promise as a tool for learning and
-  representing conceptual knowledge in language.}
+  step in learning the meanings of linguistic expressions. Number
+  knowledge is a case in point: we can only acquire concepts such as
+  seventy-three through linguistic means, and we can only know what
+  the sentence ``Seventy-three is more than twice as big as thirty-one"
+  means if we can grasp the meanings of all its component number
+  words. How, though, can machines learn systems of concepts from
+  sentences about those concepts? One approach is to estimate the
+  distribution from which those sentences are drawn, and, in doing so,
+  infer the latent concepts and relationships that best explain the
+  data. We present early work on a learning framework called Latent
+  Predicate Networks (LPNs) which learns concepts by inferring the
+  parameters of probabilistic context-sensitive grammars over
+  sentences.  We show that for a small fragment of sentences
+  expressing relationships between English number words, we can use
+  hierarchical Bayesian inference to learn LPNs that can answer simple
+  queries about previously unseen linguistic concepts within this
+  domain. These generalizations demonstrate LPNs' promise as a tool
+  for learning and representing conceptual knowledge in language.}
 
@@
 ``eight'', and the ``nine'' depends on the ``eight''. This pattern is
-a clear example of the sort of \emph{cross-serial dependencies} which
+a clear example of the \emph{cross-serial dependencies} which
 context-free grammars cannot support.
@@
 	\centering
-	\begin{minipage}[b]{\linewidth}
+	\begin{minipage}[b]{0.8\linewidth}
 1) A(a, a) :: 0.25. \\
@@
 	%\rule[0pt]{0.8\linewidth}{.5pt}
-	\begin{minipage}[b]{\linewidth}
-		\begin{lstlisting}
-values('A', [1, 2, 3, 4],
+	\begin{minipage}[b]{0.8\linewidth}
+        \fontsize{9}{10.5}\selectfont\ttfamily
+		\begin{verbatim}
+values(`A', [1, 2, 3, 4],
     [0.25, 0.30, 0.25, 0.20]).
 
-reduce('A'-[[a],[a]],1).
-reduce('A'-[[b],[b]],2).
-reduce('A'-[A2,B2],3) :- lpn('A'-[X,Y]),
+reduce(`A'-[[a],[a]],1).
+reduce(`A'-[[b],[b]],2).
+reduce(`A'-[A2,B2],3) :- lpn(`A'-[X,Y]),
     append(X,[a],A2), append(Y,[a],B2).
-reduce('A'-[A2,B2],4) :- lpn('A'-[X,Y]),
+reduce(`A'-[A2,B2],4) :- lpn(`A'-[X,Y]),
     append(X,[b],A2), append(Y,[b],B2).
@@
 lpn(P-IN) :- msw(P,V), reduce(P-IN,V).
-		\end{lstlisting}
+		\end{verbatim}
 		\subcaption{}
@@
 
-This work is also closely related to the growing field of semantic
-parsing. Both are aimed at jointly learning syntactic and semantic
-information from a body of sentences. Semantic parsers, however, learn
+LPNs are only one of many models aimed at jointly learning the
+structure and meaning of natural language expressions, with semantic
+parsers being a popular alternative. Semantic parsers, however, learn
 separate systems of syntax and semantics, often pairing Combinatory
 Categorial Grammars of syntax with an explicit logical language of
-semantics, often the untyped lambda
-calculus~\cite{berant2013semantic,liang2013learning,kwiatkowski2010inducing}. By
-contrast, our approach uses a single formalism, the latent predicate
-network, to describe both which sentences are syntactically valid and
-what sorts of semantic information those sentence imply. Semantic
-parsers also typically learn models of semantics by pairing sentences
-with an explicit logical form or real-world denotation. In our model
-it is the grammar induction itself which reveals candidate relations
-between words that are key to capturing their meaning.
+semantics, typically the untyped lambda
+calculus~\cite{berant2013semantic, liang2013learning,
+  kwiatkowski2010inducing}. By contrast, our approach uses a single
+formalism, the LPN, to describe both which sentences are syntactically
+valid and what sorts of semantic information those sentences
+imply. Semantic parsers also typically learn models of semantics by
+pairing sentences with an explicit logical form or real-world
+denotation. In our model it is the grammar induction over sentences
+alone which reveals candidate relations between words that are key to
+capturing their meaning. Future work should seek to understand the
+limits of modeling syntactic and semantic information as inseparable
+parts of as a single system, and whether there might be ways to
+flexibly combine the strengths of this approach with more traditional
+dual system models.
 
-More generally, we see this model as helping to strike a balance
-between expressivity and tractability, bridge the gap between more
-traditional PCFG grammar induction and program induction over more
-powerful languages~\cite{dechter2013bootstrap,} (CITATION other
-P.I. papers).
+Previous work has shown that generative models of probabilistic
+programs are a useful tool for understanding human concept learning
+for domains as varied as handwritten characters and structured visual
+objects~\cite{dechter2013bootstrap, Stuhlmuller, BrendensPaper}. Here,
+we continue that work by showing how some forms of both human language
+learning and concept learning may be fruitfully modeled as inference
+over a generative model of probabilistic logic programs. Future work
+should work explore how LPNs and its extensions may apply to other
+domains where generative models of probabilistic programming have
+already proven useful, as well as generalizing the current approach to
+linguistic concepts.
 

### 2014-10-21 drafts/cfp/draft.tex
Finished intro draft.

@@
 
-As Claude Shannon demonstrated with his guessing game experiment in
-1950~\cite{Shannon1950}, an adult English speaker's understanding of
-her language and her world is such that the entropy of a character in
-English text is slightly more than one bit. Plausibly, any machine
-learning system able to predict natural language text with such
-accuracy will have learned much about the concepts expressed in that
-text. This intuition motivates us to invert Shannon's experiment and
-learn linguistic concepts by learning a probabilistic generative model
-for natural language.
+Although concept learning and language acquisition have typically been
+treated as distinct problems in AI, linguistics and cognitive
+development, for a child learning to understand language, they are
+strongly coupled. People learn about many abstract concepts primarily
+through language and understanding language depends on understanding
+concepts.  Research in concept learning is often focused on concepts
+grounded in perceptual features, and while it is probably true that
+many concepts are learned via generalization from examples, some
+concepts cannot possibly be learned this way.
 
-Probabilistic Context Free Grammars (PCFGs) are popular and much
-studied generative models for natural language
-syntax~\cite{Someone}. While debate continues about their ability to
-express the syntax of natural language, it is agreed that they cannot
-express dependencies between words that depend upon each other for
-their meaning. Consider an example from the domain of sentences
-addressed in this paper: ``after twenty eight comes twenty nine.''
-The second ``twenty'' depends on the first ``twenty'' and the
-``eight'', and the ``nine'' depends on the ``eight''. This pattern is
-a clear example of the \emph{cross-serial dependencies} which
-context-free grammars cannot support.
+Number concepts are a good example: children do not learn about the
+meaning of ``seventy five'' by seeing examples of seventy five things;
+they do not know that ``seventy five'' is more than ``twenty five''
+because of their perceptual experiences of these quantities. Rather,
+children learn the meaning of ``seventy five'' (or ``a billion and
+five'') by noticing how number words are used in language, in counting
+sequences, in arithmetic excercises, etc. Other good examples of such
+abstract concepts are kinship and social relations (e.g. ``my father
+in law's grandmother''), temporal relations (``the day after last
+Thanksgiving.''), and spatial relations (``above and just to the left
+of'').
 
-Our approach, then, will be to explore the use of probabilistic
-context \emph{sensitive} grammars for modeling natural language,
-extending the tools of hierarchical grammar induction~\cite{SOMEONE} from
-context-free grammars to ones that can express not only syntactic
-relationships but semantic ones as well. In the place of latent
-syntactic categories, we will refer to latent predicates -- relations
-between sentence components -- that, when learned, should correspond
-to the concepts that underlie sentence semantics.
+Such abstract concepts share many of the properties of language
+syntax: there is an unbounded set of them, they derive their
+meanings via composition, and, although people only ever say, hear, or
+read about a small number of them, they are able to reason correctly
+about any them. There seems to be a grammar to these concepts, and
+grasping this grammar is critical to understanding their meanings and
+how to use them. This motivates our approach here, which is to apply
+the tools of probabilistic grammars more familiar from studies of
+syntax to the problem of concept aquisition.
 
-To explore this approach, we introduce a probabilisitic model called a
-Latent Predicate Network (LPN), and describe an inference procedure to
-learn its parameters using the probabilistic logic programming
-language PRISM. Finally, we study its performance when trained on
-sentences describing the relationships between numbers.
+Doing so requires overcoming some technical barriers.
+
+First, whereas context-free grammars are suitable for describing large
+swaths of language syntax, the grammars of concepts are not
+context-free. To address this, we use Range Concatentation Grammars, a
+context-sensitive grammar formalism -- one of several developed within
+linguistics -- and extend this formalism to a probabilistic model of
+sentences.
+
+Second, the categories of syntax -- the nonterminals of the grammar --
+are often assumed to be known to children innately and given to
+automated learners by human experts. The categories that underlie
+conceptual knowledge, on the other hand, are far more numerous, vary
+from domain to domain, and are unlikely to be known to the
+learner. This motivates our use of latent predicates that, through
+learning, assume the role of a domain's underlying concepts (in the
+number domain, these might correspond to the concepts of
+successorship, order of magnitude, magnitude comparison,
+exact vs. approximate, etc.).
+
+Finally, using probabilistic context-sensitive grammars with latent
+predicates brings up problems of tractability: our goal is to find a
+middle ground between expressivity and tractability. Using tools like
+PRISM -- a probabilistic logic programming system that naturally
+implements efficient dynamic programming algorithms for our models --
+we are able to explore which domains and which grammar architectures
+are a good fit for this grammar-based approach.
+
+In the rest of this paper describe early work on this approach. First, we
+present our probabilistic model of concept learning, which we call
+Latent Predicate Networks (LPNs). Then  we describe our approach to
+inference and its implementation in PRISM. Finally, we present
+preliminary experimental results in the domain of number concept
+learning that demonstrate the promise of this approach.
+
+% people learn about many
+% abstract concepts primarily through language and, conversely, these
+% abstract concepts are critical to understanding language. Consider a
+% child learning the meaning of the number words: the meaning of the
+% ``seventy five'' is learned not by seeing collections of seventy five
+% things but rather by understanding how ``seventy five'' is used in the
+% various sentences, counting sequences, arithmetic excercises, etc.
+
+% As Claude Shannon demonstrated with his
+% guessing game experiment in 1950~\cite{Shannon1950}, an adult English
+% speaker's understanding of her language and her world is such that the
+% entropy of a character in English text is slightly more than one
+% bit. Plausibly, any machine learning system able to predict natural
+% language text with such accuracy will have learned much about the
+% concepts expressed in that text. This intuition motivates us to invert
+% Shannon's experiment and learn linguistic concepts by learning a
+% probabilistic generative model for natural language.
+
+% Probabilistic Context Free Grammars (PCFGs) are popular and much
+% studied generative models for natural language
+% syntax~\cite{Someone}. While debate continues about their ability to
+% express the syntax of natural language, it is agreed that they cannot
+% express dependencies between words that depend upon each other for
+% their meaning. Consider an example from the domain of sentences
+% addressed in this paper: ``after twenty eight comes twenty nine.''
+% The second ``twenty'' depends on the first ``twenty'' and the
+% ``eight'', and the ``nine'' depends on the ``eight''. This pattern is
+% a clear example of the \emph{cross-serial dependencies} which
+% context-free grammars cannot support.
+
+% Our approach, then, will be to explore the use of probabilistic
+% context \emph{sensitive} grammars for modeling natural language,
+% extending the tools of hierarchical grammar induction~\cite{SOMEONE} from
+% context-free grammars to ones that can express not only syntactic
+% relationships but semantic ones as well. In the place of latent
+% syntactic categories, we will refer to latent predicates -- relations
+% between sentence components -- that, when learned, should correspond
+% to the concepts that underlie sentence semantics.
+
+% To explore this approach, we introduce a probabilisitic model called a
+% Latent Predicate Network (LPN), and describe an inference procedure to
+% learn its parameters using the probabilistic logic programming
+% language PRISM. Finally, we study its performance when trained on
+% sentences describing the relationships between numbers.
 

### 2014-10-21 drafts/cfp/draft.tex
rework the discussion

--- 2014-10-20 drafts/cfp/draft.tex
@@
-\documentclass[11pt, twocolumn]{article}
+\documentclass[10pt, twocolumn]{article}
 
@@
 \usepackage{titling}
+\usepackage{array}
+%\usepackage{multicol}
 
@@
 \author{Eyal Dechter,  Josh Rule, Josh Tenenbaum}
+\date{}
 
@@
 \maketitle
+\vspace{-2em}
 
@@
   step in learning the meanings of linguistic expressions. Number
-  knowledge is a case in point: we can only acquire concepts such as
+  knowledge is a case in point: we largely acquire concepts such as
   seventy-three through linguistic means, and we can only know what
-  the sentence ``Seventy-three is more than twice as big as thirty-one"
-  means if we can grasp the meanings of all its component number
-  words. How, though, can machines learn systems of concepts from
-  sentences about those concepts? One approach is to estimate the
-  distribution from which those sentences are drawn, and, in doing so,
-  infer the latent concepts and relationships that best explain the
-  data. We present early work on a learning framework called Latent
-  Predicate Networks (LPNs) which learns concepts by inferring the
-  parameters of probabilistic context-sensitive grammars over
-  sentences.  We show that for a small fragment of sentences
-  expressing relationships between English number words, we can use
-  hierarchical Bayesian inference to learn LPNs that can answer simple
-  queries about previously unseen linguistic concepts within this
-  domain. These generalizations demonstrate LPNs' promise as a tool
-  for learning and representing conceptual knowledge in language.}
+  the sentence ``Seventy-three is more than twice as big as
+  thirty-one" means if we can grasp the meanings of its component
+  number words. How do we make sense of this potential paradox? One
+  approach is to estimate the distribution from which sentences
+  are drawn, and, in doing so, infer the latent concepts and
+  relationships that best explain those sentences. We present early work on a
+  learning framework called Latent Predicate Networks (LPNs) which
+  learns concepts by inferring the parameters of probabilistic
+  context-sensitive grammars over sentences.  We show that for a small
+  fragment of sentences expressing relationships between English
+  number words, we can use hierarchical Bayesian inference to learn
+  LPNs that can answer simple queries about previously unseen
+  abstract concepts within this domain. These generalizations
+  demonstrate LPNs' promise as a tool for learning and representing
+  conceptual knowledge in language.}
 
@@
 As Claude Shannon demonstrated with his guessing game experiment in
-1950~\cite{Shannon1950}, an adult English speaker's understanding of
-her language and her world is such that the entropy of a character in
+1950~\cite{shannon1951prediction}, an adult English speaker's understanding of
+language and the world is such that the entropy of a character in
 English text is slightly more than one bit. Plausibly, any machine
-learning system able to predict natural language text with such
+learning system able to predict natural language text with human-like
 accuracy will have learned much about the concepts expressed in that
@@
 context \emph{sensitive} grammars for modeling natural language,
-extending the tools of hierarchical grammar induction~\cite{SOMEONE} from
+extending the tools of hierarchical grammar induction~\cite{johnson2006adaptor, goldwater2006contextual, DBLP:conf/emnlp/LiangPJK07} from
 context-free grammars to ones that can express not only syntactic
@@
 \textbf{Example:} The following is an RCG for the 3-copy language
-$\{www | w \in {a, b}^*\}$:
+$\{www | w \in \{a, b\}^*\}$:
 \begin{alignat*}{3}
@@
 LPNs can be encoded as a restricted and tractable subclass of
-SLPs. Specifically, we encode LPNs with PRISM PROLOG to obtain
-built-in predicates for probabilistic execution and Bayesian inference
-over SLPs (Figure \ref{fig:copy}).
+SLPs. Specifically, we encode LPNs with PRISM to obtain built-in
+predicates for probabilistic execution and Bayesian inference over
+SLPs (Figure \ref{fig:copy}). PRISM is an appropriate choice here,
+because it implements a previously related extension of the
+variational Bayesian Expectation Maximization algorithm for PCFGs to
+SLPs satisfying a number of restrictions
+\cite{cussens2001parameter}. Our LPN encodings satisfy these
+restrictions, in particular the need for a finite explanation graph
+({\it i.e.} set of all possible parse trees) for any
+utterance. Because we disallow $A_i([\,],[\,])$ as a valid clause,
+every term in the body of a clause has shorter arguments than the
+head, giving finite derivations and a finite explanation graph.
 
@@
 
-See Figures \ref{fig:parseexample} and \ref{fig:results}.
-
-\begin{figure}[t]
-		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
-		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
-                \label{fig:parseexample}
-\end{figure}
-
-\begin{figure}
-	\rule{\linewidth}{3in}
-	\caption{The Results}
-	\label{fig:results}
-\end{figure}
-
-\begin{tabular}{l l}
-Question & $N_p=4$ \\ \hline
+\begin{table}[t]
+\begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
+  Question & $K=4$ \\ \hline
   after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
@@
   after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
- after sixty nine comes \underline{\hspace{1cm}}? & sixty ten \checkmark \\
+  after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
   after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
@@
   before ninety five comes \underline{\hspace{1cm}}? &ninety four \checkmark \\
+\end{tabular}
+\caption{results of learning with 4 predicates and XXX}
+\label{tab:results}
+\end{table}
 
+See Figures \ref{fig:parseexample} and \ref{fig:results}.
+
+{\it Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec a
+  diam lectus. Sed sit amet ipsum mauris. Maecenas congue ligula ac
+  quam viverra nec consectetur ante hendrerit. Donec et mollis
+  dolor. Praesent et diam eget libero egestas mattis sit amet vitae
+  augue. Nam tincidunt congue enim, ut porta lorem lacinia
+  consectetur. Donec ut libero sed arcu vehicula ultricies a non
+  tortor. Lorem ipsum dolor sit amet, consectetur adipiscing
+  elit. Aenean ut gravida lorem. Ut turpis felis, pulvinar a semper
+  sed, adipiscing id dolor. Pellentesque auctor nisi id magna
+  consequat sagittis. Curabitur dapibus enim sit amet elit pharetra
+  tincidunt feugiat nisl imperdiet. Ut convallis libero in urna
+  ultrices accumsan. Donec sed odio eros. Donec viverra mi quis quam
+  pulvinar at malesuada arcu rhoncus. Cum sociis natoque penatibus et
+  magnis dis parturient montes, nascetur ridiculus mus. In rutrum
+  accumsan ultricies. Mauris vitae nisi at sem facilisis semper ac in
+  est.
+
+  Vivamus fermentum semper porta. Nunc diam velit, adipiscing ut
+  tristique vitae, sagittis vel odio. Maecenas convallis ullamcorper
+  ultricies. Curabitur ornare, ligula semper consectetur sagittis,
+  nisi diam iaculis velit, id fringilla sem nunc vel mi. Nam dictum,
+  odio nec pretium volutpat, arcu ante placerat erat, non tristique
+  elit urna et turpis. Quisque mi metus, ornare sit amet fermentum et,
+  tincidunt et orci. Fusce eget orci a orci congue vestibulum. Ut
+  dolor diam, elementum et vestibulum eu, porttitor vel
+  elit. Curabitur venenatis pulvinar tellus gravida ornare. Sed et.}
+
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
+		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+                \label{fig:parseexample}
+\end{figure}
+
+\begin{table}[t]
+\begin{tabular}{>{\tiny} l >{\tiny} l}
+\multicolumn{2}{>{\tiny}c}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} \\
+$A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
+$A_3$(one, two) : 0.3993 &
+$A_3$(two, three) : 0.2063 \\
+$A_3$(three, four) : 0.1093 &
+$A_3$(four, five) : 0.0734 \\
+$A_3$(five, six) : 0.0502 &
+$A_3$(six, seven) : 0.0355 \\
+$A_3$(eight, nine) : 0.0290 &
+$A_3$(seven, eight) : 0.0271 \\
+$A_2$(fifty, fifty) : 0.0375 &
+$A_2$(thirty, thirty) : 0.0361 \\
+$A_2$(eighty, eighty) : 0.0339 &
+$A_3$(null, one) : 0.0231 \\
+$A_2$(forty, forty) : 0.0332 &
+$A_2$(twenty, twenty) : 0.0310 \\
+$A_2$(seventy, seventy) : 0.0296 &
+$A_2$(sixty, sixty) : 0.0274 \\
+$A_2$(ninety, ninety) : 0.0260 &
+$A_3$(eighteen, nineteen) : 0.0064 \\
+$A_3$(sixteen, seventeen) : 0.0049 &
+$A_3$(eleven, twelve) : 0.0044 \\
+$A_3$(nine, ten) : 0.0044 &
+$A_3$(thirteen, fourteen) : 0.0044 \\
+$A_3$(fourteen, fifteen) : 0.0039 &
+$A_3$(ten, eleven) : 0.0034 \\
+$A_2$(null, fifty) : 0.0043 &
+$A_3$(eighty, null) : 0.0030 \\
+$A_3$(seventeen, eighteen) : 0.0030 &
+$A_3$(nine, sixty) : 0.0025 \\
+$A_2$(nine, seventy) : 0.0029 &
+$A_3$(nine, forty) : 0.0020 \\
+$A_2$(null, thirty) : 0.0022 &
+$A_2$(nine, ninety) : 0.0022 \\
+$A_3$(twelve, thirteen) : 0.0015 &
+$A_3$(fifteen, sixteen) : 0.0015 \\
+$A_3$(null, comes) : 0.0010 &
+$A_2$(twenty, after) : 0.0014 \\
+$A_2$(sixty, before) : 0.0007 &
+$A_3$(nineteen, comes) : 0.0005 \\
+$A_4$(nine, thirty) : 0.0010 & \\
 \end{tabular}
+\caption{XXX}
+\label{tab:grammar}
+\end{table}
 
-\begin{alignat*}{2}
-&S(X Y) \leftarrow A_1(X, Y) : 1.0000 \\
-&A_1(X Y, U V) \leftarrow A_2(X, U), A_3(V, Y) : 0.5002 \\
-&A_2(before, comes) : 0.7316 \\
-&A_4(after, comes) : 0.9990 \\
-&A_1(X Y, U V) \leftarrow A_3(Y, V), A_4(X, U) : 0.3428 \\
-&A_3(one, two) : 0.3993 \\
-&A_3(two, three) : 0.2063 \\
-&A_3(three, four) : 0.1093 \\
-&A_1(X Y, U V) \leftarrow A_1(V, Y), A_4(X, U) : 0.0796 \\
-&A_1(X Y, U V) \leftarrow A_1(Y, V), A_2(X, U) : 0.0712 \\
-&A_3(four, five) : 0.0734 \\
-&A_3(five, six) : 0.0502 \\
-&A_3(six, seven) : 0.0355 \\
-&A_3(eight, nine) : 0.0290 \\
-&A_3(seven, eight) : 0.0271 \\
-&A_2(fifty, fifty) : 0.0375 \\
-&A_2(thirty, thirty) : 0.0361 \\
-&A_2(eighty, eighty) : 0.0339 \\
-&A_3(null, one) : 0.0231 \\
-&A_2(forty, forty) : 0.0332 \\
-&A_2(twenty, twenty) : 0.0310 \\
-&A_2(seventy, seventy) : 0.0296 \\
-&A_2(sixty, sixty) : 0.0274 \\
-&A_2(ninety, ninety) : 0.0260 \\
-&A_3(eighteen, nineteen) : 0.0064 \\
-&A_3(sixteen, seventeen) : 0.0049 \\
-&A_3(eleven, twelve) : 0.0044 \\
-&A_3(nine, ten) : 0.0044 \\
-&A_3(thirteen, fourteen) : 0.0044 \\
-&A_3(fourteen, fifteen) : 0.0039 \\
-&A_3(ten, eleven) : 0.0034 \\
-&A_2(null, fifty) : 0.0043 \\
-&A_3(eighty, null) : 0.0030 \\
-&A_3(seventeen, eighteen) : 0.0030 \\
-&A_1(X Y, U V) \leftarrow A_2(Y, X), A_3(V, U) : 0.0021 \\
-&A_3(nine, sixty) : 0.0025 \\
-&A_2(nine, seventy) : 0.0029 \\
-&A_3(nine, forty) : 0.0020 \\
-&A_1(X Y, U V) \leftarrow A_1(V, Y), A_2(X, U) : 0.0013 \\
-&A_2(null, thirty) : 0.0022 \\
-&A_2(nine, ninety) : 0.0022 \\
-&A_3(twelve, thirteen) : 0.0015 \\
-&A_3(fifteen, sixteen) : 0.0015 \\
-&A_1(X Y, U V) \leftarrow A_2(V, X), A_3(Y, U) : 0.0008 \\
-&A_3(null, comes) : 0.0010 \\
-&A_1(X Y, U V) \leftarrow A_1(X, U), A_2(V, Y) : 0.0008 \\
-&A_1(X Y, U V) \leftarrow A_1(X, U), A_2(Y, V) : 0.0008 \\
-&A_2(twenty, after) : 0.0014 \\
-&A_2(sixty, before) : 0.0007 \\
-&A_3(nineteen, comes) : 0.0005 \\
-&A_4(nine, thirty) : 0.0010 \\
-&A_1(X Y, U V) \leftarrow A_1(X, U), A_4(Y, V) : 0.0004 \\
-\end{alignat*} \\
 \section{Discussion}
 
-LPNs are only one of many models aimed at jointly learning the
-structure and meaning of natural language expressions, with semantic
-parsers being a popular alternative. Semantic parsers, however, learn
-separate systems of syntax and semantics, often pairing Combinatory
-Categorial Grammars of syntax with an explicit logical language of
-semantics, typically the untyped lambda
-calculus~\cite{berant2013semantic, liang2013learning,
-  kwiatkowski2010inducing}. By contrast, our approach uses a single
-formalism, the LPN, to describe both which sentences are syntactically
-valid and what sorts of semantic information those sentences
-imply. Semantic parsers also typically learn models of semantics by
-pairing sentences with an explicit logical form or real-world
-denotation. In our model it is the grammar induction over sentences
-alone which reveals candidate relations between words that are key to
-capturing their meaning. Future work should seek to understand the
-limits of modeling syntactic and semantic information as inseparable
-parts of as a single system, and whether there might be ways to
-flexibly combine the strengths of this approach with more traditional
-dual system models.
+LPNs seems particularly well-suited to learning in domains where the
+number of concepts is infinite. In domains like these, the only way to
+understand each concept is to discover structure in the sentences that
+refer to these concepts. While number is a prime example, others
+include spatial relations, kinship and other social relations, and
+temporal relations. Future work should explore how LPNs might explain
+learning in these domains, as well as how more complex number
+relations might be learned.
 
-Previous work has shown that generative models of probabilistic
-programs are a useful tool for understanding human concept learning
-for domains as varied as handwritten characters and structured visual
-objects~\cite{dechter2013bootstrap, Stuhlmuller, BrendensPaper}. Here,
-we continue that work by showing how some forms of both human language
-learning and concept learning may be fruitfully modeled as inference
-over a generative model of probabilistic logic programs. Future work
-should work explore how LPNs and its extensions may apply to other
-domains where generative models of probabilistic programming have
-already proven useful, as well as generalizing the current approach to
-linguistic concepts.
+Many of the algorithms and formalisms used here were originally
+developed for use in other closely related areas, including logic
+programming and semantic parsing. Semantic parsing, in particular,
+seems especially related to the challenge we face here of jointly
+learning meaning and structure from
+sentences~\cite{berant2013semantic, liang2013learning,
+  kwiatkowski2010inducing}. Semantic parsing, however, tends to frame
+this challenge slightly differently than we do here, namely, by asking
+how utterances can be mapped to an explicit internal logical
+language. By contrast, we focus here on systems where meaning and
+structure seem inseparable. Understanding how these two approaches
+relate and inform one another is an interesting and open question.
+
+%% Previous work has shown that generative models of probabilistic
+%% programs are a useful tool for understanding human concept learning
+%% for domains as varied as handwritten characters and structured visual
+%% objects~\cite{dechter2013bootstrap, stuhlmuller2010learning,
+%%   lake2012concept}. Here, we continue that work by showing how some
+%% forms of both human language learning and concept learning may be
+%% fruitfully modeled as inference over a generative model of
+%% probabilistic logic programs. Future work should work explore how LPNs
+%% and its extensions may apply to other domains where generative models
+%% of probabilistic programming have already proven useful, as well as
+%% generalizing the current approach to abstract, symbolic concepts.
 

### 2014-10-21 drafts/cfp/draft.tex

@@
 
-As Claude Shannon demonstrated with his guessing game experiment in
-1950~\cite{shannon1951prediction}, an adult English speaker's understanding of
-language and the world is such that the entropy of a character in
-English text is slightly more than one bit. Plausibly, any machine
-learning system able to predict natural language text with human-like
-accuracy will have learned much about the concepts expressed in that
-text. This intuition motivates us to invert Shannon's experiment and
-learn linguistic concepts by learning a probabilistic generative model
-for natural language.
+Although concept learning and language acquisition have typically been
+treated as distinct problems in AI, linguistics and cognitive
+development, for a child learning to understand language, they are
+strongly coupled. People learn about many abstract concepts primarily
+through language and understanding language depends on understanding
+concepts.  Research in concept learning is often focused on concepts
+grounded in perceptual features, and while it is probably true that
+many concepts are learned via generalization from examples, some
+concepts cannot possibly be learned this way.
 
-Probabilistic Context Free Grammars (PCFGs) are popular and much
-studied generative models for natural language
-syntax~\cite{Someone}. While debate continues about their ability to
-express the syntax of natural language, it is agreed that they cannot
-express dependencies between words that depend upon each other for
-their meaning. Consider an example from the domain of sentences
-addressed in this paper: ``after twenty eight comes twenty nine.''
-The second ``twenty'' depends on the first ``twenty'' and the
-``eight'', and the ``nine'' depends on the ``eight''. This pattern is
-a clear example of the \emph{cross-serial dependencies} which
-context-free grammars cannot support.
+Number concepts are a good example: children do not learn about the
+meaning of ``seventy five'' by seeing examples of seventy five things;
+they do not know that ``seventy five'' is more than ``twenty five''
+because of their perceptual experiences of these quantities. Rather,
+children learn the meaning of ``seventy five'' (or ``a billion and
+five'') by noticing how number words are used in language, in counting
+sequences, in arithmetic excercises, etc. Other good examples of such
+abstract concepts are kinship and social relations (e.g. ``my father
+in law's grandmother''), temporal relations (``the day after last
+Thanksgiving.''), and spatial relations (``above and just to the left
+of'').
 
-Our approach, then, will be to explore the use of probabilistic
-context \emph{sensitive} grammars for modeling natural language,
-extending the tools of hierarchical grammar induction~\cite{johnson2006adaptor, goldwater2006contextual, DBLP:conf/emnlp/LiangPJK07} from
-context-free grammars to ones that can express not only syntactic
-relationships but semantic ones as well. In the place of latent
-syntactic categories, we will refer to latent predicates -- relations
-between sentence components -- that, when learned, should correspond
-to the concepts that underlie sentence semantics.
+Such abstract concepts share many of the properties of language
+syntax: there is an unbounded set of them, they derive their
+meanings via composition, and, although people only ever say, hear, or
+read about a small number of them, they are able to reason correctly
+about any them. There seems to be a grammar to these concepts, and
+grasping this grammar is critical to understanding their meanings and
+how to use them. This motivates our approach here, which is to apply
+the tools of probabilistic grammars more familiar from studies of
+syntax to the problem of concept aquisition.
 
-To explore this approach, we introduce a probabilisitic model called a
-Latent Predicate Network (LPN), and describe an inference procedure to
-learn its parameters using the probabilistic logic programming
-language PRISM. Finally, we study its performance when trained on
-sentences describing the relationships between numbers.
+Doing so requires overcoming some technical barriers.
+
+First, whereas context-free grammars are suitable for describing large
+swaths of language syntax, the grammars of concepts are not
+context-free. To address this, we use Range Concatentation Grammars, a
+context-sensitive grammar formalism -- one of several developed within
+linguistics -- and extend this formalism to a probabilistic model of
+sentences.
+
+Second, the categories of syntax -- the nonterminals of the grammar --
+are often assumed to be known to children innately and given to
+automated learners by human experts. The categories that underlie
+conceptual knowledge, on the other hand, are far more numerous, vary
+from domain to domain, and are unlikely to be known to the
+learner. This motivates our use of latent predicates that, through
+learning, assume the role of a domain's underlying concepts (in the
+number domain, these might correspond to the concepts of
+successorship, order of magnitude, magnitude comparison,
+exact vs. approximate, etc.).
+
+Finally, using probabilistic context-sensitive grammars with latent
+predicates brings up problems of tractability: our goal is to find a
+middle ground between expressivity and tractability. Using tools like
+PRISM -- a probabilistic logic programming system that naturally
+implements efficient dynamic programming algorithms for our models --
+we are able to explore which domains and which grammar architectures
+are a good fit for this grammar-based approach.
+
+In the rest of this paper describe early work on this approach. First, we
+present our probabilistic model of concept learning, which we call
+Latent Predicate Networks (LPNs). Then  we describe our approach to
+inference and its implementation in PRISM. Finally, we present
+preliminary experimental results in the domain of number concept
+learning that demonstrate the promise of this approach.
+
+% people learn about many
+% abstract concepts primarily through language and, conversely, these
+% abstract concepts are critical to understanding language. Consider a
+% child learning the meaning of the number words: the meaning of the
+% ``seventy five'' is learned not by seeing collections of seventy five
+% things but rather by understanding how ``seventy five'' is used in the
+% various sentences, counting sequences, arithmetic excercises, etc.
+
+% As Claude Shannon demonstrated with his
+% guessing game experiment in 1950~\cite{Shannon1950}, an adult English
+% speaker's understanding of her language and her world is such that the
+% entropy of a character in English text is slightly more than one
+% bit. Plausibly, any machine learning system able to predict natural
+% language text with such accuracy will have learned much about the
+% concepts expressed in that text. This intuition motivates us to invert
+% Shannon's experiment and learn linguistic concepts by learning a
+% probabilistic generative model for natural language.
+
+% Probabilistic Context Free Grammars (PCFGs) are popular and much
+% studied generative models for natural language
+% syntax~\cite{Someone}. While debate continues about their ability to
+% express the syntax of natural language, it is agreed that they cannot
+% express dependencies between words that depend upon each other for
+% their meaning. Consider an example from the domain of sentences
+% addressed in this paper: ``after twenty eight comes twenty nine.''
+% The second ``twenty'' depends on the first ``twenty'' and the
+% ``eight'', and the ``nine'' depends on the ``eight''. This pattern is
+% a clear example of the \emph{cross-serial dependencies} which
+% context-free grammars cannot support.
+
+% Our approach, then, will be to explore the use of probabilistic
+% context \emph{sensitive} grammars for modeling natural language,
+% extending the tools of hierarchical grammar induction~\cite{SOMEONE} from
+% context-free grammars to ones that can express not only syntactic
+% relationships but semantic ones as well. In the place of latent
+% syntactic categories, we will refer to latent predicates -- relations
+% between sentence components -- that, when learned, should correspond
+% to the concepts that underlie sentence semantics.
+
+% To explore this approach, we introduce a probabilisitic model called a
+% Latent Predicate Network (LPN), and describe an inference procedure to
+% learn its parameters using the probabilistic logic programming
+% language PRISM. Finally, we study its performance when trained on
+% sentences describing the relationships between numbers.
 

### 2014-10-21 drafts/cfp/draft.tex
Added experiments section.

@@
 
+To evaluate LPNs as a probabilistic model of concept acquisition, we
+trained an LPN with $4$ latent predicates on a set of sentences
+expressing successor and predessor relations in numbers between one
+and ninety-nine. The training set was the collection of sentences $$X
+= \{[\text{after}\, | \, \text{before}] \, \langle n \rangle \,
+\text{comes} \, \langle n+1 \rangle \,|\, n \in 1,\dots,99\}.$$ The
+lexicon was the set of number word corresponding to $1$ through $19$,
+the decades $20, \dots, 30$ and the words ``before'' and ``after.'' It
+is not difficult to manually work out an LPN that describes this
+limited domain of sentences; see Figure~\ref{fig:parseexample}, for a
+possible LPN derivation of an example sentence.
+
+Although it is difficult to know how common these kinds of sentences
+are in child-directed speech, words for small numbers are far more
+common than words for larger ones(cite CHILDES). On the other hand,
+children learning to count to large numbers rehearse the sequence. To
+approximate this distribution of evidence, we drew these sample
+sentences from a sum of a geometric distribution with parameter $0.5$
+and a uniform distribution. These components were weighted $75\%$ and
+$25\%$, respectively. We drew $2000$ examples from this distribution,
+holding out the sentences in Table~\ref{tab:results} for
+evaluation.
+
+For inference, we used default $\frac{1}{D}$ pseudocounts (where $D$ is the
+dimensionality of the Dirichlet distributions). We found that different random
+initialization for this experiment did not lead to qualitatively different
+results, though further investigation will be necessary to see how
+robust our the algorithm is to local maximuma for LPNs.
+
+We evaluated the learned model by asking for Viterbi (i.e. maximum a
+posteriori) completions of the last words of each held out test
+sentence. Table~\ref{tab:results} shows these completions. The grammar
+correctly learns much of the structure of these sentences, including
+the difference between sentences starting with ``before'' and
+``after'' and the edge cases that relate decade words like ``twenty''
+to non-decade words like ``twenty one.''
+
+To inspect visually the learned grammar, we thresholded rules
+according to the expected number of times they were used in parsing
+the training dataset. Table~\ref{tab:grammar} shows all rules with
+expected count above $1e-6$. This reduces from $2669$ to $52$ the
+number of significant rules. On inspection, predicate $A_2$ forms
+``before'' sentences, predicate $A_4$ forms ``after'' sentences,
+predicate $A_3$ is successorship recursively defined over the decades
+and ones, and predicate $A_2$ is a category for the decade words.
+
 \begin{table}[t]
@@
 \end{table}
-
-See Figures \ref{fig:parseexample} and \ref{fig:results}.
-
-{\it Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec a
-  diam lectus. Sed sit amet ipsum mauris. Maecenas congue ligula ac
-  quam viverra nec consectetur ante hendrerit. Donec et mollis
-  dolor. Praesent et diam eget libero egestas mattis sit amet vitae
-  augue. Nam tincidunt congue enim, ut porta lorem lacinia
-  consectetur. Donec ut libero sed arcu vehicula ultricies a non
-  tortor. Lorem ipsum dolor sit amet, consectetur adipiscing
-  elit. Aenean ut gravida lorem. Ut turpis felis, pulvinar a semper
-  sed, adipiscing id dolor. Pellentesque auctor nisi id magna
-  consequat sagittis. Curabitur dapibus enim sit amet elit pharetra
-  tincidunt feugiat nisl imperdiet. Ut convallis libero in urna
-  ultrices accumsan. Donec sed odio eros. Donec viverra mi quis quam
-  pulvinar at malesuada arcu rhoncus. Cum sociis natoque penatibus et
-  magnis dis parturient montes, nascetur ridiculus mus. In rutrum
-  accumsan ultricies. Mauris vitae nisi at sem facilisis semper ac in
-  est.
-
-  Vivamus fermentum semper porta. Nunc diam velit, adipiscing ut
-  tristique vitae, sagittis vel odio. Maecenas convallis ullamcorper
-  ultricies. Curabitur ornare, ligula semper consectetur sagittis,
-  nisi diam iaculis velit, id fringilla sem nunc vel mi. Nam dictum,
-  odio nec pretium volutpat, arcu ante placerat erat, non tristique
-  elit urna et turpis. Quisque mi metus, ornare sit amet fermentum et,
-  tincidunt et orci. Fusce eget orci a orci congue vestibulum. Ut
-  dolor diam, elementum et vestibulum eu, porttitor vel
-  elit. Curabitur venenatis pulvinar tellus gravida ornare. Sed et.}
 

### 2014-10-21 drafts/cfp/draft.tex
almost finished with draft

--- 2014-10-21 drafts/cfp/draft.tex
@@
   number words. How do we make sense of this potential paradox? One
-  approach is to estimate the distribution from which sentences
-  are drawn, and, in doing so, infer the latent concepts and
-  relationships that best explain those sentences. We present early work on a
+  approach is to estimate the distribution from which sentences are
+  drawn, and, in doing so, infer the latent concepts and relationships
+  that best explain those sentences. We present early work on a
   learning framework called Latent Predicate Networks (LPNs) which
@@
   number words, we can use hierarchical Bayesian inference to learn
-  LPNs that can answer simple queries about previously unseen
-  abstract concepts within this domain. These generalizations
-  demonstrate LPNs' promise as a tool for learning and representing
-  conceptual knowledge in language.}
+  LPNs that can answer simple queries about previously unseen abstract
+  concepts within this domain. These generalizations demonstrate LPNs'
+  promise as a tool for learning and representing conceptual knowledge
+  in language.}
 
@@
 treated as distinct problems in AI, linguistics and cognitive
-development, for a child learning to understand language, they are
-strongly coupled. People learn about many abstract concepts primarily
-through language and understanding language depends on understanding
-concepts.  Research in concept learning is often focused on concepts
-grounded in perceptual features, and while it is probably true that
-many concepts are learned via generalization from examples, some
-concepts cannot possibly be learned this way.
+development, they are strongly coupled for a child learning to
+understand language. More generally, people learn many abstract
+concepts primarily through language even though understanding language
+depends on understanding the underlying concepts. Research in concept
+learning is often focused on concepts grounded in perceptual features,
+and while it is almost certainly true that many concepts are learned
+via generalization from concrete examples, some concepts cannot
+possibly be learned this way.
 
@@
 
-Such abstract concepts share many of the properties of language
-syntax: there is an unbounded set of them, they derive their
-meanings via composition, and, although people only ever say, hear, or
-read about a small number of them, they are able to reason correctly
-about any them. There seems to be a grammar to these concepts, and
-grasping this grammar is critical to understanding their meanings and
-how to use them. This motivates our approach here, which is to apply
-the tools of probabilistic grammars more familiar from studies of
-syntax to the problem of concept aquisition.
+Such abstract systems of concepts share many of the properties of
+language syntax: there is an unbounded set of concepts, they derive
+their meanings via composition, and, although people only ever say,
+hear, or read about a small number of them, they are able to reason
+correctly about any them. There seems to be a grammar to these
+concepts, and grasping this grammar is critical to understanding their
+meanings and how to use them. This motivates our approach here, which
+is to apply the tools of probabilistic grammars more familiar from
+studies of syntax to the problem of concept aquisition.
 
@@
 number domain, these might correspond to the concepts of
-successorship, order of magnitude, magnitude comparison,
-exact vs. approximate, etc.).
+successorship, order of magnitude, magnitude comparison, exact
+vs. approximate, {\it etc}.).
 
-Finally, using probabilistic context-sensitive grammars with latent
-predicates brings up problems of tractability: our goal is to find a
+Finally, inducing probabilistic context-sensitive grammars with latent
+predicates threatens to be intractable: our goal is to find a
 middle ground between expressivity and tractability. Using tools like
@@
 
-In the rest of this paper describe early work on this approach. First, we
-present our probabilistic model of concept learning, which we call
-Latent Predicate Networks (LPNs). Then  we describe our approach to
-inference and its implementation in PRISM. Finally, we present
-preliminary experimental results in the domain of number concept
-learning that demonstrate the promise of this approach.
+The rest of this paper describes early work on this approach. First,
+we present Latent Predicate Networks (LPNs), our probabilistic model
+of concept learning. Then, we describe our approach to inference and
+its implementation in PRISM. Finally, we present preliminary
+experimental results in the domain of number concept learning that
+demonstrate the promise of this approach.
 
@@
 
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
+		\caption{A schematic LPN.}
+		\label{fig:architecture}
+\end{figure}
+
 An LPN is a hierarchical Bayesian model of strings
 extending the Hierarchical Dirichlet PCFG model (HD-PCFG) to
-Stochastic Range Concatenation Grammars (SRCGs).
+Probabilistic Range Concatenation Grammars (PRCGs).
 
@@
 
-We extend RCGs to SRCGs by annotating each clause $C_k \in P$ with
+We extend RCGs to PRCGs by annotating each clause $C_k \in P$ with
 probabilities $p_k$ such that for all predicates ${A \in N, \,
-  \sum_{k:head(C_k)=A} p_k = 1}$. An SRCG defines a distribution over
+  \sum_{k:head(C_k)=A} p_k = 1}$. An PRCG defines a distribution over
 strings $x$ by sampling from derivations of $S(x)$ according to the
@@
 well defined distribution as long as no probability mass is placed on
-derivations of infinite length; in this paper, we only consider SRCGs
+derivations of infinite length; in this paper, we only consider PRCGs
 with derivations of finite length, so we need not worry about this
@@
 
-An LPN is an SRCG with 1 unary predicate $S$, and $K$ densely
-connected binary predicates $A_1, \dots, A_K$ (Figure
+An LPN is an PRCG with 1 unary predicate $S$, and $K$ densely
+connected binary predicates ${A_k}_{k=1}^K$ (Figure
 \ref{fig:architecture}). Specifically, for each latent predicate
-$A_i$, we define rules such that $A_i(w_a,w_b)$ is true for each
+$A_k$, we define rules such that $A_k(w_a,w_b)$ is true for each
 possible pair of terminals, $w_a,w_b \in T$. We also define rules such
-that $A_i(XY,UV)$ is true for each possible pair of predicates,
-$A_j,A_k$, and every possible ordering of the variables $X,Y,U$ and
-$V$ across $A_j$ and $A_k$. Finally, we define $S(XY)$ to be the
-concatenation of two strings generated by $A_1$.
-
-\begin{figure}[t]
-		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
-		\caption{A schematic LPN.}
-		\label{fig:architecture}
-\end{figure}
+that $A_k(XY,UV)$ is true for each possible pair of predicates,
+$A_i,A_j$, and every possible ordering of the variables $X,Y,U$ and
+$V$ across $A_i$ and $A_j$. Finally, we define $S(XY)$ to be the
+concatenation of the two arguments of $A_1(X,Y)$.
 
@@
 \end{tabular}
-\caption{results of learning with 4 predicates and XXX}
+\caption{examples of generalization for held out sentences after learning over 4 latent predicates}
 \label{tab:results}
@@
 
-See Figures \ref{fig:parseexample} and \ref{fig:results}.
-
-{\it Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec a
-  diam lectus. Sed sit amet ipsum mauris. Maecenas congue ligula ac
-  quam viverra nec consectetur ante hendrerit. Donec et mollis
-  dolor. Praesent et diam eget libero egestas mattis sit amet vitae
-  augue. Nam tincidunt congue enim, ut porta lorem lacinia
-  consectetur. Donec ut libero sed arcu vehicula ultricies a non
-  tortor. Lorem ipsum dolor sit amet, consectetur adipiscing
-  elit. Aenean ut gravida lorem. Ut turpis felis, pulvinar a semper
-  sed, adipiscing id dolor. Pellentesque auctor nisi id magna
-  consequat sagittis. Curabitur dapibus enim sit amet elit pharetra
-  tincidunt feugiat nisl imperdiet. Ut convallis libero in urna
-  ultrices accumsan. Donec sed odio eros. Donec viverra mi quis quam
-  pulvinar at malesuada arcu rhoncus. Cum sociis natoque penatibus et
-  magnis dis parturient montes, nascetur ridiculus mus. In rutrum
-  accumsan ultricies. Mauris vitae nisi at sem facilisis semper ac in
-  est.
-
-  Vivamus fermentum semper porta. Nunc diam velit, adipiscing ut
-  tristique vitae, sagittis vel odio. Maecenas convallis ullamcorper
-  ultricies. Curabitur ornare, ligula semper consectetur sagittis,
-  nisi diam iaculis velit, id fringilla sem nunc vel mi. Nam dictum,
-  odio nec pretium volutpat, arcu ante placerat erat, non tristique
-  elit urna et turpis. Quisque mi metus, ornare sit amet fermentum et,
-  tincidunt et orci. Fusce eget orci a orci congue vestibulum. Ut
-  dolor diam, elementum et vestibulum eu, porttitor vel
-  elit. Curabitur venenatis pulvinar tellus gravida ornare. Sed et.}
+See Figures \ref{fig:parseexample}.
 
@@
 \begin{table}[t]
-\begin{tabular}{>{\tiny} l >{\tiny} l}
-\multicolumn{2}{>{\tiny}c}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} \\
-\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} \\
-$A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
-$A_3$(one, two) : 0.3993 &
-$A_3$(two, three) : 0.2063 \\
-$A_3$(three, four) : 0.1093 &
-$A_3$(four, five) : 0.0734 \\
-$A_3$(five, six) : 0.0502 &
-$A_3$(six, seven) : 0.0355 \\
-$A_3$(eight, nine) : 0.0290 &
-$A_3$(seven, eight) : 0.0271 \\
-$A_2$(fifty, fifty) : 0.0375 &
-$A_2$(thirty, thirty) : 0.0361 \\
-$A_2$(eighty, eighty) : 0.0339 &
-$A_3$(null, one) : 0.0231 \\
-$A_2$(forty, forty) : 0.0332 &
-$A_2$(twenty, twenty) : 0.0310 \\
-$A_2$(seventy, seventy) : 0.0296 &
-$A_2$(sixty, sixty) : 0.0274 \\
-$A_2$(ninety, ninety) : 0.0260 &
-$A_3$(eighteen, nineteen) : 0.0064 \\
+\begin{tabular}{>{\tiny} l >{\tiny} l >{\tiny} l}
+\multicolumn{3}{>{\tiny}c}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} \\
+\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} \\
+$A_2$(before, comes) : 0.73 & $A_4$(after, comes) : 1.0  &
+$A_3$(one, two) : 0.40 \\
+$A_3$(two, three) : 0.21 &
+$A_3$(three, four) : 0.11 &
+$A_3$(four, five) : 0.07 \\
+$A_3$(five, six) : 0.05 &
+$A_3$(six, seven) : 0.04 \\
+$A_3$(eight, nine) : 0.03 &
+$A_3$(seven, eight) : 0.03 \\
+$A_2$(fifty, fifty) : 0.04 &
+$A_2$(thirty, thirty) : 0.04 \\
+$A_2$(eighty, eighty) : 0.03 &
+$A_3$(null, one) : 0.02 \\
+$A_2$(forty, forty) : 0.03 &
+$A_2$(twenty, twenty) : 0.03 \\
+$A_2$(seventy, seventy) : 0.03 &
+$A_2$(sixty, sixty) : 0.03 \\
+$A_2$(ninety, ninety) : 0.03 &
+$A_3$(eighteen, nineteen) : 0.01 \\
 $A_3$(sixteen, seventeen) : 0.0049 &
@@
 \end{tabular}
-\caption{XXX}
+\caption{The grammar which produced Table \ref{tab:results}.}
 \label{tab:grammar}

### 2014-10-21 drafts/cfp/draft.tex

@@
 \end{table}
-
-See Figures \ref{fig:parseexample}.
 

### 2014-10-21 drafts/cfp/draft.tex
Draft changes.

@@
 \title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
-\author{Eyal Dechter,  Josh Rule, Josh Tenenbaum}
+\author{Eyal Dechter, Josh Rule, Josh Tenenbaum ADD AFFILIATION}
 \date{}
@@
 predicates threatens to be intractable: our goal is to find a
-middle ground between expressivity and tractability. Using tools like
-PRISM -- a probabilistic logic programming system that naturally
+middle ground between expressivity and tractability. Using
+PRISM~\cite{DBLP:journals/jair/SatoK01} -- a probabilistic logic programming system that naturally
 implements efficient dynamic programming algorithms for our models --
@@
 
-\subsection{Architecture}
+\subsection{Generic Architecture}
 
 An LPN is an PRCG with 1 unary predicate $S$, and $K$ densely
-connected binary predicates ${A_k}_{k=1}^K$ (Figure
+connected binary predicates $\{A_k\}_{k=1}^K$ (Figure
 \ref{fig:architecture}). Specifically, for each latent predicate
 $A_k$, we define rules such that $A_k(w_a,w_b)$ is true for each
-possible pair of terminals, $w_a,w_b \in T$. We also define rules such
-that $A_k(XY,UV)$ is true for each possible pair of predicates,
-$A_i,A_j$, and every possible ordering of the variables $X,Y,U$ and
-$V$ across $A_i$ and $A_j$. Finally, we define $S(XY)$ to be the
-concatenation of the two arguments of $A_1(X,Y)$.
+possible pair of terminals, $w_a,w_b \in T$. $T$ may include the empty
+string, but we do not allow latent predicates of the form
+$A_k(\epsilon, \epsilon)$: see Section~\ref{sec:implementation}. We also define rules such that $A_k(XY,UV)$
+is true for each possible pair of predicates, $A_i,A_j$, and every
+possible ordering of the variables $X,Y,U$ and $V$ across $A_i$ and
+$A_j$. Finally, we define $S(XY)$ to be the concatenation of the two
+arguments of $A_1(X,Y)$.
 
@@
 
-\section{Implementation}
+\section{Implementation \label{sec:implementation}}
 

### 2014-10-21 drafts/cfp/draft.tex
tweaking the tables and figures

--- 2014-10-21 drafts/cfp/draft.tex
@@
 
-\textbf{Example:} The following is an RCG for the 3-copy language
-$\{www | w \in \{a, b\}^*\}$:
-\begin{alignat*}{3}
-&S(XYZ) &&\rightarrow A(X, Y, Z)\\
-&A(aX, aY, aZ) &&\rightarrow A(X, Y, Z)\\
-&A(bX, bY, bZ) &&\rightarrow A(X, Y, Z)\\
-&A(\epsilon, \epsilon, \epsilon) &&\rightarrow 
-\end{alignat*}
+%% \textbf{Example:} The following is an RCG for the 3-copy language
+%% $\{www | w \in \{a, b\}^*\}$:
+%% \begin{alignat*}{3}
+%% &S(XYZ) &&\rightarrow A(X, Y, Z)\\
+%% &A(aX, aY, aZ) &&\rightarrow A(X, Y, Z)\\
+%% &A(bX, bY, bZ) &&\rightarrow A(X, Y, Z)\\
+%% &A(\epsilon, \epsilon, \epsilon) &&\rightarrow 
+%% \end{alignat*}
 
@@
 with derivations of finite length, so we need not worry about this
-requirement.
+requirement. See Figure \ref{fig:grammar} for an example of the context-sensitive 2-copy language $\{ww\,|w \in \{a,b\}^+\}$ as a PRCG.
 
@@
 	\end{minipage}
-	\caption{Possible encodings of the 2-copy language as (\subref{fig:grammar}) an LPN, (\subref{fig:prism}) a PRISM program.}
+	\caption{Possible encodings of the 2-copy language $\{ww\,|\, w \in \{a,b\}^+\}$ as (\subref{fig:grammar}) an LPN, (\subref{fig:prism}) a PRISM program.}
 	\label{fig:copy}
@@
 \begin{table}[t]
-\begin{tabular}{>{\tiny} l >{\tiny} l >{\tiny} l}
-\multicolumn{3}{>{\tiny}c}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} \\
-\multicolumn{3}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} \\
-$A_2$(before, comes) : 0.73 & $A_4$(after, comes) : 1.0  &
-$A_3$(one, two) : 0.40 \\
-$A_3$(two, three) : 0.21 &
-$A_3$(three, four) : 0.11 &
-$A_3$(four, five) : 0.07 \\
-$A_3$(five, six) : 0.05 &
-$A_3$(six, seven) : 0.04 \\
-$A_3$(eight, nine) : 0.03 &
-$A_3$(seven, eight) : 0.03 \\
-$A_2$(fifty, fifty) : 0.04 &
-$A_2$(thirty, thirty) : 0.04 \\
-$A_2$(eighty, eighty) : 0.03 &
-$A_3$(null, one) : 0.02 \\
-$A_2$(forty, forty) : 0.03 &
-$A_2$(twenty, twenty) : 0.03 \\
-$A_2$(seventy, seventy) : 0.03 &
-$A_2$(sixty, sixty) : 0.03 \\
-$A_2$(ninety, ninety) : 0.03 &
-$A_3$(eighteen, nineteen) : 0.01 \\
+\begin{tabular}{>{\tiny} l >{\tiny} l}
+\multicolumn{2}{>{\tiny}c}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} \\
+\multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} \\
+$A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
+$A_3$(one, two) : 0.3993 &
+$A_3$(two, three) : 0.2063 \\
+$A_3$(three, four) : 0.1093 &
+$A_3$(four, five) : 0.0734 \\
+$A_3$(five, six) : 0.0502 &
+$A_3$(six, seven) : 0.0355 \\
+$A_3$(eight, nine) : 0.0290 &
+$A_3$(seven, eight) : 0.0271 \\
+$A_2$(fifty, fifty) : 0.0375 &
+$A_2$(thirty, thirty) : 0.0361 \\
+$A_2$(eighty, eighty) : 0.0339 &
+$A_3$(null, one) : 0.0231 \\
+$A_2$(forty, forty) : 0.0332 &
+$A_2$(twenty, twenty) : 0.0310 \\
+$A_2$(seventy, seventy) : 0.0296 &
+$A_2$(sixty, sixty) : 0.0274 \\
+$A_2$(ninety, ninety) : 0.0260 &
+$A_3$(eighteen, nineteen) : 0.0064 \\
 $A_3$(sixteen, seventeen) : 0.0049 &

### 2014-10-21 drafts/cfp/draft.tex
Edits and reworking of discussion.

@@
 
-An LPN is an PRCG with 1 unary predicate $S$, and $K$ densely
-connected binary predicates $\{A_k\}_{k=1}^K$ (Figure
-\ref{fig:architecture}). Specifically, for each latent predicate
-$A_k$, we define rules such that $A_k(w_a,w_b)$ is true for each
-possible pair of terminals, $w_a,w_b \in T$. $T$ may include the empty
-string, but we do not allow latent predicates of the form
-$A_k(\epsilon, \epsilon)$: see Section~\ref{sec:implementation}. We also define rules such that $A_k(XY,UV)$
-is true for each possible pair of predicates, $A_i,A_j$, and every
-possible ordering of the variables $X,Y,U$ and $V$ across $A_i$ and
-$A_j$. Finally, we define $S(XY)$ to be the concatenation of the two
-arguments of $A_1(X,Y)$.
+An LPN is an PRCG with the following generic architecture: the
+\emph{axiom} of an LPN is a unary predicate $S$, and it has $K$
+densely connected binary predicates $\{A_k\}_{k=1}^K$ (Figure
+\ref{fig:architecture}). For each latent predicate $A_k$, we define
+rules such that $A_k(w_a,w_b)$ is true for each possible pair of
+terminals, $w_a,w_b \in T$. $T$ may include the empty string $\epsilon$, but we
+do not allow latent predicates of the form $A_k(\epsilon, \epsilon)$:
+see Section~\ref{sec:implementation}. We also define rules such that
+$A_k(XY,UV)$ is true for each possible pair of predicates, $A_i,A_j$,
+and every possible ordering of the variables $X,Y,U$ and $V$ across
+$A_i$ and $A_j$. Finally, we define $S(XY)$ to be the concatenation of
+the two arguments of $A_1(X,Y)$.
 
@@
 
-Inference in hierarchical Dirichlet processes in stochastic grammars
-and stochastic logic programs (SLPs) has been an active area of
+Bayesian inference over stochastic grammars
+and stochastic logic programs (CITE MUGGLETON) has been an active area of
 research in recent decades~\cite{DBLP:conf/emnlp/LiangPJK07,
@@
 
-lpn(P-IN) :- msw(P,V), reduce(P-IN,V).
+lpn(P-IN) :- reduce(P-IN,V), msw(P,V).
 		\end{verbatim}
@@
 
-LPNs can be encoded as a restricted and tractable subclass of
-SLPs. Specifically, we encode LPNs with PRISM to obtain built-in
-predicates for probabilistic execution and Bayesian inference over
-SLPs (Figure \ref{fig:copy}). PRISM is an appropriate choice here,
-because it implements a previously related extension of the
-variational Bayesian Expectation Maximization algorithm for PCFGs to
-SLPs satisfying a number of restrictions
-\cite{cussens2001parameter}. Our LPN encodings satisfy these
-restrictions, in particular the need for a finite explanation graph
-({\it i.e.} set of all possible parse trees) for any
-utterance. Because we disallow $A_i([\,],[\,])$ as a valid clause,
-every term in the body of a clause has shorter arguments than the
-head, giving finite derivations and a finite explanation graph.
+LPNs can be encoded as a restricted subclass of PRISM programs; this
+is very similar to how PCFGs are encoded in
+PRISM~\cite{DBLP:conf/cl/2000}. A general probabilistic logic
+programming system based on PROLOG, PRISM proves built-in predicates
+for probabilistic execution and Bayesian inference over logic programs
+with stochastic choices. There are several
+restrictions placed on PRISM programs to maintain the validity of
+their probabilistic interpretation. Most importantly, derivations of a
+probabilistic predicate cannot contain cycles. Because we disallow
+$A_i(\epsilon,\epsilon)$ as a valid clause, every term in the body of a clause
+has shorter arguments than the head, giving acyclic derivations.
 
-First, we define the distribution over each non-terminal symbol, $n
-\in N$, using the {\tt values} predicate. For each $n$ we define the
-values the symbol can take on ({\it i.e.} one for each rule), as well
-as the probability of each value. For each rule in the LPN, we also
-define a corresponding clause of the {\tt reduce} predicate to capture
-the actual relations between the left-hand and right-hand side of each
-rule. Sampling and parsing begins with a call to {\tt lpn(P-IN)},
-which chooses some rule for the left-hand side via {\tt msw(P,V)} and
-recursively reduces the input term using the chosen rule ({\tt
-reduce(P-IN,V)}).
+% First, we define the distribution over each non-terminal symbol, $n
+% \in N$, using the {\tt values} predicate. For each $n$ we define the
+% values the symbol can take on ({\it i.e.} one for each rule), as well
+% as the probability of each value. For each rule in the LPN, we also
+% define a corresponding clause of the {\tt reduce} predicate to capture
+% the actual relations between the left-hand and right-hand side of each
+% rule. Sampling and parsing begins with a call to {\tt lpn(P-IN)},
+% which chooses some rule for the left-hand side via {\tt msw(P,V)} and
+% recursively reduces the input term using the chosen rule ({\tt
+% reduce(P-IN,V)}).
 
@@
 and ones, and predicate $A_2$ is a category for the decade words.
+
+Our LPN does not learn to how to transition between the last word in a decade
+and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
+makes the intuitively reasonable generalization that ``seventy nine''
+should be followed by ``seventy ten.'' It will be interesting to see
+how common such mistakes are in children learning to count.
 
@@
 
-LPNs seems particularly well-suited to learning in domains where the
-number of concepts is infinite. In domains like these, the only way to
-understand each concept is to discover structure in the sentences that
-refer to these concepts. While number is a prime example, others
-include spatial relations, kinship and other social relations, and
-temporal relations. Future work should explore how LPNs might explain
-learning in these domains, as well as how more complex number
-relations might be learned.
+Though simple, the knowledge learned by LPNs in the simple number
+domain we explore in this paper goes beyond the expressive capacity of
+PCFGs and HMMs, and yet it avoids intractability by using a restricted
+formalism. The human ability to learn abstract concepts must also rely
+on such a compromise, and it remains to be seen to what extent LPNs
+can model this ability.
+
+This work leaves many open questions. What domains of concepts are
+amenable to being represented and learned by LPNs? Can multiple and
+varied systems of concepts be learned at once? Perhaps most
+importantly, are there generic architectures for LPNs that allow a
+larger number of predicates without an insurmountable blow-up in the
+number of rules? One idea which we are pursuing is using layered
+connectivity, much like neural networks, to limit the number of rules
+in which each predicate participates.
 
@@
   kwiatkowski2010inducing}. Semantic parsing, however, tends to frame
-this challenge slightly differently than we do here, namely, by asking
+this challenge slightly differently than we do here, by asking
 how utterances can be mapped to an explicit internal logical

### 2014-10-21 drafts/cfp/draft.tex
final draft from Josh

--- 2014-10-21 drafts/cfp/draft.tex
@@
 \usepackage{array}
-%\usepackage{multicol}
+\usepackage{authblk}
 
@@
 \title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
-\author{Eyal Dechter, Josh Rule, Josh Tenenbaum ADD AFFILIATION}
+\author{Eyal Dechter (edechter)}
+\author{Joshua Rule (rule)}
+\author{Joshua B. Tenenbaum (jbt)}
+\affil{Department of Brain and Cognitive Sciences, MIT (mit.edu)}
+
+\renewcommand\Authands{ and }
 \date{}
@@
 \maketitle
-\vspace{-2em}
+\vspace{-4em}
 
@@
   thirty-one" means if we can grasp the meanings of its component
-  number words. How do we make sense of this potential paradox? One
-  approach is to estimate the distribution from which sentences are
-  drawn, and, in doing so, infer the latent concepts and relationships
-  that best explain those sentences. We present early work on a
-  learning framework called Latent Predicate Networks (LPNs) which
-  learns concepts by inferring the parameters of probabilistic
+  number words. How do we begin to solve this problem? One approach is
+  to estimate the distribution from which sentences are drawn, and, in
+  doing so, infer the latent concepts and relationships that best
+  explain those sentences. We present early work on a learning
+  framework called Latent Predicate Networks (LPNs) which learns
+  concepts by inferring the parameters of probabilistic
   context-sensitive grammars over sentences.  We show that for a small
@@
   number words, we can use hierarchical Bayesian inference to learn
-  LPNs that can answer simple queries about previously unseen abstract
-  concepts within this domain. These generalizations demonstrate LPNs'
-  promise as a tool for learning and representing conceptual knowledge
-  in language.}
+  LPNs that can answer simple queries about previously unseen
+  relationships within this domain. These generalizations demonstrate
+  LPNs' promise as a tool for learning and representing conceptual
+  knowledge in language.}
 
@@
 and while it is almost certainly true that many concepts are learned
-via generalization from concrete examples, some concepts cannot
-possibly be learned this way.
+via generalization from concrete examples, some concepts cannot be
+learned this way.
 
@@
 five'') by noticing how number words are used in language, in counting
-sequences, in arithmetic excercises, etc. Other good examples of such
-abstract concepts are kinship and social relations (e.g. ``my father
+sequences, in arithmetic excercises, {\it etc}. Other good examples of such
+abstract concepts are kinship and social relations ({\it e.g.} ``my father
 in law's grandmother''), temporal relations (``the day after last
@@
 head is derived; otherwise, its head is derived if its body clauses
-are derived~\footnote{This description of the language of an RCG
+are derived.\footnote{This description of the language of an RCG
   technically only holds for \emph{non-combinatory} RCGs, in which the
@@
 are in child-directed speech, words for small numbers are far more
-common than words for larger ones(cite CHILDES). On the other hand,
-children learning to count to large numbers rehearse the sequence. To
-approximate this distribution of evidence, we drew these sample
-sentences from a sum of a geometric distribution with parameter $0.5$
-and a uniform distribution. These components were weighted $75\%$ and
-$25\%$, respectively. We drew $2000$ examples from this distribution,
-holding out the sentences in Table~\ref{tab:results} for
+common than words for larger ones \cite{macwhinney2000childes}. On the
+other hand, children learning to count to large numbers rehearse the
+sequence. To approximate this distribution of evidence, we drew these
+sample sentences from a sum of a geometric distribution with parameter
+$0.5$ and a uniform distribution. These components were weighted
+$75\%$ and $25\%$, respectively. We drew $2000$ examples from this
+distribution, holding out the sentences in Table~\ref{tab:results} for
 evaluation.
@@
 
-We evaluated the learned model by asking for Viterbi (i.e. maximum a
+We evaluated the learned model by asking for Viterbi ({\it i.e.} maximum a
 posteriori) completions of the last words of each held out test
@@
 sentences~\cite{berant2013semantic, liang2013learning,
-  kwiatkowski2010inducing}. Semantic parsing, however, tends to frame
-this challenge slightly differently than we do here, namely, by asking
-how utterances can be mapped to an explicit internal logical
-language. By contrast, we focus here on systems where meaning and
-structure seem inseparable. Understanding how these two approaches
-relate and inform one another is an interesting and open question.
+  kwiatkowski2010inducing, poon2009unsupervised}. Semantic parsing,
+however, tends to frame this challenge slightly differently than we do
+here, namely, by asking how utterances can be mapped to an explicit
+internal logical language. By contrast, we focus here on systems where
+meaning and structure seem inseparable. Understanding how these two
+approaches relate and inform one another is an interesting and open
+question.
 
@@
 
-\nocite{*}
-
 \end{document}

### 2014-10-21 drafts/cfp/draft.tex
Changes and fixed problem with bib file.

--- 2014-10-21 drafts/cfp/draft.tex
@@
 
-\setlength{\droptitle}{-4em}   % This is your set screw
+\setlength{\droptitle}{-6em}   % This is your set screw
 
@@
 \maketitle
-\vspace{-2em}
+\vspace{-6em}
 
@@
 		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
-		\caption{A schematic LPN.}
+		\caption{The architecture and rules of a schematic LPN.}
 		\label{fig:architecture}
@@
 
+\vspace{-1em}
 \section{Implementation \label{sec:implementation}}
@@
 \end{tabular}
-\caption{examples of generalization for held out sentences after learning over 4 latent predicates}
+\caption{Viterbi completions of held-out sentences to evaluate an LPN with 4 latent predicates in the number domain.}
 \label{tab:results}
@@
 
+\vspace{-1em}
 \section{Discussion}

### 2014-10-21 drafts/cfp/draft.tex

@@
 \maketitle
+
 \vspace{-6em}
@@
 
-Though simple, the knowledge learned by LPNs in the simple number
+Though simple, the knowledge learned by LPNs in the number
 domain we explore in this paper goes beyond the expressive capacity of
@@
 sentences~\cite{berant2013semantic, liang2013learning,
-  kwiatkowski2010inducing}. Semantic parsing, however, tends to frame
-this challenge slightly differently than we do here, by asking
-how utterances can be mapped to an explicit internal logical
-language. By contrast, we focus here on systems where meaning and
-structure seem inseparable. Understanding how these two approaches
-relate and inform one another is an interesting and open question.
+  kwiatkowski2010inducing, poon2009unsupervised}. Semantic parsing,
+however, tends to frame this challenge slightly differently than we do
+here, namely, by asking how utterances can be mapped to an explicit
+internal logical language. By contrast, we focus here on systems where
+meaning and structure seem inseparable. Understanding how these two
+approaches relate and inform one another is an interesting and open
+question.
 

### 2014-10-21 drafts/cfp/draft.tex
final from Josh

@@
 
-\setlength{\droptitle}{-6em}   % This is your set screw
+\setlength{\droptitle}{-4em}   % This is your set screw
 
@@
 \maketitle
-
-\vspace{-6em}
+\vspace{-4em}
 
@@
 		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
-		\caption{The architecture and rules of a schematic LPN.}
+		\caption{A schematic LPN.}
 		\label{fig:architecture}
@@
 
-\vspace{-1em}
 \section{Implementation \label{sec:implementation}}
@@
 trained an LPN with $4$ latent predicates on a set of sentences
-expressing successor and predessor relations in numbers between one
+expressing successor and predecessor relations in numbers between one
 and ninety-nine. The training set was the collection of sentences $$X
 = \{[\text{after}\, | \, \text{before}] \, \langle n \rangle \,
-\text{comes} \, \langle n+1 \rangle \,|\, n \in 1,\dots,99\}.$$ The
+\text{comes} \, \langle n+1 \rangle \,|\, n, n+1 \in 1,\dots,99\}.$$ The
 lexicon was the set of number word corresponding to $1$ through $19$,
@@
 \end{tabular}
-\caption{Viterbi completions of held-out sentences to evaluate an LPN with 4 latent predicates in the number domain.}
+\caption{examples of generalization for held out sentences after learning over 4 latent predicates}
 \label{tab:results}
@@
 
-\vspace{-1em}
 \section{Discussion}
 
-Though simple, the knowledge learned by LPNs in the number
+Though simple, the knowledge learned by LPNs in the simple number
 domain we explore in this paper goes beyond the expressive capacity of
@@
 sentences~\cite{berant2013semantic, liang2013learning,
-  kwiatkowski2010inducing, poon2009unsupervised}. Semantic parsing,
-however, tends to frame this challenge slightly differently than we do
-here, namely, by asking how utterances can be mapped to an explicit
-internal logical language. By contrast, we focus here on systems where
-meaning and structure seem inseparable. Understanding how these two
-approaches relate and inform one another is an interesting and open
-question.
+  kwiatkowski2010inducing}. Semantic parsing, however, tends to frame
+this challenge slightly differently than we do here, by asking how
+utterances can be mapped to an explicit internal logical language. By
+contrast, we focus here on systems where meaning and structure seem
+inseparable. Understanding how these two approaches relate and inform
+one another is an interesting and open question.
 

### 2014-10-21 drafts/cfp/draft.tex
Changes to draft.

--- 2014-10-21 drafts/cfp/draft.tex
@@
 Bayesian inference over stochastic grammars
-and stochastic logic programs (CITE MUGGLETON) has been an active area of
-research in recent decades~\cite{DBLP:conf/emnlp/LiangPJK07,
-  goldwater2006contextual, johnson2006adaptor, cussens2001parameter}.
+and stochastic logic programs has been an active area of
+research in recent decades~\cite{\cite{DBLP:journals/etai/Muggleton00}, cussens2001parameter, DBLP:conf/emnlp/LiangPJK07,
+  goldwater2006contextual, johnson2006adaptor}.
 Variational inference is a popular approach in this domain and the one
@@
 
-\section{Experiments}
+\section{Experiment}
 
@@
 \end{tabular}
-\caption{The grammar which produced Table \ref{tab:results}.}
+\caption{The 4 predicate LPN learned to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
 \label{tab:grammar}

### 2014-10-21 drafts/cfp/draft.tex
Final draft changes.

@@
 
-\setlength{\droptitle}{-6em}   % This is your set screw
+\setlength{\droptitle}{-5em}   % This is your set screw
 
@@
   seventy-three through linguistic means, and we can only know what
-  the sentence ``Seventy-three is more than twice as big as
+  the sentence ``seventy-three is more than twice as big as
   thirty-one" means if we can grasp the meanings of its component
@@
 
-Such abstract systems of concepts share many of the properties of
-language syntax: there is an unbounded set of concepts, they derive
+Such concepts share many of the properties of
+language syntax: they are unbounded in number, they derive
 their meanings via composition, and, although people only ever say,
@@
 probabilities $p_k$ such that for all predicates ${A \in N, \,
-  \sum_{k:head(C_k)=A} p_k = 1}$. An PRCG defines a distribution over
+  \sum_{k:head(C_k)=A} p_k = 1}$. A PRCG defines a distribution over
 strings $x$ by sampling from derivations of $S(x)$ according to the
@@
 
-An LPN is an PRCG with the following generic architecture: the
+An LPN is a PRCG with the following generic architecture: the
 \emph{axiom} of an LPN is a unary predicate $S$, and it has $K$
@@
 
-In words, the weights of clauses sharing head predicate $A_k$ are drawn from a Dirichlet distribution defined by $\vec{\alpha}_{A_k}$. Sentences $x_j$ are then drawn from the resultant PRCG.
+In words, the weights of clauses sharing head predicate $A_k$ are
+drawn from a Dirichlet distribution defined by
+$\vec{\alpha}_{A_k}$. Sentences $x_j$ are then drawn from the
+resulting PRCG.
 
@@
 and stochastic logic programs has been an active area of
-research in recent decades~\cite{\cite{DBLP:journals/etai/Muggleton00}, cussens2001parameter, DBLP:conf/emnlp/LiangPJK07,
+research in recent decades~\cite{DBLP:journals/etai/Muggleton00, cussens2001parameter, DBLP:conf/emnlp/LiangPJK07,
   goldwater2006contextual, johnson2006adaptor}.
@@
 we adopt here. As explained in the next section, we implemented
-inference by translating LPNs into PRISM programs and using the
+inference by translating LPNs into PRISM programs and using its
 built-in Variational Bayes Expectation-Maximization
@@
 is very similar to how PCFGs are encoded in
-PRISM~\cite{DBLP:conf/cl/2000}. A general probabilistic logic
-programming system based on PROLOG, PRISM proves built-in predicates
-for probabilistic execution and Bayesian inference over logic programs
-with stochastic choices. There are several
-restrictions placed on PRISM programs to maintain the validity of
-their probabilistic interpretation. Most importantly, derivations of a
-probabilistic predicate cannot contain cycles. Because we disallow
-$A_i(\epsilon,\epsilon)$ as a valid clause, every term in the body of a clause
-has shorter arguments than the head, giving acyclic derivations.
+PRISM~\cite{DBLP:conf/cl/2000}. See Figure~\ref{fig:copy} for an
+example. A general probabilistic logic programming system based on
+PROLOG, PRISM provides built-in predicates for probabilistic execution
+and Bayesian inference over logic programs with stochastic
+choices. There are several restrictions placed on PRISM programs to
+maintain the validity of their probabilistic interpretation. Most
+importantly, derivations of a probabilistic predicate cannot contain
+cycles. Because we disallow $A_i(\epsilon,\epsilon)$ as a valid
+clause, every term in the body of a clause has shorter arguments than
+the head, giving acyclic and finite derivations.
 
@@
 = \{[\text{after}\, | \, \text{before}] \, \langle n \rangle \,
-\text{comes} \, \langle n+1 \rangle \,|\, n, n+1 \in 1,\dots,99\}.$$ The
+\text{comes} \, \langle n+1 \rangle \,|\, n \in 1,\dots,99\},$$ where
+$\langle n \rangle$ is the number word corresponding to $n$. The
 lexicon was the set of number word corresponding to $1$ through $19$,
-the decades $20, \dots, 30$ and the words ``before'' and ``after.'' It
-is not difficult to manually work out an LPN that describes this
-limited domain of sentences; see Figure~\ref{fig:parseexample}, for a
-possible LPN derivation of an example sentence.
+the decades $20, \dots, 30$, the empty string, and the words
+``before'' and ``after.'' It is not difficult to manually work out an
+LPN that describes this limited domain of sentences; see
+Figure~\ref{fig:parseexample}, for a possible LPN derivation of an
+example sentence.
 
@@
 results, though further investigation will be necessary to see how
-robust our the algorithm is to local maximuma for LPNs.
+robust the algorithm is to local maximuma when fitting LPNs.
 
@@
 
-To inspect visually the learned grammar, we thresholded rules
-according to the expected number of times they were used in parsing
-the training dataset. Table~\ref{tab:grammar} shows all rules with
-expected count above $1e-6$. This reduces from $2669$ to $52$ the
-number of significant rules. On inspection, predicate $A_2$ forms
-``before'' sentences, predicate $A_4$ forms ``after'' sentences,
-predicate $A_3$ is successorship recursively defined over the decades
-and ones, and predicate $A_2$ is a category for the decade words.
-
-Our LPN does not learn to how to transition between the last word in a decade
-and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
-makes the intuitively reasonable generalization that ``seventy nine''
-should be followed by ``seventy ten.'' It will be interesting to see
-how common such mistakes are in children learning to count.
-
-\begin{table}[t]
-\begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
-  Question & $K=4$ \\ \hline
-  after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
-  after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
-  after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
-  after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
-  after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
-  after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
-  after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
-  after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
-  after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
-  after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
-  after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
-  after seventy nine comes \underline{\hspace{1cm}}? & seventy ten $\times$ \\
-  after ninety five comes \underline{\hspace{1cm}}? & ninety six \checkmark \\
-  before twenty three comes \underline{\hspace{1cm}}? & twenty two \checkmark \\
-  before thirty comes \underline{\hspace{1cm}}? & thirty eighty $\times$ \\
-  before thirty eight comes \underline{\hspace{1cm}}? & thirty seven \checkmark \\
-  before forty one comes \underline{\hspace{1cm}}? & forty \checkmark \\
-  before fifty three comes \underline{\hspace{1cm}}? & fifty two \checkmark \\
-  before sixty eight comes \underline{\hspace{1cm}}? & sixty seven \checkmark \\
-  before seventy two comes \underline{\hspace{1cm}}? & seventy one \checkmark \\
-  before seventy three comes \underline{\hspace{1cm}}? & seventy two \checkmark \\
-  before eighty five comes \underline{\hspace{1cm}}? & eighty four \checkmark \\
-  before ninety two comes \underline{\hspace{1cm}}? & ninety one \checkmark \\
-  before ninety three comes \underline{\hspace{1cm}}? & ninety two \checkmark \\
-  before ninety five comes \underline{\hspace{1cm}}? &ninety four \checkmark \\
-\end{tabular}
-\caption{Viterbi completions of held-out sentences to evaluate an LPN with 4 latent predicates in the number domain.}
-\label{tab:results}
-\end{table}
-
-\begin{figure}[t]
-		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
-		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
-                \label{fig:parseexample}
-\end{figure}
-
-\begin{table}[t]
+\begin{table}[h]
 \begin{tabular}{>{\tiny} l >{\tiny} l}
@@
 \end{tabular}
-\caption{The 4 predicate LPN learned to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
+\caption{The 4 predicate LPN trained to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
 \label{tab:grammar}
 \end{table}
+
+To inspect visually the learned grammar, we thresholded rules
+according to the expected number of times they were used in parsing
+the training dataset. Table~\ref{tab:grammar} shows all rules with
+expected count above $1e-6$. This reduces from $2669$ to $52$ the
+number of significant rules. On inspection, predicate $A_2$ forms
+``before'' sentences, predicate $A_4$ forms ``after'' sentences,
+predicate $A_3$ is successorship recursively defined over the decades
+and ones, and predicate $A_2$ is a category for the decade words.
+
+Our LPN does not learn to how to transition between the last word in a decade
+and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
+makes the intuitively reasonable generalization that ``seventy nine''
+should be followed by ``seventy ten.''
+
+\begin{table}[t]
+\begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
+  Question & $K=4$ \\ \hline
+  after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
+  after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
+  after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
+  after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
+  after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
+  after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
+  after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
+  after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
+  after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
+  after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
+  after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
+  after seventy nine comes \underline{\hspace{1cm}}? & seventy ten $\times$ \\
+  after ninety five comes \underline{\hspace{1cm}}? & ninety six \checkmark \\
+  before twenty three comes \underline{\hspace{1cm}}? & twenty two \checkmark \\
+  before thirty comes \underline{\hspace{1cm}}? & thirty eighty $\times$ \\
+  before thirty eight comes \underline{\hspace{1cm}}? & thirty seven \checkmark \\
+  before forty one comes \underline{\hspace{1cm}}? & forty \checkmark \\
+  before fifty three comes \underline{\hspace{1cm}}? & fifty two \checkmark \\
+  before sixty eight comes \underline{\hspace{1cm}}? & sixty seven \checkmark \\
+  before seventy two comes \underline{\hspace{1cm}}? & seventy one \checkmark \\
+  before seventy three comes \underline{\hspace{1cm}}? & seventy two \checkmark \\
+  before eighty five comes \underline{\hspace{1cm}}? & eighty four \checkmark \\
+  before ninety two comes \underline{\hspace{1cm}}? & ninety one \checkmark \\
+  before ninety three comes \underline{\hspace{1cm}}? & ninety two \checkmark \\
+  before ninety five comes \underline{\hspace{1cm}}? &ninety four \checkmark \\
+\end{tabular}
+\caption{Viterbi completions of held-out sentences to evaluate an LPN with 4 latent predicates in the number domain.}
+\label{tab:results}
+\end{table}
+
+\begin{figure}[t]
+		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
+		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+                \label{fig:parseexample}
+\end{figure}
 
@@
 learning meaning and structure from
-sentences~\cite{berant2013semantic, liang2013learning,
-  kwiatkowski2010inducing, poon2009unsupervised}. Semantic parsing,
+sentences~\cite{berant2013semantic, liang2013learning, kwiatkowski2010inducing, poon2009unsupervised}. Semantic parsing,
 however, tends to frame this challenge slightly differently than we do

### 2014-10-23 drafts/cfp/draft.tex
Added python script ipa.py to translate english to ipa automatically.

@@
 five'') by noticing how number words are used in language, in counting
-sequences, in arithmetic excercises, {\it etc}. Other good examples of such
+sequences, in arithmetic exercises, {\it etc}. Other good examples of such
 abstract concepts are kinship and social relations ({\it e.g.} ``my father
@@
 First, whereas context-free grammars are suitable for describing large
-swaths of language syntax, the grammars of concepts are not
-context-free. To address this, we use Range Concatentation Grammars, a
+swathes of language syntax, the grammars of concepts are not
+context-free. To address this, we use Range Concatenation Grammars, a
 context-sensitive grammar formalism -- one of several developed within
@@
 
-\subsection{Probabilistic Range Concatentation Grammars}
+\subsection{Probabilistic Range Concatenation Grammars}
 Range Concatenation Grammars (RCGs) are a class of string grammars
@@
 results, though further investigation will be necessary to see how
-robust the algorithm is to local maximuma when fitting LPNs.
+robust the algorithm is to local maxima when fitting LPNs.
 

### 2014-11-05 prism/assoc_list.pl
Added files to the prism subdirectory.

@@ -1,0 +1,16 @@
+%% association list
+%% an Association List is a list of key value pairs K\V. 
+
+get_assoc([K\V|_], K, V) :- !.
+get_assoc([_|Rest], K, V) :- get_assoc(Rest, K, V). 
+
+get_assocs(Assoc, Ks, Vs) :- 
+    Vs @= [V: K in Ks, [V], get_assoc(Assoc, K, V)].
+           
+
+set_assoc([], K, V, [K\V]).
+set_assoc([K\_|AssocIn], K, V, [K\V|AssocIn]) :- !.
+set_assoc([P|AssocIn], K, V, [P|AssocOut]) :- set_assoc(AssocIn, K, V, AssocOut).
+
+
+

### 2014-11-05 prism/onlineVBEM.pl

@@ -1,0 +1,164 @@
+
+:- cl('assoc_list.pl').
+
+set_ovbem_rec(Flag, Val) :- 
+    retractall(ovbem_rec(Flag, _)),
+    assert(ovbem_rec(Flag, Val)).
+
+get_ovbem_rec(Flag, Val) :- ovbem_rec(Flag, Val).
+
+ovbem_rec_assoc(Assoc) :- 
+    findall(K\V, ovbem_rec(K, V), Assoc).
+
+:- dynamic ovbem_rec/2. 
+ovbem_rec(niter, 100).
+ovbem_rec(batch_size, 10).
+ovbem_rec(data_set, []).
+ovbem_rec(held_out, []).
+ovbem_rec(iter_between, 10).
+ovbem_rec(held_out_data_expected_log_like, []).
+ovbem_rec(logfile, '.ovbem_log').
+ovbem_rec(outfile, '.ovbem_out').
+
+save_ovbem_rec :- 
+    ovbem_out(File), 
+    ovbem_rec_assoc(Assoc), 
+    save_clauses(File, [Assoc], []). 
+
+run_ovbem :- 
+    ovbem_rec_assoc(Assoc), 
+    get_assocs(Assoc, 
+               [niter, batch_size, data_set, held_out, iter_between], 
+               [NIter, BatchSize, DataSet, HeldOut, IterBetween]), 
+    ovbem_held_out(NIter, 1, BatchSize, DataSet, HeldOut, IterBetween).
+    
+ovbem_held_out(NIter, Iter, 
+               BatchSize, DataSet, HeldOut, IterBetween) :- 
+    (((NIter-Iter) > IterBetween) -> 
+        IterBetween1 = IterBetween;
+        IterBetween1 = NIter-Iter),
+    ovbem(IterBetween1, BatchSize, DataSet, Iter), 
+    Iter1 is Iter + IterBetween1,
+    free_energy(HeldOut, _, _, _, ExpectedLogLike),
+    get_ovbem_rec(held_out_data_expected_log_like, LLs), 
+    append(LLs, [ExpectedLogLike], LLs1),
+    set_ovbem_rec(held_out_data_expected_log_like, LLs1),
+    save_ovbem_rec,
+    write('NIter: '), write(NIter), nl,
+    write('Iter1: '), write(Iter1), nl,
+    (Iter < NIter -> 
[…]

### 2014-11-05 prism/posterior_loglikelihood.pl

@@ -1,0 +1,13 @@
+%% Author: Eyal Dechter
+%%
+%% Calculate the posterior dirichlet variational lowerbound of a test dataset.
+
+:- cl('util.pl').
+
+
+
+
+
+
+
+

### 2014-11-05 prism/testOvbem.pl

@@ -1,0 +1,25 @@
+:- ['train.psm'].
+:- prism([load], 'out.sys.psm').
+:- ['onlineVBEM.pl'].
+:- ['../experiments/prism/verbs0.psm'].
+
+:- set_training_flags.
+
+:- set_ovbem_rec(niter, 30), 
+   set_ovbem_rec(batch_size, 1), 
+   set_ovbem_rec(iter_between, 5).
+   
+:- findall(prove('S_1'-[X]), example(X), DataSet), 
+    set_ovbem_rec(data_set, DataSet).
+
+:- findall(prove('S_1'-[X]), example(X), DataSet), 
+   take(3, DataSet, HeldOut), 
+   write(HeldOut), nl, 
+    set_ovbem_rec(held_out, HeldOut). 
+
+main :- 
+    viterbif(prove('S_1'-[[w,a,k]])),
+    run_ovbem.
+
+    
+    

### 2014-11-05 prism/util.pl

@@ -1,0 +1,97 @@
+%% util.py
+%% Author: Eyal Dechter 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+gijoe_root(DIR) :- 
+    expand_environment('$GIJOE_ROOT', DIR).
+gijoe_bin(DIR) :- 
+    gijoe_root(B), 
+    atom_concat(B, '/bin', DIR).
+prismify(SysFile, PsmFile) :- 
+    gijoe_bin(BINDIR), 
+    write(BINDIR), nl,
+    formatAtom("~a/prismify ~a ~a", [BINDIR, SysFile, PsmFile], A),
+    format("Prismifying LPN ~a ... ~1n", [SysFile]),
+    system(A),
+    format("Done.", []).
+
+% load LPN .sys file at path. By default, load precompiled code found at
+% <path>.out. 
+load_lpn(Path) :- 
+    load_lpn([load], Path).
+
+% load LPN path with supplied options. Options are the same as the
+% those to prism(options, path).
+load_lpn(Options, Path) :-
+    format("Loading LPN in file ~a...~1n", [Path]),
+    format("Options: ~w~1n", [Options]),
+    prism(Options, Path),
+    format("Done").
+
+
+:- dynamic lpn_value_pred/2. 
+lpn_value_pred(Switch, _) :- throw(lpn_value_pred_not_found(Switch)).
+assert_lpn_value_pred(Switch, Pred) :- 
+    asserta(lpn_value_pred(Switch, Pred)).
+
+%% lpn_set_sw_... 
+%% change the values, probs, and alpha values of lpn switch
+lpn_set_sw_v(Switch, Vs) :- 
+    lpn_value_pred(Switch, P), 
+    RetractTerm =.. [P, K], 
+    AssertTerm =.. [P, Vs], 
+    retract(RetractTerm), 
+    assert(AssertTerm).
+
+lpn_set_sw_a(Switch, As) :- 
+    set_sw_a(Switch, As).
+
+lpn_set_sw_vp(Switch, Vs, Ps) :- 
+    lpn_set_sw_v(Switch, Vs), 
+    set_sw(Switch, Ps).
+
+lpn_set_sw_va(Switch, Vs, As) :- 
+    lpn_set_sw_v(Switch, Vs), 
+    set_sw_a(Switch, As).
+
+lpn_set_sw_vpa(Switch, Vs, Ps, As) :- 
+    lpn_set_sw_v(Switch, Vs), 
[…]

### 2014-11-05 prism/vis.pl

@@ -1,0 +1,138 @@
+%% %% Print a Parse Tree %% %%
+
+srsTree(X,[Olhs,Orule,Orhs]) :-
+    select(msw(P,V),X,X2),
+    select(srs(P-IN),X2,X3),
+    delete(X3,reduce(_,_),X4),
+    maplist(srsTree,X4,Orhs),
+    prismToLPN(P,V,Orule),
+    predicateToString(P-IN,Olhs).
+
+%% hack! This currently only works up to 9-ary predicates
+predicateToString(P-IN,S) :-
+    sub_atom(P,0,_,2,P2),
+    atom_lists_to_atoms(IN,A2),
+    atom_concat(P2,'(',S1),
+    atom_concat(A2,')',S2),
+    atom_concat(S1,S2,S).
+
+atom_lists_to_atoms([],'').
+atom_lists_to_atoms([X],O) :-
+    recursive_atom_concat_space(X,O).
+
+atom_lists_to_atoms([X|XS],O) :-
+    length(XS,XL), XL > 0,
+    atom_lists_to_atoms(XS,OS),
+    recursive_atom_concat_space(X,OS2),
+    atom_concat(OS2,', ',O1),
+    atom_concat(O1,OS,O).
+
+recursive_atom_concat([],'').
+recursive_atom_concat([X],X).
+recursive_atom_concat([X|XS],O) :-
+    length(XS,XL), XL > 0,
+    recursive_atom_concat(XS,OS),
+    atom_concat(X,OS,O).
+
+recursive_atom_concat_space([],'').
+recursive_atom_concat_space([X],X).
+recursive_atom_concat_space([X|XS],O) :-
+    length(XS,XL),
+    XL > 0,
+    recursive_atom_concat_space(XS,OS),
+    atom_concat(X,' ',O1),
+    atom_concat(O1,OS,O).
+
+formatSRSTree(X) :- format("~n",[]), formatSRSTree(X,'').
+formatSRSTree([LHS,RULE,RHS],SEP) :-
+    format("~a~a~n~a~a~n",[SEP,LHS,SEP,RULE]),
+    atom_concat(SEP,'    ',NEWSEP),
+    findall(X,(member(X,RHS),formatSRSTree(X,NEWSEP)),_).
+
+%% %% Print The Grammar %% %%
+
[…]

### 2014-11-05 src/GIJoe/SRS/Prismify.hs
Minor changes.

@@ -2,6 +2,7 @@ module GIJoe.SRS.Prismify where
 
 import Prelude hiding (head)
 
+
 import Data.Function (on)
 import Data.List hiding (head)
 import Data.Text (unpack)
@@ -110,3 +111,4 @@ prismify inpath outpath = do
 ------
 showQuoted x = "'" ++ show x ++ "'"
 quoted x = "'" ++ x ++ "'"
+

### 2014-11-05 prolog/onlineVBEM.pl
Removed files from prolog subdirectory (moving things to prism subdirectory seems more apt).

@@ -1,105 +1,0 @@
-ovbem(NIter, BatchSize, DataSet) :-
-    % get number of latent parameters
-    findall(Info, get_sw_a(Info), PInit),
-    findall(X, member(switch(_,_,_,X),PInit), Gs),
-    flatten(Gs,G),
-    length(G,N),
-    % initialize rate
-    UNIF is 1/N,
-    GBar @= [UNIF : _X in 1..N],
-    HBar = GBar,
-    % run the online VB-EM
-    ovbem(NIter, BatchSize, DataSet, rate(2,GBar,HBar,0.01), 1). 
-
-ovbem(NIter, _BatchSize, _DataSet, _RateInit, Iter) :- Iter > NIter, !.
-ovbem(NIter, BatchSize, DataSet, RateInit, Iter) :- 
-    ovbem1(DataSet, BatchSize, RateInit, RateFinal), 
-    NextIter is Iter+1, 
-    ovbem(NIter, BatchSize, DataSet, RateFinal, NextIter).
-
-ovbem1(DataSet, BatchSize, RateInit, RateFinal) :- 
-    length(DataSet, N),
-    % get current alphas
-    findall(Info, get_sw_a(Info), ParamsInit),
-    % sample a data point from the dataset
-    random_multiselect(DataSet, BatchSize, Batch),
-    M is round(N/BatchSize),
-    BatchOut @= [D1: D in Batch, [D1], multiply_datum_count(D, M, D1)],
-    % run vbem
-    learn(BatchOut),
-    % get output alphas
-    findall(Info, get_sw_a(Info), ParamsFinal),
-    % update learning rate
-    update_learning_rate(ParamsInit,ParamsFinal,RateInit,RateFinal),
-    RateFinal = rate(_,_,_,StepSize),
-    print_rate(RateFinal),
-    % update parameters towards ParamsFinal
-    foreach(switch(Sw, Status, Vals, As0) in ParamsInit, [As0, As1, AsOut],
-            (member(switch(Sw, Status, Vals, As1), ParamsFinal), !,
-             update_vb_params(StepSize, As0, As1, AsOut),
-             set_sw_a(Sw, AsOut))).
-
-print_rate(rate(T,G,H,R)) :-
-    sumlist(G,GSum),
[…]

### 2014-11-05 prolog/vis.pl

@@ -1,140 +1,0 @@
-%% %% Print a Parse Tree %% %%
-
-srsTree(X,[Olhs,Orule,Orhs]) :-
-    select(msw(P,V),X,X2),
-    select(srs(P-IN),X2,X3),
-    delete(X3,reduce(_,_),X4),
-    maplist(srsTree,X4,Orhs),
-    prismToSRS(P,V,Orule),
-    predicateToString(P-IN,Olhs).
-
-%% hack! This currently only works up to 9-ary predicates
-predicateToString(P-IN,S) :-
-    sub_atom(P,0,_,2,P2),
-    atom_lists_to_atoms(IN,A2),
-    atom_concat(P2,'(',S1),
-    atom_concat(A2,')',S2),
-    atom_concat(S1,S2,S).
-
-atom_lists_to_atoms([],'').
-atom_lists_to_atoms([X],O) :-
-    recursive_atom_concat_space(X,O).
-
-atom_lists_to_atoms([X|XS],O) :-
-    length(XS,XL), XL > 0,
-    atom_lists_to_atoms(XS,OS),
-    recursive_atom_concat_space(X,OS2),
-    atom_concat(OS2,', ',O1),
-    atom_concat(O1,OS,O).
-
-recursive_atom_concat([],'').
-recursive_atom_concat([X],X).
-recursive_atom_concat([X|XS],O) :-
-    length(XS,XL), XL > 0,
-    recursive_atom_concat(XS,OS),
-    atom_concat(X,OS,O).
-
-recursive_atom_concat_space([],'').
-recursive_atom_concat_space([X],X).
-recursive_atom_concat_space([X|XS],O) :-
-    length(XS,XL),
-    XL > 0,
-    recursive_atom_concat_space(XS,OS),
-    atom_concat(X,' ',O1),
-    atom_concat(O1,OS,O).
-
-formatSRSTree(X) :- format("~n",[]), formatSRSTree(X,'').
-formatSRSTree([LHS,RULE,RHS],SEP) :-
-    format("~a~a~n~a~a~n",[SEP,LHS,SEP,RULE]),
-    atom_concat(SEP,'    ',NEWSEP),
-    findall(X,(member(X,RHS),formatSRSTree(X,NEWSEP)),_).
-
-%% %% Print The Grammar %% %%
-
[…]

### 2014-11-05 drafts/nips_workshop/nips_workshop.tex

--- 2014-10-23 drafts/cfp/draft.tex
@@
-\documentclass[10pt, twocolumn]{article}
-
-\usepackage{fullpage}
+\documentclass{article} % For LaTeX2e
+\usepackage{nips14submit_e,times}
+\usepackage{hyperref}
+\usepackage{url}
+\usepackage{graphicx}
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{listings}
 \usepackage{anyfontsize}
 \usepackage{subcaption}
-\usepackage{graphicx}
 \usepackage{titling}
@@
 
-\setlength{\droptitle}{-5em}   % This is your set screw
+\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
 
-\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
-\author{Eyal Dechter (edechter)}
-\author{Joshua Rule (rule)}
-\author{Joshua B. Tenenbaum (jbt)}
-\affil{Department of Brain and Cognitive Sciences, MIT (mit.edu)}
+\author{Eyal Dechter\thanks{{\tt \{edechter, rule, jbt\} @mit.edu}}}
+\author{Joshua Rule}
+\author{Joshua B. Tenenbaum}
+\affil{Department of Brain and Cognitive Sciences, MIT}
 
-\renewcommand\Authands{ and }
-\date{}
+\newcommand{\fix}{\marginpar{FIX}}
+\newcommand{\new}{\marginpar{NEW}}
+
+\nipsfinalcopy % Uncomment for camera-ready version
 
 \begin{document}
-\vspace{-10cm}
+
 \maketitle
 
-\vspace{-6em}
-
-{\small For humans, learning abstract concepts and learning language
-  go hand in hand: we acquire abstract knowledge primarily through
-  linguistic experience, and acquiring abstract concepts is a crucial
-  step in learning the meanings of linguistic expressions. Number
-  knowledge is a case in point: we largely acquire concepts such as
-  seventy-three through linguistic means, and we can only know what
-  the sentence ``seventy-three is more than twice as big as
-  thirty-one" means if we can grasp the meanings of its component
-  number words. How do we begin to solve this problem? One approach is
-  to estimate the distribution from which sentences are drawn, and, in
-  doing so, infer the latent concepts and relationships that best
-  explain those sentences. We present early work on a learning
-  framework called Latent Predicate Networks (LPNs) which learns
-  concepts by inferring the parameters of probabilistic
-  context-sensitive grammars over sentences.  We show that for a small
-  fragment of sentences expressing relationships between English
-  number words, we can use hierarchical Bayesian inference to learn
-  LPNs that can answer simple queries about previously unseen
-  relationships within this domain. These generalizations demonstrate
-  LPNs' promise as a tool for learning and representing conceptual
-  knowledge in language.}
+\begin{abstract}
+  For humans, learning abstract concepts and learning language go hand
+  in hand: we acquire abstract knowledge primarily through linguistic
+  experience, and acquiring abstract concepts is a crucial step in
+  learning the meanings of linguistic expressions. Number knowledge is
+  a case in point: we largely acquire concepts such as seventy-three
+  through linguistic means, and we can only know what the sentence
+  ``seventy-three is more than twice as big as thirty-one" means if we
+  can grasp the meanings of its component number words. How do we
+  begin to solve this problem? One approach is to estimate the
+  distribution from which sentences are drawn, and, in doing so, infer
+  the latent concepts and relationships that best explain those
+  sentences. We present early work on a learning framework called
+  Latent Predicate Networks (LPNs) which learns concepts by inferring
+  the parameters of probabilistic context-sensitive grammars over
+  sentences.  We show that for a small fragment of sentences
+  expressing relationships between English number words, we can use
+  hierarchical Bayesian inference to learn LPNs that can answer simple
+  queries about previously unseen relationships within this
+  domain. These generalizations demonstrate LPNs' promise as a tool
+  for learning and representing conceptual knowledge in language.
+\end{abstract}
 
@@
 
-% people learn about many
-% abstract concepts primarily through language and, conversely, these
-% abstract concepts are critical to understanding language. Consider a
-% child learning the meaning of the number words: the meaning of the
-% ``seventy five'' is learned not by seeing collections of seventy five
-% things but rather by understanding how ``seventy five'' is used in the
-% various sentences, counting sequences, arithmetic excercises, etc.
-
-% As Claude Shannon demonstrated with his
-% guessing game experiment in 1950~\cite{Shannon1950}, an adult English
-% speaker's understanding of her language and her world is such that the
-% entropy of a character in English text is slightly more than one
-% bit. Plausibly, any machine learning system able to predict natural
-% language text with such accuracy will have learned much about the
-% concepts expressed in that text. This intuition motivates us to invert
-% Shannon's experiment and learn linguistic concepts by learning a
-% probabilistic generative model for natural language.
-
-% Probabilistic Context Free Grammars (PCFGs) are popular and much
-% studied generative models for natural language
-% syntax~\cite{Someone}. While debate continues about their ability to
-% express the syntax of natural language, it is agreed that they cannot
-% express dependencies between words that depend upon each other for
-% their meaning. Consider an example from the domain of sentences
-% addressed in this paper: ``after twenty eight comes twenty nine.''
-% The second ``twenty'' depends on the first ``twenty'' and the
-% ``eight'', and the ``nine'' depends on the ``eight''. This pattern is
-% a clear example of the \emph{cross-serial dependencies} which
-% context-free grammars cannot support.
-
-% Our approach, then, will be to explore the use of probabilistic
-% context \emph{sensitive} grammars for modeling natural language,
-% extending the tools of hierarchical grammar induction~\cite{SOMEONE} from
-% context-free grammars to ones that can express not only syntactic
-% relationships but semantic ones as well. In the place of latent
-% syntactic categories, we will refer to latent predicates -- relations
-% between sentence components -- that, when learned, should correspond
-% to the concepts that underlie sentence semantics.
-
-% To explore this approach, we introduce a probabilisitic model called a
-% Latent Predicate Network (LPN), and describe an inference procedure to
-% learn its parameters using the probabilistic logic programming
-% language PRISM. Finally, we study its performance when trained on
-% sentences describing the relationships between numbers.
-
 \section{Latent Predicate Networks}
@@
 \begin{figure}[t]
+  \begin{subfigure}[b]{0.5\linewidth}
     \includegraphics[width=\linewidth]{lpn/lpn.pdf}
@@
     \label{fig:architecture}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\linewidth]{parseTree/parse.pdf}
+    \caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+    \label{fig:parseexample}
+  \end{subfigure}
 \end{figure}
@@
 is the \emph{head predicate}.
-
-%% \textbf{Example:} The following is an RCG for the 3-copy language
-%% $\{www | w \in \{a, b\}^*\}$:
-%% \begin{alignat*}{3}
-%% &S(XYZ) &&\rightarrow A(X, Y, Z)\\
-%% &A(aX, aY, aZ) &&\rightarrow A(X, Y, Z)\\
-%% &A(bX, bY, bZ) &&\rightarrow A(X, Y, Z)\\
-%% &A(\epsilon, \epsilon, \epsilon) &&\rightarrow 
-%% \end{alignat*}
 
@@
 
-% Now we can describe the model we introduce here, the Latent Concept
-% Grammar (LCG). We introduce the LCG as a machine learning model to
-% learn a distribution over instances of relations over strings. Unlike
-% a PRCG, which places a distribution over terms RCGs consist of a set
-% of rewrite rules, called clauses, each of which takes a termwhich
-% operate on a set of terms (predicates applied to arguments) to produce
-% new terms. Each rewrite rule is of the form:
-
-% The vast majority of the models of concept acquisition that have been
-% proposed in the artificial intelligence, machine learning and
-% cognitive science literatures cannot represent the conceptual systems
-% that even young children learn. Take, for example, the school-age
-% child's knowledge of number. This child knows that there is no largest
-% number; that ``one million forty three'' and ``8439''denote numbers
-% but that ``one million forty billion'' and ``a82bd'' do not; that no
-% two cardinalities refer to the same quantity but that some number
-% words refer to the same cardinalities (e.g. ``two thousand'' and
-% ``twenty-one hundred''). Moreover, the child knows how to guage the
-% approximate number of elements in a collection and how to count this
-% number exactly. In short, the child's knowledge of number consists of
-% a complex system of concepts over an unbounded set of entities
-% (i.e. the numbers).
-
-% Children are not born with this kind of knowledge, and, since number
-% is a relatively recent cultural invention, nor is it likely that
-% children possess innate domain specific background knowledge that
-% makes it easier to acquire. Children only ever see a tiny fraction of
-% all the numbers they can reason about and this evidence is biased,
-% noisy, and can be unreliable.
-
-% A fundamental challenge for cognitive science -- and for any field
-% interested in understanding human-like concept acquisition -- is to
-% provide a theory of how such complex systems of concepts can be
-% learned from sparse and noisy data.
-
-% People are able to learn about complex systems of related
-% concepts from small amounts of noisy data. A compelling approach to
-% understanding such learning is to frame it as Bayesian inference over
-% probabilistic knowledge bases. Advances in context free grammar
-% induction in the past couple of decades can be seen as an instance of
-% this approach, in which
-
-% Our goal is to understand how children can obtain semantic knowledge
-% about a system of concepts from the structure of relavent perceptual
-% stimuli. How does a child learn that prefixing ``great-'' corresponds
-% to adding an additional generation in kinship relationships? How does
-% a child learn that ``just before'' is after ``a while before'' which
-% is before ``a while after''? How does a child learn that if
-% Figure~\ref{pairOfApples} is ``a pair of apples'' then
-% Figure~\ref{singleApple} is a single apple?
-
-% Our proposal is that some interesting fragments of human conceptual
-% knowledge can be represented as structured relations between
-% linguistic tokens and other perceptual inputs; that the rules that
-% define these relations implicitly define the concepts that are
-% learned; and that these rules can be learned using the tools of
-% hierarchical Bayesian inference.
-
-% For example, the concept TWENTY SEVEN is defined as the collection of
-% representations ``27'', ``twenty seven'', ``3 times 9,'' along with a
-% system of symbolic rules that translate among these representations
-% and facilitate answered various queries. According to our hypothesis,
-% the difference between TWENTY EIGHT and TWENTY SEVEN is not that they
-% are represented in the mind by distinct mental symbols but rather that
-% the rules that operate on their various representations treat them in
-% systematically different ways.
-
-% This view implies that an attempt to construct a computational account
-% of concept learning in humans needs to ground out that learning in the
-% relavent perceptual stimuli.
-
-% Contribution:
-% \begin{itemize}
-% \item We present Generate Rewrite Systems (GRS) -- a formalism that borrows
-%   heavily from Linear Context-Free Rewriting Systems (CITE)
-% \item We show how GRS, though restricted in representational capactiy,
-%   can capture substantial cognitively plausible first-order conceptual knowledge 
-% \item We propose a heirarchical Bayesian model for GRS that provides a coherent framework in which a GRS can be learned given evidence
-% \item We show how GRS can be transformed into PRISM programs and learned within that probabilstic logic programming system
-% \item We provide computational experiments that provide evidence that
-%   our model can provide an account of some kinds of complex conceptual
-%   knowledge can be inferred from noisy and sparse data
-% \end{itemize}
-
-% \section{The Model}
-
-% Suppose we are given some positive instances $y_1, \dots,
-% y_M$ of relations $R_1, \dots, R_N$. Our model assumes that each
-% positive instance $y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent
-% stochastic grammar (LSG), $\mathcal{K}$. $\mathcal{K}$ consists of
-% some fixed number $K$ of clauses (i.e. rules) $C_1, \dots, C_K$ over
-% the $N$ observable relations plus a fixed number of $H$ hidden
-% relations $T_1, \dots, T_H$. The \emph{architecture} of the model is
-% the exact specification of the parameters $K$ and $H$ and the clauses
-% $\mathcal{C}$. To each clause $C_i$ is associated a positive real
-% weights $\theta_i$.
-
-% The clauses of $\mathcal{K}$ are Horn clauses in a first order logic
-% with string concatenation. We will use PROLOG notation, in which lower
-% case letters are predicates and constants and uppercase letters are
-% variables. If two symbols are adjacent it means that the corresponding
-% strings are concatenated. For example, the following would is a valid
-% clause: $a(XY, XV) \leftarrow b(X, UV), a(X, X)$. In this rule $a$ and
-% $b$ are relations with arity two.
-
-% Let $C_1, \dots, C_{N(R)} = C(R)$ be the set of clauses whose head is
-% relation $R$ and let $\Theta_R$ be the set of corresponding
-% weights. Let $p_R(C_n) \sim \text{Discrete}(\Theta_R)$; that is, $p_R$
-% is a distribution over the clauses headed by $R$, such that each
-% clauses is sampled in proportion to its weight parameter.
-
-% A positive instance of this relation is sampled from $\mathcal{K}$ by
-% arbitrarily (i.e. uniformly at random) choosing a relation $R$ and
-% instantiating that relation by sampling a clause from $p_R$. Recursing
-% on this procedure for every atom in the body of this clause binds the
-% variables in the clause's head.
-
-% Given this assumed generative process for the evidence, learning a
-% program involves inferring the parameter $\theta$ given a prior
-% probability distribution over their values. We place an independent
-% Dirichlet distribution over the parameters of every relation. This is
-% a natural choice that enables us to promote sparse programs, i.e.,
-% programs in which most of the mass is placed on a small number of
-% predicates and rules.
-
 \subsection{Learning Model}
@@
   \vec{w}_{A_k} &\sim Dir(\vec{\alpha}_{A_k})\\
-  x_j &\underset{iid}{\sim} p_{\text{\textsc{prcg}}}(S(x_j)\,|\,\{w_{A_k}\})
+  x_j &\underset{iid}{\sim} p_{\text{\textsc{prcg}}}(S(x_j)\,|\,\{w_{A_k}\})\\
+  p(\{\vec{w}_{A_k}\}\,|\, \vec{x}, \{\vec{\alpha}_{A_k}\}) &\propto
+  \prod_j p_{\text{\textsc{prcg}}}(x_j|\{\vec{w}_{A_k}\}) \prod_{A_k}
+  p_{\text{\textsc{dir}}}(\vec{w}_{A_k}\,|\,\vec{\alpha}_{A_k})
 \end{align*}
-
-\[ p(\{\vec{w}_{A_k}\}\,|\, \vec{x}, \{\vec{\alpha}_{A_k}\}) \propto\]\[ \prod_j
-p_{\text{\textsc{prcg}}}(x_j|\{\vec{w}_{A_k}\}) \prod_{A_k} p_{\text{\textsc{dir}}}(\vec{w}_{A_k}\,|\,\vec{\alpha}_{A_k}) \]
 
@@
 
-\vspace{-1em}
 \section{Implementation \label{sec:implementation}}
-
-\lstset{
-    language=Prolog,
-    basicstyle=\fontsize{9}{10.5}\selectfont\ttfamily
-}
 
@@
 
-% First, we define the distribution over each non-terminal symbol, $n
-% \in N$, using the {\tt values} predicate. For each $n$ we define the
-% values the symbol can take on ({\it i.e.} one for each rule), as well
-% as the probability of each value. For each rule in the LPN, we also
-% define a corresponding clause of the {\tt reduce} predicate to capture
-% the actual relations between the left-hand and right-hand side of each
-% rule. Sampling and parsing begins with a call to {\tt lpn(P-IN)},
-% which chooses some rule for the left-hand side via {\tt msw(P,V)} and
-% recursively reduces the input term using the chosen rule ({\tt
-% reduce(P-IN,V)}).
-
 \section{Experiment}
@@
 \begin{table}[h]
+  \begin{subtable}[b]{0.5\linewidth}
     \begin{tabular}{>{\tiny} l >{\tiny} l}
@@
     \end{tabular}
-\caption{The 4 predicate LPN trained to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
+    \caption{The 4 predicate LPN trained to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
     \label{tab:grammar}
-\end{table}
-
-To inspect visually the learned grammar, we thresholded rules
-according to the expected number of times they were used in parsing
-the training dataset. Table~\ref{tab:grammar} shows all rules with
-expected count above $1e-6$. This reduces from $2669$ to $52$ the
-number of significant rules. On inspection, predicate $A_2$ forms
-``before'' sentences, predicate $A_4$ forms ``after'' sentences,
-predicate $A_3$ is successorship recursively defined over the decades
-and ones, and predicate $A_2$ is a category for the decade words.
-
-Our LPN does not learn to how to transition between the last word in a decade
-and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
-makes the intuitively reasonable generalization that ``seventy nine''
-should be followed by ``seventy ten.''
-
-\begin{table}[t]
+  \end{subtable}
+  \begin{subtable}[b]{0.5\linewidth}
     \begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
@@
     \label{tab:results}
+  \end{subtable}
 \end{table}
 
-\begin{figure}[t]
-		\includegraphics[width=\linewidth]{parseTree/parse.pdf}
-		\caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
-                \label{fig:parseexample}
-\end{figure}
+To inspect visually the learned grammar, we thresholded rules
+according to the expected number of times they were used in parsing
+the training dataset. Table~\ref{tab:grammar} shows all rules with
+expected count above $1e-6$. This reduces from $2669$ to $52$ the
+number of significant rules. On inspection, predicate $A_2$ forms
+``before'' sentences, predicate $A_4$ forms ``after'' sentences,
+predicate $A_3$ is successorship recursively defined over the decades
+and ones, and predicate $A_2$ is a category for the decade words.
 
-\vspace{-1em}
+Our LPN does not learn to how to transition between the last word in a decade
+and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
+makes the intuitively reasonable generalization that ``seventy nine''
+should be followed by ``seventy ten.''
+
 \section{Discussion}
@@
 
-%% Previous work has shown that generative models of probabilistic
-%% programs are a useful tool for understanding human concept learning
-%% for domains as varied as handwritten characters and structured visual
-%% objects~\cite{dechter2013bootstrap, stuhlmuller2010learning,
-%%   lake2012concept}. Here, we continue that work by showing how some
-%% forms of both human language learning and concept learning may be
-%% fruitfully modeled as inference over a generative model of
-%% probabilistic logic programs. Future work should work explore how LPNs
-%% and its extensions may apply to other domains where generative models
-%% of probabilistic programming have already proven useful, as well as
-%% generalizing the current approach to abstract, symbolic concepts.
+\subsubsection*{Acknowledgments}
+
+We should probably include acknowledgments for this paper.
+
+\subsubsection*{References}
 
 \bibliographystyle{amsplain}
-\bibliography{draft}
+\bibliography{nips_workshop}
 

### 2014-11-05 src/GIJoe/SRS/Architecture.hs
Added some untracked LPN Haskell files.

@@ -1,0 +1,90 @@
+
+module GIJoe.SRS.Architecture where
+
+import Data.List
+import System.Random.Shuffle (shuffle', shuffleM)
+import Control.Monad.Random.Class
+import Control.Monad
+
+import GIJoe.SRS.Parse
+import GIJoe.SRS.Prismify
+
+
+ipaLexicon = do xs <- readFile "./experiments/prism/englishPhones.txt"
+                return $ lines xs
+
+
+buildGrammar :: Int -- ^ number of hidden predicates
+             -> [String] -- ^ lexicon
+             -> [String] -- ^ list of observed unary predicates
+             -> FilePath
+             -> IO ()
+buildGrammar n lexicon observed outpath = do
+  writeFile outpath $ show rs
+  writeFile (outpath ++ ".psm") $ prismClauses rs
+  where rs = fromRight $ parseSystem "buildGrammar" $ systemString
+        fromRight (Right x) = x
+        fromRight (Left err) = error $ show err
+        systemString = unlines $ 
+                       (concat [rules i j k | i <- [1..n], j <-[1..n], k <- [j..n]])++
+                       [ruleLex i w v | i <- [1..n],
+                        (wi, w) <- zip [0..] lexicon,
+                        (vi, v) <- zip [0..] lexicon,
+                        vi >= wi] ++ 
+                       [ruleLex i w "null" | i <- [1..n], w <- lexicon] ++ 
+                       [ruleLex i "null" w | i <- [1..n], w <- lexicon] ++
+                       [pred ++ "(X Y) <-- A" ++ show i ++ "(X, Y)." | (pred, i) <- zip observed [1..]]
+        rules i j k | j /= k  = do
+          [v1, v2, v3, v4] <- permutations ["X", "Y", "U", "V"]
+          return $ "A" ++ show i ++ "(X Y, U V) <-- " ++
+            "A" ++ show j ++ "(" ++ v1 ++ ", " ++ v2 ++ "), " ++
+            "A" ++ show k ++ "(" ++ v3 ++ ", " ++ v4 ++ ")."
+        rules i j k | j == k  = do
+          [v1, v2, v3, v4] <- permutations ["X", "Y", "U", "V"]
+          guard $ (v1, v2) <= (v3, v4)
+          return $ "A" ++ show i ++ "(X Y, U V) <-- " ++
[…]

### 2014-11-05 src/GIJoe/SRS/Prismify-Main.hs

@@ -1,0 +1,20 @@
+
+module Main where
+
+import System.Environment
+import GIJoe.SRS.Prismify
+
+-- executable main function: 
+main = do
+  args <- getArgs
+  progName <- getProgName
+  let go fin fout = prismify fin fout
+  let prismify_usage = unlines [
+        "Usage: prismify input-sys-file output-psm-file."
+        ]
+  case args of
+    [] -> error $ unlines ["no arguments supplied.", prismify_usage]
+    [fin] -> go fin (fin ++ ".psm")
+    [fin, fout] -> go fin fout
+    _ -> error $ unlines ["too many arguments.", prismify_usage]
+    

### 2014-11-06 drafts/nips_workshop/nips_workshop.tex
significant cuts for length of NIPS paper

@@
 
-\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
+\title{Latent Predicate Networks: unsupervised learning of  probabilistic programs with context sensitive grammars}
 
@@
 \author{Joshua B. Tenenbaum}
-\affil{Department of Brain and Cognitive Sciences, MIT}
+\affil{{\normalsize Department of Brain and Cognitive Sciences, MIT}}
+\date{}
 
@@
 \begin{abstract}
+
+%%   - learning both abstract concepts and language presents an interesting chicken and egg problem
+%% 
+%%   - we can model this sort of learning a conceptual grammar for sentences
+%% 
+%%   - importantly, this grammar must be probabilistic to accomodate noisy data and represent our uncertainty about a concept's meaning
+%% 
+%%   - Here, we present early work on LPNs, showing them to be the solution we want
+%% 
+%%   - Crucially, we show how, for certain domains, learning language and abstract concepts can be done simultaneously
+%% 
+%%   - We also show that this approach naturally represents sentences as probabilistic conceptual programs.
+%% 
+%% 
+%%   While a large set of abstract concepts are necessary to understand
+%%   the utterances of a language, very few of those concepts are
+%%   initially available to children. Instead, they must be learned,
+%%   primarily {\it via} language. One approach to understanding this
+%%   joint learning problem is to model sentences as the output of
+%%   probabilistic programs constructed
+%% 
+%% 
+%% 
+%%   A large set of abstract concepts are necessary to understand the
+%%   utterances of a language, and at the same time, many of those
+%%   abstract concepts are acquired primarily {\it via} linguistic
+%%   experience. One way we might approach this sort of simultaneous
+%%   learning problem is to infer the distribution over sentences, where
+%%   the latent structure of this distributi
+%% 
   For humans, learning abstract concepts and learning language go hand
-  in hand: we acquire abstract knowledge primarily through linguistic
-  experience, and acquiring abstract concepts is a crucial step in
-  learning the meanings of linguistic expressions. Number knowledge is
-  a case in point: we largely acquire concepts such as seventy-three
-  through linguistic means, and we can only know what the sentence
-  ``seventy-three is more than twice as big as thirty-one" means if we
-  can grasp the meanings of its component number words. How do we
-  begin to solve this problem? One approach is to estimate the
-  distribution from which sentences are drawn, and, in doing so, infer
-  the latent concepts and relationships that best explain those
-  sentences. We present early work on a learning framework called
-  Latent Predicate Networks (LPNs) which learns concepts by inferring
-  the parameters of probabilistic context-sensitive grammars over
-  sentences.  We show that for a small fragment of sentences
-  expressing relationships between English number words, we can use
-  hierarchical Bayesian inference to learn LPNs that can answer simple
-  queries about previously unseen relationships within this
-  domain. These generalizations demonstrate LPNs' promise as a tool
-  for learning and representing conceptual knowledge in language.
+  in hand: we acquire much of our abstract knowledge through
+  linguistic experience, but the same abstract concepts are often
+  crucial to interpreting linguistic expressions. How can we do both
+  simultaneously? One approach is to estimate the distribution from
+  which sentences are drawn, and, in doing so, infer the latent
+  concepts and relationships that best explain those sentences. We
+  present early work on a learning framework called Latent Predicate
+  Networks (LPNs) which learns concepts by inferring the parameters of
+  probabilistic context-sensitive grammars over sentences.  We show
+  that for a small fragment of sentences expressing relationships
+  between English number words, we can use online hierarchical Bayesian
+  inference to learn LPNs that can answer simple queries about
+  previously unseen relationships within this domain. These
+  generalizations demonstrate LPNs' promise as a tool for learning and
+  representing conceptual knowledge in language.
 \end{abstract}
@@
 
-\subsection{Generic Architecture}
-
-An LPN is a PRCG with the following generic architecture: the
-\emph{axiom} of an LPN is a unary predicate $S$, and it has $K$
-densely connected binary predicates $\{A_k\}_{k=1}^K$ (Figure
-\ref{fig:architecture}). For each latent predicate $A_k$, we define
-rules such that $A_k(w_a,w_b)$ is true for each possible pair of
-terminals, $w_a,w_b \in T$. $T$ may include the empty string $\epsilon$, but we
-do not allow latent predicates of the form $A_k(\epsilon, \epsilon)$:
-see Section~\ref{sec:implementation}. We also define rules such that
-$A_k(XY,UV)$ is true for each possible pair of predicates, $A_i,A_j$,
-and every possible ordering of the variables $X,Y,U$ and $V$ across
-$A_i$ and $A_j$. Finally, we define $S(XY)$ to be the concatenation of
-the two arguments of $A_1(X,Y)$.
-
 \subsection{Learning Model}
@@
 
-\begin{figure}[t]
-	\centering
-	\begin{minipage}[b]{0.8\linewidth}
-1) A(a, a) :: 0.25. \\
-2) A(b, b) :: 0.3. \\
-3) A(X a, Y a) $\leftarrow$ A(X, Y) :: 0.25. \\
-4) A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.
-		\subcaption{}
-		\label{fig:grammar}
-	\end{minipage}
-	%\rule[0pt]{0.8\linewidth}{.5pt}
-	\begin{minipage}[b]{0.8\linewidth}
-        \fontsize{9}{10.5}\selectfont\ttfamily
-		\begin{verbatim}
-values(`A', [1, 2, 3, 4],
-    [0.25, 0.30, 0.25, 0.20]).
-
-reduce(`A'-[[a],[a]],1).
-reduce(`A'-[[b],[b]],2).
-reduce(`A'-[A2,B2],3) :- lpn(`A'-[X,Y]),
-    append(X,[a],A2), append(Y,[a],B2).
-reduce(`A'-[A2,B2],4) :- lpn(`A'-[X,Y]),
-    append(X,[b],A2), append(Y,[b],B2).
-
-lpn(P-IN) :- reduce(P-IN,V), msw(P,V).
-		\end{verbatim}
-		\subcaption{}
-		\label{fig:prism}
-	\end{minipage}
-	\caption{Possible encodings of the 2-copy language $\{ww\,|\, w \in \{a,b\}^+\}$ as (\subref{fig:grammar}) an LPN, (\subref{fig:prism}) a PRISM program.}
-	\label{fig:copy}
-\end{figure}
-
 LPNs can be encoded as a restricted subclass of PRISM programs; this
 is very similar to how PCFGs are encoded in
-PRISM~\cite{DBLP:conf/cl/2000}. See Figure~\ref{fig:copy} for an
-example. A general probabilistic logic programming system based on
-PROLOG, PRISM provides built-in predicates for probabilistic execution
-and Bayesian inference over logic programs with stochastic
-choices. There are several restrictions placed on PRISM programs to
-maintain the validity of their probabilistic interpretation. Most
-importantly, derivations of a probabilistic predicate cannot contain
-cycles. Because we disallow $A_i(\epsilon,\epsilon)$ as a valid
-clause, every term in the body of a clause has shorter arguments than
-the head, giving acyclic and finite derivations.
+PRISM~\cite{DBLP:conf/cl/2000}. A general probabilistic logic
+programming system based on PROLOG, PRISM provides built-in predicates
+for probabilistic execution and Bayesian inference over logic programs
+with stochastic choices. There are several restrictions placed on
+PRISM programs to maintain the validity of their probabilistic
+interpretation. Most importantly, derivations of a probabilistic
+predicate cannot contain cycles. Because we disallow
+$A_i(\epsilon,\epsilon)$ as a valid clause, every term in the body of
+a clause has shorter arguments than the head, giving acyclic and
+finite derivations.
 
@@
 
-\begin{table}[h]
-  \begin{subtable}[b]{0.5\linewidth}
-    \begin{tabular}{>{\tiny} l >{\tiny} l}
-      \multicolumn{2}{>{\tiny}c}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} \\
-      \multicolumn{2}{>{\tiny}c}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} \\
-      $A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
-      $A_3$(one, two) : 0.3993 &
-      $A_3$(two, three) : 0.2063 \\
-      $A_3$(three, four) : 0.1093 &
-      $A_3$(four, five) : 0.0734 \\
-      $A_3$(five, six) : 0.0502 &
-      $A_3$(six, seven) : 0.0355 \\
-      $A_3$(eight, nine) : 0.0290 &
-      $A_3$(seven, eight) : 0.0271 \\
-      $A_2$(fifty, fifty) : 0.0375 &
-      $A_2$(thirty, thirty) : 0.0361 \\
-      $A_2$(eighty, eighty) : 0.0339 &
-      $A_3$(null, one) : 0.0231 \\
-      $A_2$(forty, forty) : 0.0332 &
-      $A_2$(twenty, twenty) : 0.0310 \\
-      $A_2$(seventy, seventy) : 0.0296 &
-      $A_2$(sixty, sixty) : 0.0274 \\
-      $A_2$(ninety, ninety) : 0.0260 &
-      $A_3$(eighteen, nineteen) : 0.0064 \\
-      $A_3$(sixteen, seventeen) : 0.0049 &
-      $A_3$(eleven, twelve) : 0.0044 \\
-      $A_3$(nine, ten) : 0.0044 &
-      $A_3$(thirteen, fourteen) : 0.0044 \\
-      $A_3$(fourteen, fifteen) : 0.0039 &
-      $A_3$(ten, eleven) : 0.0034 \\
-      $A_2$(null, fifty) : 0.0043 &
-      $A_3$(eighty, null) : 0.0030 \\
-      $A_3$(seventeen, eighteen) : 0.0030 &
-      $A_3$(nine, sixty) : 0.0025 \\
-      $A_2$(nine, seventy) : 0.0029 &
-      $A_3$(nine, forty) : 0.0020 \\
-      $A_2$(null, thirty) : 0.0022 &
-      $A_2$(nine, ninety) : 0.0022 \\
-      $A_3$(twelve, thirteen) : 0.0015 &
-      $A_3$(fifteen, sixteen) : 0.0015 \\
-      $A_3$(null, comes) : 0.0010 &
-      $A_2$(twenty, after) : 0.0014 \\
-      $A_2$(sixty, before) : 0.0007 &
-      $A_3$(nineteen, comes) : 0.0005 \\
-      $A_4$(nine, thirty) : 0.0010 & \\
+  \begin{table}
+    \centering
+    \begin{tabular}{>{\tiny} l >{\tiny} l >{\tiny} l >{\tiny} l }
+      \multicolumn{2}{>{\tiny}l}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} &      $A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} & $A_3$(one, two) : 0.3993 & $A_3$(two, three) : 0.2063 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} & $A_3$(three, four) : 0.1093 & $A_3$(four, five) : 0.0734 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} & $A_3$(five, six) : 0.0502 & $A_3$(six, seven) : 0.0355 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} & $A_3$(eight, nine) : 0.0290 & $A_3$(seven, eight) : 0.0271 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} & $A_2$(fifty, fifty) : 0.0375 & $A_2$(thirty, thirty) : 0.0361 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} & $A_2$(eighty, eighty) : 0.0339 & $A_3$(null, one) : 0.0231 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} & $A_2$(forty, forty) : 0.0332 & $A_2$(twenty, twenty) : 0.0310 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} & $A_2$(seventy, seventy) : 0.0296 & $A_2$(sixty, sixty) : 0.0274 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} & $A_2$(ninety, ninety) : 0.0260 & $A_3$(eighteen, nineteen) : 0.0064 \\
+      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} & $A_3$(sixteen, seventeen) : 0.0049 & $A_3$(eleven, twelve) : 0.0044 \\
+      $A_3$(nine, ten) : 0.0044 & $A_3$(thirteen, fourteen) : 0.0044 & $A_3$(fourteen, fifteen) : 0.0039 & $A_3$(ten, eleven) : 0.0034 \\
+      $A_2$(null, fifty) : 0.0043 & $A_3$(eighty, null) : 0.0030 & $A_3$(seventeen, eighteen) : 0.0030 & $A_3$(nine, sixty) : 0.0025 \\
+      $A_2$(nine, seventy) : 0.0029 & $A_3$(nine, forty) : 0.0020 & $A_2$(null, thirty) : 0.0022 & $A_2$(nine, ninety) : 0.0022 \\
+      $A_3$(twelve, thirteen) : 0.0015 & $A_3$(fifteen, sixteen) : 0.0015 & $A_3$(null, comes) : 0.0010 & $A_2$(twenty, after) : 0.0014 \\
+      $A_2$(sixty, before) : 0.0007 & $A_3$(nineteen, comes) : 0.0005 & $A_4$(nine, thirty) : 0.0010 & \\
     \end{tabular}
@@
     \label{tab:grammar}
-  \end{subtable}
-  \begin{subtable}[b]{0.5\linewidth}
+  \end{table}
+
+\begin{figure}
+    \begin{subfigure}{0.45\linewidth}
       \begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
         Question & $K=4$ \\ \hline
-      after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
-      after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
-      after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
-      after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
-      after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
-      after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
-      after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
-      after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
-      after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
-      after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
-      after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
-      after seventy nine comes \underline{\hspace{1cm}}? & seventy ten $\times$ \\
-      after ninety five comes \underline{\hspace{1cm}}? & ninety six \checkmark \\
-      before twenty three comes \underline{\hspace{1cm}}? & twenty two \checkmark \\
-      before thirty comes \underline{\hspace{1cm}}? & thirty eighty $\times$ \\
-      before thirty eight comes \underline{\hspace{1cm}}? & thirty seven \checkmark \\
-      before forty one comes \underline{\hspace{1cm}}? & forty \checkmark \\
-      before fifty three comes \underline{\hspace{1cm}}? & fifty two \checkmark \\
-      before sixty eight comes \underline{\hspace{1cm}}? & sixty seven \checkmark \\
-      before seventy two comes \underline{\hspace{1cm}}? & seventy one \checkmark \\
-      before seventy three comes \underline{\hspace{1cm}}? & seventy two \checkmark \\
-      before eighty five comes \underline{\hspace{1cm}}? & eighty four \checkmark \\
-      before ninety two comes \underline{\hspace{1cm}}? & ninety one \checkmark \\
-      before ninety three comes \underline{\hspace{1cm}}? & ninety two \checkmark \\
-      before ninety five comes \underline{\hspace{1cm}}? &ninety four \checkmark \\
+        e.g. after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
+        e.g. after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
+        e.g. after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
+        e.g. after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
+        e.g. after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
+        e.g. after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
+        %% after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
+        %% after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
+        %% after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
+        %% after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
+        %% after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
+        %% after seventy nine comes \underline{\hspace{1cm}}? & seventy ten $\times$ \\
+        %% after ninety five comes \underline{\hspace{1cm}}? & ninety six \checkmark \\
+        %% before twenty three comes \underline{\hspace{1cm}}? & twenty two \checkmark \\
+        %% before thirty comes \underline{\hspace{1cm}}? & thirty eighty $\times$ \\
+        %% before thirty eight comes \underline{\hspace{1cm}}? & thirty seven \checkmark \\
+        %% before forty one comes \underline{\hspace{1cm}}? & forty \checkmark \\
+        %% before fifty three comes \underline{\hspace{1cm}}? & fifty two \checkmark \\
+        %% before sixty eight comes \underline{\hspace{1cm}}? & sixty seven \checkmark \\
+        %% before seventy two comes \underline{\hspace{1cm}}? & seventy one \checkmark \\
+        %% before seventy three comes \underline{\hspace{1cm}}? & seventy two \checkmark \\
+        %% before eighty five comes \underline{\hspace{1cm}}? & eighty four \checkmark \\
+        %% before ninety two comes \underline{\hspace{1cm}}? & ninety one \checkmark \\
+        %% before ninety three comes \underline{\hspace{1cm}}? & ninety two \checkmark \\
+        %% before ninety five comes \underline{\hspace{1cm}}? &ninety four \checkmark \\
       \end{tabular}
@@
       \label{tab:results}
-  \end{subtable}
-\end{table}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.45\linewidth}
+    \rule{6cm}{2.5cm}
+    \caption{PUT HELD OUT LL FIGURE HERE}
+    \label{fig:heldoutLL}
+  \end{subfigure}
+\end{figure}
 
@@
 
-\section{Discussion}
-
-Though simple, the knowledge learned by LPNs in the number
-domain we explore in this paper goes beyond the expressive capacity of
-PCFGs and HMMs, and yet it avoids intractability by using a restricted
-formalism. The human ability to learn abstract concepts must also rely
-on such a compromise, and it remains to be seen to what extent LPNs
-can model this ability.
-
-This work leaves many open questions. What domains of concepts are
-amenable to being represented and learned by LPNs? Can multiple and
-varied systems of concepts be learned at once? Perhaps most
-importantly, are there generic architectures for LPNs that allow a
-larger number of predicates without an insurmountable blow-up in the
-number of rules? One idea which we are pursuing is using layered
-connectivity, much like neural networks, to limit the number of rules
-in which each predicate participates.
-
-Many of the algorithms and formalisms used here were originally
-developed for use in other closely related areas, including logic
-programming and semantic parsing. Semantic parsing, in particular,
-seems especially related to the challenge we face here of jointly
-learning meaning and structure from
-sentences~\cite{berant2013semantic, liang2013learning, kwiatkowski2010inducing, poon2009unsupervised}. Semantic parsing,
-however, tends to frame this challenge slightly differently than we do
-here, namely, by asking how utterances can be mapped to an explicit
-internal logical language. By contrast, we focus here on systems where
-meaning and structure seem inseparable. Understanding how these two
-approaches relate and inform one another is an interesting and open
-question.
-
-\subsubsection*{Acknowledgments}
-
-We should probably include acknowledgments for this paper.
-
-\subsubsection*{References}
+\subsubsection*{Acknowledgments?}
 

### 2014-11-07 drafts/nips_workshop/nips_workshop.tex
new abstract, include Held-Out LL figure

@@
 
-%%   - learning both abstract concepts and language presents an interesting chicken and egg problem
-%% 
-%%   - we can model this sort of learning a conceptual grammar for sentences
-%% 
-%%   - importantly, this grammar must be probabilistic to accomodate noisy data and represent our uncertainty about a concept's meaning
-%% 
-%%   - Here, we present early work on LPNs, showing them to be the solution we want
-%% 
-%%   - Crucially, we show how, for certain domains, learning language and abstract concepts can be done simultaneously
-%% 
-%%   - We also show that this approach naturally represents sentences as probabilistic conceptual programs.
-%% 
-%% 
-%%   While a large set of abstract concepts are necessary to understand
-%%   the utterances of a language, very few of those concepts are
-%%   initially available to children. Instead, they must be learned,
-%%   primarily {\it via} language. One approach to understanding this
-%%   joint learning problem is to model sentences as the output of
-%%   probabilistic programs constructed
-%% 
-%% 
-%% 
-%%   A large set of abstract concepts are necessary to understand the
-%%   utterances of a language, and at the same time, many of those
-%%   abstract concepts are acquired primarily {\it via} linguistic
-%%   experience. One way we might approach this sort of simultaneous
-%%   learning problem is to infer the distribution over sentences, where
-%%   the latent structure of this distributi
-%% 
-  For humans, learning abstract concepts and learning language go hand
-  in hand: we acquire much of our abstract knowledge through
-  linguistic experience, but the same abstract concepts are often
-  crucial to interpreting linguistic expressions. How can we do both
-  simultaneously? One approach is to estimate the distribution from
-  which sentences are drawn, and, in doing so, infer the latent
-  concepts and relationships that best explain those sentences. We
-  present early work on a learning framework called Latent Predicate
-  Networks (LPNs) which learns concepts by inferring the parameters of
-  probabilistic context-sensitive grammars over sentences.  We show
-  that for a small fragment of sentences expressing relationships
-  between English number words, we can use online hierarchical Bayesian
-  inference to learn LPNs that can answer simple queries about
-  previously unseen relationships within this domain. These
-  generalizations demonstrate LPNs' promise as a tool for learning and
-  representing conceptual knowledge in language.
+  Probabilistic programs successfully capture many aspects of the
+  complex symbolic and statistical knowledge brought to bear in human
+  cognition. Children aren't born with this knowledge, however, but
+  must instead learn it, often in an unsupervised fashion. Compelling
+  probabilistic programming models of human cognition must therefore
+  also be learned. We present early work on Latent Predicate Networks
+  (LPN), a framework for unsupervised learning in a restricted space
+  of context-sensitive probabilistic programs. Specifically, we use
+  Online Variational Bayesian Expectation Maximization in
+  PRISM, a probabilistic programming system, to perform inference over
+  the set of Stochastic Logic Programs (SLP) parseable in polynomial
+  time. Using a small fragment of English sentences about number, we
+  show that LPNs can recover the relevant conceptual structure
+  necessary to generate novel sentences in this domain with high
+  probability.
+
 \end{abstract}
@@
 \section{Introduction}
+
+%% Introduction idea
+%% *** Probabilistic programs have been great models of human cognition
+%% **** one-shot learning
+%% **** linguistic pragmatics
+%% **** some other example
+%% *** one thing they lack is a good way to automatically induce full programs from data
+%% *** yet, children seem to do this all the time
+%% *** So, this motivates our approach
+%% *** But, we have to overcome some technical challenges
+%% **** many prob.prog. languages are too expressive to be tractable
+%% **** many models of concept learning use too restricted a set of primitives
+
+Probabilistic programs can successfully model many aspects of human
+cognition [CITIATONS]. Part of their strength as compared to other
+models of cognition lies in their ability to flexibly integrate both
+symbolic and statistical information within a single framework.
+
+One area probabilistic programming systems lag behind human cognition
+is in their ability to automatically induce entire
 
@@
 
-Such concepts share many of the properties of
-language syntax: they are unbounded in number, they derive
-their meanings via composition, and, although people only ever say,
-hear, or read about a small number of them, they are able to reason
-correctly about any them. There seems to be a grammar to these
-concepts, and grasping this grammar is critical to understanding their
-meanings and how to use them. This motivates our approach here, which
-is to apply the tools of probabilistic grammars more familiar from
-studies of syntax to the problem of concept aquisition.
+Such concepts share many of the properties of language syntax: they
+are unbounded in number, they derive their meanings via composition,
+and, although people only ever say, hear, or read about a small number
+of them, they are able to reason correctly about any them. There seems
+to be a grammar to these concepts, and grasping this grammar is
+critical to understanding their meanings and how to use them. This
+motivates our approach here, which is to apply the tools of
+probabilistic grammars more familiar from studies of syntax to the
+problem of concept aquisition.
+
+{\bf include somewhere:} A general probabilistic logic
+programming system based on PROLOG, PRISM provides built-in predicates
+for probabilistic execution and Bayesian inference over logic programs
+with stochastic choices.
 
@@
 
-First, whereas context-free grammars are suitable for describing large
-swathes of language syntax, the grammars of concepts are not
-context-free. To address this, we use Range Concatenation Grammars, a
-context-sensitive grammar formalism -- one of several developed within
-linguistics -- and extend this formalism to a probabilistic model of
-sentences.
+First, many formal systems of probabilistic programming are so
+expressive as to make learning generally intractable. To address this,
+we use Range Concatenation Grammar, a context-sensitive grammar
+formalism originally developed in linguistics and extend this to a
+probabilistic model of programs.
 
@@
 Finally, inducing probabilistic context-sensitive grammars with latent
-predicates threatens to be intractable: our goal is to find a
-middle ground between expressivity and tractability. Using
-PRISM~\cite{DBLP:journals/jair/SatoK01} -- a probabilistic logic programming system that naturally
-implements efficient dynamic programming algorithms for our models --
-we are able to explore which domains and which grammar architectures
-are a good fit for this grammar-based approach.
+predicates threatens to be intractable: our goal is to find a middle
+ground between expressivity and tractability. Using
+PRISM~\cite{DBLP:journals/jair/SatoK01} -- a probabilistic logic
+programming system that naturally implements efficient dynamic
+programming algorithms for our models -- we are able to explore which
+domains and which grammar architectures are a good fit for this
+grammar-based approach.
 
@@
 
-Given a collection of predicates, $\{A_k\}_{k=1}^{K}$, and a
-distribution over clauses, $\{\vec{w}_{A_k}\}$, the learning task is
-to model a set of utterances, $\{x_j\}_{j=1}^{J}$, as being generated
-according to the following distribution:
-
 \begin{align*}
@@
 
-In words, the weights of clauses sharing head predicate $A_k$ are
-drawn from a Dirichlet distribution defined by
-$\vec{\alpha}_{A_k}$. Sentences $x_j$ are then drawn from the
-resulting PRCG.
+Given a collection of predicates, $\{A_k\}_{k=1}^{K}$, and a
+distribution over clauses, $\{\vec{w}_{A_k}\}$, the learning task is
+to model a set of outputs, $\{x_j\}_{j=1}^{J}$, as being generated
+according to the above distribution. In words, the weights of clauses
+sharing head predicate $A_k$ are drawn from a Dirichlet distribution
+defined by $\vec{\alpha}_{A_k}$. Each data point $x_j$ is then drawn
+from the resulting PRCG. We use Bayes' Rule to update  weights $\{\vec{w}_{A_k}\}$ given our data $\vec{x}$.
 
-Bayesian inference over stochastic grammars
-and stochastic logic programs has been an active area of
-research in recent decades~\cite{DBLP:journals/etai/Muggleton00, cussens2001parameter, DBLP:conf/emnlp/LiangPJK07,
-  goldwater2006contextual, johnson2006adaptor}.
-Variational inference is a popular approach in this domain and the one
-we adopt here. As explained in the next section, we implemented
-inference by translating LPNs into PRISM programs and using its
-built-in Variational Bayes Expectation-Maximization
-algorithm~\cite{sato2008variational}.
+\section{Inference \label{sec:implementation}}
 
-\section{Implementation \label{sec:implementation}}
+Bayesian inference over stochastic grammars and stochastic logic
+programs has been an active area of research in recent
+decades~\cite{DBLP:journals/etai/Muggleton00, cussens2001parameter,
+  DBLP:conf/emnlp/LiangPJK07, goldwater2006contextual,
+  johnson2006adaptor}.  Variational inference is a popular approach in
+this domain and the one we adopt here. We implemented inference by
+translating LPNs into PRISM programs and using its built-in
+Variational Bayes Expectation-Maximization
+algorithm~\cite{sato2008variational}, modified for for use in an
+online, rather than batch, setting. We use the adaptive learning rate
+of \cite{ranganath2013adaptive}.
 
@@
 is very similar to how PCFGs are encoded in
-PRISM~\cite{DBLP:conf/cl/2000}. A general probabilistic logic
-programming system based on PROLOG, PRISM provides built-in predicates
-for probabilistic execution and Bayesian inference over logic programs
-with stochastic choices. There are several restrictions placed on
-PRISM programs to maintain the validity of their probabilistic
+PRISM~\cite{DBLP:conf/cl/2000}.  There are several restrictions placed
+on PRISM programs to maintain the validity of their probabilistic
 interpretation. Most importantly, derivations of a probabilistic
@@
 \begin{figure}
-    \begin{subfigure}{0.45\linewidth}
+    \begin{subfigure}[b]{0.45\linewidth}
       \begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
         Question & $K=4$ \\ \hline
-        e.g. after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
-        e.g. after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
-        e.g. after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
-        e.g. after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
-        e.g. after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
-        e.g. after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
-        %% after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
-        %% after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
-        %% after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
-        %% after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
-        %% after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
+        XXXX after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
+        XXXX after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
+        XXXX after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
+        XXXX after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
+        XXXX after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
+        XXXX after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
+        XXXX after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
+        XXXX after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
+        XXXX after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
+        XXXX after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
+        XXXX after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
         %% after seventy nine comes \underline{\hspace{1cm}}? & seventy ten $\times$ \\
@@
       \end{tabular}
-      \caption{Viterbi completions of held-out sentences to evaluate an LPN with 4 latent predicates in the number domain.}
+      \caption{}
       \label{tab:results}
@@
   \hfill
-  \begin{subfigure}{0.45\linewidth}
-    \rule{6cm}{2.5cm}
-    \caption{PUT HELD OUT LL FIGURE HERE}
+  \begin{subfigure}[b]{0.45\linewidth}
+    \includegraphics[width=\linewidth]{figures/train_number_net_0006_held_out.pdf}
+    \caption{}
     \label{fig:heldoutLL}
   \end{subfigure}
+  \caption{(\subref{tab:results}) Viterbi completions of held-out sentences to evaluate an LPN with 4 latent predicates in the number domain. (\subref{fig:heldoutLL}) PUT HELD OUT LL CAPTION HERE}
 \end{figure}

### 2014-11-07 experiments/prism/train_number_net_0006.held_out.py
Removed title from plot and bounded x axis.

@@ -18,13 +18,14 @@ data = data[1:, 1]
 
 fig = plt.figure()
 ax = fig.add_subplot(111)
-ax.plot(data, 
+ax.plot(range(1, len(data) + 1), 
+        data, 
         linewidth=2, 
         marker='o')
 ax.set_xlabel("Iteration")
 ax.set_ylabel(r"$\mathbb{E} \log p(D_{\text{test}})$")
-ax.set_title("Evaluating stochastic VBEM for LPNs")
 ax.set_ylim(top=1000)
+ax.set_xlim(right=20)
 
 # line at 0
 [lo, hi] = ax.get_xlim()

### 2014-11-07 prism/onlineVBEM.pl
Small changes.

@@ -146,6 +146,7 @@ ovbem1(DataSet, BatchSize, RateInit, RateFinal) :-
     random_multiselect(DataSet, BatchSize, Batch),
     M is round(N/BatchSize),
     BatchOut @= [D1: D in Batch, [D1], multiply_datum_count(D, M, D1)],
+    write(BatchOut), nl,
     % run vbem
     learn(BatchOut),
     % get output alphas
@@ -157,7 +158,14 @@ ovbem1(DataSet, BatchSize, RateInit, RateFinal) :-
     % update parameters towards ParamsFinal
     foreach(switch(Sw, Status, Vals, As0) in ParamsInit, [As0, As1, AsOut],
             (member(switch(Sw, Status, Vals, As1), ParamsFinal), !,
+             %% write('Updating params: '),
              update_vb_params(StepSize, As0, As1, AsOut),
+             %% write(update_vb_params(StepSize, As0, As1, AsOut)), nl,
+             nl,
+             write(switch(Sw)), nl,
+             write(as0(As0)), nl,
+             write(as1(As1)), nl,nl,
+             
              set_sw_a(Sw, AsOut))).
 
 print_rate(rate(T,G,H,R)) :-

### 2014-11-07 src/GIJoe/SRS/Architecture.hs

@@ -86,9 +86,6 @@ predicateNetwork _R _S _K lexicon observed outpath = do
             "A" ++ show (level-1) ++ "i" ++ show k ++ "(U, V).",
             "A" ++ show level ++ "i" ++ show i ++ "(X Y, U V) <-- " ++
             "A" ++ show (level-1) ++ "i" ++ show j ++ "(Y, X)," ++ 
-            "A" ++ show (level-1) ++ "i" ++ show k ++ "(V, U).",
-            "A" ++ show level ++ "i" ++ show i ++ "(X Y, U V) <-- " ++
-            "A" ++ show (level-1) ++ "i" ++ show j ++ "(Y, X)," ++ 
             "A" ++ show (level-1) ++ "i" ++ show k ++ "(V, U)."
             ]
         -- reverseRule level i j = 

### 2014-11-07 drafts/nips_workshop/nips_workshop.tex
further edits for nips workshop

@@
 derivations of infinite length; in this paper, we only consider PRCGs
-with derivations of finite length, so we need not worry about this
-requirement. See Figure \ref{fig:grammar} for an example of the context-sensitive 2-copy language $\{ww\,|w \in \{a,b\}^+\}$ as a PRCG.
+with derivations of finite length.
 
@@
 
-\begin{align*}
-  \vec{w}_{A_k} &\sim Dir(\vec{\alpha}_{A_k})\\
-  x_j &\underset{iid}{\sim} p_{\text{\textsc{prcg}}}(S(x_j)\,|\,\{w_{A_k}\})\\
-  p(\{\vec{w}_{A_k}\}\,|\, \vec{x}, \{\vec{\alpha}_{A_k}\}) &\propto
+\[ \vec{w}_{A_k} \sim Dir(\vec{\alpha}_{A_k})\hspace{3em}
+   x_j \underset{iid}{\sim} p_{\text{\textsc{prcg}}}(S(x_j)\,|\,\{w_{A_k}\}) \]
+\[  p(\{\vec{w}_{A_k}\}\,|\, \vec{x}, \{\vec{\alpha}_{A_k}\}) \propto
   \prod_j p_{\text{\textsc{prcg}}}(x_j|\{\vec{w}_{A_k}\}) \prod_{A_k}
-  p_{\text{\textsc{dir}}}(\vec{w}_{A_k}\,|\,\vec{\alpha}_{A_k})
-\end{align*}
+  p_{\text{\textsc{dir}}}(\vec{w}_{A_k}\,|\,\vec{\alpha}_{A_k}) \]
 
+  \vspace{-1em}
 Given a collection of predicates, $\{A_k\}_{k=1}^{K}$, and a
@@
 to model a set of outputs, $\{x_j\}_{j=1}^{J}$, as being generated
-according to the above distribution. In words, the weights of clauses
-sharing head predicate $A_k$ are drawn from a Dirichlet distribution
-defined by $\vec{\alpha}_{A_k}$. Each data point $x_j$ is then drawn
-from the resulting PRCG. We use Bayes' Rule to update  weights $\{\vec{w}_{A_k}\}$ given our data $\vec{x}$.
+according to the above distribution. The weights of clauses
+with head predicate $A_k$ are drawn from a Dirichlet distribution
+defined by $\vec{\alpha}_{A_k}$. Each datum $x_j$ is then drawn
+from the resulting PRCG. The probability of a set of weights given our data and Dirichlet prior is determined with Bayes' Rule.
 
@@
 
-Bayesian inference over stochastic grammars and stochastic logic
-programs has been an active area of research in recent
-decades~\cite{DBLP:journals/etai/Muggleton00, cussens2001parameter,
-  DBLP:conf/emnlp/LiangPJK07, goldwater2006contextual,
-  johnson2006adaptor}.  Variational inference is a popular approach in
-this domain and the one we adopt here. We implemented inference by
+LPNs can be encoded as a restricted subclass of PRISM programs; this
+is very similar to how PCFGs are encoded in
+PRISM~\cite{DBLP:conf/cl/2000}. Bayesian inference over stochastic
+grammars and stochastic logic programs has been an active area of
+research in recent decades~\cite{DBLP:journals/etai/Muggleton00,
+  cussens2001parameter, DBLP:conf/emnlp/LiangPJK07,
+  goldwater2006contextual, johnson2006adaptor}.  Variational inference
+is a popular approach in this domain and the one we adopt here by
 translating LPNs into PRISM programs and using its built-in
 Variational Bayes Expectation-Maximization
-algorithm~\cite{sato2008variational}, modified for for use in an
-online, rather than batch, setting. We use the adaptive learning rate
-of \cite{ranganath2013adaptive}.
-
-LPNs can be encoded as a restricted subclass of PRISM programs; this
-is very similar to how PCFGs are encoded in
-PRISM~\cite{DBLP:conf/cl/2000}.  There are several restrictions placed
-on PRISM programs to maintain the validity of their probabilistic
-interpretation. Most importantly, derivations of a probabilistic
-predicate cannot contain cycles. Because we disallow
-$A_i(\epsilon,\epsilon)$ as a valid clause, every term in the body of
-a clause has shorter arguments than the head, giving acyclic and
-finite derivations.
+algorithm~\cite{sato2008variational}. We modify it for use in an
+online setting, using the adaptive learning rate of
+\cite{ranganath2013adaptive}.
 
@@
 To evaluate LPNs as a probabilistic model of concept acquisition, we
-trained an LPN with $4$ latent predicates on a set of sentences
-expressing successor and predecessor relations in numbers between one
-and ninety-nine. The training set was the collection of sentences $$X
-= \{[\text{after}\, | \, \text{before}] \, \langle n \rangle \,
-\text{comes} \, \langle n+1 \rangle \,|\, n \in 1,\dots,99\},$$ where
-$\langle n \rangle$ is the number word corresponding to $n$. The
-lexicon was the set of number word corresponding to $1$ through $19$,
-the decades $20, \dots, 30$, the empty string, and the words
-``before'' and ``after.'' It is not difficult to manually work out an
-LPN that describes this limited domain of sentences; see
+trained an LPN with the architecture in Figure \ref{fig:architecture}
+and $K = 4$ latent predicates on a set of sentences expressing
+successor and predecessor relations in numbers between one and
+ninety-nine. The training set was the collection of sentences $$X =
+\{\text{after}\, \langle n \rangle \, \text{comes} \, \langle n+1
+\rangle \,|\, n \in 1,\dots,98\} \cup \{\text{before} \,\langle n
+\rangle \, \text{comes} \, \langle n-1 \rangle \,|\, n \in
+2,\dots,99\},$$ where $\langle n \rangle$ is the number word
+corresponding to $n$. The lexicon was the set of number word
+corresponding to $1$ through $19$, the decades $20, \dots, 90$, the
+empty string, and the words ``before'', ``after'', and ``comes''. See
 Figure~\ref{fig:parseexample}, for a possible LPN derivation of an
-example sentence.
-
-Although it is difficult to know how common these kinds of sentences
-are in child-directed speech, words for small numbers are far more
-common than words for larger ones \cite{macwhinney2000childes}. On the
-other hand, children learning to count to large numbers rehearse the
-sequence. To approximate this distribution of evidence, we drew these
+example sentence. To approximate the evidence children receive (words
+for small numbers naturally occuring more often than for larger ones,
+but children also rehearse the entire number sequence), we drew these
 sample sentences from a sum of a geometric distribution with parameter
@@
 $75\%$ and $25\%$, respectively. We drew $2000$ examples from this
-distribution, holding out the sentences in Table~\ref{tab:results} for
-evaluation.
+distribution, holding out a subset of sentences, including those in
+Table~\ref{tab:results} for evaluation.
 
@@
 
-We evaluated the learned model by asking for Viterbi ({\it i.e.} maximum a
-posteriori) completions of the last words of each held out test
+We evaluated the learned model by asking for Viterbi ({\it i.e.} maximum {\it a
+posteriori}) completions of the last words of each held out test
 sentence. Table~\ref{tab:results} shows these completions. The grammar
@@
 
-  \begin{table}
-    \centering
-    \begin{tabular}{>{\tiny} l >{\tiny} l >{\tiny} l >{\tiny} l }
-      \multicolumn{2}{>{\tiny}l}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} &      $A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} & $A_3$(one, two) : 0.3993 & $A_3$(two, three) : 0.2063 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} & $A_3$(three, four) : 0.1093 & $A_3$(four, five) : 0.0734 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} & $A_3$(five, six) : 0.0502 & $A_3$(six, seven) : 0.0355 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} & $A_3$(eight, nine) : 0.0290 & $A_3$(seven, eight) : 0.0271 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} & $A_2$(fifty, fifty) : 0.0375 & $A_2$(thirty, thirty) : 0.0361 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} & $A_2$(eighty, eighty) : 0.0339 & $A_3$(null, one) : 0.0231 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} & $A_2$(forty, forty) : 0.0332 & $A_2$(twenty, twenty) : 0.0310 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} & $A_2$(seventy, seventy) : 0.0296 & $A_2$(sixty, sixty) : 0.0274 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} & $A_2$(ninety, ninety) : 0.0260 & $A_3$(eighteen, nineteen) : 0.0064 \\
-      \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} & $A_3$(sixteen, seventeen) : 0.0049 & $A_3$(eleven, twelve) : 0.0044 \\
-      $A_3$(nine, ten) : 0.0044 & $A_3$(thirteen, fourteen) : 0.0044 & $A_3$(fourteen, fifteen) : 0.0039 & $A_3$(ten, eleven) : 0.0034 \\
-      $A_2$(null, fifty) : 0.0043 & $A_3$(eighty, null) : 0.0030 & $A_3$(seventeen, eighteen) : 0.0030 & $A_3$(nine, sixty) : 0.0025 \\
-      $A_2$(nine, seventy) : 0.0029 & $A_3$(nine, forty) : 0.0020 & $A_2$(null, thirty) : 0.0022 & $A_2$(nine, ninety) : 0.0022 \\
-      $A_3$(twelve, thirteen) : 0.0015 & $A_3$(fifteen, sixteen) : 0.0015 & $A_3$(null, comes) : 0.0010 & $A_2$(twenty, after) : 0.0014 \\
-      $A_2$(sixty, before) : 0.0007 & $A_3$(nineteen, comes) : 0.0005 & $A_4$(nine, thirty) : 0.0010 & \\
-    \end{tabular}
-    \caption{The 4 predicate LPN trained to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
-    \label{tab:grammar}
-  \end{table}
+%% Feel free to include if there's space, but this table may not be necessary.
+%%
+%%   \begin{table}
+%%     \centering
+%%     \begin{tabular}{>{\tiny} l >{\tiny} l >{\tiny} l >{\tiny} l }
+%%       \multicolumn{2}{>{\tiny}l}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} &      $A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} & $A_3$(one, two) : 0.3993 & $A_3$(two, three) : 0.2063 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} & $A_3$(three, four) : 0.1093 & $A_3$(four, five) : 0.0734 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} & $A_3$(five, six) : 0.0502 & $A_3$(six, seven) : 0.0355 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} & $A_3$(eight, nine) : 0.0290 & $A_3$(seven, eight) : 0.0271 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} & $A_2$(fifty, fifty) : 0.0375 & $A_2$(thirty, thirty) : 0.0361 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} & $A_2$(eighty, eighty) : 0.0339 & $A_3$(null, one) : 0.0231 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} & $A_2$(forty, forty) : 0.0332 & $A_2$(twenty, twenty) : 0.0310 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} & $A_2$(seventy, seventy) : 0.0296 & $A_2$(sixty, sixty) : 0.0274 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} & $A_2$(ninety, ninety) : 0.0260 & $A_3$(eighteen, nineteen) : 0.0064 \\
+%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} & $A_3$(sixteen, seventeen) : 0.0049 & $A_3$(eleven, twelve) : 0.0044 \\
+%%       $A_3$(nine, ten) : 0.0044 & $A_3$(thirteen, fourteen) : 0.0044 & $A_3$(fourteen, fifteen) : 0.0039 & $A_3$(ten, eleven) : 0.0034 \\
+%%       $A_2$(null, fifty) : 0.0043 & $A_3$(eighty, null) : 0.0030 & $A_3$(seventeen, eighteen) : 0.0030 & $A_3$(nine, sixty) : 0.0025 \\
+%%       $A_2$(nine, seventy) : 0.0029 & $A_3$(nine, forty) : 0.0020 & $A_2$(null, thirty) : 0.0022 & $A_2$(nine, ninety) : 0.0022 \\
+%%       $A_3$(twelve, thirteen) : 0.0015 & $A_3$(fifteen, sixteen) : 0.0015 & $A_3$(null, comes) : 0.0010 & $A_2$(twenty, after) : 0.0014 \\
+%%       $A_2$(sixty, before) : 0.0007 & $A_3$(nineteen, comes) : 0.0005 & $A_4$(nine, thirty) : 0.0010 & \\
+%%     \end{tabular}
+%%     \caption{The 4 predicate LPN trained to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
+%%     \label{tab:grammar}
+%%   \end{table}
 
@@
         Question & $K=4$ \\ \hline
-        XXXX after twenty comes \underline{\hspace{1cm}}? & twenty one \checkmark \\
-        XXXX after forty five comes \underline{\hspace{1cm}}? & forty six \checkmark \\
-        XXXX after forty seven comes \underline{\hspace{1cm}}? & forty eight  \checkmark \\
-        XXXX after forty nine comes \underline{\hspace{1cm}}? & forty ten $\times$ \\
-        XXXX after fifty nine comes \underline{\hspace{1cm}}? & fifty ten $\times$ \\
-        XXXX after sixty one comes \underline{\hspace{1cm}}? & sixty two \checkmark \\
-        XXXX after sixty three comes \underline{\hspace{1cm}}? & sixty four \checkmark \\
-        XXXX after sixty four comes \underline{\hspace{1cm}}? & sixty five \checkmark \\
-        XXXX after sixty five comes \underline{\hspace{1cm}}? & sixty six \checkmark \\
-        XXXX after sixty nine comes \underline{\hspace{1cm}}? & sixty ten $\times$ \\
-        XXXX after seventy three comes \underline{\hspace{1cm}}? & seventy four \checkmark \\
-        %% after seventy nine comes \underline{\hspace{1cm}}? & seventy ten $\times$ \\
-        %% after ninety five comes \underline{\hspace{1cm}}? & ninety six \checkmark \\
-        %% before twenty three comes \underline{\hspace{1cm}}? & twenty two \checkmark \\
-        %% before thirty comes \underline{\hspace{1cm}}? & thirty eighty $\times$ \\
-        %% before thirty eight comes \underline{\hspace{1cm}}? & thirty seven \checkmark \\
-        %% before forty one comes \underline{\hspace{1cm}}? & forty \checkmark \\
-        %% before fifty three comes \underline{\hspace{1cm}}? & fifty two \checkmark \\
-        %% before sixty eight comes \underline{\hspace{1cm}}? & sixty seven \checkmark \\
-        %% before seventy two comes \underline{\hspace{1cm}}? & seventy one \checkmark \\
-        %% before seventy three comes \underline{\hspace{1cm}}? & seventy two \checkmark \\
-        %% before eighty five comes \underline{\hspace{1cm}}? & eighty four \checkmark \\
-        %% before ninety two comes \underline{\hspace{1cm}}? & ninety one \checkmark \\
-        %% before ninety three comes \underline{\hspace{1cm}}? & ninety two \checkmark \\
-        %% before ninety five comes \underline{\hspace{1cm}}? &ninety four \checkmark \\
+        after twenty comes \ldots & twenty one \checkmark \\
+        after forty five comes \ldots & forty six \checkmark \\
+        after forty nine comes \ldots & forty ten $\times$ \\
+        after fifty nine comes \ldots & fifty ten $\times$ \\
+        after sixty one comes \ldots & sixty two \checkmark \\
+        before twenty three comes \ldots & twenty two \checkmark \\
+        before thirty comes \ldots & thirty eighty $\times$ \\
+        before thirty eight comes \ldots & thirty seven \checkmark \\
+        before forty one comes \ldots & forty \checkmark \\
+        before fifty three comes \ldots & fifty two \checkmark \\
+        before eighty five comes \ldots & eighty four \checkmark \\
+        %% after forty seven comes \ldots & forty eight  \checkmark \\
+        %% after sixty three comes \ldots & sixty four \checkmark \\
+        %% after sixty four comes \ldots & sixty five \checkmark \\
+        %% after sixty five comes \ldots & sixty six \checkmark \\
+        %% after sixty nine comes \ldots & sixty ten $\times$ \\
+        %% after seventy three comes \ldots & seventy four \checkmark \\
+        %% after seventy nine comes \ldots & seventy ten $\times$ \\
+        %% after ninety five comes \ldots & ninety six \checkmark \\
+        %% before sixty eight comes \ldots & sixty seven \checkmark \\
+        %% before seventy two comes \ldots & seventy one \checkmark \\
+        %% before seventy three comes \ldots & seventy two \checkmark \\
+        %% before ninety two comes \ldots & ninety one \checkmark \\
+        %% before ninety three comes \ldots & ninety two \checkmark \\
+        %% before ninety five comes \ldots &ninety four \checkmark \\
       \end{tabular}
@@
   \end{subfigure}
-  \caption{(\subref{tab:results}) Viterbi completions of held-out sentences to evaluate an LPN with 4 latent predicates in the number domain. (\subref{fig:heldoutLL}) PUT HELD OUT LL CAPTION HERE}
+  \caption{(\subref{tab:results}) Viterbi completions of held-out sentences for a 4-predicate 1-layer LPN.
+(\subref{fig:heldoutLL}) expected log likelihood on a set of held-out test data in a 35-predicate 5-layer LPN with.}
 \end{figure}

### 2014-11-07 drafts/nips_workshop/nips_workshop.tex
Rewrote introduction; fixed figure name.

--- 2014-11-07 drafts/nips_workshop/nips_workshop.tex
@@
 
-\title{Latent Predicate Networks: unsupervised learning of  probabilistic programs with context sensitive grammars}
+\title{\vspace{-1.5cm} Unsupervised learning of probabilistic programs with latent predicate networks}
 
@@
 \begin{document}
-
 \maketitle
 
+\vspace{-1cm}
 \begin{abstract}
@@
 
-Probabilistic programs can successfully model many aspects of human
-cognition [CITIATONS]. Part of their strength as compared to other
-models of cognition lies in their ability to flexibly integrate both
-symbolic and statistical information within a single framework.
+Probabilistic programs combine symbolic and statistical knowledge in way that is
+natural for humans to express. In fact, probabilistic programs have
+been proposed as models of human cognition in a variety of domains
+[CITATION]. As a suitable representation for learning, however, the flexibility
+of probabilistic programs poses a challenge. At one extreme, we can
+approach learning as fixing the symbolic structure of a probabilistic
+program and fitting some of its parameters; at the other extreme, the
+structure is unspecified and learning is performed by searching over
+valid syntactic forms. The former sacrifices the symbolic flexibility
+that probabilistic programs have to offer, and the latter is
+computationally infeasible. It is natural to suppose, then, that human
+learning charts a middle course between these extremes, restricting
+the space of probabilistic programs -- perhaps in a domain-specific
+way -- to accomodate both flexibility and tractability.
 
-One area probabilistic programming systems lag behind human cognition
-is in their ability to automatically induce entire
+Here, we present statistical learning (using the PRISM probabilistic
+logic programmming language) of a restricted type of probabilistic
+program, dubbed a \emph{latent predicate network} (LPN), for learning
+problems at the intersection of concept learning and language
+acquisition. People learn many abstract concepts primarily through
+language, and understanding language depends on understanding the
+underlying concepts. Number concepts are a good example: children do
+not learn about the meaning of ``seventy five'' by seeing examples of
+seventy five things; they do not know that ``seventy five'' is more
+than ``twenty five'' because of their perceptual experiences of these
+quantities. Rather, children learn the meaning of ``seventy five'' (or
+``a billion and five'') by noticing how number words are used in
+language, in counting sequences, in arithmetic exercises, {\it
+  etc}. Other good examples of such abstract concepts are kinship and
+social relations ({\it e.g.} ``my father in law's grandmother''),
+temporal relations (``the day after last Thanksgiving.''), and spatial
+relations (``above and just to the left of'').
 
-Although concept learning and language acquisition have typically been
-treated as distinct problems in AI, linguistics and cognitive
-development, they are strongly coupled for a child learning to
-understand language. More generally, people learn many abstract
-concepts primarily through language even though understanding language
-depends on understanding the underlying concepts. Research in concept
-learning is often focused on concepts grounded in perceptual features,
-and while it is almost certainly true that many concepts are learned
-via generalization from concrete examples, some concepts cannot be
-learned this way.
-
-Number concepts are a good example: children do not learn about the
-meaning of ``seventy five'' by seeing examples of seventy five things;
-they do not know that ``seventy five'' is more than ``twenty five''
-because of their perceptual experiences of these quantities. Rather,
-children learn the meaning of ``seventy five'' (or ``a billion and
-five'') by noticing how number words are used in language, in counting
-sequences, in arithmetic exercises, {\it etc}. Other good examples of such
-abstract concepts are kinship and social relations ({\it e.g.} ``my father
-in law's grandmother''), temporal relations (``the day after last
-Thanksgiving.''), and spatial relations (``above and just to the left
-of'').
-
-Such concepts share many of the properties of language syntax: they
-are unbounded in number, they derive their meanings via composition,
-and, although people only ever say, hear, or read about a small number
-of them, they are able to reason correctly about any them. There seems
-to be a grammar to these concepts, and grasping this grammar is
-critical to understanding their meanings and how to use them. This
-motivates our approach here, which is to apply the tools of
-probabilistic grammars more familiar from studies of syntax to the
-problem of concept aquisition.
-
-{\bf include somewhere:} A general probabilistic logic
-programming system based on PROLOG, PRISM provides built-in predicates
-for probabilistic execution and Bayesian inference over logic programs
-with stochastic choices.
-
-Doing so requires overcoming some technical barriers.
-
-First, many formal systems of probabilistic programming are so
-expressive as to make learning generally intractable. To address this,
-we use Range Concatenation Grammar, a context-sensitive grammar
-formalism originally developed in linguistics and extend this to a
-probabilistic model of programs.
-
-Second, the categories of syntax -- the nonterminals of the grammar --
-are often assumed to be known to children innately and given to
-automated learners by human experts. The categories that underlie
-conceptual knowledge, on the other hand, are far more numerous, vary
-from domain to domain, and are unlikely to be known to the
-learner. This motivates our use of latent predicates that, through
-learning, assume the role of a domain's underlying concepts (in the
+We use LPNs in an attempt to address the issues of tractibility and
+symbolic flexibility. We address the issue of tractability by
+extending Range Concatenation Grammars -- a context-sensitive grammar
+formalism originally developed in linguistic -- with a probabilistic
+interpretation, resulting in probabilistic programs in which
+inference, and thus learning, is relatively tractable. We address the
+issue of structural flexibility by building in latent predicates,
+which, through learning, assume the role of a domains underlying concepts (in the
 number domain, these might correspond to the concepts of
@@
 
-Finally, inducing probabilistic context-sensitive grammars with latent
-predicates threatens to be intractable: our goal is to find a middle
-ground between expressivity and tractability. Using
-PRISM~\cite{DBLP:journals/jair/SatoK01} -- a probabilistic logic
-programming system that naturally implements efficient dynamic
-programming algorithms for our models -- we are able to explore which
-domains and which grammar architectures are a good fit for this
-grammar-based approach.
+% Such concepts share many of the properties of language syntax: they
+% are unbounded in number, they derive their meanings via composition,
+% and, although people only ever say, hear, or read about a small number
+% of them, they are able to reason correctly about any them. There seems
+% to be a grammar to these concepts, and grasping this grammar is
+% critical to understanding their meanings and how to use them. This
+% motivates our approach here, which is to apply the tools of
+% probabilistic grammars more familiar from studies of syntax to the
+% problem of concept aquisition.
+
+% {\bf include somewhere:} A general probabilistic logic
+% programming system based on PROLOG, PRISM provides built-in predicates
+% for probabilistic execution and Bayesian inference over logic programs
+% with stochastic choices.
+
+% Doing so requires overcoming some technical barriers.
+
+% First, many formal systems of probabilistic programming are so
+% expressive as to make learning generally intractable. To address this,
+% we use Range Concatenation Grammar, a context-sensitive grammar
+% formalism originally developed in linguistics and extend this to a
+% probabilistic model of programs.
+
+% Second, the categories of syntax -- the nonterminals of the grammar --
+% are often assumed to be known to children innately and given to
+% automated learners by human experts. The categories that underlie
+% conceptual knowledge, on the other hand, are far more numerous, vary
+% from domain to domain, and are unlikely to be known to the
+% learner. This motivates our use of latent predicates that, through
+% learning, assume the role of a domain's underlying concepts (in the
+% number domain, these might correspond to the concepts of
+% successorship, order of magnitude, magnitude comparison, exact
+% vs. approximate, {\it etc}).
+
+% Finally, inducing probabilistic context-sensitive grammars with latent
+% predicates threatens to be intractable: our goal is to find a middle
+% ground between expressivity and tractability. Using
+% PRISM~\cite{DBLP:journals/jair/SatoK01} -- a probabilistic logic
+% programming system that naturally implements efficient dynamic
+% programming algorithms for our models -- we are able to explore which
+% domains and which grammar architectures are a good fit for this
+% grammar-based approach.
 
 The rest of this paper describes early work on this approach. First,
-we present Latent Predicate Networks (LPNs), our probabilistic model
-of concept learning. Then, we describe our approach to inference and
-its implementation in PRISM. Finally, we present preliminary
-experimental results in the domain of number concept learning that
-demonstrate the promise of this approach.
+we present Latent Predicate Networks (LPNs). Then, we describe our
+approach to inference and its implementation in PRISM. Finally, we
+present preliminary experimental results in the domain of number
+concept learning.
 
@@
 with derivations of finite length, so we need not worry about this
-requirement. See Figure \ref{fig:grammar} for an example of the context-sensitive 2-copy language $\{ww\,|w \in \{a,b\}^+\}$ as a PRCG.
+requirement. See Figure \ref{fig:grammar} for an example of the
+context-sensitive 2-copy language $\{ww\,|w \in \{a,b\}^+\}$ as a
+PRCG.
 

### 2014-11-10 drafts/nips_workshop/nips_workshop.tex
Final changes before NIPS workshop submission.

@@
 
-\author{Eyal Dechter\thanks{{\tt \{edechter, rule, jbt\} @mit.edu}}}
-\author{Joshua Rule}
-\author{Joshua B. Tenenbaum}
+\author{Eyal Dechter,  Joshua Rule,  Joshua B. Tenenbaum\thanks{{\tt \{edechter, rule, jbt\} @mit.edu}}}
 \affil{{\normalsize Department of Brain and Cognitive Sciences, MIT}}
@@
   Probabilistic programs successfully capture many aspects of the
-  complex symbolic and statistical knowledge brought to bear in human
-  cognition. Children aren't born with this knowledge, however, but
-  must instead learn it, often in an unsupervised fashion. Compelling
-  probabilistic programming models of human cognition must therefore
-  also be learned. We present early work on Latent Predicate Networks
-  (LPN), a framework for unsupervised learning in a restricted space
-  of context-sensitive probabilistic programs. Specifically, we use
-  Online Variational Bayesian Expectation Maximization in
-  PRISM, a probabilistic programming system, to perform inference over
-  the set of Stochastic Logic Programs (SLP) parseable in polynomial
-  time. Using a small fragment of English sentences about number, we
-  show that LPNs can recover the relevant conceptual structure
-  necessary to generate novel sentences in this domain with high
-  probability.
+  complex symbolic and statistical knowledge that human beings
+  possess. Children are not born with this knowledge and must learn
+  it, often in an unsupervised fashion. How can we flexibly and
+  tractably learn probabilistic programs? We present early work on
+  learning probabilistic programs with latent predicate networks
+  (LPNs), a framework for unsupervised learning in a restricted space
+  of probabilistic context-sensitive grammars. Specifically, we encode
+  LPNs in the probabilistic logic programming language PRISM and use
+  stochastic variational EM to infer its parameters. Using a small
+  fragment of English sentences about number, we show that LPNs can
+  recover much of the relevant conceptual structure necessary to
+  complete novel sentences in this domain with high probability.
 
@@
 
-Probabilistic programs combine symbolic and statistical knowledge in way that is
-natural for humans to express. In fact, probabilistic programs have
-been proposed as models of human cognition in a variety of domains
-[CITATION]. As a suitable representation for learning, however, the flexibility
-of probabilistic programs poses a challenge. At one extreme, we can
+Probabilistic programs combine symbolic and statistical knowledge in a
+way that is natural for humans to communicate and reason about. In
+fact, probabilistic programs have been proposed as models of human
+cognition in a variety of
+domains~\cite{DBLP:journals/cogsr/StuhlmullerG14, lake2012concept}. As
+a suitable representation for learning, however, the flexibility of
+probabilistic programs poses a challenge. At one extreme, we can
 approach learning as fixing the symbolic structure of a probabilistic
@@
 
-Here, we present statistical learning (using the PRISM probabilistic
-logic programmming language) of a restricted type of probabilistic
-program, dubbed a \emph{latent predicate network} (LPN), for learning
-problems at the intersection of concept learning and language
-acquisition. People learn many abstract concepts primarily through
+Here, we propose \emph{latent predicate networks} (LPNs) as such a
+middle course. LPNs maintain tractability by using Range Concatenation
+Grammars -- a context-sensitive grammar formalism originally developed
+in linguistics -- as the core knowledge representation. They
+maintain symbolic flexibility by building in latent predicates, which,
+through learning, come to represent underlying concepts in the
+learning domain.
+
+Learning problems at the intersection of concept learning and language
+acquisiton are good test-cases for LPNs. Such problems are important
+because people learn many abstract concepts primarily through
 language, and understanding language depends on understanding the
-underlying concepts. Number concepts are a good example: children do
-not learn about the meaning of ``seventy five'' by seeing examples of
-seventy five things; they do not know that ``seventy five'' is more
+underlying concepts. Number concepts are a good example this: children
+do not learn about the meaning of ``seventy five'' by seeing examples
+of seventy five things; they do not know that ``seventy five'' is more
 than ``twenty five'' because of their perceptual experiences of these
@@
 social relations ({\it e.g.}  ``my father in law's grandmother''),
-temporal relations (``the day after last Thanksgiving.''), and spatial
+temporal relations (``the day after last Thanksgiving''), and spatial
 relations (``above and just to the left of'').
 
-We use LPNs in an attempt to address the issues of tractibility and
-symbolic flexibility. We address the issue of tractability by
-extending Range Concatenation Grammars -- a context-sensitive grammar
-formalism originally developed in linguistic -- with a probabilistic
-interpretation, resulting in probabilistic programs in which
-inference, and thus learning, is relatively tractable. We address the
-issue of structural flexibility by building in latent predicates,
-which, through learning, assume the role of a domains underlying concepts (in the
-number domain, these might correspond to the concepts of
-successorship, order of magnitude, magnitude comparison, exact
-vs. approximate, {\it etc}.).
+The rest of this paper describes our early work on LPNs. First, we
+present LPNs. Then, we describe our approach to hierarchical Bayesian
+inference for LPNs and its implementation in PRISM, a probabilistic
+logic programmming
+language~\cite{DBLP:journals/jair/SatoK01}. Finally, we present
+preliminary experimental results in the domain of number concept
+learning.
 
-% Such concepts share many of the properties of language syntax: they
-% are unbounded in number, they derive their meanings via composition,
-% and, although people only ever say, hear, or read about a small number
-% of them, they are able to reason correctly about any them. There seems
-% to be a grammar to these concepts, and grasping this grammar is
-% critical to understanding their meanings and how to use them. This
-% motivates our approach here, which is to apply the tools of
-% probabilistic grammars more familiar from studies of syntax to the
-% problem of concept aquisition.
-
-% {\bf include somewhere:} A general probabilistic logic
-% programming system based on PROLOG, PRISM provides built-in predicates
-% for probabilistic execution and Bayesian inference over logic programs
-% with stochastic choices.
-
-% Doing so requires overcoming some technical barriers.
-
-% First, many formal systems of probabilistic programming are so
-% expressive as to make learning generally intractable. To address this,
-% we use Range Concatenation Grammar, a context-sensitive grammar
-% formalism originally developed in linguistics and extend this to a
-% probabilistic model of programs.
-
-% Second, the categories of syntax -- the nonterminals of the grammar --
-% are often assumed to be known to children innately and given to
-% automated learners by human experts. The categories that underlie
-% conceptual knowledge, on the other hand, are far more numerous, vary
-% from domain to domain, and are unlikely to be known to the
-% learner. This motivates our use of latent predicates that, through
-% learning, assume the role of a domain's underlying concepts (in the
-% number domain, these might correspond to the concepts of
-% successorship, order of magnitude, magnitude comparison, exact
-% vs. approximate, {\it etc}).
-
-% Finally, inducing probabilistic context-sensitive grammars with latent
-% predicates threatens to be intractable: our goal is to find a middle
-% ground between expressivity and tractability. Using
-% PRISM~\cite{DBLP:journals/jair/SatoK01} -- a probabilistic logic
-% programming system that naturally implements efficient dynamic
-% programming algorithms for our models -- we are able to explore which
-% domains and which grammar architectures are a good fit for this
-% grammar-based approach.
-
-The rest of this paper describes early work on this approach. First,
-we present Latent Predicate Networks (LPNs). Then, we describe our
-approach to inference and its implementation in PRISM. Finally, we
-present preliminary experimental results in the domain of number
-concept learning.
-
-\section{Latent Predicate Networks}
+\section{Latent predicate networks}
 
@@
     \includegraphics[width=\linewidth]{lpn/lpn.pdf}
-    \caption{The architecture and rules of a schematic LPN.}
+    \caption{}
     \label{fig:architecture}
@@
     \includegraphics[width=\linewidth]{parseTree/parse.pdf}
-    \caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+    \caption{}
     \label{fig:parseexample}
   \end{subfigure}
+  \caption{(\subref{fig:architecture}) The architecture and rules of a schematic single-layer LPN.
+(\subref{fig:parseexample}) A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
+
 \end{figure}
 
-An LPN is a hierarchical Bayesian model of strings
-extending the Hierarchical Dirichlet PCFG model (HD-PCFG) to
+An LPN is a hierarchical Bayesian model of strings extending the
+Hierarchical Dirichlet PCFG model~\cite{johnson2007bayesian} to
 Probabilistic Range Concatenation Grammars (PRCGs).
@@
 \psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
-the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
+the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)})$, where $A \in
 N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
@@
 well defined distribution as long as no probability mass is placed on
-derivations of infinite length; in this paper, we only consider PRCGs
+derivations of infinite length; here, we only consider PRCGs
 with derivations of finite length.
@@
 \subsection{Learning Model}
-
 \[ \vec{w}_{A_k} \sim Dir(\vec{\alpha}_{A_k})\hspace{3em}
@@
 
-  \vspace{-1em}
-Given a collection of predicates, $\{A_k\}_{k=1}^{K}$, and a
-distribution over clauses, $\{\vec{w}_{A_k}\}$, the learning task is
-to model a set of outputs, $\{x_j\}_{j=1}^{J}$, as being generated
-according to the above distribution. The weights of clauses
-with head predicate $A_k$ are drawn from a Dirichlet distribution
-defined by $\vec{\alpha}_{A_k}$. Each datum $x_j$ is then drawn
-from the resulting PRCG. The probability of a set of weights given our data and Dirichlet prior is determined with Bayes' Rule.
+  \vspace{-1em} Given a collection of predicates, $\{A_k\}_{k=1}^{K}$,
+  and a distribution over clauses, $\{\vec{w}_{A_k}\}$, the learning
+  task is to model a set of outputs, $\{x_j\}_{j=1}^{J}$, as being
+  generated according to the above distribution. The weights of
+  clauses with head predicate $A_k$ are drawn from a Dirichlet
+  distribution defined by $\vec{\alpha}_{A_k}$. Each datum $x_j$ is
+  then drawn from the resulting PRCG. The probability of a set of
+  weights given our data and Dirichlet prior is determined with Bayes'
+  rule.
 
@@
 
-LPNs can be encoded as a restricted subclass of PRISM programs; this
-is very similar to how PCFGs are encoded in
+LPNs can be encoded as a restricted subclass of PRISM programs in a
+very similar way to how PCFGs are encoded in
 PRISM~\cite{DBLP:conf/cl/2000}. Bayesian inference over stochastic
@@
 translating LPNs into PRISM programs and using its built-in
-Variational Bayes Expectation-Maximization
-algorithm~\cite{sato2008variational}. We modify it for use in an
-online setting, using the adaptive learning rate of
-\cite{ranganath2013adaptive}.
+Variational Bayes Expectation-Maximization (VBEM)
+algorithm~\cite{sato2008variational}.
 
-\section{Experiment}
+For large data sets, VBEM is slow and, as implemented in PRISM,
+too memory intensive to scale. This is because the parse forests of
+every dataset must either be evaluated on each iteration or they must
+be stored in memory between iterations.  Therefore, we also
+implemented a stochastic VBEM algorithm adapted for PRISM programs,
+following~\cite{ranganath2013adaptive,
+  DBLP:journals/jmlr/HoffmanBWP13}.
+
+\begin{figure}
+    \begin{subfigure}{0.45\linewidth}
+      \begin{tabular}{>{\footnotesize} l >{\footnotesize} l >{\footnotesize} l}
+        Question & $K=4$ \\ \hline
+        after twenty comes \ldots & twenty one & \checkmark \\
+        after forty five comes \ldots & forty six & \checkmark \\
+        after forty nine comes \ldots & forty ten & $\times$ \\
+        after fifty nine comes \ldots & fifty ten & $\times$ \\
+        after sixty one comes \ldots & sixty two & \checkmark \\
+        before twenty three comes \ldots & twenty two & \checkmark \\
+        before thirty comes \ldots & thirty eighty & $\times$ \\
+        before thirty eight comes \ldots & thirty seven & \checkmark \\
+        before forty one comes \ldots & forty & \checkmark \\
+        before fifty three comes \ldots & fifty two & \checkmark \\
+        before eighty five comes \ldots & eighty four & \checkmark \\
+        %% after forty seven comes \ldots & forty eight  \checkmark \\
+        %% after sixty three comes \ldots & sixty four \checkmark \\
+        %% after sixty four comes \ldots & sixty five \checkmark \\
+        %% after sixty five comes \ldots & sixty six \checkmark \\
+        %% after sixty nine comes \ldots & sixty ten $\times$ \\
+        %% after seventy three comes \ldots & seventy four \checkmark \\
+        %% after seventy nine comes \ldots & seventy ten $\times$ \\
+        %% after ninety five comes \ldots & ninety six \checkmark \\
+        %% before sixty eight comes \ldots & sixty seven \checkmark \\
+        %% before seventy two comes \ldots & seventy one \checkmark \\
+        %% before seventy three comes \ldots & seventy two \checkmark \\
+        %% before ninety two comes \ldots & ninety one \checkmark \\
+        %% before ninety three comes \ldots & ninety two \checkmark \\
+        %% before ninety five comes \ldots &ninety four \checkmark \\
+      \end{tabular}
+      \caption{}
+      \label{tab:results}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.50\linewidth}
+    \vspace{1em}
+    \includegraphics[width=\linewidth]{figures/train_number_net_0006_held_out.pdf}
+    \caption{}
+    \label{fig:heldoutLL}
+  \end{subfigure}
+  \caption{(\subref{tab:results}) Viterbi completions of held-out sentences for a 4-predicate 1-layer LPN.
+(\subref{fig:heldoutLL}) expected log likelihood on a set of held-out test data in a 5-layer LPN with 5 predicates per layer.}
+\end{figure}
+
+\section{Preliminary Experiments}
 
 To evaluate LPNs as a probabilistic model of concept acquisition, we
-trained an LPN with the architecture in Figure \ref{fig:architecture}
-and $K = 4$ latent predicates on a set of sentences expressing
-successor and predecessor relations in numbers between one and
-ninety-nine. The training set was the collection of sentences $$X =
-\{\text{after}\, \langle n \rangle \, \text{comes} \, \langle n+1
-\rangle \,|\, n \in 1,\dots,98\} \cup \{\text{before} \,\langle n
-\rangle \, \text{comes} \, \langle n-1 \rangle \,|\, n \in
-2,\dots,99\},$$ where $\langle n \rangle$ is the number word
+trained a single layer LPN with the architecture in Figure
+\ref{fig:architecture} and $K = 4$ latent predicates on a set of
+sentences expressing successor and predecessor relations in numbers
+between one and ninety-nine. The training set was the collection of
+sentences $X = \{\text{after}\, \langle n \rangle \, \text{comes} \,
+\langle n+1 \rangle \,|\, n \in 1,\dots,98\} \cup \{\text{before}
+\,\langle n \rangle \, \text{comes} \, \langle n-1 \rangle \,|\, n \in
+2,\dots,99\},$ where $\langle n \rangle$ is the number words
 corresponding to $n$. The lexicon was the set of number word
@@
 empty string, and the words ``before'', ``after'', and ``comes''. See
-Figure~\ref{fig:parseexample}, for a possible LPN derivation of an
-example sentence. To approximate the evidence children receive (words
-for small numbers naturally occuring more often than for larger ones,
-but children also rehearse the entire number sequence), we drew these
-sample sentences from a sum of a geometric distribution with parameter
-$0.5$ and a uniform distribution. These components were weighted
-$75\%$ and $25\%$, respectively. We drew $2000$ examples from this
-distribution, holding out a subset of sentences, including those in
-Table~\ref{tab:results} for evaluation.
+Figure~\ref{fig:parseexample} for a possible LPN derivation of an
+example sentence.
+%  To approximate the evidence children receive (words
+% for small numbers naturally occuring more often than for larger ones,
+% but children also rehearse the entire number sequence), we drew these
+% sample sentences from a sum of a geometric distribution with parameter
+% $0.5$ and a uniform distribution. These components were weighted
+% $75\%$ and $25\%$, respectively.
+We trained the LPN using $2000$ examples, holding out a subset of
+sentences -- including those in Table~\ref{tab:results} -- for evaluation.
 
-For inference, we used default $\frac{1}{D}$ pseudocounts (where $D$ is the
-dimensionality of the Dirichlet distributions). We found that different random
-initialization for this experiment did not lead to qualitatively different
-results, though further investigation will be necessary to see how
-robust the algorithm is to local maxima when fitting LPNs.
+% For inference, we used default $\frac{1}{D}$ pseudocounts (where $D$ is the
+% dimensionality of the Dirichlet distributions). We found that different random
+% initialization for this experiment did not lead to qualitatively different
+% results, though further investigation will be necessary to see how
+% robust the algorithm is to local maxima when fitting LPNs.
 
-We evaluated the learned model by asking for Viterbi ({\it i.e.} maximum {\it a
-posteriori}) completions of the last words of each held out test
-sentence. Table~\ref{tab:results} shows these completions. The grammar
-correctly learns much of the structure of these sentences, including
-the difference between sentences starting with ``before'' and
-``after'' and the edge cases that relate decade words like ``twenty''
-to non-decade words like ``twenty one.''
+We evaluated the learned model by asking for Viterbi ({\it i.e.}
+maximum {\it a posteriori}) completions of the last words of each held
+out test sentence. Table~\ref{tab:results} shows some of these
+completions. The grammar correctly learns much of the structure of
+these sentences, including the difference between sentences starting
+with ``before'' and ``after'' and the edge cases that relate decade
+words like ``twenty'' to non-decade words like ``twenty one.''
+
+Separately, we trained an LPN with five layers of five predicates each
+on a larger dataset including multiple forms of each sentence
+(\textit{e.g.} ``twenty comes before twenty one'' and ``before twenty one comes
+twenty''). Figure~\ref{fig:heldoutLL} shows the the expected log
+likelihood of held out data as a function of algorithm iteration. The
+expected log likelihood is a lower bound on the log expected
+probability. Its convergence to nearly zero suggests that the trained
+LPN discovers the underlying structure of the sentences.
+
+\section{Conclusion}
+LPNs allow for tractable learning over a restricted subset of
+probabilistic programs. Our preliminary results suggest that LPNs may
+enable learning at the intersection of concept learning and language
+acquisition, a domain for which an integration of statistical and
+symbolic knowledge is particularly promising.
 
@@
 
-\begin{figure}
-    \begin{subfigure}[b]{0.45\linewidth}
-      \begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
-        Question & $K=4$ \\ \hline
-        after twenty comes \ldots & twenty one \checkmark \\
-        after forty five comes \ldots & forty six \checkmark \\
-        after forty nine comes \ldots & forty ten $\times$ \\
-        after fifty nine comes \ldots & fifty ten $\times$ \\
-        after sixty one comes \ldots & sixty two \checkmark \\
-        before twenty three comes \ldots & twenty two \checkmark \\
-        before thirty comes \ldots & thirty eighty $\times$ \\
-        before thirty eight comes \ldots & thirty seven \checkmark \\
-        before forty one comes \ldots & forty \checkmark \\
-        before fifty three comes \ldots & fifty two \checkmark \\
-        before eighty five comes \ldots & eighty four \checkmark \\
-        %% after forty seven comes \ldots & forty eight  \checkmark \\
-        %% after sixty three comes \ldots & sixty four \checkmark \\
-        %% after sixty four comes \ldots & sixty five \checkmark \\
-        %% after sixty five comes \ldots & sixty six \checkmark \\
-        %% after sixty nine comes \ldots & sixty ten $\times$ \\
-        %% after seventy three comes \ldots & seventy four \checkmark \\
-        %% after seventy nine comes \ldots & seventy ten $\times$ \\
-        %% after ninety five comes \ldots & ninety six \checkmark \\
-        %% before sixty eight comes \ldots & sixty seven \checkmark \\
-        %% before seventy two comes \ldots & seventy one \checkmark \\
-        %% before seventy three comes \ldots & seventy two \checkmark \\
-        %% before ninety two comes \ldots & ninety one \checkmark \\
-        %% before ninety three comes \ldots & ninety two \checkmark \\
-        %% before ninety five comes \ldots &ninety four \checkmark \\
-      \end{tabular}
-      \caption{}
-      \label{tab:results}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{figures/train_number_net_0006_held_out.pdf}
-    \caption{}
-    \label{fig:heldoutLL}
-  \end{subfigure}
-  \caption{(\subref{tab:results}) Viterbi completions of held-out sentences for a 4-predicate 1-layer LPN.
-(\subref{fig:heldoutLL}) expected log likelihood on a set of held-out test data in a 35-predicate 5-layer LPN with.}
-\end{figure}
+% To inspect visually the learned grammar, we thresholded rules
+% according to the expected number of times they were used in parsing
+% the training dataset. Table~\ref{tab:grammar} shows all rules with
+% expected count above $1e-6$. This reduces from $2669$ to $52$ the
+% number of significant rules. On inspection, predicate $A_2$ forms
+% ``before'' sentences, predicate $A_4$ forms ``after'' sentences,
+% predicate $A_3$ is successorship recursively defined over the decades
+% and ones, and predicate $A_2$ is a category for the decade words.
 
-To inspect visually the learned grammar, we thresholded rules
-according to the expected number of times they were used in parsing
-the training dataset. Table~\ref{tab:grammar} shows all rules with
-expected count above $1e-6$. This reduces from $2669$ to $52$ the
-number of significant rules. On inspection, predicate $A_2$ forms
-``before'' sentences, predicate $A_4$ forms ``after'' sentences,
-predicate $A_3$ is successorship recursively defined over the decades
-and ones, and predicate $A_2$ is a category for the decade words.
+% Our LPN does not learn to how to transition between the last word in a decade
+% and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
+% makes the intuitively reasonable generalization that ``seventy nine''
+% should be followed by ``seventy ten.''
 
-Our LPN does not learn to how to transition between the last word in a decade
-and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
-makes the intuitively reasonable generalization that ``seventy nine''
-should be followed by ``seventy ten.''
-
-\subsubsection*{Acknowledgments?}
-
-\bibliographystyle{amsplain}
+\bibliographystyle{plain}
 \bibliography{nips_workshop}
 
+% \nocite{*}
+
 \end{document}

### 2014-11-10 experiments/prism/train_number_net_0006.held_out.py

@@ -10,7 +10,7 @@ plt.rcParams['text.latex.preamble']=\
                 [r"\usepackage{amssymb, amsmath}"]
 
 plt.rc('text', usetex=True)
-plt.rc('font', family='serif')
+plt.rc('font', **{'family': 'serif', 'size': 16})
 
 data = np.loadtxt("train_number_net_0006.held_out.csv", delimiter=",")
 data.sort(axis=0)

### 2014-11-10 experiments/prism/NumberDist.py
Cleaning up directories.

@@ -1,23 +1,0 @@
-
-import numpy as np
-
-frac_first_19=0.75
-total_count = 4000
-exponential_decay_constant=1/2.
-
-if __name__=="__main__":
-    xs = np.random.geometric(p=exponential_decay_constant, 
-                             size=np.floor(total_count*frac_first_19))
-    xs = xs[xs<100]
-    xs = xs[xs > 0]
-    
-    ys = np.random.choice(range(1, 100), size=np.floor(total_count-xs.size))
-
-    zs = np.concatenate([xs, ys])
-    
-    (cnts, bins) = np.histogram(zs, range(1, 101))
-    print cnts, bins
-    
-    
-
-

### 2014-11-10 experiments/prism/main.pl

@@ -1,22 +1,0 @@
-main(Gs,Foutp,Fouta) :-
-    set_prism_flag(restart,1),
-    set_prism_flag(learn_mode,vb),
-    set_prism_flag(viterbi_mode,vb),
-    set_prism_flag(default_sw_a,uniform),
-    set_prism_flag(log_scale,on),
-    learn(Gs),
-    save_sw(Foutp),
-    save_sw_a(Fouta).
-
-getTrain(F,Gs) :-
-    load_clauses(F,ALL,[]),
-    findall(X,member(train(X),ALL),Gs).
-
-getTest(F,Gs) :-
-    load_clauses(F,ALL,[]),
-    findall(X,member(test(X),ALL),Gs).
-
-# srs(P-IN) :-
-#     msw(P,V),
-#     reduce(P-IN,V).
-

### 2014-11-10 experiments/prism/numbers.pl

@@ -1,99 +1,0 @@
-number_word(1, [one]).
-number_word(2, [two]).
-number_word(3, [three]).
-number_word(4, [four]).
-number_word(5, [five]).
-number_word(6, [six]).
-number_word(7, [seven]).
-number_word(8, [eight]).
-number_word(9, [nine]).
-number_word(10, [ten]).
-number_word(11, [eleven]).
-number_word(12, [twelve]).
-number_word(13, [thirteen]).
-number_word(14, [fourteen]).
-number_word(15, [fifteen]).
-number_word(16, [sixteen]).
-number_word(17, [seventeen]).
-number_word(18, [eighteen]).
-number_word(19, [nineteen]).
-number_word(20, [twenty]).
-number_word(21, [twenty,one]).
-number_word(22, [twenty,two]).
-number_word(23, [twenty,three]).
-number_word(24, [twenty,four]).
-number_word(25, [twenty,five]).
-number_word(26, [twenty,six]).
-number_word(27, [twenty,seven]).
-number_word(28, [twenty,eight]).
-number_word(29, [twenty,nine]).
-number_word(30, [thirty]).
-number_word(31, [thirty,one]).
-number_word(32, [thirty,two]).
-number_word(33, [thirty,three]).
-number_word(34, [thirty,four]).
-number_word(35, [thirty,five]).
-number_word(36, [thirty,six]).
-number_word(37, [thirty,seven]).
-number_word(38, [thirty,eight]).
-number_word(39, [thirty,nine]).
-number_word(40, [forty]).
-number_word(41, [forty,one]).
-number_word(42, [forty,two]).
-number_word(43, [forty,three]).
-number_word(44, [forty,four]).
-number_word(45, [forty,five]).
-number_word(46, [forty,six]).
-number_word(47, [forty,seven]).
-number_word(48, [forty,eight]).
-number_word(49, [forty,nine]).
-number_word(50, [fifty]).
-number_word(51, [fifty,one]).
-number_word(52, [fifty,two]).
-number_word(53, [fifty,three]).
-number_word(54, [fifty,four]).
-number_word(55, [fifty,five]).
-number_word(56, [fifty,six]).
-number_word(57, [fifty,seven]).
-number_word(58, [fifty,eight]).
-number_word(59, [fifty,nine]).
-number_word(60, [sixty]).
-number_word(61, [sixty,one]).
-number_word(62, [sixty,two]).
-number_word(63, [sixty,three]).
[…]

### 2014-11-10 experiments/prism/onlineVBEM.pl

@@ -1,90 +1,0 @@
-
-ovbem(NIter, BatchSize, DataSet, Eps) :- 
-    ovbem(NIter, BatchSize, DataSet, 1, Eps). 
-
-ovbem(NIter, BatchSize, DataSet, Iter, Eps) :- Iter > NIter, !.
-ovbem(NIter, BatchSize, DataSet, Iter, Eps) :- 
-    Tau0 is 0,
-    Kappa is 0.8,
-    StepSize is (Tau0 + Iter) ** (-Kappa), 
-    ovbem1(DataSet, BatchSize, StepSize),
-    save_ovbem_alphas(Iter),
-    get_delta_ovbem_alphas(Iter, Delta),
-    write('Iter: '), write(Iter), nl,
-    write('StepSize: '), write(StepSize), nl,
-    write('Normed Delta Alpha: '), write(Delta), nl,
-    ((Iter>1, Delta < Eps) -> true; 
-     (Iter1 is Iter+1, 
-      ovbem(NIter, BatchSize, DataSet, Iter1, Eps))).
-
-ovbem1(DataSet, BatchSize, StepSize) :- 
-    length(DataSet, N),
-    % get current alphas
-    findall(Info, 
-            get_sw_a(Info), 
-            ParamsInit), 
-    % sample a data point from the dataset
-    random_multiselect(DataSet, BatchSize, Batch),
-    M is round(N/BatchSize),
-    BatchOut @= [D1: D in Batch, [D1], multiply_datum_count(D, M, D1)],
-    % run vbem
-    learn(BatchOut), 
-    % get output alphas
-    findall(Info, 
-            get_sw_a(Info), 
-            ParamsFinal),
-    % update parameters towards ParamsFinal
-    foreach(switch(Sw, Status, Vals, As0) in ParamsInit, [As0, As1, AsOut],
-            (member(switch(Sw, Status, Vals, As1), ParamsFinal), !,
-             update_vb_params(StepSize, As0, As1, AsOut), 
-             set_sw_a(Sw, AsOut))). 
-
-update_vb_params(_, [], [], []).
-update_vb_params(StepSize, [A0|As0], [A1|As1], [AOut|AsOut]) :-
-    AOut is (1-StepSize)*A0 + StepSize*A1, 
-    update_vb_params(StepSize, As0, As1, AsOut).
-
[…]

### 2014-11-10 experiments/prism/prune.pl

@@ -1,35 +1,0 @@
-%% take a psm srs system and prune out rules
-
-write_sw_vs(Sw, Vs, Clause) :- 
-    Clause=values(Sw, Vs).
-
-write_sw_ps(Sw, Ps, Clause) :- 
-    Clause = (:- set_sw(Sw, Ps)).
-
-write_sw_as(Sw, As, Clause) :- 
-    Clause = (:- set_sw_a(Sw, As)). 
-
-write_sw(Sw, [Vs, Ps, As], Clauses) :- 
-    write_sw_vs(Sw, Vs, C1), 
-    write_sw_ps(Sw, Ps, C2), 
-    write_sw_as(Sw, As, C3), 
-    Clauses = [C1, C2, C3].
-
-delete_switches([], []). 
-delete_switches([C|CsIn], CsOut) :- 
-    C = values(_, _), !, delete_switches(CsIn, CsOut).
-delete_switches([C|CsIn], [C|CsOut]) :- 
-    C = values(_, _), !, delete_switches(CsIn, CsOut).
-
-replace_switches(ClausesIn, Switches, ClausesOut) :- 
-    delete_switches(ClausesIn, Clauses1), 
-    NewSwitches @= [Clauses : [Sw, Vs, Ps, As] in Switches, 
-                    [Clauses],
-                    write_sw(Sw, [Vs, Ps, As], Clauses)], 
-    append(NewSwitches, Clauses1, ClausesOut). 
-    
-    
-
- 
-    
-

### 2014-11-10 experiments/prism/run_parallel.sh

@@ -1,29 +1,0 @@
-#!/bin/bash
-
-NumExamples=(1700);
-NumReps=1;
-NumPred=(4);
-SaveDir="data/saturday_"
-args=()
-
-for nEx in ${NumExamples[*]}
-do
-    for rep in $(eval echo {1..$NumReps})
-    do
-        for nPred in ${NumPred[*]}
-        do
-            args+=("$nEx")
-            args+=("$rep")
-            args+=("$nPred")
-        done
-    done
-done
-
-CMD="parallel -n 3 --noswap --resume-failed --joblog ./parallel_log --delay 10 upprism srs {} $SaveDir &> logs/overnight/saturday_log_{1}_{2}_{3}.log ::: ${args[*]}"
-echo $CMD
-`$CMD`
-
-
-
-
-

### 2014-11-10 experiments/prism/train_number_net_0001.pl

@@ -1,38 +1,0 @@
-
-:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
-   cl(P).
-
-%% load .sys.psm file.
-:- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
-   prism([load], P).
-
-%% load data set
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest.pl', P),
-   cl(P).
-
-:- set_training_flags.
-
-%% ovbem config
-:- set_ovbem_rec(niter, 1000), 
-   set_ovbem_rec(batch_size, 1), 
-   set_ovbem_rec(iter_between, 5).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest.log', P),
-   set_ovbem_rec(logfile, P).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest.data', P),
-   set_ovbem_rec(outfile, P).
-   
-:- findall(X, train(_, X), DataSet), 
-   set_ovbem_rec(data_set, DataSet).
-
-:- findall(X, test(_, X), HeldOutAll), 
-   take(10, HeldOutAll, HeldOut), 
-   write(HeldOut), nl, 
-   set_ovbem_rec(held_out, HeldOut). 
-
-%% run training
-main :- run_ovbem.
-
-
-    

### 2014-11-10 experiments/prism/train_number_net_0002.pl

@@ -1,40 +1,0 @@
-
-:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
-   cl(P).
-
-%% load .sys.psm file.
-:- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
-   prism([load], P).
-
-%% load data set
-%% this includes before and after sentences with 20 held out test sentences
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
-   cl(P).
-
-:- set_training_flags.
-
-%% ovbem config
-:- set_ovbem_rec(niter, 1000), 
-   set_ovbem_rec(batch_size, 1), 
-   set_ovbem_rec(iter_between, 5).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0002.log', P),
-   set_ovbem_rec(logfile, P).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0002.data', P),
-   set_ovbem_rec(outdir, P), 
-   initialize_outdir.
-   
-:- findall(X, train(_, X), DataSet), 
-   set_ovbem_rec(data_set, DataSet).
-
-:- findall(X, test(_, X), HeldOutAll), 
-   take(1, HeldOutAll, HeldOut), 
-   write(HeldOut), nl, 
-   set_ovbem_rec(held_out, HeldOut). 
-
-%% run training
-main :- run_ovbem.
-
-
-    

### 2014-11-10 experiments/prism/train_number_net_0003.pl

@@ -1,42 +1,0 @@
-
-:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
-   cl(P).
-
-%% load .sys.psm file.
-:- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
-   prism([load], P),
-   set_ovbem_rec(sys_psm_file, P).
-
-%% load data set
-%% this includes before and after sentences with 20 held out test sentences
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
-   cl(P).
-
-:- set_training_flags.
-
-%% ovbem config
-:- set_ovbem_rec(niter, 1000), 
-   set_ovbem_rec(batch_size, 1), 
-   set_ovbem_rec(iter_between, 5).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0003.log', P),
-   set_ovbem_rec(logfile, P).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0003.data', P),
-   set_ovbem_rec(outdir, P), 
-   initialize_outdir.
-   
-:- findall(X, train(_, X), DataSet), 
-   set_ovbem_rec(data_set, DataSet).
-
-:- findall(X, test(_, X), HeldOutAll), 
-   take(10, HeldOutAll, HeldOut), 
-   write(HeldOut), nl, 
-   set_ovbem_rec(held_out, HeldOut). 
-
-%% run training
-main :- 
-    run_ovbem.
-
-
-    

### 2014-11-10 experiments/prism/train_number_net_0004.pl

@@ -1,37 +1,0 @@
-
-:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
-   cl(P).
-
-%% load .sys.psm file.
-:- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
-   prism([load], P),
-   set_ovbem_rec(sys_psm_file, P).
-
-%% load data set
-%% this includes before and after sentences with 20 held out test sentences
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
-   cl(P).
-
-:- set_training_flags.
-
-%% ovbem config
-:- set_ovbem_rec(niter, 100), 
-   set_ovbem_rec(batch_size, 10), 
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0004.log', P),
-   set_ovbem_rec(logfile, P).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0004.data', P),
-   set_ovbem_rec(outdir, P), 
-   initialize_outdir.
-   
-:- findall(count(X, 100), train(_, X), DataSet), 
-   set_ovbem_rec(data_set, DataSet).
-
-%% run training
-main :- 
-    set_training_flags,
-    run_ovbem.
-
-
-    

### 2014-11-10 experiments/prism/train_number_net_0005.pl

@@ -1,38 +1,0 @@
-
-:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
-   cl(P).
-
-%% load .sys.psm file.
-load_sys_psm :- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
-                prism([load], P),
-                set_ovbem_rec(sys_psm_file, P).
-
-%% load data set
-%% this includes before and after sentences with 20 held out test sentences
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
-   cl(P).
-
-:- set_training_flags.
-
-%% ovbem config
-:- set_ovbem_rec(niter, 100), 
-   set_ovbem_rec(batch_size, 10).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0005.log', P),
-   set_ovbem_rec(logfile, P).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0005.data', P),
-   set_ovbem_rec(outdir, P).
-   
-:- findall(count(X, 100), train(_, X), DataSet), 
-   set_ovbem_rec(data_set, DataSet).
-
-%% run training
-main :- 
-    initialize_outdir,
-    load_sys_psm, 
-    set_training_flags,
-    run_ovbem.
-
-
-    

### 2014-11-10 experiments/prism/train_number_net_0006.held_out.py

@@ -1,36 +1,0 @@
-import numpy as np
-from matplotlib import pyplot as plt
-
-from matplotlib import rc
-# rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
-## for Palatino and other serif fonts use:
-#rc('font',**{'family':'serif','serif':['Palatino']})
-rc('text', usetex=True)
-plt.rcParams['text.latex.preamble']=\
-                [r"\usepackage{amssymb, amsmath}"]
-
-plt.rc('text', usetex=True)
-plt.rc('font', **{'family': 'serif', 'size': 16})
-
-data = np.loadtxt("train_number_net_0006.held_out.csv", delimiter=",")
-data.sort(axis=0)
-data = data[1:, 1]
-
-fig = plt.figure()
-ax = fig.add_subplot(111)
-ax.plot(range(1, len(data) + 1), 
-        data, 
-        linewidth=2, 
-        marker='o')
-ax.set_xlabel("Iteration")
-ax.set_ylabel(r"$\mathbb{E} \log p(D_{\text{test}})$")
-ax.set_ylim(top=1000)
-ax.set_xlim(right=20)
-
-# line at 0
-[lo, hi] = ax.get_xlim()
-ax.plot([lo, hi], [0, 0], 'k--')
-
-fig.savefig("train_number_net_0006.held_out.pdf")
-
-plt.show()

### 2014-11-10 experiments/prism/train_number_net_0006.pl

@@ -1,40 +1,0 @@
-
-:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
-   cl(P).
-
-%% load .sys.psm file.
-load_sys_psm :- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_5_5_5.sys.psm', 
-                                   P),
-                prism([load], P),
-                set_ovbem_rec(sys_psm_file, P).
-
-%% load data set
-%% this includes before and after sentences with 20 held out test sentences
-:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
-   cl(P).
-
-:- set_training_flags.
-
-%% ovbem config
-:- set_ovbem_rec(niter, 100), 
-   set_ovbem_rec(batch_size, 10).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0007.log', P),
-   set_ovbem_rec(logfile, P).
-
-:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0007.data', P),
-   set_ovbem_rec(outdir, P).
-   
-:- findall(count(X, 1), train(_, X), DataSet), 
-   set_ovbem_rec(data_set, DataSet).
-
-%% run training
-main :- 
-    initialize_outdir,
-    load_sys_psm, 
-    set_training_flags,
-    run_ovbem.
-
-
-    
-          

### 2014-11-10 experiments/prism/vis.pl

@@ -1,139 +1,0 @@
-%% %% Print a Parse Tree %% %%
-
-srsTree(X,[Olhs,Orule,Orhs]) :-
-    select(msw(P,V),X,X2),
-    select(srs(P-IN),X2,X3),
-    delete(X3,reduce(_,_),X4),
-    maplist(srsTree,X4,Orhs),
-    prismToLPN(P,V,Orule),
-    predicateToString(P-IN,Olhs).
-
-%% hack! This currently only works up to 9-ary predicates
-predicateToString(P-IN,S) :-
-    sub_atom(P,0,_,2,P2),
-    atom_lists_to_atoms(IN,A2),
-    atom_concat(P2,'(',S1),
-    atom_concat(A2,')',S2),
-    atom_concat(S1,S2,S).
-
-atom_lists_to_atoms([],'').
-atom_lists_to_atoms([X],O) :-
-    recursive_atom_concat_space(X,O).
-
-atom_lists_to_atoms([X|XS],O) :-
-    length(XS,XL), XL > 0,
-    atom_lists_to_atoms(XS,OS),
-    recursive_atom_concat_space(X,OS2),
-    atom_concat(OS2,', ',O1),
-    atom_concat(O1,OS,O).
-
-recursive_atom_concat([],'').
-recursive_atom_concat([X],X).
-recursive_atom_concat([X|XS],O) :-
-    length(XS,XL), XL > 0,
-    recursive_atom_concat(XS,OS),
-    atom_concat(X,OS,O).
-
-recursive_atom_concat_space([],'').
-recursive_atom_concat_space([X],X).
-recursive_atom_concat_space([X|XS],O) :-
-    length(XS,XL),
-    XL > 0,
-    recursive_atom_concat_space(XS,OS),
-    atom_concat(X,' ',O1),
-    atom_concat(O1,OS,O).
-
-formatSRSTree(X) :- format("~n",[]), formatSRSTree(X,'').
-formatSRSTree([LHS,RULE,RHS],SEP) :-
-    format("~a~a~n~a~a~n",[SEP,LHS,SEP,RULE]),
-    atom_concat(SEP,'    ',NEWSEP),
-    findall(X,(member(X,RHS),formatSRSTree(X,NEWSEP)),_).
-
-%% %% Print The Grammar %% %%
-
[…]

### 2014-11-10 prolog/time.pl

@@ -1,0 +1,15 @@
+
+nAs(1, [a]). 
+nAs(N, [a|Xs]) :- 
+    N1 is N-1,
+    nAs(N1, Xs).
+
+go(Times) :- 
+   Times @= [T:N in 1..15, 
+             [Start, X, End, T],
+             (cputime(Start), 
+              nAs(N, X), 
+              viterbi(prove('S_1'-[X])), 
+              cputime(End), 
+              T is (End-Start)/1000)].
+              

### 2014-11-10 drafts/Number.tex
Added some untracked files."

--- 2014-08-15 drafts/ApproximatePCFG/ApproximatePCFG.tex
@@
 
+\usepackage{graphicx}
+
+\usepackage{natbib}
+
+\makeatletter
+\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}
+\makeatother
+
+\newcommand{\derives}{\xRightarrow{*}}
+
+\title{Understanding number: a case study in learning systems of concepts}
 \author{Eyal Dechter}
@@
 \begin{document}
-The Inside-Outside algorithm is a popular algorithm for computing the
-MAP marginals of PCFG rules given a sentence or collection of
-sentences. The algorithm's computational complexity is $O(n^3 K)$
-where $n$ is the length of the sentence and $K$ is the the number of
-rules in the PCFG. The cubic algorithm makes inference tractable for
-relatively long sentences -- containing perhaps a few dozen terminals
--- but makes it impossible for this algorithm to be used in an online
-fashion, or to be used on grammars that span multiple levels of
-linguistic abstration (for example, imagine trying to learn
-morphological and syntactic grammar at the same time with the
-Inside-Outside algorithm).
+\maketitle
+\date
 
-Our goal here is to develop an algorithm that approximates the MAP
-marginals of a PCFG given data but that can run much faster; further,
-we would like to be able to put bounds on the approximation. The core
-intuition is that if we know that the probability of some symbol
-dominating a span cannot be more than some very small amount, we
-shouldn't have to calculate that amount accurately; instead, we should
-be able to substitute this small upperbound in our calculations,
-absolving us of exploring parts of the space of parses that are very
-unlikely to be relevant.
+\section{Abstract}
+How do people learn, represent, and reason about systems of inter-related concepts?
 
-In what follows we will assume that we have a PCFG $G$ in Chomsky
-normal form with start symbol $S$, non-terminals $\mathcal{A}$,
-terminals $\mathcal{T}$, and rules $\mathcal{R}$, where we associate
-with each rule $r \in \mathcal{R}$ a potential $\psi(r)$. We also
-assume that we are given a sentence $x = x_1, \dots, x_N$.
-
-For every triple $(A, i, j)$ where $A$ is a nonterminal and $1 \leq i
-\leq j \leq N$, the Inside-Outside algorithm computes an alpha and
-beta value. The alpha and beta values are defined according to the
-following recurrence:
-
-\begin{align}
-\alpha(A, i, i) &= \psi(A \rightarrow x_i)\\
-\alpha(A, i, j) &= \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
-  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)\\
-\beta(A, 1, N) &= 1 \\
-\beta(A, i, j) &= \sum_{k=1}^{i-1} \sum_{B \rightarrow C A \in \mathcal{R}}
-  \beta(B, k, j) \psi(B \rightarrow C A) \alpha(C, k, i-1)\\ 
-  &+  \sum_{k=j+1}^{N} \sum_{B \rightarrow A C \in \mathcal{R} : A \neq C}
-  \beta(B, i, k) \psi(B \rightarrow A C) \alpha(C, j+1, k) 
-\end{align}
-
-Suppose we have an approximate $\alpha'$ of $\alpha$ with the
-property that for some small $\epsilon > 0$ and for all $A, i, j$:
-
-\begin{align}
-\alpha'(A, i, i) &= \alpha(A, i, i)\\ 
-\alpha(A, i, j) &\leq \alpha'(A, i, j) \leq (1+\epsilon)  \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
-  \psi(A \rightarrow B C ) \alpha'(B, i, k) \alpha'(C, k+1, j).\\
-\end{align}
-
-By how much can $\alpha'$ over estimate $\alpha$? Let $E(A, i,
-j)=\frac{\alpha'(A, i, j)}{\alpha(A, i, j)}$ be the measure of this
-over-estimate. We have that
-
-\begin{align}
-E_{\alpha}(A, i, i) &= 1\\
-E_{\alpha}(A, i,j) &= \frac{\alpha'(A, i, j)}{\alpha(A, i, j)}\\
-          &\leq (1+\epsilon) \frac{ \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
-  \psi(A \rightarrow B C ) \alpha'(B, i, k) \alpha'(C, k+1, j)}
-           {\sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
-  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)}\\
-          &=  (1+\epsilon) \frac{ \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
-  \psi(A \rightarrow B C ) E_{\alpha}(B, i, k) \alpha(B, i, k) E_{\alpha}(C, k+1, j) \alpha(C, k+1, j)}
-           {\sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
-  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)}\\
-         &= (1+\epsilon) \max_{k\in \{i, \dots, j-1\}, A \rightarrow B C \in \mathcal{R} }
-            E_{\alpha}(B, i, k) E_{\alpha}(C, k+1, j).
-\end{align}
-
-Let $\hat{E}_{\alpha}(d) = \max_{A\in\mathcal{A}, i \in \{1, \dots, N-d \}}
-E_{\alpha}(A, i, i+d)$ be the maximum upper bound above over all spans of
-length $d$. Then we have that
-
-\begin{align}
-\hat{E}_{\alpha}(1) &= 1\\
-\hat{E}_{\alpha}(d) &= (1+\epsilon) \max_{s \in \{1, \dots, d-1\}} \hat{E}_{\alpha}(s) \hat{E}_{\alpha}(d-s).\\
-\end{align}
-
-Solving this recurrence, gives $\hat{E}_{\alpha}(d) = (1 + \epsilon)^{d-1}$. To
-re-iterate, what this means is that if we have such an $\alpha'$ then
-for all $A, i, j$, $\alpha'(A, i, j) \leq (1 + \epsilon)^{j - i}
-\alpha(A, i, j)$.
-
-We can now develop an analog $\beta'$ of $\beta$ which has the same
-dependency on $\alpha'$ that $\beta$ has on $\alpha$. Let
-$\hat{E}_{\beta}(d)$ be an upper bound on $\frac{\beta'(A, i,
-  i+d)}{\beta(A, i, i+d)}$ for all $A$ and $i$. Then a similar
-analysis as above shows that one such $\hat{E}_\beta$ is
-
-\begin{align}
-\hat{E}_{\beta}(N) &= 1\\
-\hat{E}_{\beta}(N-k) &= (1+\epsilon)^{N-k-1} \max_{s \in \{N-k+1, \dots, N\}} \hat{E}_{\beta}(s).
-\end{align}
-
-Solving this recurrence yields $\hat{E}_{\beta}(d) = (1+\epsilon)^{\frac{(N-d)(N+d-1)}{2}}$.
-
-In the Inside-Outside algorithm, the merit $\mu(A, i, j) = \alpha(A,
-i, j) \beta(A, i, j)$ is the unnormalized probability that nonterminal
-$A$ yields span $(i, j)$. Define $\mu'(A, i, j) = \alpha'(A, i, j)
-\beta(A, i, j)$. By the results above, we have that the ratio of the
-actual merit to our approximation is upper bounded by
-$(1+\epsilon)^{\frac{(N+d-2)(N-d+1)}{2}}$, where, as before, $d$ is
-the length of the span $(i, j)$.
-
-\section{Approximate Inside}
-Here we describe the algorithm ApproxInside which computes the
-$\alpha'$ values.
-
-\begin{algorithm}
-\caption{The Approximate Inside Algorithm \label{alg:approxInside}}
-\begin{algorithmic}[1]
-\Procedure {ApproxInside}{$s$, $G=(S, bRules, uRules)$, $U$, $\epsilon$}
-  \State \Comment $s$ sentences of length $n$
-  \State \Comment $G$ grammar 
-  \State \Comment $S$ start symbol
-  \State \Comment $U$ upper bound, function from $(A, i, j) \mapsto \mathbf{R}$
-  \State $\alpha \leftarrow \Call{InitAlpha}{G, 1, n}$ \Comment Initialize table of $\alpha$ 
-  \State \Return \Call{$GO_{OR}$}{$S$, $1$, $n$}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{The Approximate Inside Algorithm (cont)}
-\begin{algorithmic}[2]
-\Procedure {$GO_{OR}$}{$A$, $i$, $j$}
-  \State \Comment $A$ nonterminal symbol
-  \State $children \leftarrow \Call{GetChildren}{A, i, k}$
-  \State $lb \leftarrow 0 $ \Comment Initialize lower bound on OR node to 0
-  \For{$node \in children$}{}{}
-    \If{$node$ is unary of form $A \rightarrow t, i, i$ and $t = s_i$}
-      \State $v \leftarrow \psi(A \rightarrow t)$
-    \ElsIf{$node$ is binary}
-      \State $v \leftarrow $\;\Call{$GO_{AND}$}{$node$, $lb$}
-    \EndIf
-    \State $lb \leftarrow lb + v$ \Comment Update lower bound
-  \EndFor
-  \State $\alpha[A, i, j] \leftarrow lb$ \Comment Insert $\alpha$ value into table.
-\EndProcedure
-\Statex
-\Procedure {$GO_{AND}$}{$A \rightarrow B\, C$, $i$, $k$,$j$, $lb$}
-  \State $w \leftarrow \psi(A \rightarrow B\, C)$ 
-  \State $v_{\ell} \leftarrow \alpha[B, i, k]$ 
-  \State $v_{r} \leftarrow \alpha[C, k+1, j]$
-  \State $u_{\ell} \leftarrow U[B, i, k]$ 
-  \State $u_{r} \leftarrow U[C, k+1, j]$
-  \If{neither $v_{\ell}$ nor $v_r$ is $NIL$} \Comment If both values are already computed, just use them
-    \State \Return $w * v_{\ell} * v_r$
-  \ElsIf{ $v_r \neq NIL$}  \Comment If the right value is already computed, compute left value
-    \If {$u_\ell \leq \frac{\epsilon * lb}{w * v_r}$}
-      \State $v_\ell \leftarrow u_\ell$
-    \Else
-      \State $v_\ell \leftarrow $\;\Call{$GO_{OR}$}{$B$, $i$, $k$}
-    \EndIf
-  \ElsIf{ $v_\ell \neq NIL$}  \Comment If the left value is already computed, compute right value
-    \If {$u_r \leq \frac{\epsilon * lb}{w * v_\ell}$}
-      \State $v_r \leftarrow u_r$
-    \Else
-      \State $v_r \leftarrow $\;\Call{$GO_{OR}$}{$C$, $k+1$, $j$}
-    \EndIf
-  \Else \Comment If neither values yet computed, compute both
-    \If {$u_\ell u_r \leq \frac{\epsilon * lb}{w}$}
-      \State $v_r \leftarrow u_r$
-      \State $v_\ell \leftarrow u_\ell$
-    \Else
-      \State $v_\ell \leftarrow $\;\Call{$GO_{OR}$}{$B$, $i$, $k$}
-      \State $v_r \leftarrow $\;\Call{$GO_{OR}$}{$C$, $k+1$, $j$}
-    \EndIf
-  \EndIf
-  \State \Return $w * v_{\ell} * v_r$
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{The Approximate Outside Algorithm.\label{alg:approxOutside}}
-\begin{algorithmic}[2]
-\Procedure {$ApproxOutside$}{$s$, $G$, $S$}
-  \State $n \leftarrow length(s)$
-  \State $\beta[S, 1, n] \leftarrow 1$ \Comment initialize $\beta$ recursion, unless $n==1$ in which case we are done
-  \State $open \leftarrow [(S, 1, n)]$ \Comment initialize list that will hold nodes to be processed
-  \While {$open \neq NIL$}
-  \State $children \leftarrow []$ \Comment initialize list that holds next level of nodes to be processed
-   \For {$(A, i, j) \in open$ \Comment $i \neq j$}
-    \For {$A \leftarrow B\, C \in rules(G)$} \Comment iterate over binary rules headed by $A$
-      \For {$k \leftarrow i, \dots, j - 1$}
-        \If {$\alpha[B, i, k] \neq 0$}
-          \State if $\beta[C, k+1, j] = NIL$, set to $0$
-          \State $\beta[C, k+1, j] \leftarrow \beta[C, k+1, j] + \psi(A \rightarrow B C) \alpha[B, i, k] \beta[A, i, j]$
-          \State add $(C, k+1, j)$ to $children$
-        \EndIf
-
-        \If {$\alpha[C, k+1, j] \neq 0$}
-          \State if $\beta[B, i, k] = NIL$, set to $0$
-          \State $\beta[B, i, k] \leftarrow \beta[B, i, k] + \psi(A \rightarrow B C) \alpha[C, k+1, j] \beta[A, i, j]$
-          \State add $(B, i, k)$ to $children$
-        \EndIf 
-      \EndFor 
-   \EndFor
-   \EndFor
-   \State $open \leftarrow children$
-  \EndWhile    
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{The Approximate Outside Algorithm.\label{alg:approxOutside}}
-\begin{algorithmic}[2]
-\Procedure {$ApproxOutside$}{$s$, $G$, $S$}
-  \State $OPEN \leftarrow []$ \Comment initialize OPEN FIFO queue
-  \State $n \leftarrow length(s)$
-  \State $\beta(S, 1, n) \leftarrow 1$
-  \State \Call{AddChildrenToOpen}{$S, 1, n, \beta(S, 1, n)$}
-  \While {$OPEN$ is not empty}
-    \State $node \leftarrow pop(OPEN)$ where $rule_{node} = (A \rightarrow B C, i, k, j) = fst(node)$ and $pa_\beta= snd(node)$
-    \State $a_\ell \leftarrow \alpha[B, i, k]$
-    \State $a_r \leftarrow \alpha[C, k+1, j]$
-    \If {$a_r \neq NIL$}
-      \If {$\beta[B, i, k] = NIL$}
-         $\beta[B, i, k] \leftarrow 0$
-      \EndIf
-      \State $\beta[B, i, k] \leftarrow \beta[B, i, k] + pa_\beta a_\ell \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
-      \State \Call{AddChildrenToOpen}{$B, i, k, \beta(B, i, k)$}
-    \EndIf
-    \If {$a_l \neq NIL$}
-      \If {$\beta[C, k+1, j] = NIL$}
-         $\beta[C, k+1, j] \leftarrow 0$
-      \EndIf
-      \State $\beta[C, k+1, j] \leftarrow \beta[C, k+1, j] + pa_\beta a_r \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
-      \State \Call{AddChildrenToOpen}{$C, k+1, j, \beta(C, k+1, j)$}
-    \EndIf
-  \EndWhile  
-\EndProcedure
-
-\Statex
-\Procedure {AddChildrenToOPEN}{A, i, j, betaVal}
-  \State $\{c_i\}\leftarrow \Call{GetChildren}{A, i, j}$ \Comment Get AND-node children
-  \State insert pairs $\{(c_i, betaVal\}_i)$ into OPEN
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\noindent \textbf{Claim: Algorithm~\ref{alg:approxInside} correctly computes the $\beta'$ values.}
-\noindent TODO
-
-\noindent \textbf{Claim: Algorithm~\ref{alg:approxOutside} correctly computes the $\beta'$ values.}
-\noindent TODO
-
-\section{A simple upper bound}
-We will present a simple upper bound on a slightly restricted set of
-PCFGs which is based on the following intuition: if I am a nonterminal
-$A$ and I know that I need to yield a string $x_1 \dots x_n$, then
-this is going to be at least as hard as starting at $A$ and generating
-any one of the elements in $x_1 \dost x_n$. That is, the highest
-probability single word in my yield will have a higher probability of
-being produced, then any longer span.
-
-Suppose our PCFG has potentials less than or equal to one. This is the
-case in the commons situation in which the PCFG is normalized, or if,
-as in variational approximations to the Hierarchical Diirichlet PCFG,
-our PCFG is unnormalized with values summing to less than one. Now we
-define a function $g(A, w, m)$ for nonterminal $A$, terminal $w$ and
-integer $w>0$, which we claim is greater the weight our PCFG would
-assign to any parse tree rooted at $A$ and yielding any span of length
-$m$ containing $w$. Specifically, consider any such parse tree, and
-set all weights not on the path between $A$ and $w$ to 1. The total
-weight assigned to the parse tree is the product of the weights on its
-edges, so the resulting parse tree will have a strictly greater score.
-
-\begin{algorithm}
-\caption{An O(n^2}
-\begin{algorithmic}[2]
-\Procedure {$GO_{OR}$}{$A$, $i$, $j$}
-
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
+\bibliographystyle{bibliographystyle}
+\bibliography{main}
 

### 2014-11-10 drafts/note1.tex
Added some untracked files."

Notes on meeting with Josh T., Josh R. on number concept learning
8/27/2014

- How do children learn number concepts that go beyond counting and
the cardinality principle?

- In particular, how do children represent and learn the set of
complimentary relations and routines that support number knowledge
over time?

- Relations of interest: 
1) Next word in number list: e.g. Next(``three hundred and forty five'', ``three hundred and forty six'')
2) Approximate magnitude: e.g. AM("eighty six", ~80)
3) Exact cardinatlity: e.g. Exact("eight six", 86)

others: 
1) Is one number greater than another: LexGreater(forty five, twenty seven)
2) Number as digits: Digit(forty five, 45)
3) Approximate forms of addition and multiplication.

- How do we account for different stages of number knowledge development: 
Stage 1: (ages 1-4) pre-cp knowledge, numbers 1-10
Stage 2: (ages 4-6) two-digit numbers
Stage 3: (ages >6) adult like extension to larger numbers

### 2014-11-10 experiments/prism/datasets/allExamples2.pl
Added some untracked files.

@@ -1,0 +1,198 @@
+train(1, srs('S_1'-[[after,one,comes,two]])).
+train(2, srs('S_1'-[[after,two,comes,three]])).
+train(3, srs('S_1'-[[after,three,comes,four]])).
+train(4, srs('S_1'-[[after,four,comes,five]])).
+train(5, srs('S_1'-[[after,five,comes,six]])).
+train(6, srs('S_1'-[[after,six,comes,seven]])).
+train(7, srs('S_1'-[[after,seven,comes,eight]])).
+train(8, srs('S_1'-[[after,eight,comes,nine]])).
+train(9, srs('S_1'-[[after,nine,comes,ten]])).
+train(10, srs('S_1'-[[after,ten,comes,eleven]])).
+train(11, srs('S_1'-[[after,eleven,comes,twelve]])).
+train(12, srs('S_1'-[[after,twelve,comes,thirteen]])).
+train(13, srs('S_1'-[[after,thirteen,comes,fourteen]])).
+train(14, srs('S_1'-[[after,fourteen,comes,fifteen]])).
+train(15, srs('S_1'-[[after,fifteen,comes,sixteen]])).
+train(16, srs('S_1'-[[after,sixteen,comes,seventeen]])).
+train(17, srs('S_1'-[[after,seventeen,comes,eighteen]])).
+train(18, srs('S_1'-[[after,eighteen,comes,nineteen]])).
+train(19, srs('S_1'-[[after,nineteen,comes,twenty]])).
+train(20, srs('S_1'-[[after,twenty,comes,twenty,one]])).
+train(21, srs('S_1'-[[after,twenty,one,comes,twenty,two]])).
+train(22, srs('S_1'-[[after,twenty,two,comes,twenty,three]])).
+train(23, srs('S_1'-[[after,twenty,three,comes,twenty,four]])).
+train(24, srs('S_1'-[[after,twenty,four,comes,twenty,five]])).
+train(25, srs('S_1'-[[after,twenty,five,comes,twenty,six]])).
+train(26, srs('S_1'-[[after,twenty,six,comes,twenty,seven]])).
+train(27, srs('S_1'-[[after,twenty,seven,comes,twenty,eight]])).
+train(28, srs('S_1'-[[after,twenty,eight,comes,twenty,nine]])).
+train(29, srs('S_1'-[[after,twenty,nine,comes,thirty]])).
+train(30, srs('S_1'-[[after,thirty,comes,thirty,one]])).
[…]

### 2014-11-10 experiments/prism/datasets/num_trainTest.pl

@@ -1,0 +1,199 @@
+train(1, prove('S_1'-[[twenty,six,is,after,twenty,five]])).
+train(2, prove('S_1'-[[sixty,five,is,after,sixty,four]])).
+train(3, prove('S_1'-[[ninety,is,after,eighty,nine]])).
+train(4, prove('S_1'-[[after,forty,eight,is,forty,nine]])).
+train(5, prove('S_1'-[[thirty,is,after,twenty,nine]])).
+train(6, prove('S_1'-[[after,thirty,six,is,thirty,seven]])).
+train(7, prove('S_1'-[[after,sixty,three,is,sixty,four]])).
+train(8, prove('S_1'-[[forty,three,is,after,forty,two]])).
+train(9, prove('S_1'-[[ninety,two,is,after,ninety,one]])).
+train(10, prove('S_1'-[[after,forty,is,forty,one]])).
+train(11, prove('S_1'-[[three,is,after,two]])).
+train(12, prove('S_1'-[[after,six,is,seven]])).
+train(13, prove('S_1'-[[seventeen,is,after,sixteen]])).
+train(14, prove('S_1'-[[after,seventeen,is,eighteen]])).
+train(15, prove('S_1'-[[twenty,is,after,nineteen]])).
+train(16, prove('S_1'-[[after,sixty,five,is,sixty,six]])).
+train(17, prove('S_1'-[[after,seventy,two,is,seventy,three]])).
+train(18, prove('S_1'-[[after,eighty,three,is,eighty,four]])).
+train(19, prove('S_1'-[[fifty,two,is,after,fifty,one]])).
+train(20, prove('S_1'-[[thirty,three,is,after,thirty,two]])).
+train(21, prove('S_1'-[[after,sixty,one,is,sixty,two]])).
+train(22, prove('S_1'-[[ninety,seven,is,after,ninety,six]])).
+train(23, prove('S_1'-[[thirty,one,is,after,thirty]])).
+train(24, prove('S_1'-[[after,seventy,eight,is,seventy,nine]])).
+train(25, prove('S_1'-[[after,eighty,one,is,eighty,two]])).
+train(26, prove('S_1'-[[after,eighteen,is,nineteen]])).
+train(27, prove('S_1'-[[thirty,eight,is,after,thirty,seven]])).
+train(28, prove('S_1'-[[forty,seven,is,after,forty,six]])).
+train(29, prove('S_1'-[[fifty,seven,is,after,fifty,six]])).
[…]

### 2014-11-10 experiments/prism/datasets/num_trainTest_2.pl

@@ -1,0 +1,395 @@
+train(1, prove('S_1'-[[sixty,three,is,before,sixty,four]])).
+train(2, prove('S_1'-[[fifty,one,is,after,fifty]])).
+train(3, prove('S_1'-[[before,thirty,five,is,thirty,four]])).
+train(4, prove('S_1'-[[before,thirty,seven,is,thirty,six]])).
+train(5, prove('S_1'-[[seventy,one,is,before,seventy,two]])).
+train(6, prove('S_1'-[[after,eighty,five,is,eighty,six]])).
+train(7, prove('S_1'-[[after,fifty,nine,is,sixty]])).
+train(8, prove('S_1'-[[ninety,three,is,after,ninety,two]])).
+train(9, prove('S_1'-[[after,eleven,is,twelve]])).
+train(10, prove('S_1'-[[sixty,four,is,after,sixty,three]])).
+train(11, prove('S_1'-[[fifty,seven,is,after,fifty,six]])).
+train(12, prove('S_1'-[[after,fifty,one,is,fifty,two]])).
+train(13, prove('S_1'-[[after,sixty,nine,is,seventy]])).
+train(14, prove('S_1'-[[eighty,one,is,before,eighty,two]])).
+train(15, prove('S_1'-[[before,eighty,nine,is,eighty,eight]])).
+train(16, prove('S_1'-[[eighty,nine,is,before,ninety]])).
+train(17, prove('S_1'-[[before,twenty,three,is,twenty,two]])).
+train(18, prove('S_1'-[[before,seventy,one,is,seventy]])).
+train(19, prove('S_1'-[[after,fifty,three,is,fifty,four]])).
+train(20, prove('S_1'-[[sixty,six,is,after,sixty,five]])).
+train(21, prove('S_1'-[[before,sixty,is,fifty,nine]])).
+train(22, prove('S_1'-[[after,four,is,five]])).
+train(23, prove('S_1'-[[ten,is,after,nine]])).
+train(24, prove('S_1'-[[before,sixty,one,is,sixty]])).
+train(25, prove('S_1'-[[before,twenty,five,is,twenty,four]])).
+train(26, prove('S_1'-[[before,thirty,six,is,thirty,five]])).
+train(27, prove('S_1'-[[twenty,eight,is,after,twenty,seven]])).
+train(28, prove('S_1'-[[twenty,three,is,before,twenty,four]])).
+train(29, prove('S_1'-[[before,ninety,is,eighty,nine]])).
[…]

### 2014-11-10 experiments/prism/analysis/train_number_net_0006.held_out.py
Added untracked files.

@@ -1,0 +1,36 @@
+import numpy as np
+from matplotlib import pyplot as plt
+
+from matplotlib import rc
+# rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+## for Palatino and other serif fonts use:
+#rc('font',**{'family':'serif','serif':['Palatino']})
+rc('text', usetex=True)
+plt.rcParams['text.latex.preamble']=\
+                [r"\usepackage{amssymb, amsmath}"]
+
+plt.rc('text', usetex=True)
+plt.rc('font', **{'family': 'serif', 'size': 16})
+
+data = np.loadtxt("train_number_net_0006.held_out.csv", delimiter=",")
+data.sort(axis=0)
+data = data[1:, 1]
+
+fig = plt.figure()
+ax = fig.add_subplot(111)
+ax.plot(range(1, len(data) + 1), 
+        data, 
+        linewidth=2, 
+        marker='o')
+ax.set_xlabel("Iteration")
+ax.set_ylabel(r"$\mathbb{E} \log p(D_{\text{test}})$")
+ax.set_ylim(top=1000)
+ax.set_xlim(right=20)
+
+# line at 0
+[lo, hi] = ax.get_xlim()
+ax.plot([lo, hi], [0, 0], 'k--')
+
+fig.savefig("train_number_net_0006.held_out.pdf")
+
+plt.show()

### 2014-11-10 experiments/prism/resources/conjugate.py

@@ -1,0 +1,45 @@
+#!/usr/bin/env python
+
+import sys
+sys.path.append("../../misc")
+
+from os.path import exists
+from pattern.en import *
+import ipa
+
+conjugations = [
+    {"tense" : PRESENT,
+     "number" : SINGULAR},
+    {"tense" : PAST,
+     "number" : SINGULAR},
+    {"tense" : PRESENT,
+     "number" : PLURAL}, 
+    {"tense" : PRESENT, 
+     "aspect" : PROGRESSIVE}]
+
+def mkConjugations(verbs):
+    if not(isinstance(verbs, list)):
+        verbs = [verbs]
+    out = []
+    for verb in verbs: 
+        for conj in conjugations: 
+            out.append(conjugate(verb, **conj))
+    return out
+
+def topHundredConjugations(useIPA=True):
+    verbFile = "./verbs.txt"
+    assert(exists(verbFile))
+    
+    fin = open("verbs.txt", "r")
+    top_hundred = [line.strip() for line in fin.readlines()]
+    fin.close()
+    
+    conjs = mkConjugations(top_hundred)
+
+    if useIPA:
+        return ipa.ipa(conjs)
+    return conjs
+
+
+
+

### 2014-11-10 experiments/prism/resources/number_corpus/NumberHist.hs

@@ -1,0 +1,12 @@
+
+
+module Main where
+
+
+
+
+main = do
+  ss <- readFile "../numbers.txt"
+  corpus <- readFile "allSentences"
+  let numbers = lines ss
+  

### 2014-11-10 experiments/prism/resources/number_corpus/get_childes_numbers.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+# extract number words from child directed speech corpus 
+# CHILDES treebank for CDS taken from http://www.socsci.uci.edu/~lpearl/CoLaLab/CHILDESTreebank/childestreebank.html
+# extract number words (category CD) using tregex
+
+TREGEX=~/Downloads/stanford-tregex-2014-08-27/tregex.sh
+CORPRA_DIR="/Users/edechter/Downloads/CHILDESTreebank-curr/"
+PATTERN="CD=number"
+
+PARSED_CORPRA=`ls ${CORPRA_DIR}/*.parsed`
+echo $PARSED_CORPRA
+
+OUT=""
+for f in $PARSED_CORPRA
+do
+    CMD="${TREGEX} $PATTERN $f"
+    echo -e "Processing $f: \n$CMD"
+    OUT="$OUT `$CMD`"
+    echo $OUT
+    echo -e "Done.\n"
+done
+

### 2014-11-10 experiments/prism/resources/process_ec2_data.py

@@ -1,0 +1,38 @@
+
+
+import numpy as np
+from matplotlib import pyplot as plt
+import pandas as pd
+
+ec2_data = np.loadtxt("./ec2_data.csv", delimiter=",")
+
+df = pd.DataFrame(ec2_data, columns = ["nExamples", "nPreds", "Acc"])
+df = df[df["nExamples"] <600]
+
+grouped = df.groupby(["nPreds", "nExamples"])
+
+
+
+grouped_mean = grouped.mean()
+grouped_std = grouped.std()
+grouped_max = grouped.max()
+grouped_min = grouped.min()
+
+fig = plt.figure()
+ax = fig.add_subplot(111)
+for npred in [1,2,3,4]:
+    ax.hold(True)
+    ax.bar(grouped_max.loc[npred].index.values+10*npred - 30, 
+           grouped_max.loc[npred].values, 
+           width=10,
+                # yerr=grouped_std.loc[npred].values/10,
+           label="%d Predicates"%npred,
+           color=[.2*i for i in [npred, npred, npred]],
+           linewidth=2
+    )
+    ax.set_ylim([0, 1])
+
+plt.legend()
+    
+
+

### 2014-11-10 experiments/prism/resources/run_parallel.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+NumExamples=(1700);
+NumReps=1;
+NumPred=(4);
+SaveDir="data/saturday_"
+args=()
+
+for nEx in ${NumExamples[*]}
+do
+    for rep in $(eval echo {1..$NumReps})
+    do
+        for nPred in ${NumPred[*]}
+        do
+            args+=("$nEx")
+            args+=("$rep")
+            args+=("$nPred")
+        done
+    done
+done
+
+CMD="parallel -n 3 --noswap --resume-failed --joblog ./parallel_log --delay 10 upprism srs {} $SaveDir &> logs/overnight/saturday_log_{1}_{2}_{3}.log ::: ${args[*]}"
+echo $CMD
+`$CMD`
+
+
+
+
+

### 2014-11-10 experiments/prism/scripts/train_number_net_0002.pl

@@ -1,0 +1,40 @@
+
+:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
+   cl(P).
+
+%% load .sys.psm file.
+:- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
+   prism([load], P).
+
+%% load data set
+%% this includes before and after sentences with 20 held out test sentences
+:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
+   cl(P).
+
+:- set_training_flags.
+
+%% ovbem config
+:- set_ovbem_rec(niter, 1000), 
+   set_ovbem_rec(batch_size, 1), 
+   set_ovbem_rec(iter_between, 5).
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0002.log', P),
+   set_ovbem_rec(logfile, P).
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0002.data', P),
+   set_ovbem_rec(outdir, P), 
+   initialize_outdir.
+   
+:- findall(X, train(_, X), DataSet), 
+   set_ovbem_rec(data_set, DataSet).
+
+:- findall(X, test(_, X), HeldOutAll), 
+   take(1, HeldOutAll, HeldOut), 
+   write(HeldOut), nl, 
+   set_ovbem_rec(held_out, HeldOut). 
+
+%% run training
+main :- run_ovbem.
+
+
+    

### 2014-11-10 experiments/prism/scripts/train_number_net_0004.pl

@@ -1,0 +1,37 @@
+
+:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
+   cl(P).
+
+%% load .sys.psm file.
+:- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
+   prism([load], P),
+   set_ovbem_rec(sys_psm_file, P).
+
+%% load data set
+%% this includes before and after sentences with 20 held out test sentences
+:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
+   cl(P).
+
+:- set_training_flags.
+
+%% ovbem config
+:- set_ovbem_rec(niter, 100), 
+   set_ovbem_rec(batch_size, 10), 
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0004.log', P),
+   set_ovbem_rec(logfile, P).
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0004.data', P),
+   set_ovbem_rec(outdir, P), 
+   initialize_outdir.
+   
+:- findall(count(X, 100), train(_, X), DataSet), 
+   set_ovbem_rec(data_set, DataSet).
+
+%% run training
+main :- 
+    set_training_flags,
+    run_ovbem.
+
+
+    

### 2014-11-10 experiments/prism/scripts/train_number_net_0006.pl

@@ -1,0 +1,40 @@
+
+:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
+   cl(P).
+
+%% load .sys.psm file.
+load_sys_psm :- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_5_5_5.sys.psm', 
+                                   P),
+                prism([load], P),
+                set_ovbem_rec(sys_psm_file, P).
+
+%% load data set
+%% this includes before and after sentences with 20 held out test sentences
+:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
+   cl(P).
+
+:- set_training_flags.
+
+%% ovbem config
+:- set_ovbem_rec(niter, 100), 
+   set_ovbem_rec(batch_size, 10).
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0007.log', P),
+   set_ovbem_rec(logfile, P).
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0007.data', P),
+   set_ovbem_rec(outdir, P).
+   
+:- findall(count(X, 1), train(_, X), DataSet), 
+   set_ovbem_rec(data_set, DataSet).
+
+%% run training
+main :- 
+    initialize_outdir,
+    load_sys_psm, 
+    set_training_flags,
+    run_ovbem.
+
+
+    
+          

### 2014-11-10 experiments/prism/scripts/train_number_net_0008.pl

@@ -1,0 +1,38 @@
+
+:- expand_environment('$GIJOE_ROOT/prism/train.pl', P), 
+   cl(P).
+
+%% load .sys.psm file.
+load_sys_psm :- expand_environment('$GIJOE_ROOT/experiments/prism/number_net_10_5_5.sys.psm', P),
+                prism([load], P),
+                set_ovbem_rec(sys_psm_file, P).
+
+%% load data set
+%% this includes before and after sentences with 20 held out test sentences
+:- expand_environment('$GIJOE_ROOT/experiments/prism/num_trainTest_2.pl', P),
+   cl(P).
+
+:- set_training_flags.
+
+%% ovbem config
+:- set_ovbem_rec(niter, 5000), 
+   set_ovbem_rec(batch_size, 1).
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0008.log', P),
+   set_ovbem_rec(logfile, P).
+
+:- expand_environment('$GIJOE_ROOT/experiments/prism/train_number_net_0008.data', P),
+   set_ovbem_rec(outdir, P).
+   
+:- findall(count(X, 1), train(_, X), DataSet), 
+   set_ovbem_rec(data_set, DataSet).
+
+%% run training
+main :- 
+    initialize_outdir,
+    load_sys_psm, 
+    set_training_flags,
+    run_ovbem.
+
+
+    

### 2014-11-10 src/GIJoe/SRS/LCG2Prism.hs
Added untracked files.

@@ -1,0 +1,4 @@
+
+module GIJoe.SRS.LCG where
+
+

### 2014-11-10 src/GIJoe/SRS/Resolution.hs

@@ -1,0 +1,44 @@
+{-# Language FlexibleContexts #-}
+
+module GIJoe.SRS.Resolution where
+
+import Control.Monad.Error.Class
+
+import Data.HashMap.Strict (HashMap)
+import qualified Data.HashMap.Strict as Map
+
+type Map = HashMap
+
+-- | Implementation of resolution algorithm for function-free
+-- Horn clause logic with string concatenation. 
+
+data Var = Var String deriving (Show, Eq, Ord)
+type Sym = String
+data Element = ElVar Var | ElSym Sym deriving (Show, Eq)
+type ElementStr = [Element]
+data Predicate = Predicate {predName :: String,
+                            predArity :: Int} deriving (Show, Eq)
+                                                       
+data Term = Term {termPred :: Predicate,
+                  termArgs :: [Elements]} deriving (Show, Eq)
+
+type Subst = Map Var Term
+
+nullSubst :: Subst
+nullSubst = Map.empty
+
+data UnificationError = ConflictError Term Term
+                        | OccursCheckError Var Term deriving Eq
+
+instance Show UnificationError where 
+    show (ConflictError t1 t2) = "UnificationError.ConflictError: attempting to unify_ " ++ show t1 ++ " and " ++ show t2 ++ "."
+    show (OccursCheckError var term) = "UnificationError.OccursCheckError:: variable " ++ show var ++ " cannot unify with term " ++ show term ++ "."
+                                                             
+mgu :: MonadError UnificationError m => Term -> Term -> m Subst
+-- | Most general unifier of two terms
+mgu (Term _P xs) (Term _Q ys) | _P /= _Q = throwError $ ConflictError _P _Q
+                              | otherwise = foldM_ mgu_
+       
+
+
+

### 2014-11-10 src/GIJoe/SRS/Verbs.hs

@@ -1,0 +1,59 @@
+
+module Verbs where
+
+import Data.List
+import System.Random.Shuffle (shuffle', shuffleM)
+import Control.Monad.Random.Class
+import Control.Monad
+
+
+import GIJoe.SRS.Parse
+import GIJoe.SRS.Prismify
+import GIJoe.SRS.Prismify
+
+ipaLexicon = do xs <- readFile "./experiments/prism/englishPhones.txt"
+                return $ lines xs
+
+
+buildGrammar :: Int -- ^ number of hidden predicates
+             -> [String] -- ^ lexicon
+             -> [String] -- ^ list of observed unary predicates
+             -> FilePath
+             -> IO ()
+buildGrammar n lexicon observed outpath = do
+  writeFile outpath $ show rs
+  writeFile (outpath ++ ".psm") $ prismClauses rs
+  where rs = fromRight $ parseSystem "buildGrammar" $ systemString
+        fromRight (Right x) = x
+        fromRight (Left err) = error $ show err
+        systemString = unlines $ 
+                       (concat [rules i j k | i <- [1..n], j <-[1..n], k <- [j..n]])++
+                       [ruleLex i w v | i <- [1..n],
+                        (wi, w) <- zip [0..] lexicon,
+                        (vi, v) <- zip [0..] lexicon,
+                        vi >= wi] ++ 
+                       [ruleLex i w "null" | i <- [1..n], w <- lexicon] ++ 
+                       [ruleLex i "null" w | i <- [1..n], w <- lexicon] ++
+                       [pred ++ "(X Y) <-- A" ++ show i ++ "(X, Y)." | (pred, i) <- zip observed [1..]]
+        rules i j k | j /= k  = do
+          [v1, v2, v3, v4] <- permutations ["X", "Y", "U", "V"]
+          return $ "A" ++ show i ++ "(X Y, U V) <-- " ++
+            "A" ++ show j ++ "(" ++ v1 ++ ", " ++ v2 ++ "), " ++
+            "A" ++ show k ++ "(" ++ v3 ++ ", " ++ v4 ++ ")."
+        rules i j k | j == k  = do
+          [v1, v2, v3, v4] <- permutations ["X", "Y", "U", "V"]
+          guard $ (v1, v2) <= (v3, v4)
[…]

### 2014-11-10 haskell/GIJoe/ApproximateInference/BoundaryProbabilities.hs
Removed everything except for LPN stuff.

@@ -1,151 +1,0 @@
-{-# Language BangPatterns, LambdaCase, RankNTypes #-}
-
-module GIJoe.ApproximateInference.BoundaryProbabilities
-       (mkOmegaTable,
-        mkGammaThetaTables,
-        mkLambdaTable, 
-        OmegaTable,
-        GammaTable,
-        ThetaTable,
-        LambdaTable
-       ) where
-
-import Prelude hiding (sum)
-import GIJoe.Grammar
-import GIJoe.Types
-
-import GIJoe.ApproximateInference.HashTable (HashTable)
-import qualified GIJoe.ApproximateInference.HashTable as HashTable
-
-import Data.Maybe
-
-import Control.Monad
-import Control.Monad.Trans
-import Control.Monad.ST
-
-import Numeric.LinearAlgebra.HMatrix
-import Data.Packed.Matrix
-import Numeric.Log
-
-import Data.HashMap.Strict (HashMap)
-import qualified Data.HashMap.Strict as HashMap
-
-import Control.Monad.Loop
-
-import Data.List hiding (sum)
-
-import Debug.Trace
-
-type OmegaTable s = HashTable s (Symbol, Symbol, Symbol) Double'
-type GammaTable s = HashTable s (Symbol, Symbol) Double'
-type ThetaTable s = HashTable s (Symbol, Symbol) Double'
-
-
-mkOmegaTable :: Grammar -> ST s (HashTable s (Symbol, Symbol, Symbol) Double')
-mkOmegaTable gr = trace ("Making OmegaTable") $ do
-  (gamma_ht, theta_ht) <- mkGammaThetaTables_ gr si
-  omega_ht <- HashTable.newSized $ (numSymbols gr)^3
-  exec_ $ do 
-    _A <- forEach $ nonterminals gr 
-    x  <- forEach $ terminals gr
-    y  <- forEach $ terminals gr
-    let !v0 = if x == y
-                 then sum $ map weight $ getUnaryRulesBySymbols gr _A x
-                 else 0
-    lift $ HashTable.insertWith omega_ht (+) (_A, x, y) v0
-    let rs = binaryRulesHeadedBy gr _A
-    BinaryRule _ _B _C w <- forEach rs
-    lift (HashTable.lookup gamma_ht (_B, x)) >>= \case
-      Nothing -> return ()
-      Just gamma_Bx ->
-        lift (HashTable.lookup theta_ht (_C, y) ) >>= \case
-          Nothing -> return ()
-          Just theta_Cy ->
[…]

### 2014-11-10 haskell/GIJoe/ApproximateInference/DList.hs

@@ -1,60 +1,0 @@
-
-module GIJoe.ApproximateInference.DList where
-
-import Prelude hiding (head, reverse, concat, break, null)
-import qualified Prelude
-
-import Debug.Trace
-
-data BTree a = Node (BTree a) (BTree a) |
-               Leaf a deriving (Show, Ord, Eq)
-
-type DList a = BTree [a] 
-
-empty :: DList a
-empty = Leaf []
-
-null :: DList a -> Bool
-null (Node l r) = null l && null r
-null (Leaf []) = True
-null _ = False
-
--- | Add element
-cons :: a -> DList a -> DList a
-cons x (Leaf xs) = Leaf (x:xs)
-cons x (Node l r) = Node (cons x l) r
-
--- | Pop off first element
-pop :: DList a -> (a, DList a)
-pop dlist = case pop' dlist of
-  Nothing -> error "pop: empty list"
-  Just res -> res
-  where pop' (Leaf (x:xs)) = Just (x, Leaf xs)
-        pop' (Node l r) = case pop' l of
-          Just (x, l') -> Just (x, concat l' r)
-          Nothing -> pop' r
-        pop' (Leaf []) = Nothing
-
-        -- debug x = trace "pop'!" x
-
-
--- | Remove a single element from head of list
-head :: DList a -> a
-head (Leaf xs) = Prelude.head xs
-head (Node l r) = head l
-
--- | Return everything but first element
-tail :: DList a -> DList a
-tail xs = let (_, xs) = pop xs in xs
-
--- | Concatenate two DLists
-concat :: DList a -> DList a -> DList a
-concat = Node
-
--- | Reverse a DList
-reverse :: DList a -> DList a
-reverse (Leaf xs) = Leaf (Prelude.reverse xs)
-reverse (Node l r) = Node (reverse r) (reverse l) -- this enables reverse without evaluating intermediate lists
-
-fromList :: [a] -> DList a
-fromList = Leaf

### 2014-11-10 haskell/GIJoe/ApproximateInference/LazyDequeue.hs

@@ -1,70 +1,0 @@
-
-module GIJoe.ApproximateInference.LazyDequeue
-       (LazyDequeue(..),
-        empty,
-        enqueue,
-        enqueueMany,
-        dequeue,
-        fromList,
-        toList,
-        null)
-       where
-
-import Prelude hiding (null)
-import GIJoe.ApproximateInference.DList (DList)
-import qualified GIJoe.ApproximateInference.DList as DList
-
-import Debug.Trace
-
-data LazyDequeue a = LazyDequeue (DList a) (DList a) deriving (Show, Eq, Ord)
-
--- | An empty queue
-empty :: LazyDequeue a
-{-# INLINE empty #-}
-empty = LazyDequeue DList.empty DList.empty
-
-null :: LazyDequeue a -> Bool
-null (LazyDequeue l r) = DList.null l && DList.null r
-
-enqueue :: a -> LazyDequeue a -> LazyDequeue a
-{-# INLINE enqueue #-}
-enqueue a (LazyDequeue l r) = LazyDequeue (DList.cons a l) r
-
-enqueueMany :: [a] -> LazyDequeue a -> LazyDequeue a
-{-# INLINE enqueueMany #-}
-enqueueMany xs q = let (LazyDequeue l r) = shift q 
-  in LazyDequeue l (DList.concat (DList.fromList xs) r)
-
--- | Dequeue a single element. If the right list is empty, reverses
--- the first list and sticks the result in the second list.
-dequeue :: LazyDequeue a -> (a, LazyDequeue a)
-dequeue (LazyDequeue l r) | DList.null r = dequeue $ shift (LazyDequeue l r)
-                          | otherwise    = (a, LazyDequeue l r')
-  where (a, r') = DList.pop r
-
-toList :: Show a => LazyDequeue a -> [a]
-toList q | null q = []
-         | otherwise = let (x, q') = dequeue q in x: toList q'
-
-fromList :: [a] -> LazyDequeue a
-{-# INLINE fromList #-}
-fromList xs = enqueueMany xs empty
-
-------------------------
--- Private functions
-
-shift :: LazyDequeue a -> LazyDequeue a
-shift (LazyDequeue l r) = LazyDequeue DList.empty (DList.concat r (DList.reverse l))
-
-
-
-                                 
-
-
-
-
-
-
-
-
-

### 2014-11-10 haskell/GIJoe/Chart.hs

@@ -1,28 +1,0 @@
-
-module GIJoe.Chart where
-
-import Control.Monad.Logic
-
-import Control.Monad.State.Strict
-import Control.Monad.Identity
-
-
-type StateLogicT s m a = LogicT (StateT s m) a
-
-runStateListT :: MonadState s m => LogicT m a -> s -> m [a]
-runStateListT logT s = observeAllT logT 
-
-toStateT :: MonadState s m => LogicT m a -> m [a]
-toStateT t = do s <- get
-                as <- runStateListT t s
-                return as
-
--- test :: StateListT Int Identity (Int, Int, Int)
--- test = do i <- liftList [0..5]
---           j <- liftList [i..5]
---           modify $ \s -> s `div` 2
---           s <- get
---           return $ (i, j, s)
-
-
-

### 2014-11-10 haskell/GIJoe/ComplexityTest.hs

@@ -1,32 +1,0 @@
-
-module Main where
-
-import Prelude hiding (lookup)
-
-import GIJoe.Grammar
-import GIJoe.Parse
-import GIJoe.InsideOutside
-
-import System.Random
-import Control.Monad.Random
-import Control.Monad
-import Control.Monad.Trans
-import Data.List hiding (lookup)
-import Data.Function
-
-main = do
-  let n = 50
-  let xs = stringToSentence $ ('a':) $ concat $ replicate (n-1) " a"
-  let lexicon = [T "a"]
-  let _K = 5
-  let _S = N 0 Nothing
-  let gr = initGrammar _S lexicon _K
-  let alphaChart = {-# SCC "alphaChart" #-} alphas gr 1000000 xs
-  putStrLn "Complexity Test"
-  putStrLn $ "K: " ++ show _K
-  putStrLn $ "n: " ++ show n
-  putStrLn $ "xs: " ++ show xs
-  forM [(i, j, t) | i <-[1..n], j<-[i..n], t <- allNonTerminals gr] $ \(i, j, t) ->
-    putStrLn $ "alpha " ++ show i ++ " " ++ show j ++ " n " ++ show t  ++ " : " ++ show (lookup i j t alphaChart)
-  -- let betaChart = {-# SCC "betaChart" #-} betas gr _S 1000000 xs alphaChart      
-  -- forM [1..n] $ \i -> putStrLn $ "betas 1 n _S: " ++ show (lookup i i _S betaChart)

### 2014-11-10 haskell/GIJoe/EM.hs

@@ -1,10 +1,0 @@
-
-
-module GIJoe.EM where
-
-
-import GIJoe.Types
-import GIJoe.Utils
-import GIJoe.Chart
-import GIJoe.Grammar
-

### 2014-11-10 haskell/GIJoe/Grammar.hs

@@ -1,506 +1,0 @@
-{-# Language BangPatterns, ParallelListComp, FlexibleInstances, DeriveGeneric #-}
-
-module GIJoe.Grammar where
-
-import GIJoe.Types
-import GIJoe.Utils
-import GIJoe.Chart
-
-import Prelude hiding (sum, lookup)
-import qualified Prelude
-
-import Numeric.Log 
-
-import Control.Monad
-import Data.List.Split (splitOn)
-import Data.List (foldl', sortBy, sort, maximumBy, foldl1', nub)
-import Control.Monad.Logic
-import Data.Function (on)
-
-import Control.DeepSeq
-
-import Control.Monad.State.Strict
-import Control.Monad.Random
-import Control.Monad.Identity
-import Control.Parallel.Strategies 
-import qualified Control.DeepSeq as DeepSeq (NFData, rnf)
-import Numeric.SpecFunctions (digamma)
-
-import Data.HashMap.Strict (HashMap, (!))
-import qualified Data.HashMap.Strict as HashMap
-
-import Data.IntMap.Strict (IntMap)
-import qualified Data.IntMap.Strict as IntMap
-
-import Data.IntSet (IntSet) 
-import qualified Data.IntSet as IntSet
-
-import Data.PQueue.Prio.Max (MaxPQueue)
-import qualified Data.PQueue.Prio.Max as MaxPQueue
-
-import GHC.Generics (Generic)
-import Data.Hashable
-
-import Data.Tree
-import Data.Tree.Pretty 
-
-import Debug.Trace
-
-import Data.Sequence (Seq, (><))
-import qualified Data.Sequence as Seq
-
-import qualified Data.MemoCombinators as Memo
-
-data Symbol = N {nonTermIndex :: Int, nonTermName :: (Maybe String)} --  non terminal
-            | T String -- terminal
-            deriving (Ord, Eq, Generic)
-                     
-instance Show Symbol where
-  show (N _ (Just s)) = s
-  show (N i Nothing) = "N" ++ show i
-  show (T s) = s
-
-instance NFData Symbol where
-  rnf (N i t) = i `seq` t `seq` N i t `seq` ()
-  rnf (T s) = s `seq` T s `seq` ()
-
-
-instance Hashable Symbol where
-
-memoSymbol :: Memo.Memo Symbol
-memoSymbol f = table (m (f . uncurry N)) (m' (f . T)) f  
-  where m = Memo.pair Memo.integral (Memo.maybe (Memo.list Memo.char))
[…]

### 2014-11-10 haskell/GIJoe/Induction.hs

@@ -1,48 +1,0 @@
-
-
-module GIJoe.Induction where
-
-import GIJoe.Grammar
-import GIJoe.Parse
-
-import System.Random
-import Control.Monad.Random
-import Control.Monad
-import Control.Monad.Trans
-import Data.List
-import Data.Function
-
-
-
-_ROOT = N 0 Nothing
-short = stringToSentence "a dog ate a mouse"
-shortCorpus = [short]
-pruneLimit = 5
-
-main = do
-  let corpus_size = 100
-  liangGrammar <- readGrammar "liangGrammar.txt"
-  
-  corpus <- replicateM corpus_size (sample liangGrammar _S)
-  let lexicon = nub $ concat corpus
-      _K = 10
-      gr = initGrammar _ROOT lexicon _K
-  gr' <- evalRandIO $ randomizeGrammar gr
---   print $ loglikelihoodCorpus gr' _ROOT pruneLimit corpus 
---   let p = parse gr' _ROOT (head corpus) 
-  let emlog = em gr' _ROOT pruneLimit corpus 10 0
---   putStr $ unlines $ map show $ sortBy (compare `on` weight) $ grammarRules $ fst $ last emlog
---   print $ map snd emlog
-  print $ map snd emlog
-  let gr_final = fst . last $ emlog
-  -- Test learned grammar by getting MAP parses on random samples from true grammar
-  test_corpus <- replicateM 5 (sample liangGrammar _S)
-  sequence $ do
-    xs <- test_corpus
-    let (prs, _) = withCharts gr0 _S pruneLimit xs $ getMapParse
-    return $ print prs
-  return ()
-
-          
-
-

### 2014-11-10 haskell/GIJoe/InsideOutside.hs

@@ -1,520 +1,0 @@
-{-# Language BangPatterns, ParallelListComp, FlexibleInstances, DeriveGeneric #-}
-
-module GIJoe.InsideOutside where
-
-import GIJoe.Types
-import GIJoe.Utils
-import GIJoe.Chart
-import GIJoe.Grammar
-
-
-import Prelude hiding (sum, lookup)
-import qualified Prelude
-
-import Numeric.Log 
-
-import Control.Monad
-import Data.List.Split (splitOn)
-import Data.List (foldl', sortBy, sort, maximumBy, foldl1')
-import qualified Data.Foldable as Foldable (toList)
-import Control.Monad.Logic
-import Data.Function (on)
-
-import Control.Monad.State.Strict
-import Control.Monad.Random
-import Control.Monad.Identity
-import Control.Parallel.Strategies 
-import qualified Control.DeepSeq as DeepSeq (NFData, rnf)
-import Numeric.SpecFunctions (digamma)
-
-import Data.HashMap.Strict (HashMap, (!))
-import qualified Data.HashMap.Strict as HashMap
-
-import Data.IntMap.Strict (IntMap)
-import qualified Data.IntMap.Strict as IntMap
-
-import Data.IntSet (IntSet) 
-import qualified Data.IntSet as IntSet
-
-import GHC.Generics (Generic)
-import Data.Hashable
-
-import Data.Tree
-import Data.Tree.Pretty 
-
-import Debug.Trace
-
-import qualified Data.Sequence as Seq
-
-type Charts a = [Chart a]
-newtype Chart a = Chart {unChart :: HashMap (Int, Int) (HashMap Symbol a)}
-                    deriving Show
-
-lookup :: Int -> Int -> Symbol -> Chart a -> Maybe a
-lookup i j sym (Chart m) = do m_sym <- HashMap.lookup (i, j) m
-                              HashMap.lookup sym m_sym
-
-{-# INLINE lookupDefault #-}
-lookupDefault :: Int -> Int -> Symbol -> a -> Chart a  -> a
-lookupDefault i j sym def chart = case lookup i j sym chart of
-  Nothing -> def
-  Just a -> a
-
-insert :: Show a => Int -> Int -> Symbol -> a -> Chart a -> Chart a
-insert i j sym a (Chart m_ij) = Chart m_ij'
-  where m_ij' =
-          case HashMap.lookup (i, j) m_ij of
[…]

### 2014-11-10 haskell/GIJoe/MCFG.hs

@@ -1,41 +1,0 @@
-{-# LANGUAGE TemplateHaskell #-}
-
-module GIJoe.MCFG where
-
-import Control.Lens
-import Data.List (intercalate)
-import Numeric.Log
-
-type Double' = Log Double
-
-data Var = Var {_varInt :: Int} deriving (Eq, Ord)
-makeLenses ''Var
-
-instance Show Var where
-  show (Var i) = "x" ++ show i
-  
-type Symbol = String
-data Atom = Atom {_atomPred :: Symbol,
-                  _atomVars :: [Var]} deriving (Eq, Ord)
-makeLenses ''Atom                                               
-instance Show Atom where
-  show (Atom s vs) = s ++ "(" ++ intercalate " " (map show vs) ++ ")"
-  
-data Rule = Rule {_lhs :: Atom,
-                  _rhs :: [Atom]} deriving (Eq, Ord)
-makeLenses ''Rule                                           
-instance Show Rule where
-  show (Rule l rs) = show l ++ " <- " ++ intercalate " " (map show rs)
-
-data WeightedRule = WRule {_wRule :: Rule,
-                           _wRuleWeight :: Double'} deriving Show
-makeLenses ''WeightedRule
-
-data MCFG = MCFG {_start :: Symbol,
-                  _rules :: [WeightedRule]} deriving Show
-makeLenses ''MCFG
-
- 
-
-                                                     
-

### 2014-11-10 haskell/GIJoe/Parse.hs

@@ -1,80 +1,0 @@
-
-
-module GIJoe.Parse where
-
-import GIJoe.Grammar
-
-
-import Control.Monad
-import Numeric.Log
-import qualified Data.HashMap.Strict as HashMap
-
-import Text.Parsec.Prim hiding (parse, State)
-import Text.Parsec hiding (parse, State)
-import Text.ParserCombinators.Parsec.Number
-
-
-readGrammar :: FilePath -> IO Grammar
-readGrammar fpath = do xs <- readFile fpath
-                       case parseGrammar fpath xs of
-                         Right gr -> return gr
-                         Left err -> error $ show err
-
-writeGrammar fpath info gr = writeFile fpath (unlines (comment:ls))
-  where ls = map go (grammarRules gr)
-        go (BinaryRule h l r w) =
-          "$" ++ show h ++ " --> " ++ "$" ++ show l ++ " " ++ "$" ++ show r ++ " :: " ++ show w
-        go (UnaryRule h c w) =
-          "$" ++ show h ++ " --> " ++ show c ++ " :: " ++ show w
-        comment = '#':info
-                         
-parseGrammar :: String -- ^ source name
-            -> String -- ^ grammar content
-            -> Either ParseError Grammar
-parseGrammar source xs = runParser grammarParser (HashMap.empty, 0) source xs
-
-comment = do spaces
-             char '#'
-             _ <- manyTill anyToken eol
-             return ()
-             
-emptyLine = spaces >> eol >> return ()
-
-eol = char '\n'
-
-grammarParser = do rs <- manyTill lineParser eof
-                   return $! grammarFromRules rs
-
-lineParser = skipMany (comment <|> emptyLine) >> ruleParser
-
-ruleParser    = do spaces
-                   h <- nonterminal
-                   spaces
-                   string "-->"
-                   spaces
-                   ss <- symbols
-                   spaces
-                   string "::"
-                   spaces
-                   d <- floating2 False
-                   many (char ' ')
-                   optional eol
-                   case ss of
-                     (s:[]) -> return $! UnaryRule h s (Exp (log d))
-                     (s1:s2:[]) -> return $! BinaryRule h s1 s2 (Exp (log d))
-                     _ -> unexpected $ "Too many symbols on left hnd side of rule."
-
-nonterminal = do char '$'
-                 xs <- many1 alphaNum
-                 mp <- liftM fst getState                 
-                 i <- liftM snd getState
[…]

### 2014-11-10 haskell/GIJoe/RunEmail.hs

@@ -1,43 +1,0 @@
-
-module Main where
-
-import GIJoe.Grammar
-import GIJoe.Parse
-import GIJoe.InsideOutside
-
-import System.Random
-import Control.Monad.Random
-import Control.Monad
-import Control.Monad.Trans
-import Data.List
-import Data.Function
-
-main = do
-  xs <- readFile
-       "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/emails.txt"
-  let corpus = take 1 $  map stringToSentence . lines $ xs
-  let _N0 = N 0 Nothing
-  let lexicon = nub $ concat corpus
-      _K = 5
-      gr = initGrammar _N0 lexicon _K
-      prune_limit = 1000000
-  gr' <- evalRandIO $ randomizeGrammar gr
-  let emlog = em gr' _N0 100000 corpus 1 1e-3
-  print $ map snd emlog
-  let gr_final = fst . last $ emlog
-  let gr_out_path = "/Users/edechter/Dropbox/Projects/GrammarInduction/data/grammars/emailGrammar1.grammar"
-
-  writeGrammar gr_out_path "grammar induction, K=10, 30 iterations" gr_final
-
-
-  -- ys <- readFile
-  --      "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/emails.txt"
-  -- let test_corpus = map stringToSentence . lines $ ys
-  -- sequence $ do
-  --   xs <- test_corpus
-  --   let (prs, _) = withCharts gr_final _N0 prune_limit xs $ getMapParse
-  --   return $ print prs
-  -- return ()
-
-          
-

### 2014-11-10 haskell/GIJoe/RunNumber.hs

@@ -1,44 +1,0 @@
-
-module Main where
-
-import GIJoe.Grammar
-import GIJoe.Parse
-
-import System.Random
-import Control.Monad.Random
-import Control.Monad
-import Control.Monad.Trans
-import Data.List
-import Data.Function
-
-
-
-
-main = do
-  xs <- readFile
-       "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/simpleNumbers.txt"
-  let corpus = map stringToSentence . lines $ xs
-  let _N0 = N 0 Nothing
-  let lexicon = nub $ concat corpus
-      _K = 10
-      gr = initGrammar _N0 lexicon _K
-      prune_limit = 1000000
-  gr' <- evalRandIO $ randomizeGrammar gr
-  let emlog = em gr' _N0 100000 corpus 30 1e-3
-  print $ map snd emlog
-  let gr_final = fst . last $ emlog
-  let gr_out_path = "/Users/edechter/Dropbox/Projects/GrammarInduction/data/grammars/test1.grammar"
-
-  writeGrammar gr_out_path "grammar induction, K=10, 30 iterations" gr_final
-
-
-  ys <- readFile
-       "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/simpleNumbersTest.txt"
-  let test_corpus = map stringToSentence . lines $ ys
-  sequence $ do
-    xs <- test_corpus
-    let (prs, _) = withCharts gr_final _N0 prune_limit xs $ getMapParse
-    return $ print prs
-  return ()
-
-          

### 2014-11-10 haskell/GIJoe/RunNumberOld.hs

@@ -1,39 +1,0 @@
-
-module Main where
-
-import GIJoe.Grammar
-import GIJoe.Induction hiding (main)
-
-import System.Random
-import Control.Monad.Random
-import Control.Monad
-import Control.Monad.Trans
-import Data.List
-import Data.Function
-
-
-
-main = do
-  xs <- readFile
-        "/Users/edechter/Dropbox/Projects/GrammarInduction/data/corpuses/simpleNumbers.txt"
-  let corpus = map stringToSentence . lines $ xs
-  print corpus
-  let lexicon = nub $ concat corpus
-      _K = 6
-      gr = initGrammar _ROOT lexicon _K
-      prune_limit = 1000000000
-  print lexicon
-  gr' <- evalRandIO $ randomizeGrammar gr
-  print gr'
-  let emlog = em gr' _ROOT prune_limit corpus 10 0
-  print $ map snd emlog
-  let gr_final = fst . last $ emlog
-
-  let test_corpus = corpus
-  sequence $ do
-    xs <- test_corpus
-    let (prs, _) = withCharts gr_final _ROOT pruneLimit xs $ getMapParse
-    return $ print prs
-  return ()
-
-          

### 2014-11-10 haskell/GIJoe/Types.hs

@@ -1,8 +1,0 @@
-
-module GIJoe.Types (
-  Double'
-  ) where
-
-import Numeric.Log
-
-type Double' = Log Double

### 2014-11-10 haskell/GIJoe/Utils.hs

@@ -1,56 +1,0 @@
-{-# Language BangPatterns, ParallelListComp, FlexibleInstances, DeriveGeneric #-}
-
-module GIJoe.Utils where
-
-import GIJoe.Types
-
-import Prelude hiding (sum)
-import Numeric.Log
-import qualified Numeric.Sum as Sum
-import Numeric.SpecFunctions (digamma, logGammaL)
-import Data.Hashable
-import Data.List (foldl', sortBy, maximumBy, foldl1')
-
-import Data.HashMap.Strict (HashMap)
-import qualified Data.HashMap.Strict as HashMap
-
-
-fromDouble' :: Double' -> Double
-{-# INLINE fromDouble' #-}
-fromDouble' (Exp lnX) = exp lnX
-
-toDouble' :: Double -> Double'
-{-# INLINE toDouble' #-}
-toDouble' x = Exp (log x)
-
-{-# RULES "toDouble'/fromDouble'" forall x. toDouble' (fromDouble' x) = x #-}
-{-# RULES "fromDouble'/toDouble'" forall x. fromDouble' (toDouble' x) = x #-}
-
-fromListAccum :: (Eq a, Hashable a) => [(a, b)] -> (b -> b -> b) -> HashMap a b -> HashMap a b
-fromListAccum pairs f mp = foldl' (\m (r, c) -> HashMap.insertWith f r c m) mp pairs
-
--- Stats Functions --------
-klDirichlet :: [Double'] -> [Double'] -> Double'
-klDirichlet qs' ps' = toDouble' $ 
-                        logGammaL qs_tot - logGammaL ps_tot
-                        + Sum.sum Sum.kbn [logGammaL p - logGammaL q | (p, q) <- zip ps qs]
-                        + Sum.sum Sum.kbn [ (q - p) * (digamma q - digamma_q_tot) | (p, q) <- zip ps qs]
-  where !qs_tot = Sum.sum Sum.kbn qs
-        !ps_tot = Sum.sum Sum.kbn ps
-        !digamma_q_tot = digamma qs_tot
-        !qs = map fromDouble' qs'
-        !ps = map fromDouble' ps'
-
-entropyDirichlet :: [Double] -> Double
-entropyDirichlet !qs = logMultinomialBeta qs
-                       + (qs_tot - _K) * digamma qs_tot
-                       - Sum.sum Sum.kbn [(q - 1) * digamma q | q <- qs]
-  where !qs_tot = Sum.sum Sum.kbn qs
[…]
